### 2017-03-28 Doc/classicthesis/Chapters/Chapter01.tex
classic thesis init

%************************************************
\chapter{Introduction}\label{ch:introduction}
%************************************************

This bundle for \LaTeX\ has two goals:
\begin{enumerate}
    \item Provide students with an easy-to-use template for their
    Master's
    or PhD thesis. (Though it might also be used by other types of
    authors
    for reports, books, etc.)
    \item Provide a classic, high-quality typographic style that is
    inspired by \citeauthor{bringhurst:2002}'s ``\emph{The Elements of
    Typographic Style}'' \citep{bringhurst:2002}.
    \marginpar{\myTitle \myVersion}
\end{enumerate}
The bundle is configured to run with a \emph{full} 
MiK\TeX\ or \TeX Live\footnote{See the file \texttt{LISTOFFILES} for
needed packages. Furthermore, \texttt{classicthesis} 
works with most other distributions and, thus, with most systems 
\LaTeX\ is available for.} 
installation right away and, therefore, it uses only freely available 
fonts. (Minion fans can easily adjust the style to their needs.)

People interested only in the nice style and not the whole bundle can
now use the style stand-alone via the file \texttt{classicthesis.sty}.
This works now also with ``plain'' \LaTeX.

As of version 3.0, \texttt{classicthesis} can also be easily used with 
\mLyX\footnote{\url{http://www.lyx.org}} thanks to Nicholas Mariette 
and Ivo Pletikosić. The \mLyX\ version of this manual will contain
more information on the details.

This should enable anyone with a basic knowledge of \LaTeXe\ or \mLyX\ to
produce beautiful documents without too much effort. In the end, this
is my overall goal: more beautiful documents, especially theses, as I
am tired of seeing so many ugly ones.

The whole template and the used style is released under the
\acsfont{GNU} General Public License.

If you like the style then I would appreciate a postcard:
\begin{center}
 André Miede \\
 Detmolder Straße 32 \\
 31737 Rinteln \\
 Germany
\end{center}
The postcards I received so far are available at:
\begin{center}
 \url{http://postcards.miede.de}
\end{center}
\marginpar{A well-balanced line width improves the legibility of
the text. That's what typography is all about, right?}
So far, many theses, some books, and several other publications have 
been typeset successfully with it. If you are interested in some
typographic details behind it, enjoy Robert Bringhurst's wonderful book.
% \citep{bringhurst:2002}.

\paragraph{Important Note:} Some things of this style might look
unusual at first glance, many people feel so in the beginning.
However, all things are intentionally designed to be as they are,
especially these:
\begin{itemize}
    \item No bold fonts are used. Italics or spaced small caps do the
    job quite well.
    \item The size of the text body is intentionally shaped like it
    is. It supports both legibility and allows a reasonable amount of
    information to be on a page. And, no: the lines are not too short.
    \item The tables intentionally do not use vertical or double
    rules. See the documentation for the \texttt{booktabs} package for
    a nice discussion of this topic.\footnote{To be found online at 
    \url{http://mirror.ctan.org/macros/latex/contrib/booktabs/}.}
    \item And last but not least, to provide the reader with a way
    easier access to page numbers in the table of contents, the page
    numbers are right behind the titles. Yes, they are \emph{not}
    neatly aligned at the right side and they are \emph{not} connected
    with dots that help the eye to bridge a distance that is not
    necessary. If you are still not convinced: is your reader
    interested in the page number or does she want to sum the numbers
    up?
\end{itemize}
Therefore, please do not break the beauty of the style by changing
these things unless you really know what you are doing! Please.

\paragraph{Yet Another Important Note:} Since \texttt{classicthesis}'
first release in 2006, many things have changed in the \LaTeX\ world. 
Trying to keep up-to-date, \texttt{classicthesis} grew and evolved 
into many directions, trying to stay (some kind of) stable and be 
compatible with its port to \mLyX. However, there are still many 
remains from older times in the code, many dirty workarounds here and 
there, and several other things I am absolutely not proud of (for 
example my unwise combination of \acsfont{KOMA} and 
\texttt{titlesec} etc.).
\graffito{An outlook into the future of \texttt{classicthesis}.}

Currently, I am looking into how to completely re-design and 
re-implement \texttt{classicthesis} making it easier to maintain and 
to use. As a general idea, \texttt{classicthesis.sty} should be 
developed and distributed separately from the template bundle itself. 
Excellent spin-offs such as \texttt{arsclassica} could also be 
integrated (with permission by their authors) as format configurations. 
Also, current trends of \texttt{microtype}, \texttt{fontspec}, etc. 
should be included as well. As I am not really into deep 
\LaTeX\ programming, 
I will reach out to the \LaTeX\ community for their expertise and help.

\section{Organization}
A very important factor for successful thesis writing is the
organization of the material. This template suggests a structure as
the following:
\begin{itemize}
    \marginpar{You can use these margins for summaries of the text
    body\dots}
    \item\texttt{Chapters/} is where all the ``real'' content goes in
    separate files such as \texttt{Chapter01.tex} etc.
 %  \item\texttt{Examples/} is where you store all listings and other
 %  examples you want to use for your text.
    \item\texttt{FrontBackMatter/} is where all the stuff goes that
    surrounds the ``real'' content, such as the acknowledgments,
    dedication, etc.
    \item\texttt{gfx/} is where you put all the graphics you use in
    the thesis. Maybe they should be organized into subfolders
    depending on the chapter they are used in, if you have a lot of
    graphics.
    \item\texttt{Bibliography.bib}: the Bib\TeX\ database to organize
    all the references you might want to cite.
    \item\texttt{classicthesis.sty}: the style definition to get this
    awesome look and feel. Does not only work with this thesis template
    but also on its own (see folder \texttt{Examples}). Bonus: works
    with both \LaTeX\ and \textsc{pdf}\LaTeX\dots and \mLyX.
    \item\texttt{ClassicThesis.tcp} a \TeX nicCenter project file.
    Great tool and it's free!
    \item\texttt{ClassicThesis.tex}: the main file of your thesis
    where all gets bundled together.
    \item\texttt{classicthesis-config.tex}: a central place to load all 
    nifty packages that are used. %In there, you can also activate 
    %backrefs in order to have information in the bibliography about 
    %where a source was cited in the text (\ie, the page number).

    \emph{Make your changes and adjustments here.} This means that you  
    specify here the options you want to load \texttt{classicthesis.sty} 
    with. You also adjust the title of your thesis, your name, and all 
    similar information here. Refer to \autoref{sec:custom} for more 
    information.

        This had to change as of version 3.0 in order to enable an easy 
        transition from the ``basic'' style to \mLyX.

\end{itemize}
In total, this should get you started in no time.

\clearpage
\section{Style Options}\label{sec:options}
There are a couple of options for \texttt{classicthesis.sty} that
allow for a bit of freedom concerning the layout:
\marginpar{\dots or your supervisor might use the margins for some
    comments of her own while reading.}
\begin{itemize}
    \item General:
        \begin{itemize}
            \item\texttt{drafting}: prints the date and time at the bottom of
    each page, so you always know which version you are dealing with.
    Might come in handy not to give your Prof. that old draft.
        \end{itemize}

    \item Parts and Chapters:
        \begin{itemize}
            \item\texttt{parts}: if you use Part divisions for your document,
    you should choose this option. (Cannot be used together with 
    \texttt{nochapters}.)

            \item\texttt{nochapters}: allows to use the look-and-feel with 
    classes that do not use chapters, \eg, for articles. Automatically
    turns off a couple of other options: \texttt{eulerchapternumbers}, 
    \texttt{linedheaders}, \texttt{listsseparated}, and \texttt{parts}.

        \item\texttt{linedheaders}: changes the look of the chapter
        headings a bit by adding a horizontal line above the chapter
        title. The chapter number will also be moved to the top of the
        page, above the chapter title.

        \end{itemize}

  \item Typography:
        \begin{itemize}
            \item\texttt{eulerchapternumbers}: use figures from Hermann Zapf's
            Euler math font for the chapter numbers. By default, old style
            figures from the Palatino font are used.

            \item\texttt{beramono}: loads Bera Mono as typewriter font. 
            (Default setting is using the standard CM typewriter font.)

            \item\texttt{eulermath}: loads the awesome Euler fonts for math. 
            Pala\-tino is used as default font.

            \item\texttt{pdfspacing}: makes use of pdftex' letter spacing
            capabilities via the \texttt{microtype} package.\footnote{Use 
            \texttt{microtype}'s \texttt{DVIoutput} option to generate
            DVI with pdftex.} This fixes some serious issues regarding 
            math formul\ae\ etc. (\eg, ``\ss'') in headers.

            \item\texttt{minionprospacing}: uses the internal \texttt{textssc}
            command of the \texttt{MinionPro} package for letter spacing. This 
            automatically enables the \texttt{minionpro} option, overriding
            \texttt{pdfspacing}.

        \end{itemize}

    \item Table of Contents:
        \begin{itemize}
             \item\texttt{tocaligned}: aligns the whole table of contents on
            the left side. Some people like that, some don't.

            \item\texttt{dottedtoc}: sets pagenumbers flushed right in the 
            table of contents.

            \item\texttt{manychapters}: if you need more than nine chapters for 
        your document, you might not be happy with the spacing between the 
        chapter number and the chapter title in the Table of Contents. 
        This option allows for additional space in this context. 
        However, it does not look as ``perfect'' if you use
        \verb|\parts| for structuring your document.

        \end{itemize}

    \item Floats:
        \begin{itemize}
    \item\texttt{listings}: loads the \texttt{listings} package (if not 
    already done) and configures the List of Listings accordingly.

    \item\texttt{floatperchapter}: activates numbering per chapter for
    all floats such as figures, tables, and listings (if used).

        \item\texttt{subfig}(\texttt{ure}): is passed to the \texttt{tocloft} 
        package to enable compatibility with the \texttt{subfig}(\texttt{ure}) 
        package. Use this option if you want use \texttt{classicthesis} with the
        \texttt{subfig} package.

%    \item\texttt{listsseparated}: will add extra space between table
%    and figure entries of different chapters in the list of tables or
%    figures, respectively. % Deprecated as of version 2.9.
        \end{itemize}

%   \item\texttt{a5paper}: adjusts the page layout according to the
%    global \texttt{a5paper} option (\emph{experimental} feature).
%    \item\texttt{minionpro}: sets Robert Slimbach's Minion as the 
%    main font of the document. The textblock size is adjusted 
%    accordingly.

   \end{itemize}
The best way to figure these options out is to try the different
possibilities and see what you and your supervisor like best.

In order to make things easier, \texttt{classicthesis-config.tex} 
contains some useful commands that might help you.

\section{Customization}\label{sec:custom}
%(As of v3.0, the Classic Thesis Style for \LaTeX{} and \mLyX{} share
%the same two \texttt{.sty} files.)
This section will show you some hints how to adapt 
\texttt{classicthesis} to your needs.

The file \texttt{classicthesis.sty}
contains the core functionality of the style and in most cases will
be left intact, whereas the file \texttt{classic\-thesis-config.tex}
is used for some common user customizations.

The first customization you are about to make is to alter the document
title, author name, and other thesis details. In order to do this, replace
the data in the following lines of \texttt{classicthesis-config.tex:}%
\marginpar{Modifications in \texttt{classic\-thesis-config.tex}%
}

\begin{lstlisting}
    % **************************************************
    % 2. Personal data and user ad-hoc commands
    % **************************************************
    \newcommand{\myTitle}{A Classic Thesis Style\xspace} 
    \newcommand{\mySubtitle}{An Homage to...\xspace} 
\end{lstlisting}

Further customization can be made in \texttt{classicthesis-config.tex}
by choosing the options to \texttt{classicthesis.sty} 
(see~\autoref{sec:options}) in a line that looks like this:

\begin{lstlisting}
    \PassOptionsToPackage{eulerchapternumbers,drafting,listings,subfig,eulermath,parts}{classicthesis}
\end{lstlisting}

Many other customizations in \texttt{classicthesis-config.tex} are
possible, but you should be careful making changes there, since some
changes could cause errors.

Finally, changes can be made in the file \texttt{classicthesis.sty},%
\marginpar{Modifications in \texttt{classicthesis.sty}%
} although this is mostly not designed for user customization. The
main change that might be made here is the text-block size, for example,
to get longer lines of text.

\section{Issues}\label{sec:issues}
This section will list some information about problems using
\texttt{classic\-thesis} in general or using it with other packages.

Beta versions of \texttt{classicthesis} can be found at Bitbucket:
\begin{center}
    \url{https://bitbucket.org/amiede/classicthesis/}
\end{center}
There, you can also post serious bugs and problems you encounter.

\subsection*{Compatibility with the \texttt{glossaries} Package}
If you want to use the \texttt{glossaries} package, take care of loading it 
with the following options:
\begin{lstlisting}
    \usepackage[style=long,nolist]{glossaries}
\end{lstlisting}
Thanks to Sven Staehs for this information.

\subsection*{Compatibility with the (Spanish) \texttt{babel} Package}
Spanish languages need an extra option in order to work with this template:
\begin{lstlisting}
    \usepackage[spanish,es-lcroman]{babel}
\end{lstlisting}
Thanks to an unknown person for this information (via the issue reporting).

\paragraph{Further information for using \texttt{classicthesis} with Spanish (in addition to the above)}
In the file \texttt{ClassicThesis.tex} activate the language: 
\begin{lstlisting}
    \selectlanguage{spanish}
\end{lstlisting}

If there are issues changing \verb|\tablename|, \eg, using this:
\begin{lstlisting}
    \renewcommand{\tablename}{Tabla}
\end{lstlisting}

This can be solved by passing \texttt{es-tabla} parameter to \texttt{babel}:
\begin{lstlisting}
    \PassOptionsToPackage{es-tabla,spanish,es-lcroman,english}{babel}
    \usepackage{babel}
\end{lstlisting}

But it is also necessary to set \texttt{spanish} in the \verb|\documentclass|.

Thanks to Alvaro Jaramillo Duque for this information.

\subsection*{Compatibility with the \texttt{pdfsync} Package}
Using the \texttt{pdfsync} package leads to linebreaking problems with the \texttt{graffito} command. 
Thanks to Henrik Schumacher for this information.

\section{Future Work}
So far, this is a quite stable version that served a couple of people
well during their thesis time. However, some things are still not as
they should be. Proper documentation in the standard format is still
missing. In the long run, the style should probably be published
separately, with the template bundle being only an application of the
style. Alas, there is no time for that at the moment\dots it could be
a nice task for a small group of \LaTeX nicians.

Please do not send me email with questions concerning \LaTeX\ or the
template, as I do not have time for an answer. But if you have
comments, suggestions, or improvements for the style or the template
in general, do not hesitate to write them on that postcard of yours.

\section{Beyond a Thesis}
The layout of \texttt{classicthesis.sty} can be easily used without the
framework of this template. A few examples where it was used to typeset 
an article, a book or a curriculum vitae can be found in the folder 
\texttt{Examples}. The examples have been tested with  
\texttt{latex} and \texttt{pdflatex} and are easy to compile. To 
encourage you even more, PDFs built from the sources can be found in the 
same folder. 
%(It might be necessary to adjust the path to 
%\texttt{classicthesis.sty} and \texttt{Bibliography.bib} within the 
%examples.)

%\lstinputlisting[caption=An Article]%
    %{Examples/classicthesis-article.tex}
    %
%\lstinputlisting[caption=A Book]%
    %{Examples/classicthesis-book.tex}
%
%\lstinputlisting[caption=A Curriculum Vit\ae]%
    %{Examples/classicthesis-cv.tex}

\section{License}
\paragraph{GNU General Public License:} This program is free software;
you can redistribute it and/or modify
 it under the terms of the \acsfont{GNU} General Public License as
 published by
 the Free Software Foundation; either version 2 of the License, or
 (at your option) any later version.

 This program is distributed in the hope that it will be useful,
 but \emph{without any warranty}; without even the implied warranty of
 \emph{merchant\-ability} or \emph{fitness for a particular purpose}.
 See the
 \acsfont{GNU} General Public License for more details.

 You should have received a copy of the \acsfont{GNU} General
 Public License
 along with this program; see the file \texttt{COPYING}.  If not,
 write to
 the Free Software Foundation, Inc., 59 Temple Place - Suite 330,
 Boston, MA 02111-1307, USA.

%*****************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************

### 2017-03-28 Doc/classicthesis/Chapters/Chapter02.tex
classic thesis init

%*****************************************
\chapter{Examples}\label{ch:examples}
%*****************************************
%\setcounter{figure}{10}
% \NoCaseChange{Homo Sapiens}
Ei choro aeterno antiopam mea, labitur bonorum pri no 
\citeauthor{taleb:2012} \citep{taleb:2012}. His no decore
nemore graecis. In eos meis nominavi, liber soluta vim cu. Sea commune
suavitate interpretaris eu, vix eu libris efficiantur.

\section{A New Section}
Illo principalmente su nos. Non message \emph{occidental} angloromanic
da. Debitas effortio simplificate sia se, auxiliar summarios da que,
se avantiate publicationes via. Pan in terra summarios, capital
interlingua se que. Al via multo esser specimen, campo responder que
da. Le usate medical addresses pro, europa origine sanctificate nos
se.

Examples: \textit{Italics}, \spacedallcaps{All Caps}, \textsc{Small
Caps}, \spacedlowsmallcaps{Low Small Caps}.

Acronym testing: \ac{UML} -- \acs{UML} -- \acf{UML} -- \acp{UML} -- \ac{IoT}

\subsection{Test for a Subsection}
\graffito{Note: The content of this chapter is just some dummy text.
It is not a real language.}
Lorem ipsum at nusquam appellantur his, ut eos erant homero
concludaturque. Albucius appellantur deterruisset id eam, vivendum
partiendo dissentiet ei ius. Vis melius facilisis ea, sea id convenire
referrentur, takimata adolescens ex duo. Ei harum argumentum per. Eam
vidit exerci appetere ad, ut vel zzril intellegam interpretaris.

Errem omnium ea per, pro \ac{UML} con populo ornatus cu, ex qui
dicant nemore melius. No pri diam iriure euismod. Graecis eleifend
appellantur quo id. Id corpora inimicus nam, facer nonummy ne pro,
kasd repudiandae ei mei. Mea menandri mediocrem dissentiet cu, ex
nominati imperdiet nec, sea odio duis vocent ei. Tempor everti
appareat cu ius, ridens audiam an qui, aliquid admodum conceptam ne
qui. Vis ea melius nostrum, mel alienum euripidis eu.

Ei choro aeterno antiopam mea, labitur bonorum pri no. His no decore
nemore graecis. In eos meis nominavi, liber soluta vim cu.

\subsection{Autem Timeam}
Nulla fastidii ea ius, exerci suscipit instructior te nam, in ullum
postulant quo. Congue quaestio philosophia his at, sea odio autem
vulputate ex. Cu usu mucius iisque voluptua. Sit maiorum propriae at,
ea cum \ac{API} primis intellegat. Hinc cotidieque reprehendunt eu
nec. Autem timeam deleniti usu id, in nec nibh altera.

%Equidem detraxit cu nam, vix eu delenit periculis. Eos ut vero
%constituto, no vidit propriae complectitur sea. Diceret nonummy in
%has, no qui eligendi recteque consetetur. Mel eu dictas suscipiantur,
%et sed placerat oporteat. At ipsum electram mei, ad aeque atomorum
%mea.
%
%Ei solet nemore consectetuer nam. Ad eam porro impetus, te choro omnes
%evertitur mel. Molestie conclusionemque vel at.

\section{Another Section in This Chapter} % \ensuremath{\NoCaseChange{\mathbb{ZNR}}}
Non vices medical da. Se qui peano distinguer demonstrate, personas
internet in nos. Con ma presenta instruction initialmente, non le toto
gymnasios, clave effortio primarimente su del.\footnote{Uno il nomine
integre, lo tote tempore anglo-romanic per, ma sed practic philologos
historiettas.}

Sia ma sine svedese americas. Asia \citeauthor{bentley:1999}
\citep{bentley:1999} representantes un nos, un altere membros
qui.\footnote{De web nostre historia angloromanic.} Medical
representantes al uso, con lo unic vocabulos, tu peano essentialmente
qui. Lo malo laborava anteriormente uso.

\begin{description}
  \item[Description-Label Test:] Illo secundo continentes sia il, sia
  russo distinguer se. Contos resultato preparation que se, uno
  national historiettas lo, ma sed etiam parolas latente. Ma unic
  quales sia. Pan in patre altere summario, le pro latino resultato.
    \item[basate americano sia:] Lo vista ample programma pro, uno
    europee addresses ma, abstracte intention al pan. Nos duce infra
    publicava le. Es que historia encyclopedia, sed terra celos
    avantiate in. Su pro effortio appellate, o.
\end{description}

Tu uno veni americano sanctificate. Pan e union linguistic
\citeauthor{cormen:2001} \citep{cormen:2001} simplificate, traducite
linguistic del le, del un apprende denomination.

\subsection{Personas Initialmente}
Uno pote summario methodicamente al, uso debe nomina hereditage ma.
Iala rapide ha del, ma nos esser parlar. Maximo dictionario sed al.

\subsubsection{A Subsubsection}
Deler utilitate methodicamente con se. Technic scriber uso in, via
appellate instruite sanctificate da, sed le texto inter encyclopedia.
Ha iste americas que, qui ma tempore capital. \citeauthor{dueck:trio} \citep{dueck:trio}

\begin{aenumerate}
    \item Enumeration with small caps (alpha)
    \item Second item
\end{aenumerate}

\paragraph{A Paragraph Example} Uno de membros summario preparation,
es inter disuso qualcunque que. Del hodie philologos occidental al,
como publicate litteratura in web. Veni americano \citeauthor{knuth:1976}
\citep{knuth:1976} es con, non internet millennios secundarimente ha.
Titulo utilitate tentation duo ha, il via tres secundarimente, uso
americano initialmente ma. De duo deler personas initialmente. Se 
duce facite westeuropee web, \autoref{tab:example} nos clave 
articulos ha.

Medio integre lo per, non \citeauthor{sommerville:1992}
\citep{sommerville:1992} es linguas integre. Al web altere integre
periodicos, in nos hodie basate. Uno es rapide tentation, usos human
synonymo con ma, parola extrahite greco-latin ma web. Veni signo
rapide nos da.

%Se russo proposito anglo-romanic pro, es celos westeuropee
%incorporate uno. Il web unic periodicos. Que usate scientia ma, sed
%tres unidirectional al, asia personas duo de. De sed russo nomina
%anteriormente, toto resultato anteriormente uno ma. Non se signo
%romanic technologia, un medio millennios con.

%Major facto sia es, con o titulo maximo international. Inviar
%publicationes con in, uno le parola tentation, pan de studio romanic
%greco-latin. Tu duo titulo debitas latente, que vista programma ma.
%Non tote tres germano se, lo parola periodicos non.

\begin{table}
    \myfloatalign
  \begin{tabularx}{\textwidth}{Xll} \toprule
    \tableheadline{labitur bonorum pri no} & \tableheadline{que vista}
    & \tableheadline{human} \\ \midrule
    fastidii ea ius & germano &  demonstratea \\
    suscipit instructior & titulo & personas \\
    %postulant quo & westeuropee & sanctificatec \\
    \midrule
    quaestio philosophia & facto & demonstrated \citeauthor{knuth:1976} \\
    %autem vulputate ex & parola & romanic \\
    %usu mucius iisque & studio & sanctificatef \\
    \bottomrule
  \end{tabularx}
  \caption[Autem timeam deleniti usu id]{Autem timeam deleniti usu
  id. \citeauthor{knuth:1976}}  \label{tab:example}
\end{table}

\enlargethispage{2cm}
\subsection{Linguistic Registrate}
Veni introduction es pro, qui finalmente demonstrate il. E tamben
anglese programma uno. Sed le debitas demonstrate. Non russo existe o,
facite linguistic registrate se nos. Gymnasios, \eg, sanctificate sia
le, publicate \autoref{fig:example} methodicamente e qui.

Lo sed apprende instruite. Que altere responder su, pan ma, \ie, signo
studio. \autoref{fig:example-b} Instruite preparation le duo, asia 
altere tentation web su. Via unic facto rapide de, iste questiones 
methodicamente o uno, nos al.

\begin{figure}[bth]
        \myfloatalign
        \subfloat[Asia personas duo.]
        {\includegraphics[width=.45\linewidth]{gfx/example_1}} \quad
        \subfloat[Pan ma signo.]
        {\label{fig:example-b}%
         \includegraphics[width=.45\linewidth]{gfx/example_2}} \\
        \subfloat[Methodicamente o uno.]
        {\includegraphics[width=.45\linewidth]{gfx/example_3}} \quad
        \subfloat[Titulo debitas.]
        {\includegraphics[width=.45\linewidth]{gfx/example_4}}
        \caption[Tu duo titulo debitas latente]{Tu duo titulo debitas
        latente. \ac{DRY}}\label{fig:example}
\end{figure}

%*****************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************

### 2017-03-28 Doc/classicthesis/FrontBackmatter/Abstract.tex
classic thesis init

%*******************************************************
% Abstract
%*******************************************************
%\renewcommand{\abstractname}{Abstract}
\pdfbookmark[1]{Abstract}{Abstract}
\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Abstract}
Short summary of the contents in English\dots a great guide by 
Kent Beck how to write good abstracts can be found here:  
\begin{center}
\url{https://plg.uwaterloo.ca/~migod/research/beckOOPSLA.html}
\end{center}

\vfill

%\begin{otherlanguage}{ngerman}
%\pdfbookmark[1]{Zusammenfassung}{Zusammenfassung}
%\chapter*{Zusammenfassung}
%Kurze Zusammenfassung des Inhaltes in deutscher Sprache\dots 
%\end{otherlanguage}

\endgroup

\vfill

### 2017-03-28 Doc/classicthesis/FrontBackmatter/Bibliography.tex
classic thesis init

%********************************************************************
% Bibliography
%*******************************************************
% work-around to have small caps also here in the headline
\manualmark
\markboth{\spacedlowsmallcaps{\bibname}}{\spacedlowsmallcaps{\bibname}} % work-around to have small caps also
%\phantomsection 
\refstepcounter{dummy}
\addtocontents{toc}{\protect\vspace{\beforebibskip}} % to have the bib a bit from the rest in the toc
\addcontentsline{toc}{chapter}{\tocEntry{\bibname}}
\label{app:bibliography}
\printbibliography

### 2017-03-31 Doc/classicthesis/Chapters/Chapter01.tex
a few bib
udoc

@@
 %************************************************
+
+In recent years some new concepts have appeared in common people's vocabulary, like \textit{machine learning}, \textit{big data}, \textit{artificial intelligence}, \textit{automation}, etc., but there are two in particular that we are going to focus and try to combine: \ac{IoT} and Internet Security \& Privacy.
+
+The \ac{IoT} is a term with a wide range of interpretations \citep{Atzori20102787}, but a breaf definition could be the set of devices, mainly resource constrained, that are interconnected between them in order to achieve a goal.
+This includes from lampposts with proximity sensors that talk to each other in order to light up part of the street when a passerby walks by, to a sensor on your clothes that tells the washing machine how much detergent to use.
+
+Security \& Privacy, thanks to organizations like \href{https://wikileaks.org/}{WikiLeaks}, are now taken in consideration by any technology consumer, not only professionals.
+People are conscious about what their data can be used for, and demand more control over it.
+
+And IoT has proved to not address neither security nor privacy, with recent events like the Mirai botnet DDoS attack on October 2016, considered the biggest DDoS in history \citep{jeyanthi:2017}.
+
+\section{Motivation}
+
+\section{Challenges}
+
+\section{Goals}
+
+\section{Outline of this thesis}
 

### 2017-04-04 Doc/classicthesis/Chapters/Chapter01.tex
doc

@@
 Security \& Privacy, thanks to organizations like \href{https://wikileaks.org/}{WikiLeaks}, are now taken in consideration by any technology consumer, not only professionals.
-People are conscious about what their data can be used for, and demand more control over it.
+People are conscious about what their data can be used for, demanding more control over it.
 
-And IoT has proved to not address neither security nor privacy, with recent events like the Mirai botnet DDoS attack on October 2016, considered the biggest DDoS in history \citep{jeyanthi:2017}.
+And IoT has proved to not address neither security nor privacy, with recent events like the Mirai botnet DDoS attack on October 2016, considered the biggest DDoS in history \citep{jeyanthi:2017}, or like the multiple vulnerabilities affecting baby monitors \citep{rapid7babycam}.
+
+A recent approach to address the problem of privacy is the \textit{strong anonymity},  that conceals our personal details while letting us continue to operate online as a clearly defined individual \citep{stronganonymity}.
+One very promising way to achieve this is using \acp{ZKP}
+\marginpar{To read more about \acs{ZKP} aside the introduction done in this thesis, you can read my Mathematics thesis \citep{TODO}.}
+, cryptographic methods that allows to proof knowledge of data without disclosing it.
+Furthermore, IBM has been developing a cryptographic protocol suite for privacy-preserving authentication and transfer of certified attributes based on \ac{ZKP}, called Identity Mixer, Idemix for short \citep{idemixurl}.
+
+The goal of this project is to integrate Idemix with the \ac{IoT}.
+It will be done using the ABC4Trust's \ac{P2ABCE}, a framework that defines common architecture, policy language and data artifacts, but based on either IBM's Idemix or Microsoft's U-Prove \citep{p2abcurl}.
+This gives us a standardized language to exchange Idemix's messages between \ac{IoT} devices and usual PCs.
 
@@
 \section{Outline of this thesis}
-
-This bundle for \LaTeX\ has two goals:
-\begin{enumerate}
-    \item Provide students with an easy-to-use template for their
-    Master's
-    or PhD thesis. (Though it might also be used by other types of
-    authors
-    for reports, books, etc.)
-    \item Provide a classic, high-quality typographic style that is
-    inspired by \citeauthor{bringhurst:2002}'s ``\emph{The Elements of
-    Typographic Style}'' \citep{bringhurst:2002}.
-    \marginpar{\myTitle \myVersion}
-\end{enumerate}
-The bundle is configured to run with a \emph{full} 
-MiK\TeX\ or \TeX Live\footnote{See the file \texttt{LISTOFFILES} for
-needed packages. Furthermore, \texttt{classicthesis} 
-works with most other distributions and, thus, with most systems 
-\LaTeX\ is available for.} 
-installation right away and, therefore, it uses only freely available 
-fonts. (Minion fans can easily adjust the style to their needs.)
-
-People interested only in the nice style and not the whole bundle can
-now use the style stand-alone via the file \texttt{classicthesis.sty}.
-This works now also with ``plain'' \LaTeX.
-
-As of version 3.0, \texttt{classicthesis} can also be easily used with 
-\mLyX\footnote{\url{http://www.lyx.org}} thanks to Nicholas Mariette 
-and Ivo Pletikosić. The \mLyX\ version of this manual will contain
-more information on the details.
-
-This should enable anyone with a basic knowledge of \LaTeXe\ or \mLyX\ to
-produce beautiful documents without too much effort. In the end, this
-is my overall goal: more beautiful documents, especially theses, as I
-am tired of seeing so many ugly ones.
-
-The whole template and the used style is released under the
-\acsfont{GNU} General Public License.
-
-If you like the style then I would appreciate a postcard:
-\begin{center}
- André Miede \\
- Detmolder Straße 32 \\
- 31737 Rinteln \\
- Germany
-\end{center}
-The postcards I received so far are available at:
-\begin{center}
- \url{http://postcards.miede.de}
-\end{center}
-\marginpar{A well-balanced line width improves the legibility of
-the text. That's what typography is all about, right?}
-So far, many theses, some books, and several other publications have 
-been typeset successfully with it. If you are interested in some
-typographic details behind it, enjoy Robert Bringhurst's wonderful book.
-% \citep{bringhurst:2002}.
-
-\paragraph{Important Note:} Some things of this style might look
-unusual at first glance, many people feel so in the beginning.
-However, all things are intentionally designed to be as they are,
-especially these:
-\begin{itemize}
-    \item No bold fonts are used. Italics or spaced small caps do the
-    job quite well.
-    \item The size of the text body is intentionally shaped like it
-    is. It supports both legibility and allows a reasonable amount of
-    information to be on a page. And, no: the lines are not too short.
-    \item The tables intentionally do not use vertical or double
-    rules. See the documentation for the \texttt{booktabs} package for
-    a nice discussion of this topic.\footnote{To be found online at 
-    \url{http://mirror.ctan.org/macros/latex/contrib/booktabs/}.}
-    \item And last but not least, to provide the reader with a way
-    easier access to page numbers in the table of contents, the page
-    numbers are right behind the titles. Yes, they are \emph{not}
-    neatly aligned at the right side and they are \emph{not} connected
-    with dots that help the eye to bridge a distance that is not
-    necessary. If you are still not convinced: is your reader
-    interested in the page number or does she want to sum the numbers
-    up?
-\end{itemize}
-Therefore, please do not break the beauty of the style by changing
-these things unless you really know what you are doing! Please.
-
-\paragraph{Yet Another Important Note:} Since \texttt{classicthesis}'
-first release in 2006, many things have changed in the \LaTeX\ world. 
-Trying to keep up-to-date, \texttt{classicthesis} grew and evolved 
-into many directions, trying to stay (some kind of) stable and be 
-compatible with its port to \mLyX. However, there are still many 
-remains from older times in the code, many dirty workarounds here and 
-there, and several other things I am absolutely not proud of (for 
-example my unwise combination of \acsfont{KOMA} and 
-\texttt{titlesec} etc.).
-\graffito{An outlook into the future of \texttt{classicthesis}.}
-
-Currently, I am looking into how to completely re-design and 
-re-implement \texttt{classicthesis} making it easier to maintain and 
-to use. As a general idea, \texttt{classicthesis.sty} should be 
-developed and distributed separately from the template bundle itself. 
-Excellent spin-offs such as \texttt{arsclassica} could also be 
-integrated (with permission by their authors) as format configurations. 
-Also, current trends of \texttt{microtype}, \texttt{fontspec}, etc. 
-should be included as well. As I am not really into deep 
-\LaTeX\ programming, 
-I will reach out to the \LaTeX\ community for their expertise and help.
-
-\section{Organization}
-A very important factor for successful thesis writing is the
-organization of the material. This template suggests a structure as
-the following:
-\begin{itemize}
-    \marginpar{You can use these margins for summaries of the text
-    body\dots}
-    \item\texttt{Chapters/} is where all the ``real'' content goes in
-    separate files such as \texttt{Chapter01.tex} etc.
- %  \item\texttt{Examples/} is where you store all listings and other
- %  examples you want to use for your text.
-    \item\texttt{FrontBackMatter/} is where all the stuff goes that
-    surrounds the ``real'' content, such as the acknowledgments,
-    dedication, etc.
-    \item\texttt{gfx/} is where you put all the graphics you use in
-    the thesis. Maybe they should be organized into subfolders
-    depending on the chapter they are used in, if you have a lot of
-    graphics.
-    \item\texttt{Bibliography.bib}: the Bib\TeX\ database to organize
-    all the references you might want to cite.
-    \item\texttt{classicthesis.sty}: the style definition to get this
-    awesome look and feel. Does not only work with this thesis template
-    but also on its own (see folder \texttt{Examples}). Bonus: works
-    with both \LaTeX\ and \textsc{pdf}\LaTeX\dots and \mLyX.
-    \item\texttt{ClassicThesis.tcp} a \TeX nicCenter project file.
-    Great tool and it's free!
-    \item\texttt{ClassicThesis.tex}: the main file of your thesis
-    where all gets bundled together.
-    \item\texttt{classicthesis-config.tex}: a central place to load all 
-    nifty packages that are used. %In there, you can also activate 
-    %backrefs in order to have information in the bibliography about 
-    %where a source was cited in the text (\ie, the page number).
-
-    \emph{Make your changes and adjustments here.} This means that you  
-    specify here the options you want to load \texttt{classicthesis.sty} 
-    with. You also adjust the title of your thesis, your name, and all 
-    similar information here. Refer to \autoref{sec:custom} for more 
-    information.
-
-        This had to change as of version 3.0 in order to enable an easy 
-        transition from the ``basic'' style to \mLyX.
-
-\end{itemize}
-In total, this should get you started in no time.
-
-\clearpage
-\section{Style Options}\label{sec:options}
-There are a couple of options for \texttt{classicthesis.sty} that
-allow for a bit of freedom concerning the layout:
-\marginpar{\dots or your supervisor might use the margins for some
-    comments of her own while reading.}
-\begin{itemize}
-    \item General:
-        \begin{itemize}
-            \item\texttt{drafting}: prints the date and time at the bottom of
-    each page, so you always know which version you are dealing with.
-    Might come in handy not to give your Prof. that old draft.
-        \end{itemize}
-
-    \item Parts and Chapters:
-        \begin{itemize}
-            \item\texttt{parts}: if you use Part divisions for your document,
-    you should choose this option. (Cannot be used together with 
-    \texttt{nochapters}.)
-
-            \item\texttt{nochapters}: allows to use the look-and-feel with 
-    classes that do not use chapters, \eg, for articles. Automatically
-    turns off a couple of other options: \texttt{eulerchapternumbers}, 
-    \texttt{linedheaders}, \texttt{listsseparated}, and \texttt{parts}.
-
-        \item\texttt{linedheaders}: changes the look of the chapter
-        headings a bit by adding a horizontal line above the chapter
-        title. The chapter number will also be moved to the top of the
-        page, above the chapter title.
-
-        \end{itemize}
-
-  \item Typography:
-        \begin{itemize}
-            \item\texttt{eulerchapternumbers}: use figures from Hermann Zapf's
-            Euler math font for the chapter numbers. By default, old style
-            figures from the Palatino font are used.
-
-            \item\texttt{beramono}: loads Bera Mono as typewriter font. 
-            (Default setting is using the standard CM typewriter font.)
-
-            \item\texttt{eulermath}: loads the awesome Euler fonts for math. 
-            Pala\-tino is used as default font.
-
-            \item\texttt{pdfspacing}: makes use of pdftex' letter spacing
-            capabilities via the \texttt{microtype} package.\footnote{Use 
-            \texttt{microtype}'s \texttt{DVIoutput} option to generate
-            DVI with pdftex.} This fixes some serious issues regarding 
-            math formul\ae\ etc. (\eg, ``\ss'') in headers.
-
-            \item\texttt{minionprospacing}: uses the internal \texttt{textssc}
-            command of the \texttt{MinionPro} package for letter spacing. This 
-            automatically enables the \texttt{minionpro} option, overriding
-            \texttt{pdfspacing}.
-
-        \end{itemize}
-
-    \item Table of Contents:
-        \begin{itemize}
-             \item\texttt{tocaligned}: aligns the whole table of contents on
-            the left side. Some people like that, some don't.
-
-            \item\texttt{dottedtoc}: sets pagenumbers flushed right in the 
-            table of contents.
-
-            \item\texttt{manychapters}: if you need more than nine chapters for 
-        your document, you might not be happy with the spacing between the 
-        chapter number and the chapter title in the Table of Contents. 
-        This option allows for additional space in this context. 
-        However, it does not look as ``perfect'' if you use
-        \verb|\parts| for structuring your document.
-
-        \end{itemize}
-
-    \item Floats:
-        \begin{itemize}
-    \item\texttt{listings}: loads the \texttt{listings} package (if not 
-    already done) and configures the List of Listings accordingly.
-
-    \item\texttt{floatperchapter}: activates numbering per chapter for
-    all floats such as figures, tables, and listings (if used).
-
-        \item\texttt{subfig}(\texttt{ure}): is passed to the \texttt{tocloft} 
-        package to enable compatibility with the \texttt{subfig}(\texttt{ure}) 
-        package. Use this option if you want use \texttt{classicthesis} with the
-        \texttt{subfig} package.
-
-%    \item\texttt{listsseparated}: will add extra space between table
-%    and figure entries of different chapters in the list of tables or
-%    figures, respectively. % Deprecated as of version 2.9.
-        \end{itemize}
-
-%   \item\texttt{a5paper}: adjusts the page layout according to the
-%    global \texttt{a5paper} option (\emph{experimental} feature).
-%    \item\texttt{minionpro}: sets Robert Slimbach's Minion as the 
-%    main font of the document. The textblock size is adjusted 
-%    accordingly.
-
-   \end{itemize}
-The best way to figure these options out is to try the different
-possibilities and see what you and your supervisor like best.
-
-In order to make things easier, \texttt{classicthesis-config.tex} 
-contains some useful commands that might help you.
-
-\section{Customization}\label{sec:custom}
-%(As of v3.0, the Classic Thesis Style for \LaTeX{} and \mLyX{} share
-%the same two \texttt{.sty} files.)
-This section will show you some hints how to adapt 
-\texttt{classicthesis} to your needs.
-
-The file \texttt{classicthesis.sty}
-contains the core functionality of the style and in most cases will
-be left intact, whereas the file \texttt{classic\-thesis-config.tex}
-is used for some common user customizations.
-
-The first customization you are about to make is to alter the document
-title, author name, and other thesis details. In order to do this, replace
-the data in the following lines of \texttt{classicthesis-config.tex:}%
-\marginpar{Modifications in \texttt{classic\-thesis-config.tex}%
-}
-
-\begin{lstlisting}
-    % **************************************************
-    % 2. Personal data and user ad-hoc commands
-    % **************************************************
-    \newcommand{\myTitle}{A Classic Thesis Style\xspace} 
-    \newcommand{\mySubtitle}{An Homage to...\xspace} 
-\end{lstlisting}
-
-Further customization can be made in \texttt{classicthesis-config.tex}
-by choosing the options to \texttt{classicthesis.sty} 
-(see~\autoref{sec:options}) in a line that looks like this:
-
-\begin{lstlisting}
-    \PassOptionsToPackage{eulerchapternumbers,drafting,listings,subfig,eulermath,parts}{classicthesis}
-\end{lstlisting}
-
-Many other customizations in \texttt{classicthesis-config.tex} are
-possible, but you should be careful making changes there, since some
-changes could cause errors.
-
-Finally, changes can be made in the file \texttt{classicthesis.sty},%
-\marginpar{Modifications in \texttt{classicthesis.sty}%
-} although this is mostly not designed for user customization. The
-main change that might be made here is the text-block size, for example,
-to get longer lines of text.
-
-\section{Issues}\label{sec:issues}
-This section will list some information about problems using
-\texttt{classic\-thesis} in general or using it with other packages.
-
-Beta versions of \texttt{classicthesis} can be found at Bitbucket:
-\begin{center}
-    \url{https://bitbucket.org/amiede/classicthesis/}
-\end{center}
-There, you can also post serious bugs and problems you encounter.
-
-\subsection*{Compatibility with the \texttt{glossaries} Package}
-If you want to use the \texttt{glossaries} package, take care of loading it 
-with the following options:
-\begin{lstlisting}
-    \usepackage[style=long,nolist]{glossaries}
-\end{lstlisting}
-Thanks to Sven Staehs for this information.
-
-\subsection*{Compatibility with the (Spanish) \texttt{babel} Package}
-Spanish languages need an extra option in order to work with this template:
-\begin{lstlisting}
-    \usepackage[spanish,es-lcroman]{babel}
-\end{lstlisting}
-Thanks to an unknown person for this information (via the issue reporting).
-
-\paragraph{Further information for using \texttt{classicthesis} with Spanish (in addition to the above)}
-In the file \texttt{ClassicThesis.tex} activate the language: 
-\begin{lstlisting}
-    \selectlanguage{spanish}
-\end{lstlisting}
-
-If there are issues changing \verb|\tablename|, \eg, using this:
-\begin{lstlisting}
-    \renewcommand{\tablename}{Tabla}
-\end{lstlisting}
-
-This can be solved by passing \texttt{es-tabla} parameter to \texttt{babel}:
-\begin{lstlisting}
-    \PassOptionsToPackage{es-tabla,spanish,es-lcroman,english}{babel}
-    \usepackage{babel}
-\end{lstlisting}
-
-But it is also necessary to set \texttt{spanish} in the \verb|\documentclass|.
-
-Thanks to Alvaro Jaramillo Duque for this information.
-
-\subsection*{Compatibility with the \texttt{pdfsync} Package}
-Using the \texttt{pdfsync} package leads to linebreaking problems with the \texttt{graffito} command. 
-Thanks to Henrik Schumacher for this information.
-
-\section{Future Work}
-So far, this is a quite stable version that served a couple of people
-well during their thesis time. However, some things are still not as
-they should be. Proper documentation in the standard format is still
-missing. In the long run, the style should probably be published
-separately, with the template bundle being only an application of the
-style. Alas, there is no time for that at the moment\dots it could be
-a nice task for a small group of \LaTeX nicians.
-
-Please do not send me email with questions concerning \LaTeX\ or the
-template, as I do not have time for an answer. But if you have
-comments, suggestions, or improvements for the style or the template
-in general, do not hesitate to write them on that postcard of yours.
-
-\section{Beyond a Thesis}
-The layout of \texttt{classicthesis.sty} can be easily used without the
-framework of this template. A few examples where it was used to typeset 
-an article, a book or a curriculum vitae can be found in the folder 
-\texttt{Examples}. The examples have been tested with  
-\texttt{latex} and \texttt{pdflatex} and are easy to compile. To 
-encourage you even more, PDFs built from the sources can be found in the 
-same folder. 
-%(It might be necessary to adjust the path to 
-%\texttt{classicthesis.sty} and \texttt{Bibliography.bib} within the 
-%examples.)
-
-%\lstinputlisting[caption=An Article]%
-    %{Examples/classicthesis-article.tex}
-    %
-%\lstinputlisting[caption=A Book]%
-    %{Examples/classicthesis-book.tex}
-%
-%\lstinputlisting[caption=A Curriculum Vit\ae]%
-    %{Examples/classicthesis-cv.tex}
-
-\section{License}
-\paragraph{GNU General Public License:} This program is free software;
-you can redistribute it and/or modify
- it under the terms of the \acsfont{GNU} General Public License as
- published by
- the Free Software Foundation; either version 2 of the License, or
- (at your option) any later version.
-
- This program is distributed in the hope that it will be useful,
- but \emph{without any warranty}; without even the implied warranty of
- \emph{merchant\-ability} or \emph{fitness for a particular purpose}.
- See the
- \acsfont{GNU} General Public License for more details.
-
- You should have received a copy of the \acsfont{GNU} General
- Public License
- along with this program; see the file \texttt{COPYING}.  If not,
- write to
- the Free Software Foundation, Inc., 59 Temple Place - Suite 330,
- Boston, MA 02111-1307, USA.
-
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************

### 2017-04-04 Doc/classicthesis/Chapters/Chapter02.tex
doc

@@
 %*****************************************
-\chapter{Examples}\label{ch:examples}
+\chapter{State of the art}\label{ch:stateoftheart}
 %*****************************************
-%\setcounter{figure}{10}
-% \NoCaseChange{Homo Sapiens}
-Ei choro aeterno antiopam mea, labitur bonorum pri no 
-\citeauthor{taleb:2012} \citep{taleb:2012}. His no decore
-nemore graecis. In eos meis nominavi, liber soluta vim cu. Sea commune
-suavitate interpretaris eu, vix eu libris efficiantur.
 
-\section{A New Section}
-Illo principalmente su nos. Non message \emph{occidental} angloromanic
-da. Debitas effortio simplificate sia se, auxiliar summarios da que,
-se avantiate publicationes via. Pan in terra summarios, capital
-interlingua se que. Al via multo esser specimen, campo responder que
-da. Le usate medical addresses pro, europa origine sanctificate nos
-se.
+In this chapter we present the two dimensions of this project: the \ac{IoT} development state and an introduction to IBM's privacy-preserving solution, Idemix.
 
-Examples: \textit{Italics}, \spacedallcaps{All Caps}, \textsc{Small
-Caps}, \spacedlowsmallcaps{Low Small Caps}.
+\section{Internet of Things}
 
-Acronym testing: \ac{UML} -- \acs{UML} -- \acf{UML} -- \acp{UML} -- \ac{IoT}
-
-\subsection{Test for a Subsection}
-\graffito{Note: The content of this chapter is just some dummy text.
-It is not a real language.}
-Lorem ipsum at nusquam appellantur his, ut eos erant homero
-concludaturque. Albucius appellantur deterruisset id eam, vivendum
-partiendo dissentiet ei ius. Vis melius facilisis ea, sea id convenire
-referrentur, takimata adolescens ex duo. Ei harum argumentum per. Eam
-vidit exerci appetere ad, ut vel zzril intellegam interpretaris.
-
-Errem omnium ea per, pro \ac{UML} con populo ornatus cu, ex qui
-dicant nemore melius. No pri diam iriure euismod. Graecis eleifend
-appellantur quo id. Id corpora inimicus nam, facer nonummy ne pro,
-kasd repudiandae ei mei. Mea menandri mediocrem dissentiet cu, ex
-nominati imperdiet nec, sea odio duis vocent ei. Tempor everti
-appareat cu ius, ridens audiam an qui, aliquid admodum conceptam ne
-qui. Vis ea melius nostrum, mel alienum euripidis eu.
-
-Ei choro aeterno antiopam mea, labitur bonorum pri no. His no decore
-nemore graecis. In eos meis nominavi, liber soluta vim cu.
-
-\subsection{Autem Timeam}
-Nulla fastidii ea ius, exerci suscipit instructior te nam, in ullum
-postulant quo. Congue quaestio philosophia his at, sea odio autem
-vulputate ex. Cu usu mucius iisque voluptua. Sit maiorum propriae at,
-ea cum \ac{API} primis intellegat. Hinc cotidieque reprehendunt eu
-nec. Autem timeam deleniti usu id, in nec nibh altera.
-
-%Equidem detraxit cu nam, vix eu delenit periculis. Eos ut vero
-%constituto, no vidit propriae complectitur sea. Diceret nonummy in
-%has, no qui eligendi recteque consetetur. Mel eu dictas suscipiantur,
-%et sed placerat oporteat. At ipsum electram mei, ad aeque atomorum
-%mea.
-%
-%Ei solet nemore consectetuer nam. Ad eam porro impetus, te choro omnes
-%evertitur mel. Molestie conclusionemque vel at.
-
-\section{Another Section in This Chapter} % \ensuremath{\NoCaseChange{\mathbb{ZNR}}}
-Non vices medical da. Se qui peano distinguer demonstrate, personas
-internet in nos. Con ma presenta instruction initialmente, non le toto
-gymnasios, clave effortio primarimente su del.\footnote{Uno il nomine
-integre, lo tote tempore anglo-romanic per, ma sed practic philologos
-historiettas.}
-
-Sia ma sine svedese americas. Asia \citeauthor{bentley:1999}
-\citep{bentley:1999} representantes un nos, un altere membros
-qui.\footnote{De web nostre historia angloromanic.} Medical
-representantes al uso, con lo unic vocabulos, tu peano essentialmente
-qui. Lo malo laborava anteriormente uso.
-
-\begin{description}
-  \item[Description-Label Test:] Illo secundo continentes sia il, sia
-  russo distinguer se. Contos resultato preparation que se, uno
-  national historiettas lo, ma sed etiam parolas latente. Ma unic
-  quales sia. Pan in patre altere summario, le pro latino resultato.
-    \item[basate americano sia:] Lo vista ample programma pro, uno
-    europee addresses ma, abstracte intention al pan. Nos duce infra
-    publicava le. Es que historia encyclopedia, sed terra celos
-    avantiate in. Su pro effortio appellate, o.
-\end{description}
-
-Tu uno veni americano sanctificate. Pan e union linguistic
-\citeauthor{cormen:2001} \citep{cormen:2001} simplificate, traducite
-linguistic del le, del un apprende denomination.
-
-\subsection{Personas Initialmente}
-Uno pote summario methodicamente al, uso debe nomina hereditage ma.
-Iala rapide ha del, ma nos esser parlar. Maximo dictionario sed al.
-
-\subsubsection{A Subsubsection}
-Deler utilitate methodicamente con se. Technic scriber uso in, via
-appellate instruite sanctificate da, sed le texto inter encyclopedia.
-Ha iste americas que, qui ma tempore capital. \citeauthor{dueck:trio} \citep{dueck:trio}
-
-\begin{aenumerate}
-    \item Enumeration with small caps (alpha)
-    \item Second item
-\end{aenumerate}
-
-\paragraph{A Paragraph Example} Uno de membros summario preparation,
-es inter disuso qualcunque que. Del hodie philologos occidental al,
-como publicate litteratura in web. Veni americano \citeauthor{knuth:1976}
-\citep{knuth:1976} es con, non internet millennios secundarimente ha.
-Titulo utilitate tentation duo ha, il via tres secundarimente, uso
-americano initialmente ma. De duo deler personas initialmente. Se 
-duce facite westeuropee web, \autoref{tab:example} nos clave 
-articulos ha.
-
-Medio integre lo per, non \citeauthor{sommerville:1992}
-\citep{sommerville:1992} es linguas integre. Al web altere integre
-periodicos, in nos hodie basate. Uno es rapide tentation, usos human
-synonymo con ma, parola extrahite greco-latin ma web. Veni signo
-rapide nos da.
-
-%Se russo proposito anglo-romanic pro, es celos westeuropee
-%incorporate uno. Il web unic periodicos. Que usate scientia ma, sed
-%tres unidirectional al, asia personas duo de. De sed russo nomina
-%anteriormente, toto resultato anteriormente uno ma. Non se signo
-%romanic technologia, un medio millennios con.
-
-%Major facto sia es, con o titulo maximo international. Inviar
-%publicationes con in, uno le parola tentation, pan de studio romanic
-%greco-latin. Tu duo titulo debitas latente, que vista programma ma.
-%Non tote tres germano se, lo parola periodicos non.
-
-\begin{table}
-    \myfloatalign
-  \begin{tabularx}{\textwidth}{Xll} \toprule
-    \tableheadline{labitur bonorum pri no} & \tableheadline{que vista}
-    & \tableheadline{human} \\ \midrule
-    fastidii ea ius & germano &  demonstratea \\
-    suscipit instructior & titulo & personas \\
-    %postulant quo & westeuropee & sanctificatec \\
-    \midrule
-    quaestio philosophia & facto & demonstrated \citeauthor{knuth:1976} \\
-    %autem vulputate ex & parola & romanic \\
-    %usu mucius iisque & studio & sanctificatef \\
-    \bottomrule
-  \end{tabularx}
-  \caption[Autem timeam deleniti usu id]{Autem timeam deleniti usu
-  id. \citeauthor{knuth:1976}}  \label{tab:example}
-\end{table}
-
-\enlargethispage{2cm}
-\subsection{Linguistic Registrate}
-Veni introduction es pro, qui finalmente demonstrate il. E tamben
-anglese programma uno. Sed le debitas demonstrate. Non russo existe o,
-facite linguistic registrate se nos. Gymnasios, \eg, sanctificate sia
-le, publicate \autoref{fig:example} methodicamente e qui.
-
-Lo sed apprende instruite. Que altere responder su, pan ma, \ie, signo
-studio. \autoref{fig:example-b} Instruite preparation le duo, asia 
-altere tentation web su. Via unic facto rapide de, iste questiones 
-methodicamente o uno, nos al.
-
-\begin{figure}[bth]
-        \myfloatalign
-        \subfloat[Asia personas duo.]
-        {\includegraphics[width=.45\linewidth]{gfx/example_1}} \quad
-        \subfloat[Pan ma signo.]
-        {\label{fig:example-b}%
-         \includegraphics[width=.45\linewidth]{gfx/example_2}} \\
-        \subfloat[Methodicamente o uno.]
-        {\includegraphics[width=.45\linewidth]{gfx/example_3}} \quad
-        \subfloat[Titulo debitas.]
-        {\includegraphics[width=.45\linewidth]{gfx/example_4}}
-        \caption[Tu duo titulo debitas latente]{Tu duo titulo debitas
-        latente. \ac{DRY}}\label{fig:example}
-\end{figure}
-
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
+\section{Idemix}

### 2017-04-05 Doc/classicthesis/Chapters/Chapter01.tex
state of the art: IoT

@@
 One very promising way to achieve this is using \acp{ZKP}
-\marginpar{To read more about \acs{ZKP} aside the introduction done in this thesis, you can read my Mathematics thesis \citep{TODO}.}
+\marginpar{To read more about \acs{ZKP} aside the introduction done in this thesis, you can read my Mathematics thesis \citep{tfgmates}.}
 , cryptographic methods that allows to proof knowledge of data without disclosing it.

### 2017-04-05 Doc/classicthesis/Chapters/Chapter02.tex
state of the art: IoT

@@
 
+The development for Internet of Things depends heavily on each target device.
+We can differentiate two big groups: those with enough processing power to act like an usual computer, and those constrained devices that can't perform arbitrary tasks, sometimes called \textit{embedded}.
+
+We can consider in the first category powerful ARM devices like Raspberry Pi 3, with a 64-bit architecture, 4 CPU cores, 1GB of RAM, which can even compile its own binaries, run the \textit{Java Virtual Machine}, etc., working in practice like any other computer.
+These kind of devices do not present any major difficulty in terms of research.
+
+What we will consider to be a more \textit{pure} \ac{IoT} device will be the constrained ones, where it's not trivial to develop any algorithm and run it successfully.
+
+Very known devices fall into this category, like Arduino, powered by Atmel's AVR ATmega328 8-bit microprocessors, with 32KB of program flash memory, 2KB of SRAM, 16MHz of CPU \citep{ATmega328}.
+It seems clear that memory and computation power are a very big issue to deal with when developing to this devices.
+
+A step above in power we can find ESP8266, the most famous Espressif's microcontroller, with built-in WiFi antenna, a Tensilica 16bit RISC microcontroller at 80MHz, 50Kb of RAM, and 1MB flash memory \citep{ESP8266}.
+The possibility of direct WiFi connectivity is its best selling point, putting the \textit{Internet} in \acl{IoT}.
+
+In another level of power we have microcontrollers usually found in routers, but used in many other applications, like the On-Board Units (OBU) used in Vehicular ad hoc networks (VANET).
+Characteristics in this range vary around a single core 32-bit CPU, at some hundreds MHz, with tens or hundreds MB of RAM and flash memory, which places them near the first \ac{IoT} mentioned category.
+
+Although one can code in assembly language for these microcontrollers, there exist C compilers, and many frameworks to build firmware binaries: Arduino Core, Contiki, propietary SDKs, Mongoose OS, ThreadX OS (Real-Time OS), OpenWrt, LEDE, etc.
+Each firmware targets specific ranges of devices, depending on processing power and memory limitations.
+For example, Arduino and Contiki aim for microcontrollers like Atmel's ATmega and TI's MSP430, but can also be used in ESP8266, a more powerful microcontroller.
+
+In particular OpenWrt and LEDE (a fork of OpenWrt) are based on Linux, with optimized library binaries, providing many packages through \textit{opkg} \citep{opkg}.
+To compile C/C++ code, build the firmware or packages, a complete build system and cross-compiler toolchain can be installed in a x86 host, and using Makefiles select the target hardware \citep{openwrtbuildsystem}.
+
+Devices running OpenWRT and other Linux distros are in the limit between \ac{IoT} categories, but the need of a cross-compiler marks that they belong to the second category.
+
+%%%
+Starting a big project development for \ac{IoT} aiming the most constrained devices may not be a good idea.
+The lack of usual operative system tools (like POSIX), I/O, or even threads can make debugging a tedious task.
+With good programming practices one can start from the top and slowly end at the bottom with reliable code.
+
+For this reason the current \ac{PoC} is developed on LEDE, Linux Embedded Development Environment \citep{ledeproject}, using the Onion Omega2 development board, a Mediatek MT688 microcontroller \citep{MT7688} with a 32-bit MIPS 24KEc CPU at 580MHz, 64MB of RAM and 16MB of flash and built-in WiFi.
+This development board uses LEDE as its firmware, but its CPU is also listed as compatible with ThreadX OS \citep{THREADX}, a Real-Time Operative System for embedded devices.
+
+The \ac{PoC} will take advantage of the Linux system using files and sockets like in any other Linux desktop distribution, so we can focus on the project itself rather than the specific platform APIs for storage and connectivity.
+
+%%%
+
 \section{Idemix}

### 2017-04-06 Doc/classicthesis/Chapters/Chapter02.tex
vendehumos

@@
 \section{Idemix}
+
+\paragraph{TODO: Move to motivation}
+
+%%%%
+The problem of Internet privacy has been approached by securing the transmission channel (e.g.
+SSL/TLS) and the data stored in both ends (strict access policies, local encryption, etc.).
+In the end, the data exists in two entities, the owner of the data and the service provider.
+The owner is the most interested in securing his data, and can apply as many measures as he wants, but only on his side of the table.
+The service provider that stores the user data needs it to provide the service, and a successful attack would reveal many users data, aside from how many measures each one used to protect it.
+The case of PlayStation Network outage in 2011 \citep{PSN2011} affected 77 million accounts, with suspected credit card fraud, is an example of this kind of attacks.
+
+\hfil
+
+%%%%
+
+Other solutions are based on minimal disclosure.
+Standards like OAuth offer secure delegated access to the user information and when registering to a new service, the user can give a key to access only the data they want from another trusted service.
+This lets the service provider to work with the OAuth server, offering the same service as before without knowing as much data.
+
+But this is only minimizes how many services have our data. Our OAuth provider could be attacked, revealing all our data, or our service provider, revealing now less data.
+
+%%%%
+\hfill
+
+IBM proposes a step forward using \aclp{ZKP}:
+
+%%%%
+
+\begin{center}
+	\textit{If your personal data is never collected, it cannot be stolen}.
+\end{center}
+
+\paragraph{TODO: Poner descripción + formal de Idemix, luego P2ABCE, y meter que en P2ABCE había una implementación de sc basada en C}

### 2017-04-06 Doc/classicthesis/Chapters/Chapter03.tex
vendehumos

%************************************************
\chapter{Objectives and Methodology}\label{ch:objandmeth}
%************************************************

%TODO resumir organización del capítulos

\section{Project description}

\section{Working methodology}

\section{Development environment}

% C as pure as possible aiming to lower powerful devices without C++ support or std libraries
% CMake for portability and easy to config cross-compilation
% Java for P2ABCE
% Docker for building environments, both maven and LEDE

% HW: Omega2 with LEDE
%	  Raspberry Pi 3 with Raspbian

### 2017-04-07 Doc/classicthesis/Chapters/Chapter03.tex
dev env

@@
 
+\subsection{Hardware}
+
+To test our design in a realistic but easy  to work with deployment, we used as the \ac{IoT} device an Onion Omega2 development board, and as the delegation server a Raspberry Pi 3.
+
+\paragraph{Onion Omega2:}\marginpar{
+	\vskip0pt
+	\includegraphics[width=\linewidth]{gfx/Omega2}
+
+	Onion Omega2
+} A device that falls inside the second category of IoT, powerful enough to run a familiar Linux environment, where we can develop and debug the first \acp{PoC} without troubling ourselves with non-related to the project problems.
+
+Nonetheless, the Omega2 needs fine tunning to start operating, and basic knowledge of electronics is vital to make it work. The two main things to begin with Omega2 are:
+\begin{itemize}
+	\item A reliable 3.3V with a maximum of 800mA power supply (a USB with a step-down circuit works fine), with quality soldering and wires to avoid unwanted resistances.
+	\item A Serial to USB adapter wired to the TX and RX UART pins to use the Serial Terminal, in case WiFi doesn't work and no SSH is available, and because the connection is more reliable in case of wireless interferences.
+\end{itemize}
+
+\begin{table}[h]
+	\myfloatalign
+	\begin{tabularx}{0.75\textwidth}{ll} \toprule
+		MCU & Mediatek MT688 \citep{MT7688} \\
+		CPU & MIPS32 24KEc 580MHz \\
+		RAM & 64MB \\
+		Storage & 16MB \\
+		Firmware & LEDE (OpenWRT fork distro) \\
+		Connectivity & Wifi b/g/n \\
+		Power & 3.3V 300mA \\
+		\bottomrule
+	\end{tabularx}
+	\caption[Onion Omega 2 Specifications]{Onion Omega2 Specifications.}
+	\label{tab:Omega2Specs}
+\end{table}
+
+\paragraph{Raspberry Pi 3:}\marginpar{
+	\vskip0pt
+	\includegraphics[width=\linewidth]{gfx/RPi3}
+
+	Raspberry Pi 3
+} Another familiar environment, powerful enough to debug and hold the delegated \ac{P2ABCE} Java services (User, Verifier, \dots) of P2ABC with its 1GB of RAM, and with two network interfaces it's perfect to work as the gateway for the IoT devices to the Internet.
+
+Only a microSD with enough space to burn the binary with the OS is needed to plug\&play with the Raspberry Pi.
+We use Raspbian, a stable Debian based distro, recommended by the Raspberry Pi designers, and ready to use with the \ac{P2ABCE} compiled \textit{.jar} services.
+
+\begin{table}[h]
+	\myfloatalign
+	\begin{tabularx}{0.75\textwidth}{ll} \toprule
+		CPU & ARMv8 64bit quad-core 1.2GHz \\
+		RAM & 1GB \\
+		Storage & microSD \\
+		Firmware & Raspbian (Debian based distro) \\
+		Connectivity & Wifi n + Ethernet \\
+		Power & 5V 2A \\
+		\bottomrule
+	\end{tabularx}
+	\caption[Raspberry Pi 3 Specifications]{Raspberry Pi 3 Specifications.}
+	\label{tab:RPi3Specs}
+\end{table}
+
 % C as pure as possible aiming to lower powerful devices without C++ support or std libraries
+% LibGMP and OpenSSL for PoC
 % CMake for portability and easy to config cross-compilation
 % Java for P2ABCE
-% Docker for building environments, both maven and LEDE
+% Docker for building environments, both maven and LEDE SDK
 
-% HW: Omega2 with LEDE
-%	  Raspberry Pi 3 with Raspbian
+\subsection{Software}
+
+The development is divided between the \ac{IoT} device code and the \ac{P2ABCE} services.
+
+\hfil
+
+The P2ABCE is already written in Java, and few modifications will be done to the code in comparison to the existing project size, so we will continue using Java with the P2ABCE part.
+
+\hfil
+
+All IoT devices, have a C cross-compiler, some even a C++ cross-compiler.
+The worst case scenario is that one must write assembly code, and that code will be specific of that target, so we won't consider them.
+If now we focus on the most constrained devices, we could find out that some can't compile C++, some may not have many common libraries, and that the memory limitations they face make practically impossible to use dynamic memory, if we want to avoid very possible execution malfunctions.
+
+For that reason, the developed code for IoT devices must be written with standard C without using dynamic memory.
+
+\hfil
+
+A project with thousands of lines of code can't be written in a single file. And to manage the compilation of multiple files, organized in various directories, we will use CMake.
+
+CMake has many advantages over Makefiles:
+
+\begin{itemize}
+	\item Cross-platform. It works in many systems, and more specifically, in Linux it generates Makefiles.
+	\item Simpler syntax.
+Adding a library, files to compile, set definitions, etc.
+can be done with one CMake command, with rich documentation on the project's \href{https://cmake.org/cmake/help/latest/}{website}.
+	\item Cross-compilation.
+With only a \href{http://www.vtk.org/Wiki/CMake_Cross_Compiling#The_toolchain_file}{\small{CMAKE TOOLCHAIN}} file, CMake sets up automatically the cross-compilation with Makefiles and the C/C++ cross-compiler provided.
+\end{itemize}
+
+\hfil
+
+Although the ideal final code is pure C without external libraries or dynamic memory, the \ac{PoC} uses three major libraries:
+
+\begin{itemize}
+	\item OpenSSL: Provides reliable and tested AES and SHA256 implementations.
+	\item LibGMP: Provides multiprecision integer modular arithmetic.
+	\item cJSON: Provides a JSON parser to store and read the status in a human readable way.
+\end{itemize}
+
+These three libraries are used to implement different interfaces in the project, and C implementations of these interfaces should replace the external libraries in the future.
+
+\hfil
+
+Finally, we use Docker to deploy the compilation environments:
+
+\paragraph{P2ABCE environment} A container with OpenJDK 7 and Maven installed, with the Idemix maven plugins installed following the project \href{https://github.com/p2abcengine/p2abcengine/wiki/How-to-Build-the-ABC-Engine}{instructions} to use Idemix as the Engine for P2ABCE.
+
+\paragraph{LEDE SDK environment} A container with CMake and the LEDE SDK \citep{ledeproject} installed and configured for the Omega2 target.
+
+The Dockerfiles can be found in the Appendix.
+%TODO

### 2017-05-01 Doc/classicthesis/Chapters/Draft.tex
draft

%************************************************
\chapter{Draft}\label{ch:draft}
%************************************************

\section{APDU}

% Reader image, APDU protocol,

\section{P2ABCE existing code}

In the \ac{P2ABCE} repository \citep{p2abcurl} is available the project's code, divided in two solutions: a complete P2ABCE implementation in Java and a Multos Smartcard implementation as companion for the project.

The Java code is managed by a Maven project, structured using various known design patterns, but not of our interest. The structure we are actually interested in are the REST Services and their use of the Components classes, in which the smartcards are included.

P2ABCE project is based on the concept of smartcards to store the credentials, logical or physical. An interface is defined to communicate with these smartcards, and then different implementations allow to use either \textit{Software Smartcards} or \textit{Hardware Smartcards}.

The \textit{SoftwareSmartcard} class implements the interface in Java, suitable for tests and self-stored smartcards that any application using P2ABCE may need.

The \textit{HardwareSmartcard} class uses the standard APDU messages [TODO:ref] to interact with smartcards. P2ABCE defines for every method in the mentioned interface, the necessary APDU instructions, and currently relies on \textit{javax.smartcardio} abstract classes (implemented by Oracle in their JRE) to communicate with the smartcard reader. This way, it doesn't matter what manufacturer issues the smartcard, or if it's an Android device, if they support the APDU API, P2ABCE will work with them.

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/p2abceBasicUML}
	\end{center}
	\caption{Basic P2ABCE structure}
	\label{fig:p2abceBasicUML}
\end{figure}

As a PoC the P2ABCE project includes the ABC4Trust Card Lite, an implementation for ML3-36K-R1 Multos Smartcards. The code is written in C, but is very dependent on the Multos framework, aside from numerous bugs and bad coding habits. The code is structured in two files, \textit{main.h} and \textit{main.c}, with 557 and 5157 lines of code respectively. The \textit{main()} function is a 2635 lines long \textit{switch-case} with practically no comments.

At this stage, we have two options to implement our IoT device compatible with P2ABCE:

\begin{itemize}
	\item Implement in C the \textit{Smartcard} interface used by P2ABCE architecture, and using a communication protocol to call the methods from the machine running the P2ABCEngine.
	\item Present the IoT device as a hardware smartcard, using the APDU protocol (already defined, standard and with minimal overload). Providing a \textit{javax.smartcardio} ``IoT implementation'', the already existing \textit{HardwareSmartcard} class can work 
\end{itemize}

\section{Multos}

### 2017-05-02 Doc/classicthesis/Chapters/Draft.tex
draft APDU + Multos

@@
 
-\section{APDU}
+\section{Smart card APDU}
 
-% Reader image, APDU protocol,
+To communicate the smart cards and the reader an standardized protocol is specified in ISO/IEC 7816-4 \citep{APDUISO}.
 
-\section{P2ABCE existing code}
+The messages, also kown as \ac{APDU}, are divided in APDU Commands and APDU Responses.
+
+\textbf{APDU Commands} consist in 4 mandatory bytes (CLA, INS, P1, P2), and an optional payload.
+
+\begin{itemize}
+	\item CLA byte: Instruction class. Denotes if the command is interindustry standard or proprietary.
+	\item INS byte: Instruction code. Indicates the specific command.
+	\item P1, P2 bytes: Instruction parameters.
+	\item Lc, 0-3 bytes: Command data length.
+	\item Command data: Lc bytes of data.
+	\item Le, 0-3 bytes: Expected response data length.
+\end{itemize}
+
+This way, minimal number of bytes are needed to transmit commands to the smart card, allowing manufacturer's personalization of the smart card behavior and capabilities along with standard operations.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=0.55\linewidth]{gfx/APDUCommand}
+	\end{center}
+	\caption{APDU Command}
+	\label{fig:APDUCommand}
+\end{figure}
+
+\textbf{APDU Responses} are generated inside the smart card, always as an answer to an APDU Command. They consist on an optional payload and two mandatory status bytes.
+
+\begin{itemize}
+	\item Response data: At most Le bytes of data.
+	\item SW1-SW2 bytes: Status bytes. Encode the exit status of the instruction.
+\end{itemize}
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=0.55\linewidth]{gfx/APDUResponse}
+	\end{center}
+	\caption{APDU Response}
+	\label{fig:APDUResponse}
+\end{figure}
+
+\hfil
+
+The transmission protocol varies between different types of readers and smart cards (e.g.
+chip, contact-less), but what is common between every smart card interaction, is the \textit{APDU Command-Response Dialogue}.
+As long as the smart card has a power supply, it can maintain the dynamic memory in RAM between APDU Commands, what allows to do in two or more commmands complex operations, transmit more bytes than a single APDU can admit, etc.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=0.75\linewidth]{gfx/APDUdialog}
+	\end{center}
+	\caption{APDU Command-Response Dialogue}
+	\label{fig:APDUdialog}
+\end{figure}
+
+\section{P2ABCE}
 
@@
 The code is written in C, but is very dependent on the Multos framework, aside from numerous bugs and bad coding habits.
-The code is structured in two files, \textit{main.h} and \textit{main.c}, with 557 and 5157 lines of code respectively.
-The \textit{main()} function is a 2635 lines long \textit{switch-case} with practically no comments.
 
@@
 \begin{itemize}
-	\item Implement in C the \textit{Smartcard} interface used by P2ABCE architecture, and using a communication protocol to call the methods from the machine running the P2ABCEngine.
-	\item Present the IoT device as a hardware smartcard, using the APDU protocol (already defined, standard and with minimal overload).
-Providing a \textit{javax.smartcardio} ``IoT implementation'', the already existing \textit{HardwareSmartcard} class can work
+	\item Implement in C the \textit{Smartcard} interface used by P2ABCE architecture, and use some communication protocol to remotelly call the methods from the machine running the P2ABC Engine.
+	\item Present the IoT device as a hardware smart card, using the APDU protocol (already defined, standard and with minimal overload).
+Providing a \textit{javax.smartcardio} ``IoT implementation'' to communicate with the IoT device through a transmission protocol, the already existing \textit{HardwareSmartcard} class can work with the new \textit{IoTSmartcard} in the IoT device.
 \end{itemize}
 
-\section{Multos}
+%TODO : hablar más de los servicios REST
+
+\section{MULTOS}
+
+MULTOS is a multi-application smart card operative system, which provides a custom developing environment, with rich documentation \citep{MultosTechLib}.
+MULTOS smart cards communicate like any other smart card following the standard, but internally offers a very specific architecture, affecting the way one must code applications for it.
+
+In this section we will present the main characteristics of a MULTOS smart card that shaped the ABC4Trust Card Lite code and that we had to be aware of when adapting it to IoT devices.
+
+\paragraph{MULTOS programming languages} A native assembly language called MEL, C and, to a lesser extent, Java, are the available languages to code for MULTOS.
+In our case, ABC4T Card Lite uses MEL and C.
+
+\paragraph{MULTOS Workflow}
+
+Most of the transmission and communication process is done by MULTOS core, and it then selects, based on the CLA byte of the APDU, the application to load.
+This application is what most developers will only worry about, and is where their $main()$ function will start.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/multosWorkflow}
+	\end{center}
+	\caption{MULTOS workflow}
+	\label{fig:multosWorkflow}
+\end{figure}
+
+The application uses then the \textit{multos.h} file that declares multiple global variables already loaded with the needed data, including the APDU Command bytes.
+
+Now the developer is in charge of checking what instruction was sent and if the APDU has the expected ISO Case.
+If everything is ok, code what needs to be run and write in specific data space the APDU Response bytes, call \textit{multosExit()} and MULTOS will be in charge to send the APDU Response.
+
+In summary, our application starts with all data loaded and exits without worrying how to send the answer.
+A very comfortable workflow that we must now implement for our IoT device if we would want ABC4T Card Lite code to work.
+
+\paragraph{MULTOS Memory Layout}
+
+Each application in MULTOS has access to a specific memory layout, divided in different categories:
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/multosMemLay}
+	\end{center}
+	\caption{MULTOS Memory Layout}
+	\label{fig:multosMemLay}
+\end{figure}
+
+The Code Space is where the application code is stored.
+The Data Space is divided in Static memory, Public memory and Dynamic memory.
+
+\textbf{Static memory} are the application variables declared after the specific \textit{\#pragma melstatic} compiler directive.
+These variables are stored in the non-volatile EEPROM, and any write is assured to be saved because they are not loaded into RAM.
+
+\textbf{Public memory} can be seen as the input/output buffer for applications and MULTOS system.
+The APDU header appears at the top of Public, and command data at the bottom.
+The application writes then the APDU Response bytes in Public, at specific position (see \autoref{fig:multosPubMem}).
+To declare variables in this data space, the \textit{\#pragma melpublic} directive is available.
+
+\textbf{Dynamic memory} works like usual program memory, with Session Data storing global variables and the Stack.
+The limited size of RAM in IoT devices and smart cards makes the use of dynamic memory not advisable.
+The compiler directive to use Session Data is \textit{\#pragma melsession}.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/multosPubMem}
+	\end{center}
+	\caption{MULTOS Public Memory Data Map}
+	\label{fig:multosPubMem}
+\end{figure}
+
+\hfil
+
+With regards to primitive types, to avoid confusion with their sizes, MULTOS defines and uses the following data types specified in \autoref{fig:multosDataTypes}.
+It's important to notice that MULTOS is Big Endian
+\marginpar{\includegraphics[width=\linewidth]{gfx/Big-Endian}\\Big-Endian, Wikipedia}
+and when storing structures there is no padding between defined variables, unlike modern compilers that perform data structure alignment \citep{dataStructAlign} for performance.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/multosDataTypes}
+	\end{center}
+	\caption{MULTOS Data Types}
+	\label{fig:multosDataTypes}
+\end{figure}
+
+\paragraph{MULTOS Standard C-API}
+
+A collection of more than a hundred functions are provided for arithmetic, cryptography, memory and smart card operations.
+The \textit{multos.h} interface provides access to these functions, that ultimately call their respective primitive instructions in assembly code.
+The primitive instructions are but a system call with an operation code, loading data in the needed registers.
+Therefore,  no implementation for these tools is available, nor in C, nor in assembly code.
+
+Nevertheless, the C-API documentation \citep{MultosTechLib} provides rich description for each function.
+
+\section{ABC4Trust Card Lite}
+
+The code from the P2ABCE project for a MULTOS smart card has some good and bad points to have in count:
+
+The best asset of this code is that it's written in C aiming to a very constrained device, similar in computational power to typical IoT devices, and very limited memory.
+
+Some \textit{tricks} include using \textit{union} for data types that will be stored on the same data location, but at different moments (e.g.
+depending on APDU Command INS byte) or strong use of pointers and \textit{memcpy} calls to copy \textit{structs} with multiple variables as arrays of bytes.
+
+Among the many drawbacks we could mention the awful coding, the dependency on MULTOS framework and bugs.
+
+The code is structured in two files, \textit{main.h} and \textit{main.c}, with 557 and 5157 lines of code respectively.
+
+The file \textit{main.h} is mostly a reimplementation in assembly MEL of some MULTOS' C-API functions available in \textit{multos.h}.
+
+The \textit{main.c} consists on near 600 lines of variables and data structures declarations, followed by the \textit{main()} function, a 2635 lines long \textit{switch-case} with practically no comments, to conclude with the implementation of thirty functions called \textit{Subroutines} at the end of the file.
+
+This gives an idea of the problematic to maintain or even understand the code.
+But once one studies MULTOS framework and applies many refactoring techniques to ABC4T Card's code, this becomes the best starting point for the IoT version.
+
+\section{IoT smart card code}
+
+After many design decisions in the process to adapt the original ABC4T Card Lite code to pure C over an standard architecture machine, in this section we present the current PoC code, most important decisions, workflow, and future work.
+
+\section{IoT and P2ABCE}
+
+In this section we will define how an IoT device will be integrated in the P2ABCE environment, being totally compatible with any other system using P2ABCE, addressing the power and memory constrains IoT devices face.
+
+\hfil
+
+Our main goal is to make an IoT device capable to act as an User or Verifier in the P2ABCE architecture.
+For this, the device should manage complex XML schemas, perform cryptographic ZKP operations and communicate with the Verifier or Prover with which it's interacting.
+
+The communication is already solved by the \textit{Internet} capabilities of IoT devices.
+
+Our concern are the XML files exchanged and the cryptographic operations involving secret keys that must remain private to the IoT device.
+
+Here is where we look at the P2ABCE architecture more closely, and the concept of smart cards shows a solution for the second issue.
+Even in the case we were to implement all P2ABCE inside an IoT device, we would have to implement support for software smart cards, to keep the secret inside the IoT device.
+We will start building the house from the ground, implementing the smart card operations inside the IoT device.
+
+Now that in our design we have the smart card, we need to address the first point, XML data. %TODO

### 2017-05-03 Doc/classicthesis/Chapters/Draft.tex
draft - ABC4T card code + from P2ABCE to IoT

@@
 %************************************************
+
+% TODO
+% From Idemix To P2ABCE: reasons
 
@@
 
-The code from the P2ABCE project for a MULTOS smart card has some good and bad points to have in count:
+P2ABCE provides a smart card reference implementation, ABC4Trust Card Lite \citep{ABC4TCardLite}. It supports device-bound U-Prove and Idemix, and virtually any discrete logarithm based pABC system.
 
-The best asset of this code is that it's written in C aiming to a very constrained device, similar in computational power to typical IoT devices, and very limited memory.
+Version 1.2 is based on MULTOS ML3 cards, with approximately 64KB of EEPROM (non-volatile memory), 1KB of RAM and an Infineon SLE 78 microcontroller, a 16-bit based CPU aimed for chip cards.
 
-Some \textit{tricks} include using \textit{union} for data types that will be stored on the same data location, but at different moments (e.g.
-depending on APDU Command INS byte) or strong use of pointers and \textit{memcpy} calls to copy \textit{structs} with multiple variables as arrays of bytes.
+The card stores the user's private key $x$ and any \ac{BLOB} that the P2ABCE may need (like user's credentials).
+Then P2ABCE delegates the cryptographic operations on the smart card, that operates with $x$.
 
-Among the many drawbacks we could mention the awful coding, the dependency on MULTOS framework and bugs.
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/ABC4TCardLite}
+	\end{center}
+	\caption{ABC4Trust Card Lite}
+	\label{fig:ABC4TCardLite}
+\end{figure}
+
+The cryptographic operations performed by the smartcard are the modular exponentiation and addition that discrete logarithm ZKPs are based on.
+
+\hfil
+
+The code is available from the P2ABCE project and has some good and bad points to have in count:
+
+The best asset of this code is that it's written in C aiming to a very constrained device, similar in computational power to many IoT devices, and very limited memory.
+
+Some \textit{tricks} in the code include using \textit{union} data types for variables that will be stored on the same data location, but at different moments (e.g.
+depending on APDU Command INS byte), minimizing this way the use of RAM and making code readability better; or strong use of pointers and \textit{memcpy} calls to copy structures with multiple variables as arrays of bytes.
+
+Among the many drawbacks, we could highlight the awful coding, the strong dependency on MULTOS framework and some bugs found.
 
@@
 
-The file \textit{main.h} is mostly a reimplementation in assembly MEL of some MULTOS' C-API functions available in \textit{multos.h}.
+The file \textit{main.h} is mostly a reimplementation in assembly MEL of some MULTOS functionality already offered with latest \textit{multos.h}.
 
-The \textit{main.c} consists on near 600 lines of variables and data structures declarations, followed by the \textit{main()} function, a 2635 lines long \textit{switch-case} with practically no comments, to conclude with the implementation of thirty functions called \textit{Subroutines} at the end of the file.
+The \textit{main.c} consists on near 600 lines of variables and data structures declarations, followed by the \textit{main()} function, a 2635 lines long \textit{switch-case} with practically no comments, and to conclude, the implementation of thirty functions called \textit{Subroutines} at the end of the file.
 
 This gives an idea of the problematic to maintain or even understand the code.
-But once one studies MULTOS framework and applies many refactoring techniques to ABC4T Card's code, this becomes the best starting point for the IoT version.
-
-\section{IoT smart card code}
-
-After many design decisions in the process to adapt the original ABC4T Card Lite code to pure C over an standard architecture machine, in this section we present the current PoC code, most important decisions, workflow, and future work.
+But once one studies MULTOS framework in deep and applies many refactoring techniques to ABC4T Card's code, this becomes the best starting point for the IoT version.
 
@@
 
-Our concern are the XML files exchanged and the cryptographic operations involving secret keys that must remain private to the IoT device.
+Our concern are the data artifacts exchanged as XML and the cryptographic operations involving secret keys that must remain private to the IoT device.
 
@@
 
-Now that in our design we have the smart card, we need to address the first point, XML data. %TODO
+Now that in our design we have the smart card, we need to address the first point, XML squemas.
+We understand with \textit{XML squemas} both managing the XML syntax and the whole process to generate a proper answer, that is, basically, the crypto engine that relies on Idemix and the smart card to hold the secret information.
+
+Taking in consideration Idemix is currently provided in Java as a considerably big project, and version 3.0.36 still hasn't got official documentation, the task to port the Idemix crypto engine to IoT could be done in the future, if the chosen devices have enough storage and capacity.
+And even in that case, we would need to implement P2ABCE User and Verifier's architecture to completely free ourselves from an external P2ABCE machine.
+
+The final port would be so big, many IoT devices would fall out of the requisites to run it, failing in our initial objective.
+
+\hfil
+
+After this analysis of the P2ABCE, we conclude that the mandatory requisite for any IoT device that wants to work with this system and keep its private keys in it, is to implement the smart card functionality, and delegate the rest of the operations on a machine capable of running P2ABCE, until that functionality is implemented for the IoT device.
+
+\hfil
+
+This architecture is not really such an original idea.
+For example, IPv6 involves managing 128 bits per address and large headers, and many use cases only need IoT devices to communicate inside a private network.
+That's why many of them use 6LoWPAN to compress packets or use smaller address sizes.
+To communicate a 6LoWPAN with the Internet, a proxy is needed to transform 6LoWPAN packets to IPv6.
+
+\hfil
+
+Therefore, the IoT device now has a \textbf{duality} in its functions, because it is the User that starts any interaction with other systems, and it's also the smart card that P2ABCE delegates for crypto operations.
+It can also be seen as a \textbf{double delegation}.
+The IoT device delegates on the external P2ABCE server to manage the protocol, and the P2ABCE server delegates on the IoT, acting now as a smart card, for the cryptography.
+
+\hfil
+
+We find here two challenges: how to delegate from the IoT device to the P2ABCE delegation server, and how to transmit them and the APDUs to the IoT device.
+
+Currently P2ABCE offers various REST web services to run different roles in P2ABCE system: User Service, Issuer Service, Verification Service, etc.
+An application that integrates P2ABCE can make use of this services in the same machine or implement the functionality using the core components written in Java, the same ones the REST services use.
+Our PoC machine, the Omega2, can make REST calls easily, but other devices may use \ac{CoAP}, and in that case, the P2ABCE REST services should be rewritten to offer CoAP support.
+The commands needed to delegate to the P2ABCE delegation server will be the same to operate with the REST services.
+This way, the first issue is solved.
+
+The transmission of the messages will depend on the specific use case, capabilities and resources available.
+If the delegation server is connected, for example, through RS-232 serial with the IoT device, and physically inaccessible, in the same way an IoT device on its own would be protected, the communication is simple, and not far away from the Arduino Yun idea of combining two devices, one more powerful but to use only when needed.
+But if the IoT device and the delegation server are apart, or more than one IoT device delegates to it, then the transmission must be secured.
+They could use 6LoWPAN to talk to each other (the delegation service could be deployed in the proxy) and then secure communications with existing solutions, like with pre-shared symmetric keys, certificates for authentication and authorization, etc., it depends on each particular deployment.
+
+At the end of the day, this is all about usual security in IoT. Many other studies focus on this matter, so we will assume it can be done, and will focus on what's new, P2ABCE in IoT.
+
+\hfil
+
+% TODO: imagen con servicios desplegados/actores en cada máquina
+
+To sum up, our IoT device will act as User (Prover or Verifier) keeping its secrets in a software smart card.
+When it starts an interaction with other actor of the P2ABCE system (Issuer, Verifier, etc.), the IoT device will delegate with a remote call (using REST in our PoC) to a P2ABCE delegation server, attaching the XML file and the necessary information for the server to send the APDUs to the software smart card (in our PoC using TCP sockets, giving the IP and listening port).
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/DelegationProving}
+	\end{center}
+	\caption{IoT Delegation in P2ABCE for Proving.}
+	\label{fig:DelegationProving}
+\end{figure}
+
+\hfil
+
+This simple design keeps the benefit of a 100\% compatible P2ABCE deployment, and the integration of IoT devices to the P2ABCE ecosystem.
+
+In the future, more functionality currently delegated in the P2ABCE server can be implemented in the IoT device, if its resources allow it.
+For example, in a M2M environment, where an IoT device can act as User and Verifier, the verification consists on sending a Presentation Policy, and verifying the Presentation Token, which implies less logic than generating it as the User.
+Therefore, the implementation of the Verifier functionality would reduce significantly the need of a delegation server, but as we said, managing complex XML squemas is not something many IoT devices could do.
+
+\section{IoT Smart Card}
+
+After many design decisions in the process to adapt the original ABC4T Card Lite code to pure C, working over a more usual architecture machine, in this section, we present the current \ac{PoC} code, most important decisions, workflow execution, and future work.
+
+First, let's define what a \textit{more usual architecture} is.
+If we remember the MULTOS section, the framework gives an application a very specific memory layout and entry and output points of execution, that could be seen as a single process execution machine.
+Many IoT devices work like a computer, with multiple processes or threads, without pre-loaded data on startup (like the APDU MULTOS loads for the application), a non-volatile memory for data and code, maybe a basic file system in this memory, and RAM with the program's stack, heap, data and code.
+
+Our PoC is tested on a Linux system, and we will give instructions on how to adapt each part to work with other typical IoT systems.
+For example, other IoT devices may work like MULTOS and let access variables in non-volatile memory during execution, and in that case, the port should be changed according to these particularities.
+
+\hfil
+
+\paragraph{Project structure}
+
+% CMake for managing the cross-compilation, libraries, tests, etc.
+
+\hfil
+
+\paragraph{Core smart card}
+
+% TODO: diagrama con los ficheros de common/core
+
+After refactoring the original ABC4Trust Card's code, most of it fell in what we will call the core of the smart card.
+
+All types and variable definitions and the APDU handling is done in this code.
+
+However, the ABC4Trust's code depended on the MULTOS C-API for the input/output of data, modular arithmetic, and even AES128 and SHA256 cryptography.
+
+A characteristic of MULTOS C-API is that every function name starts with \textit{multos}, but as we said, the \textit{main.h} file implemented equivalent functions to some available in \textit{multos.h}.
+Our first step was to replace the \textit{main.h} functions for the standard ones in the C-API.
+Then, we implemented, following the C-API documentation, the functions from \textit{multos.h} (only the used ones) changing their names from \textit{multosFoo()} to \textit{mFoo()} for readability and emphasize that they were no longer from MULTOS.
+
+Future changes in the code may refactor it so there's no longer need for the MULTOS framework functions.
+
+\paragraph{Interfaces}
+
+To implement MULTOS functions, we needed to use some libraries, so we defined a facade to isolate the implementation of the core smart card from our different options, that could vary depending on the hardware or the system used by the IoT device.
+
+The use of a facade lets us, for example, change the implementation of modular arithmetic with a hardware optimized version, or a future more lightweight library, or our very own software implementation using the same data types that the core uses, minimizing the data transformations needed.
+
+Taking a step forward, we make the core smart card totally independent of any library, only on our interfaces. This means that typical standard C libraries, like \textit{stdlib}, or
+
+% TODO : diagrama con cada interfaz definida, y otro con cada implementación utilizada (indicar third party src)
+
+\paragraph{External libraries}
+
+% TODO : en el workflow se explican cómo se pasa del main al core, etc

### 2017-05-04 Doc/classicthesis/Chapters/Draft.tex
iot sc code components

@@
 
-\paragraph{Project structure}
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/P2ABCE-IoT-bw}
+	\end{center}
+	\caption{P2ABCE-IoT.}
+	\label{fig:P2ABCE-IoT}
+\end{figure}
 
-% CMake for managing the cross-compilation, libraries, tests, etc.
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/P2ABCE-IoT-color}
+	\end{center}
+	\caption{P2ABCE-IoT-color.}
+	\label{fig:P2ABCE-IoT-color}
+\end{figure}
+
+\subsection{Code structure}
+
+We divide the project in three different sections with the objective of enhancing maintainability, improving future changes, ports, fixes, etc.
+
+The first section is what could be called as the core of the smart card, the second one the interface for the tools the core need and may depend on the platform, and finally third party libraries, that in may be empty if the interfaces implementation doesn't need any.
+
+In our PoC we used CMake to manage the project, due to the cross-compilation tools, integration with multiple IDEs and tests.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/IoTCScomponents-color}
+	\end{center}
+	\caption{IoT Smart Card Code Structure.}
+	\label{fig:IoTCScomponents-color}
+\end{figure}
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/IoTCScomponents-bw}
+	\end{center}
+	\caption{IoT Smart Card Code Structure.}
+	\label{fig:IoTCScomponents-bw}
+\end{figure}
 
@@
 
-% TODO: diagrama con los ficheros de common/core
+The smart card logic lies in this section, the concepts of APDU Commands, what instructions are defined in P2ABCE smart cards and how to process them and generate proper APDU Responses.
+
+Changes in the APDU protocol for P2ABCE must be done here, independently of the target platform.
 
@@
 All types and variable definitions and the APDU handling is done in this code.
-
 However, the ABC4Trust's code depended on the MULTOS C-API for the input/output of data, modular arithmetic, and even AES128 and SHA256 cryptography.
@@
 
-Taking a step forward, we make the core smart card totally independent of any library, only on our interfaces. This means that typical standard C libraries, like \textit{stdlib}, or
+Taking a step forward, we make the core smart card totally independent of any library, only on our interfaces.
+This means that typical C libraries, like the standard \textit{stdlib.h}, or  \textit{string.h} are also behind the facade, in case some IoT system doesn't support them.
+The main goal we go after with this decision is that future developers adapting the code to a specific platform need to make no change to the \textit{core smart card}'s code, only to the interfaces implementation.
 
-% TODO : diagrama con cada interfaz definida, y otro con cada implementación utilizada (indicar third party src)
+\paragraph{External utilities}
 
-\paragraph{External libraries}
+If the IoT system offers well tested libraries that could aid in the interfaces implementation, or we simply found a pure C implementation for the task, these third party libraries belong to this section.
 
-% TODO : en el workflow se explican cómo se pasa del main al core, etc
+In our PoC, we use two ANSI C libraries, for base64 and JSON, and two shared libraries available in as packages in LEDE, GMPLib and OpenSSL.
+The last two libraries offer more functionality than we need, hence, it's desired in a production code to implement \textit{Modular Arithmetic} and \textit{Cryptography} interfaces with more lightweight alternatives.
+
+For example, Atmel's ATAES132A \citep{ATAES132A}
+\marginpar{\includegraphics[width=\linewidth]{gfx/atmel} \\ Atmel's cryptography chips.}
+offers a serial chip for secure key storage, AES128 execution and random number generation.
+Another serial chip like ESP8266 offers WiFi connectivity, typically used with Arduino, and can also perform AES encryption.
+For random number generation, a technique used with Contiki devices is to read from sensors aleatory data and use it as seed.
+All these alternatives depend on the target device, but are all valid.
+The \textit{interfaces} and \textit{external utilities} sections allow for a clean and fast port of the code.
+
+\subsection{PoC Workflow}
+
+% PoC BIOSC : TCP sockets, JSON serialization
+
+% PoC IoT javax.smartcardio : protocolo de 2 bytes + APDU
+
+% 1. Boot: deserialize sc status
+% 2. Listening: open tcp socket
+% 3. The IoT device delegates on the P2ABCE for some task and sends the IP and Port of the IoT Smart Card.
+% 4. An APDU Command arrives : read in 2 bytes the length of the apdu and

### 2017-05-05 Doc/classicthesis/Chapters/Draft.tex
draft describir workflow

@@
 
+\hfil
+
 \begin{figure}[bth]
@@
 \end{figure}
+
+Originally, the Lc and Le bytes had only 1 byte if present, restricting the payload data to be at most 256 bytes long.
+An extension to the protocol changed the meaning of a Lc or Le 0x00 byte (256 bytes long payload), so when the byte corresponding to Lc or Le started with 0x00, the next two bytes where the real length.
+With this, an Extended APDU lets up to 65536 bytes of data.
+
+The problem here is that not all readers or smart cards support extended APDUs.
+Originally, to send more than 256 bytes of data, a new APDU Command instruction was defined so the smart card stored the payload in a buffer, until other APDU Command made use of it.
+To send more data in an APDU Response, the status bytes were set to: SW1=0x61 and SW2 to the remaining bytes to send.
+Because a smart card can't send APDU Commands, the card terminal then sends a GET RESPONSE, a special APDU Command, with Le set to the number of bytes specified in SW2.
+Iterating this process, the smart card could send as many bytes as it wanted.
 
@@
 	\end{center}
-	\caption{P2ABCE-IoT.}
+	\caption{IoT in P2ABCE deployment diagram.}
 	\label{fig:P2ABCE-IoT}
@@
 	\end{center}
-	\caption{P2ABCE-IoT-color.}
+	\caption{IoT in P2ABCE deployment diagram.}
 	\label{fig:P2ABCE-IoT-color}
@@
 % 4. An APDU Command arrives : read in 2 bytes the length of the apdu and
+
+In a real deployment, we would offer a full library with an API to other processes to delegate on the P2ABC Engine, that automatizes the listening and security, that we presented in the deployment diagram \textbf{TODO:ref-img} as the \textit{P2ABCE IoT Toolkit}.
+
+But to test the IoT smart card, we use the \textit{curl} command for REST delegation calls and the BIOSC (Basic Input Output Smart Card) for APDU communication.
+
+To transmit the APDU messages in our PoC we use a simple protocol, consisting in one first byte for the instruction: receive an APDU Command or close the connection.
+In the first case, then we read two header bytes with the length of the APDU Command or Response to receive, followed by those APDU bytes.
+The message is sent over TCP for a reliable transmission (concept of session, packet retransmission, reordering, etc.).
+
+We lack of any security (authentication and authorization) that a real system should implement.
+It is vital to authenticate the delegation service, to authorize it to make APDU Commands, and the same with the IoT device, to prevent attacks.
+This belongs to \textit{usual} security, as we already said, and for that reason it's not in the PoC.
+
+\paragraph{P2ABCE REST Delegation}
+
+The only P2ABCE Service that needed to be modified was the User Service. We added the REST call
+
+\begin{center}
+	\textit{/initIoTsmartcard/{issuerParametersUid}?host=\&port=}
+\end{center}
+
+where we communicate the P2ABCE server that a IoT Smart Card is accessible via \textit{host} and \textit{port}.
+Then a new \textit{HardwareSmartcard} object is stored in the P2ABC Engine, but instead of the \textit{javax.smartcardio} Oracle's \textit{CardTerminal} implementation, we use our \textit{IoTsmartcardio} implementation for the \textit{HardwareSmartcard} constructor.
+
+The rest of P2ABCE code is unchanged and will work as if a real smart card was in use. \textit{IoTsmartcardio} implementation will transmit the APDUs through the TCP socket with the format mentioned.
+
+\hfil
+
+Our test consists on the tutorial available in P2ABCE's repository. From the Omega2's terminal we run \textit{curl} commands to send REST petitions. For example:
+
+\begin{verbatim}
+$ curl -X POST --header 'Content-Type: text/xml'
+  'http://localhost:9200/user/initIoTsmartcard/http%3A%2F%2
+  Fticketcompany%2FMyFavoriteSoccerTeam%2Fissuance%3Aidemix
+  ?host=192.168.3.1&port=8888'
+\end{verbatim}
+
+The IoT device will only manage XML as data files to exchange between the third party actor and the P2ABCE delegation service, without parsing them.
+
+After a REST call to the User Service, the P2ABC Engine will talk to the IoT smart card process.
+
+\paragraph{APDU transmission with BIOSC}
+
+% TODO : diagrama del software
+
+Before sending the REST message, the Omega2 must be listening in the specified port.
+At start up, BIOSC reads from a JSON file the status of the smart card (credentials in the BLOB, private keys, PIN and PUK codes, etc.), then opens the TCP socket and listens in a loop.
+
+When the delegation server sends an APDU Command following the simple protocol, BIOSC stores in a byte array the APDU Command bytes and calls the \textit{handle\_APDU} function in \textit{Core Smart Card}.
+
+The \textit{APDU\_handler} will parse the APDU bytes and check both CLA and INS bytes.
+For each possible INS the is a function that must always finish calling either \textit{mExit()} or \textit{output\_large\_data()}.
+
+The \textit{mExit} function is the reimplementation of MULTOS C-API exit functionality, that finishes the application execution returning to the MULTOS OS, that will send the APDU Response bytes to the terminal.
+This led in the original ABC4Trust Card's code to some tricky situations.
+Imagine a function that ends with an if-else expression.
+Many times we save a line writing something like
+
+\begin{verbatim}
+if(condition)
+    return a;
+return b;
+\end{verbatim}
+
+instead of the complete
+
+\begin{verbatim}
+if(condition)
+    return a;
+else
+    return b;
+\end{verbatim}
+
+This is a swift example of what the \textit{multosExit} function led to, because it is called in both the \textit{switch-case} processing the APDU instruction, and the \textit{Subroutines}.
+For example, the PIN code check was programmed in a way every reason to fail finished the application execution, with different error codes.
+Another example, checking if a credential is stored, if not, fail and exit inside the \textit{readBlob} subroutine.
+
+In our standard architecture machine, if we call a subroutine, it must end, and return the control to the handling function, and so the \textit{APDU\_handler} must return to the listening loop of BIOSC.
+
+Having all that in count, the \textit{mExit} function is not implemented as the documentation specifies, but only saves the smart card status serializing it in the JSON file again, and once it's saved, parses the APDU Response and sends it to the socket.
+
+The \textit{output\_large\_data} function is a tool used in MULTOS smart cards that don't support Extended APDUs.
+The APDU Command handler saves in a buffer the data to send, and its size.
+Then, \textit{output\_large\_data} will manage the buffer expecting future GET RESPONSE commands from the P2ABCE server.
+
+A future change to the project may include support for Extended APDUs, avoiding the use of multiple GET RESPONSE.
+To do this, the \textit{output\_large\_data} function and the \textit{IoTsmartcardio} implementation of \textit{javax.smartcardio} must support the feature.

### 2017-05-05 Code/util_sc/BIOSC.c

@@ -11,18 +11,17 @@
 /********************************************************************/
 
 
-#include <smartcard_utils_interface/sc_status_io_util.h>
 #include <smartcard_common/APDU_handler.h>
 #include <smartcard_common/global_vars.h>
 #include <smartcard_utils_interface/serialize_util.h>
 #include <stdio.h>
 #include <stdlib.h>
 #include <sys/socket.h>
-#include <p2abc_iot_toolkit_include/smartcard_utils_interface/error_codes.h>
+#include <p2abc_iot_toolkit_include/error_codes.h>
 #include <netinet/in.h>
 #include <unistd.h>
 #include <macrologger.h>
-#include <p2abc_iot_toolkit_include/smartcard_common/abc4T_types.h>
+#include <smartcard_utils_interface/APDU_IO_util.h>
 #include <p2abc_iot_toolkit_include/smartcard_utils_interface/crypto_util.h>
 #include <p2abc_iot_toolkit_include/smartcard_common/subroutines.h>
 
@@ -49,10 +48,9 @@ void receive_commands(){
             // Read APDU
             if(read(connfd, apdu_bytes, apdu_len)<=0)
                 return;
-            // Interpret the bytes
-            deserialize_APDU_command(apdu_bytes, apdu_len);
+
             // Handle the APDU Command
-            handle_APDU();
+            handle_APDU(apdu_bytes, apdu_len);
         }
     }
     LOG_DEBUG("Finish connection with FF command");

### 2017-05-05 Code/util_sc/p2abc_iot_toolkit_include/smartcard_common/APDU_handler.h

@@ -6,7 +6,7 @@ extern "C" {
 #endif
 
 
-void handle_APDU();
+void handle_APDU(BYTE *apdu_bytes, int length);
 
 #ifdef TEST_PATRAS_MODE       //TODO: remove if not necessary
 void handle_INS_GET_KX_AND_DEVICE_KEY(void);

### 2017-05-05 Code/util_sc/p2abc_iot_toolkit_include/smartcard_utils_interface/error_codes.h

@@ -1,22 +1,0 @@
-#ifndef __ERROR_CODES_H
-#define __ERROR_CODES_H
-
-#ifdef __cplusplus
-extern "C" {
-#endif
-
-#define ERROR_CANT_MALLOC                       11
-#define ERROR_SOCKET                            12
-#define ERROR_CANT_OPEN_FILE                    21
-#define ERROR_CANT_READ_FILE                    22
-#define ERROR_BAD_JSON_VALUE_LENGTH             23
-#define ERROR_BAD_JSON_ARRAY_LENGTH             24
-#define ERROR_APDU_TOO_SHORT                    31
-#define ERROR_APDU_TOO_LONG                     32
-#define ERROR_OPENSSL                           41
-
-#ifdef __cplusplus
-}
-#endif
-
-#endif // __ERROR_CODES_H

### 2017-05-05 Code/util_sc/p2abc_iot_toolkit_include/smartcard_utils_interface/mExit_util.h

@@ -1,30 +1,0 @@
-#ifndef __MEXIT_UTILH
-#define __MEXIT_UTILH
-
-#ifdef __cplusplus
-extern "C" {
-#endif
-
-/********************************************************************/
-/*************************** mExit Util  ****************************/
-/********************************************************************/
-
-
-
-
-void save_status();
-
-
-/**
- * Takes the values in the APDU variables and public data
- * and outputs it as APDU response.
- */
-void output_apdu_response();
-
-
-
-#ifdef __cplusplus
-}
-#endif
-
-#endif //__MEXIT_UTIL_H

### 2017-05-05 Code/util_sc/p2abc_iot_toolkit_include/smartcard_utils_interface/sc_status_io_util.h

@@ -1,30 +1,0 @@
-#ifndef __SC_STATUS_IO_UTIL_H
-#define __SC_STATUS_IO_UTIL_H
-
-#ifdef __cplusplus
-extern "C" {
-#endif
-
-
-/********************************************************************/
-/************************ SC STATUS IO UTIL *************************/
-/********************************************************************/
-
-/********************************************************************/
-/** Initializes the static data of the smartcard.                   */
-/********************************************************************/
-/** THIS INTERFACE MAY BE CHANGED FOR TARGET PLATFORM               */
-/** AND INITIALIZATION/STORING METHOD                               */
-/********************************************************************/
-
-// Using Json Files
-char * json_file;
-void init_smartcard_from_json_file(char * smartcard_status_file);
-void save_smartcard_to_json_file(char * json_file);
-
-
-#ifdef __cplusplus
-}
-#endif
-
-#endif //__SC_STATUS_IO_UTIL_H

### 2017-05-05 Code/util_sc/p2abc_iot_toolkit_include/smartcard_utils_interface/serialize_util.h

@@ -19,6 +19,10 @@ extern "C" {
 /** CODE AND THE SERIALIZATION METHOD.                              */
 /********************************************************************/
 
+
+void save_status();
+
+
 /**
  * Current implementation:
  * Allocates memory and returns the pointer to the
@@ -35,11 +39,10 @@ char* serialize_smartcard_status();
  */
 void deserialize_smartcard_status(unsigned char * ascii);
 
-void serialize_APDU_response(unsigned char* ap_r, int * buf_len);
-
-void deserialize_APDU_command(unsigned char * apdu_bytes, int length);
-
-
+// Using Json Files
+char * json_file;
+void init_smartcard_from_json_file(char * smartcard_status_file);
+void save_smartcard_to_json_file(char * json_file);
 
 
 #ifdef __cplusplus

### 2017-05-05 Code/util_sc/p2abc_iot_toolkit_src/smartcard_common/APDU_handler.c

@@ -6,9 +6,12 @@
 #include <smartcard_common/subroutines.h>
 #include <smartcard_common/defs_ins.h>
 #include <macrologger.h>
-#include <smartcard_common/abc4T_types.h>
+#include <smartcard_utils_interface/APDU_IO_util.h>
 
-void handle_APDU() {
+void handle_APDU(BYTE *apdu_bytes, int length) {
+
+    // Interpret the bytes
+    parse_APDU_command(apdu_bytes, length);
 
     /* Check class in APDU. */
     if (CLA != ABC_CLA && INS != INS_GET_RESPONSE){

### 2017-05-05 Code/util_sc/p2abc_iot_toolkit_src/smartcard_common/m_adapted_API.c

@@ -4,7 +4,8 @@
 #include <smartcard_utils_interface/crypto_util.h>
 #include <smartcard_utils_interface/arithmetic_util.h>
 #include <smartcard_common/global_vars.h>
-#include <smartcard_utils_interface/mExit_util.h>
+#include <smartcard_utils_interface/APDU_IO_util.h>
+#include <smartcard_utils_interface/serialize_util.h>
 
 
 void mAESECBDecipher (BYTE *cipherText, BYTE *plainText, BYTE keyLength, BYTE *key){

### 2017-05-05 Code/util_sc/p2abc_iot_toolkit_src/smartcard_utils_interface/mExit_util.c

@@ -1,31 +1,0 @@
-#include <smartcard_utils_interface/mExit_util.h>
-///////////////////////////////////////////////////////////////////////
-/********************************************************************/
-
-/*   Implementation that writes to connfd the APDU bytes            */
-/*   and delegates to sc_status_io to save the sc status to         */
-/*   the same json file from start                                  */
-/********************************************************************/
-#include <smartcard_common/defs_types.h>
-#include <smartcard_utils_interface/sc_status_io_util.h>
-#include <smartcard_common/APDU_types.h>
-#include <smartcard_common/global_vars.h>
-#include <unistd.h>
-#include <macrologger.h>
-#include <smartcard_common/subroutines.h>
-
-void output_apdu_response(){
-    WORD APDUlength = La + 2;
-    APDUlength = rectifyWordEndianness(APDUlength);
-    write(connfd, &APDUlength, 2);
-    write(connfd, apdu_data.dataout, La);
-    write(connfd, &SW1, 1);
-    write(connfd, &SW2, 1);
-    fsync(connfd);
-    LOG_BYTES(apdu_data.dataout, La, "APDU Response");
-    LOG_DEBUG("SW = %02X%02X", SW1, SW2);
-}
-
-void save_status(){
-    save_smartcard_to_json_file(json_file);
-}

### 2017-05-05 Code/util_sc/p2abc_iot_toolkit_src/smartcard_utils_interface/sc_status_io_util.c

@@ -1,41 +1,0 @@
-#include <smartcard_utils_interface/sc_status_io_util.h>
-#include <smartcard_utils_interface/serialize_util.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <smartcard_utils_interface/error_codes.h>
-
-
-void init_smartcard_from_json_file(char * json_file){
-    // 1. Open smartcard file
-    FILE * f = fopen(json_file, "rb");
-    if(f==NULL) exit(ERROR_CANT_OPEN_FILE);
-    // 2. Save to array of char
-    fseek(f, 0, SEEK_END);
-    long fsize = ftell(f);
-    fseek(f, 0, SEEK_SET);  //same as rewind(f);
-    char * string = (char*)malloc(fsize + 1);
-    if(string==NULL) exit(ERROR_CANT_MALLOC);
-    if(fread(string, 1, fsize, f) < fsize) exit(ERROR_CANT_READ_FILE);
-    string[fsize] = 0;
-    // 3. Close smartcard file
-    fclose(f);
-    // 4. Call serialize_util.h deserialize_smartcart_status
-    deserialize_smartcard_status(string);
-    // 5. Free allocated string
-    free(string);
-}
-
-void save_smartcard_to_json_file(char * json_file){
-    //TODO save write
-    // 1. Open smartcard file
-    FILE * f = fopen(json_file, "w+");
-    if(f==NULL) exit(ERROR_CANT_OPEN_FILE);
-    // 2. Save to array of char
-    char * json_string = serialize_smartcard_status();
-    if(json_string==NULL) exit(ERROR_CANT_MALLOC);
-    fputs(json_string, f);
-    // 3. Close smartcard file
-    fclose(f);
-    // 4. Free allocated string
-    free(json_string);
-}

### 2017-05-05 Code/util_sc/p2abc_iot_toolkit_src/smartcard_utils_interface/serialize_util.c

@@ -7,10 +7,55 @@
 #include <smartcard_external_utilities/base64.h>
 #include <smartcard_external_utilities/cJSON.h>
 #include <smartcard_utils_interface/system_funcs.h>
-#include <smartcard_utils_interface/error_codes.h>
+#include <error_codes.h>
 #include <string.h>
-#include <macrologger.h>
-#include <smartcard_common/APDU_types.h>
+
+
+void save_status(){
+    save_smartcard_to_json_file(json_file);
+}
+
+
+// USE JSON FILES
+
+void init_smartcard_from_json_file(char * json_file){
+    // 1. Open smartcard file
+    FILE * f = fopen(json_file, "rb");
+    if(f==NULL) exit(ERROR_CANT_OPEN_FILE);
+    // 2. Save to array of char
+    fseek(f, 0, SEEK_END);
+    long fsize = ftell(f);
+    fseek(f, 0, SEEK_SET);  //same as rewind(f);
+    char * string = (char*)malloc(fsize + 1);
+    if(string==NULL) exit(ERROR_CANT_MALLOC);
+    if(fread(string, 1, fsize, f) < fsize) exit(ERROR_CANT_READ_FILE);
+    string[fsize] = 0;
+    // 3. Close smartcard file
+    fclose(f);
+    // 4. Call serialize_util.h deserialize_smartcart_status
+    deserialize_smartcard_status(string);
+    // 5. Free allocated string
+    free(string);
+}
+
+void save_smartcard_to_json_file(char * json_file){
+    //TODO secure write : simulate atomic writing
+    // 1. Open smartcard file
+    FILE * f = fopen(json_file, "w+");
+    if(f==NULL) exit(ERROR_CANT_OPEN_FILE);
+    // 2. Save to array of char
+    char * json_string = serialize_smartcard_status();
+    if(json_string==NULL) exit(ERROR_CANT_MALLOC);
+    fputs(json_string, f);
+    // 3. Close smartcard file
+    fclose(f);
+    // 4. Free allocated string
+    free(json_string);
+}
+
+
+
+// AUX FUNCTIONS FOR JSON SERIALIZATION
 
 void serialize_BYTE(cJSON * object, char * name, BYTE value){
     char temp_hex_string[3];
@@ -58,6 +103,12 @@ void deserialize_BYTE_ARRAY(cJSON * object, char * name, BYTE * val, WORD exp_le
         free(temp_unbase64);
[…]

### 2017-05-05 Code/util_sc_tests/apdu_tests/handle_APDU_tests.cpp

@@ -4,8 +4,7 @@
 #include <smartcard_common/APDU_handler.h>
 #include <smartcard_common/defs_ins.h>
 #include <smartcard_common/subroutines.h>
-#include <smartcard_utils_interface/sc_status_io_util.h>
-
+#include <smartcard_utils_interface/serialize_util.h>
 
 TEST(test_INS_PIN_TRIALS_LEFT, test1) {
     BYTE pin1[] = {0x12, 0x34, 0x56, 0x78};

### 2017-05-09 Code/util_sc/p2abc_iot_toolkit_include/smartcard_common/APDU_handler.h
fix

@@ -1,6 +1,8 @@
 #ifndef __APDU_HANDLER_H
 #define __APDU_HANDLER_H
 
+#include "defs_types.h"
+
 #ifdef __cplusplus
 extern "C" {
 #endif

### 2017-05-11 Doc/benchmarks/testNet.py
bench

@@ -1,0 +1,125 @@
+#!/usr/bin/python
+import socket
+import binascii
+import struct
+
+
+
+def sendCommand(comm):
+    global s
+    l = len(comm)
+    c = bytearray([0x01])
+    c += struct.pack("!H", l)
+    c += comm
+    s.send(c)
+
+
+def getResponse(l):
+    global s
+    b = bytearray(s.recv(l))
+    #print(binascii.hexlify(b))
+
+
+# create a socket object
+s = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+
+
+# get local machine name
+#host = socket.gethostname()
+#host = "192.168.33.96"
+host = "192.168.1.49"
+
+port = 8888
+
+# connection to hostname on the port.
+s.connect((host, port))
+
+#
+# '''
+# 01          :   command recvAPDU
+# 0004        :   length of APDU
+# BC060000    :   APDU bytes
+#                 INS_SET_WORKING_MODE
+# '''
+# test1 = bytearray([0xbc, 0x06, 0x00, 0x00])
+# sendCommand(test1)
+# getResponse()
+#
+#
+# '''
+# 01          :   command recvAPDU
+# 0005        :   length of APDU
+# BC840000ff  :   APDU bytes
+#                 INS_GET_INFO
+# '''
+# test2 = bytearray([0xbc, 0x84, 0x00, 0x00, 0xff])
+#
+# sendCommand(test2)
+# getResponse()
+#
+
+from time import time
+start = time()
+for i in range(1000):
+    comm = bytearray([0xBC, 0x8E, 0x00, 0x00, 0x00])
+    sendCommand(comm)
+    getResponse(len(comm))
+
+
+
+    comm = bytearray([0xBC, 0x02, 0x00, 0x00, 0x01])
+    sendCommand(comm)
[…]

### 2017-05-12 Doc/classicthesis/Chapters/Draft.tex
draft bench

@@
 To do this, the \textit{output\_large\_data} function and the \textit{IoTsmartcardio} implementation of \textit{javax.smartcardio} must support the feature.
+
+\section{Performance Evaluation}
+
+\subsection{Testbed description}
+
+% Hardware, network,
+
+\subsubsection{P2ABCE setting}
+
+To test the correct execution of the IoT smart card, we will use the ABC system from the tutorial in the P2ABCE Wiki \citep{p2abcurlwiki}.
+It is based on a soccer club which wishes to issue VIP-tickets to a soccer match.
+The VIP-member number in the ticket is inspectable for a lottery, ie.
+after the game, a random presentation token is inspected and the winning member is notified.
+
+First the various entities are \textbf{setup}, where several artifacts are generated and distributed. Then a ticket credential containing the following attributes is issued:
+
+\begin{verbatim}
+First name: John
+Last name: Dow
+Birthday: 1985-05-05Z
+Member number: 23784638726
+Matchday: 2013-08-07Z
+\end{verbatim}
+
+During \textbf{issuance}, a \textit{scope exclusive pseudonym} is established and the newly issued credential is bound to this pseudonym.
+This ensures that the ticket credential can not be used without the smart card.
+
+Then \textbf{presentation} is performed.
+The \textit{presentation policy} specifies that the member number is inspectable and a predicate ensures that the matchday is in fact $2013-08-07Z$.
+This last part ensures that a ticket issued for another match can not be used.
+
+The ticket holder was lucky and his presentation token was chosen in the lottery. The presentation token is therefore inspected.
+
+\hfil
+
+\subsubsection{Execution environment}
+
+First we will execute the test in our development machine (laptop).
+After asserting that the services work as expected, we then run the test in a Raspberry Pi 3, exactly like in the laptop.
+Finally, we will deploy the IoT smart card in a Omega2 and the delegation services in the Raspberry Pi 3.
+After every test, we checked that the issuing and proving were successful, in case a cryptographic error appeared in the implementation.
+
+Lets have a closer look at the hardware of each device:
+
+\paragraph{Development laptop}\marginpar{
+	\vskip0pt
+	\includegraphics[width=0.8\linewidth]{gfx/DELL}
+
+	DELL XPS 15
+}
+Our device is a DELL XPS 15, with a Core i7-6700HQ at 3.5GHz quad core processor and 32GB of DDR4 RAM, running Ubuntu 16.10.
+
+This is a powerful machine that can simulate the performance of many servers and clients that would implement P2ABCE in a real environment, giving a reference point for performance comparisons.
+
+\paragraph{Raspberry Pi 3}\marginpar{
+	\vskip0pt
+	\includegraphics[width=0.8\linewidth]{gfx/RPi3}
+
+	Raspberry Pi 3
+} A familiar environment, powerful enough to debug and hold the delegated P2ABCE Java services of P2ABCE with its 1GB of RAM, and with two network interfaces, perfect to work as the gateway for the IoT devices to the Internet.
+
+Only a microSD with enough space to burn the OS is needed to plug\&play with the Raspberry Pi.
+We use Raspbian, a stable Debian based distro, recommended by the Raspberry Pi designers, and ready to use with the P2ABCE compiled \textit{self-contained .jar} services.
+
+\begin{table}[h]
+	\myfloatalign
+	\begin{tabularx}{0.75\textwidth}{ll} \toprule
+		CPU & ARMv8 64bit quad-core @1.2GHz \\
+		RAM & 1GB \\
+		Storage & microSD \\
+		Firmware & Raspbian (Debian based distro) \\
+		Connectivity & Wifi n + Ethernet \\
+		Power & 5V 2A \\
+		\bottomrule
+	\end{tabularx}
+	\caption[Raspberry Pi 3 Specifications]{Raspberry Pi 3 Specifications.}
+	\label{tab:RPi3Specs}
+\end{table}
+
+\paragraph{Onion Omega2}\marginpar{
+	\vskip0pt
+	\includegraphics[width=\linewidth]{gfx/Omega2}
+
+	Onion Omega2
+} A device that falls inside the category of IoT, powerful enough to run a Linux environment, LEDE, where we can develop and debug the first PoC without troubling ourselves with problems not related to the project itself.
+
+\begin{table}[h]
+	\myfloatalign
+	\begin{tabularx}{0.75\textwidth}{ll} \toprule
+		MCU & Mediatek MT688 \citep{MT7688} \\
+		CPU & MIPS32 24KEc 580MHz \\
+		RAM & 64MB \\
+		Storage & 16MB \\
+		Firmware & LEDE (OpenWRT fork distro) \\
+		Connectivity & Wifi b/g/n \\
+		Power & 3.3V 300mA \\
+		\bottomrule
+	\end{tabularx}
+	\caption[Onion Omega 2 Specifications]{Onion Omega2 Specifications.}
+	\label{tab:Omega2Specs}
+\end{table}
+
+Nonetheless, the Omega2 needs fine tunning to start operating, and basic knowledge of electronics to make it work. The two main things to begin with Omega2 are:
+\begin{itemize}
+	\item A reliable 3.3V with a maximum of 800mA power supply, e.g.
+a USB2.0 with a step-down circuit, with quality soldering and wires to avoid unwanted resistances.
+The Omega2 will usually use up to 350mA, when the WiFi module is booting up.
+The mean consumption is about 200mA.
+	\item A Serial to USB adapter wired to the TX and RX UART pins to use the Serial Terminal, to avoid the use of SSH over WiFi.
+\end{itemize}
+
+Also, we will need a cross-compiler to generate the smart card binary.
+Because we will use GMP and OpenSSL as shared libraries, the best option is to use the LEDE SDK.
+The SDK manages the available packages and generates the GCC toolchain.
+Also, using CMake,
+%TODO
+
+\paragraph{The network} In our third scenario, the Raspberry Pi 3 and the Omega2 will talk to each other over TCP.
+This implies possible network delays depending on the quality of the connection.
+The Raspberry Pi 3 is connected over Ethernet to a switch with WiFi access point.
+The Omega2 is connected over WiFi n to said AP.
+To ensure the delay wasn't significant, we measured 6000 APDU messages, and the results show that the mean transmission time is less than half a millisecond per APDU.
+As we will see in the results section, this network time is negligible.
+
+\subsection{The test code}
+
+There are three pieces of software that conform the test: the P2ABCE services, the IoT smart card, and a shell script automatizing the REST calls to all the services.
+
+\paragraph{P2ABCE services} This is the common part to our three sets.
+The services are compiled in a self-contained Jetty web server, or in WAR format, ready to be deployed in a server like Tomcat.
+We use the same JAR files with the Jetty web server for the PC and Raspberry Pi.
+
+We modified the User Service code to measure the execution time for each REST method.
+The time stamp starts inside the method and ends before it's \textit{return} instruction.
+This way, we filter the time the server spends processing the HTTP protocol and deciding which Java method to call.
+
+\paragraph{IoT smart card} Our C implementation of the P2ABCE smart card, compiled for the Omega2, with BIOSC listening on port 8888 for the APDU messages.
+
+We tested the execution in two modes, a full logging where every step was printed in terminal, and another one with no logging.
+With the first mode, we can check a proper execution, every byte exchanged, and with the second one, we measure the execution without unnecessary I/O.
+
+\paragraph{Shell script} To orchestrate all the services we use a simple script that performs the REST calls using  \textit{curl}.
+Here we perform the mentioned steps: setup of the P2ABCE system, issuance of the credential and a prove for the presentation policy.
+
+In the setup, the system parameters are generated, indicating key sizes of 1024 bits, the ones currently supported by the ABC4Trust and IoT smart cards.
+Then the system parameters and public keys from the services (issuer, inspector, revocation authority) are also exchanged.
+
+In the issuance, a two step protocol is performed by the User.
+The Issuer and the User send each other the XML data with the cryptographic information, and finish the protocol with a credential issued inside a smart card, software (Java) or hardware (physical or IoT).
+
+Finally, a Verifier sends the User a \textit{presentation policy}. The User generates a \textit{presentation token} for the Verifier and Inspector.
+
+%%%%%
+There are two script versions, one where every REST call is done from the same machine, therefore avoiding the exchange of the XML files by other means (like \textit{scp}); the other one is ready to be executed on a real setup, two scripts, one for the Omega2, where the REST calls are the delegation on the Raspberry Pi 3, and other for the PC, running the Issuer and Verifier as third party entities.
+The XML files generated by each service can be sent with the same protocol the IoT device and the Verifier, for example, would communicate in the real world.
+This way it's clearer that the \textit{curl} commands are the delegation protocol for the IoT device, and then the IoT device can send the XML as it wishes, over the Internet, Bluetooth, NFC, etc.
+%%%%%
+
+\subsection{Results}

### 2017-05-15 Doc/classicthesis/Chapters/Draft.tex
results

@@
 
+\paragraph{Future work in tests} The lack of a physical MULTOS smart card precludes us to load and test the ABC4Trust code and measure the time P2ABCE needed using the \textit{HardwareSmartcard} class.
+This would be really interesting because for a single method from the \textit{Smartcard} interface, \textit{HardwareSmartcard} needs to send various APDUs, but \textit{SoftwareSmartcard} can perform the operations immediately.
+
 \subsection{The test code}
@@
 \subsection{Results}
+
+After 15 executions for each scenario (laptop, RPi3, Omega2), we take the means and compare each step.
+
+It is worth noting first that during the test, the measured use of the CPU showed that P2ABCE does not benefit of multi core processors, therefore, it only uses one of the four cores in the Raspberry Pi 3.
+
+\paragraph{The network}
+To test the network we sent six thousand APDUs to the Omega2, but instead of calling the \textit{APDU handler}, the Omega2 responded with the same bytes back.
+This way, the Omega2 only performed the simple BIOSC protocol, reading and writing the TCP socket.
+The APDUs had multiple sizes, taken from the most common APDUs logged in a successful execution. The test showed that our test network speed was around $0.014$ ms per byte.
+
+\paragraph{The setup}
+Because the Omega2 doesn't intervene until the creation of the smart card, the times measured in the second and third scenarios are practically identical.
+
+% TODO : table y gráfico
+
+As we can see, the laptop is about ten times faster than Raspberry Pi 3, but considering that the highest time is less than two and a half seconds, and that the setup is done only once, this isn't a worrisome problem.
+
+\paragraph{Creation of the smart card}
+Here we create a \textit{SoftwareSmartcard} or a \textit{HardwareSmartcard} object that the User service will use in the following REST calls.
+
+The REST method to create a \textit{SoftwareSmartcard} is $/createSmartcard$, and to create a \textit{HardwareSmartcard}, using the \textit{IoTsmartcardio} implementation, we use $/initIoTsmartcard$.
+
+% TODO: gráfico y tabla
+
+\begin{figure}[bth]
+	\myfloatalign
+	\subfloat[Times and relative speedup]
+	{\includegraphics[width=.45\linewidth]{gfx/graphics/createSCtable}} \quad
+	\subfloat[Comparison graph]
+	{\includegraphics[width=.4\linewidth]{gfx/graphics/createSC}} \\
+	\caption{Create smart card times (seconds)}
+	\label{fig:createSmartCard:graph}
+\end{figure}
+
+The RPi3 is about 16 times slower than the laptop in the creation of the \textit{SoftwareSmartcard}, but almost 9 times faster than the setup of the smart card in the Omega2 using APDUs.
+This gives us that the laptop is 145 times faster than the combination of RPi3 and Omega2.
+
+Again, this operation is done only once per device, and includes commands from the creation of the PIN and PUK of the smart card, to storing the system parameters of P2ABCE, equivalent to the previous setup step.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=0.8\linewidth]{gfx/APDUsInitIoTSC}
+	\end{center}
+	\caption{Init IoT Smart Card APDU Commands exchanged.}
+	\label{fig:APDUsInitIoTSC}
+\end{figure}
+
+This is the first interaction between the RPi3 and the Omega2 IoT smart card, using 30 APDU Commands, and Responses, as shown in \autoref{fig:APDUsInitIoTSC}, with a total of 1109 bytes exchanged.
+From our network benchmark, using TCP sockets, the delay in the transmission is only around 15 and 20 ms, negligible as we said compared to the almost 20 seconds the operation lasts.
+
+\paragraph{Issuing of the credential}
+%TODO
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/IssuanceInteraction}
+	\end{center}
+	\caption{Issuance interaction.}
+	\label{fig:IssuanceInteraction}
+\end{figure}
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=0.7\linewidth]{gfx/IssuanceAPDUs}
+	\end{center}
+	\caption{Issuance APDU Commands.}
+	\label{fig:IssuanceAPDUs}
+\end{figure}
+
+\paragraph{Presentation}
+%TODO
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=0.7\linewidth]{gfx/APDUsProving}
+	\end{center}
+	\caption{Proving APDU Commands.}
+	\label{fig:APDUsProving}
+\end{figure}
+
+To ensure that all the process was successful, it's enough to check if the Verifier and the Inspector returned XML files, accepting the prove, or an error code.
+Of course, every execution measured in the test was successful.

### 2017-05-16 Doc/classicthesis/Chapters/Draft.tex
last interactions
results finish

@@
 
-% TODO : table y gráfico
+\begin{figure}[bth]
+	\myfloatalign
+	\subfloat[Times and relative speedup]
+	{\includegraphics[width=\linewidth]{gfx/graphics/setuptable}} \quad
+	\subfloat[Comparison graph]
+	{\includegraphics[width=.8\linewidth]{gfx/graphics/setup}} \\
+	\caption{Setup times (milliseconds)}
+	\label{fig:setup:graph}
+\end{figure}
 
@@
 
-% TODO: gráfico y tabla
-
 \begin{figure}[bth]
@@
 	\subfloat[Times and relative speedup]
-	{\includegraphics[width=.45\linewidth]{gfx/graphics/createSCtable}} \quad
+	{\includegraphics[width=0.5\linewidth]{gfx/graphics/createSCtable}} \quad
 	\subfloat[Comparison graph]
-	{\includegraphics[width=.4\linewidth]{gfx/graphics/createSC}} \\
+	{\includegraphics[width=0.45\linewidth]{gfx/graphics/createSC}} \\
 	\caption{Create smart card times (seconds)}
@@
 
-The RPi3 is about 16 times slower than the laptop in the creation of the \textit{SoftwareSmartcard}, but almost 9 times faster than the setup of the smart card in the Omega2 using APDUs.
-This gives us that the laptop is 145 times faster than the combination of RPi3 and Omega2.
+From \autoref{fig:createSmartCard:graph} we see that the RPi3 is about 16 times slower than the laptop in the creation of the \textit{SoftwareSmartcard}, but almost 9 times faster than the setup of the smart card in the Omega2 using APDUs.
+This gives us that the laptop is 145 times faster than the combination of RPi3 and Omega2 in our IoT deployment.
+But looking at the times, this process lasts up to 20 seconds, making it something feasible.
 
@@
 
-This is the first interaction between the RPi3 and the Omega2 IoT smart card, using 30 APDU Commands, and Responses, as shown in \autoref{fig:APDUsInitIoTSC}, with a total of 1109 bytes exchanged.
+This is the first interaction between the RPi3 and the Omega2 IoT smart card, \textbf{using 30 APDU Commands, and Responses}, as shown in \autoref{fig:APDUsInitIoTSC}, with a total of $1109$ bytes exchanged.
 From our network benchmark, using TCP sockets, the delay in the transmission is only around 15 and 20 ms, negligible as we said compared to the almost 20 seconds the operation lasts.
@@
 \paragraph{Issuing of the credential}
-%TODO
+The issuance is done in three steps for the User service, shown in \autoref{fig:IssuanceInteraction} with a red note showing the start of each step for the User delegation, in green for the interactions between Issuer and User, and the darker red is the Identity service, choosing the first available identity to use.
+The arrows in the figure show the REST calls performed during the test, where the IoT device acts as User, the RPi3 hosts the P2ABCE delegation services and the laptop is the Issuer.
+
+The three delegation steps and the REST method called are:
+
+\begin{center}
+	\begin{tabular}{l|l}
+	First issuance protocol step & $/issuanceProtocolStep$ \\
+	Second issuance protocol step  & $/issuanceProtocolStepUi$ \\
+	\textit{(end of first step for the User)} & \\
+	Third issuance protocol step & $/issuanceProtocolStep$ \\
+	\textit{(second step for the User)}  & \\
+\end{tabular}
+\end{center}
+
 \begin{figure}[bth]
@@
 
-\paragraph{Presentation}
-%TODO
+As we can see, the three REST calls to the delegation User service involve communication with the smart card. We show in \autoref{fig:IssuanceAPDUs} the APDU Commands for each REST call.
+There are 45 APDU Commands in total, $3197$ bytes exchanged, that would introduce a latency of 45ms in the network, negligible.
+
+\begin{figure}[bth]
+	\myfloatalign
+	\subfloat[Times (ms) and relative speedup]
+	{\includegraphics[width=\linewidth]{gfx/graphics/issuancetable}} \quad
+	\subfloat[Comparison graph]
+	{\includegraphics[width=.8\linewidth]{gfx/graphics/issuance}} \\
+	\caption{Issuance times (milliseconds)}
+	\label{fig:issuance:graph}
+\end{figure}
+
+In \autoref{fig:issuance:graph} we have the times spent in each REST call.
+The laptop shows again to be many times faster than the other two scenarios, but the times are again feasible even for the IoT environment.
+
+Lets compare the Raspberry Pi 3 and the Omega2 executions.
+There is a correlation between the number of APDU Commands needed in each step with the increment in time when using the IoT smart card, that is, how much the delegation server.
+
+The first one only involved one APDU, with $33$ bytes total (Command and Response), and times are almost identical.
+The second call needed $34$ APDUs, with $1623$ bytes, and the increase in time is around tree times slower than the RPi3 on its own.
+The third call used $20$ APDUs, $1541$ bytes, and makes the IoT scenario almost 7 times slower.
+
+The analysis shows where there are more cryptographic operations involving the Omega2, and because the amount of data exchanged is minimal, the difference in processing power between Omega2 and Raspberry Pi 3 is clear.
+
+\paragraph{Presentation token}
+The final step of the test involves a Prove, or Presentation in P2ABCE, where the Verifier sends the User or Prover the Presentation Policy, and the User answers with the Presentation Token, without more steps.
+In \autoref{fig:ProvingInteraction}, using the same colors as in the Issuance interaction, we can see the delegation messages done by the Omega2.
+
+To ensure that all the process was successful, it's enough to check if the Verifier and the Inspector returned XML files, accepting the prove, or an error code.
+Of course, every execution measured in the test was successful.
+
+In \autoref{fig:APDUsProving} the APDU Commands for each step are shown, $28$ in total, $1939$ bytes, $27$ms approx. in the network.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/ProvingInteraction}
+	\end{center}
+	\caption{Proving interaction.}
+	\label{fig:ProvingInteraction}
+\end{figure}
 
@@
 
-To ensure that all the process was successful, it's enough to check if the Verifier and the Inspector returned XML files, accepting the prove, or an error code.
-Of course, every execution measured in the test was successful.
+Again, as shown in \autoref{fig:proving:graph}, there is a correlation between the number of APDU Commands used, the work the IoT smart card must perform, and the time measured.
+The $20$ APDU Commands in the first call make the IoT deployment almost $8$ times slower than the Raspberry Pi 3; but with only $8$ APDU Commands, the second one is less than $1.5$ times slower.
+
+Nonetheless, it's significant the difference in performance between the laptop and the Raspberry Pi 3 in the last REST call, more than $40$ times slower, even using the \textit{SoftwareSmartcard}.
+
+\begin{figure}[bth]
+	\myfloatalign
+	\subfloat[Times (ms) and relative speedup]
+	{\includegraphics[width=.8\linewidth]{gfx/graphics/provingtable}} \quad
+	\subfloat[Comparison graph]
+	{\includegraphics[width=.7\linewidth]{gfx/graphics/proving}} \\
+	\caption{Proving times (milliseconds)}
+	\label{fig:proving:graph}
+\end{figure}
+
+Unlike the previous steps, the Presentation or Proving is done more than once, being the key feature of ZKP protocols.
+The laptop performs a prove in less than one second, the RPi3 needs $15$ seconds, but our P2ABCE IoT deployment needs $15$ seconds for the first step, and $18$s for the second step, $33$ seconds total to generate a Presentation Token.
+
+\hfil
+
+\paragraph{Memory usage on the Omega2}
+Using the tool \textit{time -v} we can get a lot of useful information about a program, once it finishes.
+In our case, the binary with BIOSC and the smart card logic starts as an empty smart card, goes through the described process, and then we can stop it, as the User Service won't use it anymore.
+
+After another round of tests, now using \textit{time -v}, the field named \textit{Maximum resident set size (kbytes)} shows the \textbf{maximum} size of RAM used by the process since its launch.
+In our case, this involves the use of static memory for the \textit{global variables} of the smart card logic, and the dynamic memory used by the third party libraries, like GMPlib, OpenSSL and cJSON.
+
+GMP and OpenSSL always allocate the data in their own ADT, what involves copying the arrays of bytes representing the big modular integers from the cryptographic operations.
+cJSON, used in the serialization of the smart card for storage, and debug being human readable, stores a copy of every saved variable in the JSON tree structure, then creates a string (array of char) with the JSON, that the user can write to a file.
+
+Understanding the many bad uses of memory done in this PoC is important for future improvements and ports.
+A custom modular library using the same array of bytes that the smart card logic, a binary serialization, and many improvements, are our future work.
+
+With all that said, the mean of the maximum memory usage measured is $6569.6$ kbytes.
+Compared to the $64$MB of RAM available in the Omega2, our PoC could be executed in more constrained devices, given the system is compatible.
+
+%%%%%%%%%%%%%%%%%%%%%%%
+
+% Consideraciones futuras: implementar directamente Smartcard interface vs seguir usando APDUs:  minimizan uso de red, facilitan protocolo de comunicación de las APDUs, son estándar, mejoran el mantenimiento del código: P2ABCE puede cambiar las interfaces, pero mantendrá compatibilidad con las smartcards, si implementamos Smartcard.java por cada cambio en la interfaz deberemos actualizar cada IoT smart card.

### 2017-05-17 Doc/classicthesis/Chapters/Chapter03.tex
objectives

@@
 
-\section{Project description}
+\section{Project objectives}
+
+We will divide the project in various objectives, starting from the principal goal, dividing our work in different categories.
+
+\hfil
+
+\paragraph{Idemix and the IoT}
+We part from this main objective, integrating IBM's privacy preserving system in the IoT environment.
+
+\paragraph{Analysis} Studying the state of the art for Idemix and the IoT, analyzing related projects like ABC4Trust's \ac{P2ABCE}, papers with similar approaches, and our task is to consider the best fitting solutions from where to start.
+
+\paragraph{Design} After studying the existing works, we must give a formal solution to be implemented.
+This includes the theoretical architecture and steps to take, from the software to implement, to the hardware to use, and any simplifications to leave as future work.
+
+Our \textbf{software} objectives include making it easily maintainable, structured and extensible, from the IoT and original project perspectives, that is, the IoT solution must be compatible with any non-IoT solutions, now and in the future with minimal effort, given new IoT systems or changes in the cryptographic protocols.
+
+Our \textbf{hardware} objective is to use as constrained as possible devices, but taking in count how early in the project we are, that means, using now IoT devices that ease testing and development, having in mind for the future the other devices.
+
+\paragraph{Validity and evaluation} Deploy in a \textbf{real scenario}, not simulators, a working \ac{PoC}, checking that it works as expected, and measuring its performance.
 

### 2017-05-18 Doc/classicthesis/Chapters/Design.tex
structure

--- 2017-05-12 Doc/classicthesis/Chapters/Draft.tex
@@
 %************************************************
-\chapter{Draft}\label{ch:draft}
+\chapter{Design and implementation}\label{ch:design}
 %************************************************
 
-% TODO
-% From Idemix To P2ABCE: reasons
+%TODO resumir organización del capítulo
 
-\section{Smart card APDU}
+\section{Preliminaries about smart cards}
 
-To communicate the smart cards and the reader an standardized protocol is specified in ISO/IEC 7816-4 \citep{APDUISO}.
+\subsection{Smart Card Communication Protocol}
 
-The messages, also kown as \ac{APDU}, are divided in APDU Commands and APDU Responses.
+To communicate the smart cards and the reader the ISO/IEC 7816-4 \citep{APDUISO} specifies a standardized protocol .
+
+The messages, also kown as \acp{APDU}, are divided in APDU Commands and APDU Responses.
 
@@
 chip, contact-less), but what is common between every smart card interaction, is the \textit{APDU Command-Response Dialogue}.
-As long as the smart card has a power supply, it can maintain the dynamic memory in RAM between APDU Commands, what allows to do in two or more commmands complex operations, transmit more bytes than a single APDU can admit, etc.
+As long as the smart card has a power supply, it can maintain the memory in RAM between APDU Commands, what allows to do in two or more steps complex operations, transmit more bytes than a single APDU can admit, etc.
 
@@
 
-Originally, the Lc and Le bytes had only 1 byte if present, restricting the payload data to be at most 256 bytes long.
+Originally, the Lc and Le bytes had only 1 byte, if present, restricting the payload data to be at most 256 bytes long.
 An extension to the protocol changed the meaning of a Lc or Le 0x00 byte (256 bytes long payload), so when the byte corresponding to Lc or Le started with 0x00, the next two bytes where the real length.
-With this, an Extended APDU lets up to 65536 bytes of data.
+With this, an Extended APDU lets up to $65536$ bytes of data.
 
-The problem here is that not all readers or smart cards support extended APDUs.
-Originally, to send more than 256 bytes of data, a new APDU Command instruction was defined so the smart card stored the payload in a buffer, until other APDU Command made use of it.
-To send more data in an APDU Response, the status bytes were set to: SW1=0x61 and SW2 to the remaining bytes to send.
-Because a smart card can't send APDU Commands, the card terminal then sends a GET RESPONSE, a special APDU Command, with Le set to the number of bytes specified in SW2.
-Iterating this process, the smart card could send as many bytes as it wanted.
+The problem here, is that not all readers or smart cards support extended APDUs.
+Originally, to send more than 256 bytes of data in an APDU Command, a \textit{Put Data} instruction is defined, so the smart card stores the payload in a buffer, until other APDU Command indicates how to use it.
 
-\section{P2ABCE}
+To send more data in an APDU Response, the status bytes are set to: SW1=0x61 and SW2 to the remaining bytes to send.
+Because a smart card can't send APDU Commands, the card terminal must send a \textit{GET RESPONSE}, a special APDU Command, with Le set to the number of bytes specified in SW2.
+Iterating this process, the smart card can send as many bytes as it wants as Response.
 
-In the \ac{P2ABCE} repository \citep{p2abcurl} is available the project's code, divided in two solutions: a complete P2ABCE implementation in Java and a Multos Smartcard implementation as companion for the project.
+With the introduction of Extended APDUs, this technique is no longer needed.
 
-The Java code is managed by a Maven project, structured using various known design patterns, but not of our interest.
-The structure we are actually interested in are the REST Services and their use of the Components classes, in which the smartcards are included.
-
-P2ABCE project is based on the concept of smartcards to store the credentials, logical or physical.
-An interface is defined to communicate with these smartcards, and then different implementations allow to use either \textit{Software Smartcards} or \textit{Hardware Smartcards}.
-
-The \textit{SoftwareSmartcard} class implements the interface in Java, suitable for tests and self-stored smartcards that any application using P2ABCE may need.
-
-The \textit{HardwareSmartcard} class uses the standard APDU messages [TODO:ref] to interact with smartcards.
-P2ABCE defines for every method in the mentioned interface, the necessary APDU instructions, and currently relies on \textit{javax.smartcardio} abstract classes (implemented by Oracle in their JRE) to communicate with the smartcard reader.
-This way, it doesn't matter what manufacturer issues the smartcard, or if it's an Android device, if they support the APDU API, P2ABCE will work with them.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/p2abceBasicUML}
-	\end{center}
-	\caption{Basic P2ABCE structure}
-	\label{fig:p2abceBasicUML}
-\end{figure}
-
-As a PoC the P2ABCE project includes the ABC4Trust Card Lite, an implementation for ML3-36K-R1 Multos Smartcards.
-The code is written in C, but is very dependent on the Multos framework, aside from numerous bugs and bad coding habits.
-
-At this stage, we have two options to implement our IoT device compatible with P2ABCE:
-
-\begin{itemize}
-	\item Implement in C the \textit{Smartcard} interface used by P2ABCE architecture, and use some communication protocol to remotelly call the methods from the machine running the P2ABC Engine.
-	\item Present the IoT device as a hardware smart card, using the APDU protocol (already defined, standard and with minimal overload).
-Providing a \textit{javax.smartcardio} ``IoT implementation'' to communicate with the IoT device through a transmission protocol, the already existing \textit{HardwareSmartcard} class can work with the new \textit{IoTSmartcard} in the IoT device.
-\end{itemize}
-
-%TODO : hablar más de los servicios REST
-
-\section{MULTOS}
+\subsection{MULTOS}
 
@@
 Nevertheless, the C-API documentation \citep{MultosTechLib} provides rich description for each function.
-
-\section{ABC4Trust Card Lite}
-
-P2ABCE provides a smart card reference implementation, ABC4Trust Card Lite \citep{ABC4TCardLite}. It supports device-bound U-Prove and Idemix, and virtually any discrete logarithm based pABC system.
-
-Version 1.2 is based on MULTOS ML3 cards, with approximately 64KB of EEPROM (non-volatile memory), 1KB of RAM and an Infineon SLE 78 microcontroller, a 16-bit based CPU aimed for chip cards.
-
-The card stores the user's private key $x$ and any \ac{BLOB} that the P2ABCE may need (like user's credentials).
-Then P2ABCE delegates the cryptographic operations on the smart card, that operates with $x$.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/ABC4TCardLite}
-	\end{center}
-	\caption{ABC4Trust Card Lite}
-	\label{fig:ABC4TCardLite}
-\end{figure}
-
-The cryptographic operations performed by the smartcard are the modular exponentiation and addition that discrete logarithm ZKPs are based on.
-
-\hfil
-
-The code is available from the P2ABCE project and has some good and bad points to have in count:
-
-The best asset of this code is that it's written in C aiming to a very constrained device, similar in computational power to many IoT devices, and very limited memory.
-
-Some \textit{tricks} in the code include using \textit{union} data types for variables that will be stored on the same data location, but at different moments (e.g.
-depending on APDU Command INS byte), minimizing this way the use of RAM and making code readability better; or strong use of pointers and \textit{memcpy} calls to copy structures with multiple variables as arrays of bytes.
-
-Among the many drawbacks, we could highlight the awful coding, the strong dependency on MULTOS framework and some bugs found.
-
-The code is structured in two files, \textit{main.h} and \textit{main.c}, with 557 and 5157 lines of code respectively.
-
-The file \textit{main.h} is mostly a reimplementation in assembly MEL of some MULTOS functionality already offered with latest \textit{multos.h}.
-
-The \textit{main.c} consists on near 600 lines of variables and data structures declarations, followed by the \textit{main()} function, a 2635 lines long \textit{switch-case} with practically no comments, and to conclude, the implementation of thirty functions called \textit{Subroutines} at the end of the file.
-
-This gives an idea of the problematic to maintain or even understand the code.
-But once one studies MULTOS framework in deep and applies many refactoring techniques to ABC4T Card's code, this becomes the best starting point for the IoT version.
-
-\section{IoT and P2ABCE}
-
-In this section we will define how an IoT device will be integrated in the P2ABCE environment, being totally compatible with any other system using P2ABCE, addressing the power and memory constrains IoT devices face.
-
-\hfil
-
-Our main goal is to make an IoT device capable to act as an User or Verifier in the P2ABCE architecture.
-For this, the device should manage complex XML schemas, perform cryptographic ZKP operations and communicate with the Verifier or Prover with which it's interacting.
-
-The communication is already solved by the \textit{Internet} capabilities of IoT devices.
-
-Our concern are the data artifacts exchanged as XML and the cryptographic operations involving secret keys that must remain private to the IoT device.
-
-Here is where we look at the P2ABCE architecture more closely, and the concept of smart cards shows a solution for the second issue.
-Even in the case we were to implement all P2ABCE inside an IoT device, we would have to implement support for software smart cards, to keep the secret inside the IoT device.
-We will start building the house from the ground, implementing the smart card operations inside the IoT device.
-
-Now that in our design we have the smart card, we need to address the first point, XML squemas.
-We understand with \textit{XML squemas} both managing the XML syntax and the whole process to generate a proper answer, that is, basically, the crypto engine that relies on Idemix and the smart card to hold the secret information.
-
-Taking in consideration Idemix is currently provided in Java as a considerably big project, and version 3.0.36 still hasn't got official documentation, the task to port the Idemix crypto engine to IoT could be done in the future, if the chosen devices have enough storage and capacity.
-And even in that case, we would need to implement P2ABCE User and Verifier's architecture to completely free ourselves from an external P2ABCE machine.
-
-The final port would be so big, many IoT devices would fall out of the requisites to run it, failing in our initial objective.
-
-\hfil
-
-After this analysis of the P2ABCE, we conclude that the mandatory requisite for any IoT device that wants to work with this system and keep its private keys in it, is to implement the smart card functionality, and delegate the rest of the operations on a machine capable of running P2ABCE, until that functionality is implemented for the IoT device.
-
-\hfil
-
-This architecture is not really such an original idea.
-For example, IPv6 involves managing 128 bits per address and large headers, and many use cases only need IoT devices to communicate inside a private network.
-That's why many of them use 6LoWPAN to compress packets or use smaller address sizes.
-To communicate a 6LoWPAN with the Internet, a proxy is needed to transform 6LoWPAN packets to IPv6.
-
-\hfil
-
-Therefore, the IoT device now has a \textbf{duality} in its functions, because it is the User that starts any interaction with other systems, and it's also the smart card that P2ABCE delegates for crypto operations.
-It can also be seen as a \textbf{double delegation}.
-The IoT device delegates on the external P2ABCE server to manage the protocol, and the P2ABCE server delegates on the IoT, acting now as a smart card, for the cryptography.
-
-\hfil
-
-We find here two challenges: how to delegate from the IoT device to the P2ABCE delegation server, and how to transmit them and the APDUs to the IoT device.
-
-Currently P2ABCE offers various REST web services to run different roles in P2ABCE system: User Service, Issuer Service, Verification Service, etc.
-An application that integrates P2ABCE can make use of this services in the same machine or implement the functionality using the core components written in Java, the same ones the REST services use.
-Our PoC machine, the Omega2, can make REST calls easily, but other devices may use \ac{CoAP}, and in that case, the P2ABCE REST services should be rewritten to offer CoAP support.
-The commands needed to delegate to the P2ABCE delegation server will be the same to operate with the REST services.
-This way, the first issue is solved.
-
-The transmission of the messages will depend on the specific use case, capabilities and resources available.
-If the delegation server is connected, for example, through RS-232 serial with the IoT device, and physically inaccessible, in the same way an IoT device on its own would be protected, the communication is simple, and not far away from the Arduino Yun idea of combining two devices, one more powerful but to use only when needed.
-But if the IoT device and the delegation server are apart, or more than one IoT device delegates to it, then the transmission must be secured.
-They could use 6LoWPAN to talk to each other (the delegation service could be deployed in the proxy) and then secure communications with existing solutions, like with pre-shared symmetric keys, certificates for authentication and authorization, etc., it depends on each particular deployment.
-
-At the end of the day, this is all about usual security in IoT. Many other studies focus on this matter, so we will assume it can be done, and will focus on what's new, P2ABCE in IoT.
-
-\hfil
-
-% TODO: imagen con servicios desplegados/actores en cada máquina
-
-To sum up, our IoT device will act as User (Prover or Verifier) keeping its secrets in a software smart card.
-When it starts an interaction with other actor of the P2ABCE system (Issuer, Verifier, etc.), the IoT device will delegate with a remote call (using REST in our PoC) to a P2ABCE delegation server, attaching the XML file and the necessary information for the server to send the APDUs to the software smart card (in our PoC using TCP sockets, giving the IP and listening port).
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/DelegationProving}
-	\end{center}
-	\caption{IoT Delegation in P2ABCE for Proving.}
-	\label{fig:DelegationProving}
-\end{figure}
-
-\hfil
-
-This simple design keeps the benefit of a 100\% compatible P2ABCE deployment, and the integration of IoT devices to the P2ABCE ecosystem.
-
-In the future, more functionality currently delegated in the P2ABCE server can be implemented in the IoT device, if its resources allow it.
-For example, in a M2M environment, where an IoT device can act as User and Verifier, the verification consists on sending a Presentation Policy, and verifying the Presentation Token, which implies less logic than generating it as the User.
-Therefore, the implementation of the Verifier functionality would reduce significantly the need of a delegation server, but as we said, managing complex XML squemas is not something many IoT devices could do.
-
-\section{IoT Smart Card}
-
-After many design decisions in the process to adapt the original ABC4T Card Lite code to pure C, working over a more usual architecture machine, in this section, we present the current \ac{PoC} code, most important decisions, workflow execution, and future work.
-
-First, let's define what a \textit{more usual architecture} is.
-If we remember the MULTOS section, the framework gives an application a very specific memory layout and entry and output points of execution, that could be seen as a single process execution machine.
-Many IoT devices work like a computer, with multiple processes or threads, without pre-loaded data on startup (like the APDU MULTOS loads for the application), a non-volatile memory for data and code, maybe a basic file system in this memory, and RAM with the program's stack, heap, data and code.
-
-Our PoC is tested on a Linux system, and we will give instructions on how to adapt each part to work with other typical IoT systems.
-For example, other IoT devices may work like MULTOS and let access variables in non-volatile memory during execution, and in that case, the port should be changed according to these particularities.
-
-\hfil
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/P2ABCE-IoT-bw}
-	\end{center}
-	\caption{IoT in P2ABCE deployment diagram.}
-	\label{fig:P2ABCE-IoT}
-\end{figure}
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/P2ABCE-IoT-color}
-	\end{center}
-	\caption{IoT in P2ABCE deployment diagram.}
-	\label{fig:P2ABCE-IoT-color}
-\end{figure}
-
-\subsection{Code structure}
-
-We divide the project in three different sections with the objective of enhancing maintainability, improving future changes, ports, fixes, etc.
-
-The first section is what could be called as the core of the smart card, the second one the interface for the tools the core need and may depend on the platform, and finally third party libraries, that in may be empty if the interfaces implementation doesn't need any.
-
-In our PoC we used CMake to manage the project, due to the cross-compilation tools, integration with multiple IDEs and tests.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/IoTCScomponents-color}
-	\end{center}
-	\caption{IoT Smart Card Code Structure.}
-	\label{fig:IoTCScomponents-color}
-\end{figure}
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/IoTCScomponents-bw}
-	\end{center}
-	\caption{IoT Smart Card Code Structure.}
-	\label{fig:IoTCScomponents-bw}
-\end{figure}
-
-\hfil
-
-\paragraph{Core smart card}
-
-The smart card logic lies in this section, the concepts of APDU Commands, what instructions are defined in P2ABCE smart cards and how to process them and generate proper APDU Responses.
-
-Changes in the APDU protocol for P2ABCE must be done here, independently of the target platform.
-
-After refactoring the original ABC4Trust Card's code, most of it fell in what we will call the core of the smart card.
-
-All types and variable definitions and the APDU handling is done in this code.
-However, the ABC4Trust's code depended on the MULTOS C-API for the input/output of data, modular arithmetic, and even AES128 and SHA256 cryptography.
-
-A characteristic of MULTOS C-API is that every function name starts with \textit{multos}, but as we said, the \textit{main.h} file implemented equivalent functions to some available in \textit{multos.h}.
-Our first step was to replace the \textit{main.h} functions for the standard ones in the C-API.
-Then, we implemented, following the C-API documentation, the functions from \textit{multos.h} (only the used ones) changing their names from \textit{multosFoo()} to \textit{mFoo()} for readability and emphasize that they were no longer from MULTOS.
-
-Future changes in the code may refactor it so there's no longer need for the MULTOS framework functions.
-
-\paragraph{Interfaces}
-
-To implement MULTOS functions, we needed to use some libraries, so we defined a facade to isolate the implementation of the core smart card from our different options, that could vary depending on the hardware or the system used by the IoT device.
-
-The use of a facade lets us, for example, change the implementation of modular arithmetic with a hardware optimized version, or a future more lightweight library, or our very own software implementation using the same data types that the core uses, minimizing the data transformations needed.
-
-Taking a step forward, we make the core smart card totally independent of any library, only on our interfaces.
-This means that typical C libraries, like the standard \textit{stdlib.h}, or  \textit{string.h} are also behind the facade, in case some IoT system doesn't support them.
-The main goal we go after with this decision is that future developers adapting the code to a specific platform need to make no change to the \textit{core smart card}'s code, only to the interfaces implementation.
-
-\paragraph{External utilities}
-
-If the IoT system offers well tested libraries that could aid in the interfaces implementation, or we simply found a pure C implementation for the task, these third party libraries belong to this section.
-
-In our PoC, we use two ANSI C libraries, for base64 and JSON, and two shared libraries available in as packages in LEDE, GMPLib and OpenSSL.
-The last two libraries offer more functionality than we need, hence, it's desired in a production code to implement \textit{Modular Arithmetic} and \textit{Cryptography} interfaces with more lightweight alternatives.
-
-For example, Atmel's ATAES132A \citep{ATAES132A}
-\marginpar{\includegraphics[width=\linewidth]{gfx/atmel} \\ Atmel's cryptography chips.}
-offers a serial chip for secure key storage, AES128 execution and random number generation.
-Another serial chip like ESP8266 offers WiFi connectivity, typically used with Arduino, and can also perform AES encryption.
-For random number generation, a technique used with Contiki devices is to read from sensors aleatory data and use it as seed.
-All these alternatives depend on the target device, but are all valid.
-The \textit{interfaces} and \textit{external utilities} sections allow for a clean and fast port of the code.
-
-\subsection{PoC Workflow}
-
-% PoC BIOSC : TCP sockets, JSON serialization
-
-% PoC IoT javax.smartcardio : protocolo de 2 bytes + APDU
-
-% 1. Boot: deserialize sc status
-% 2. Listening: open tcp socket
-% 3. The IoT device delegates on the P2ABCE for some task and sends the IP and Port of the IoT Smart Card.
-% 4. An APDU Command arrives : read in 2 bytes the length of the apdu and
-
-In a real deployment, we would offer a full library with an API to other processes to delegate on the P2ABC Engine, that automatizes the listening and security, that we presented in the deployment diagram \textbf{TODO:ref-img} as the \textit{P2ABCE IoT Toolkit}.
-
-But to test the IoT smart card, we use the \textit{curl} command for REST delegation calls and the BIOSC (Basic Input Output Smart Card) for APDU communication.
-
-To transmit the APDU messages in our PoC we use a simple protocol, consisting in one first byte for the instruction: receive an APDU Command or close the connection.
-In the first case, then we read two header bytes with the length of the APDU Command or Response to receive, followed by those APDU bytes.
-The message is sent over TCP for a reliable transmission (concept of session, packet retransmission, reordering, etc.).
-
-We lack of any security (authentication and authorization) that a real system should implement.
-It is vital to authenticate the delegation service, to authorize it to make APDU Commands, and the same with the IoT device, to prevent attacks.
-This belongs to \textit{usual} security, as we already said, and for that reason it's not in the PoC.
-
-\paragraph{P2ABCE REST Delegation}
-
-The only P2ABCE Service that needed to be modified was the User Service. We added the REST call
-
-\begin{center}
-	\textit{/initIoTsmartcard/{issuerParametersUid}?host=\&port=}
-\end{center}
-
-where we communicate the P2ABCE server that a IoT Smart Card is accessible via \textit{host} and \textit{port}.
-Then a new \textit{HardwareSmartcard} object is stored in the P2ABC Engine, but instead of the \textit{javax.smartcardio} Oracle's \textit{CardTerminal} implementation, we use our \textit{IoTsmartcardio} implementation for the \textit{HardwareSmartcard} constructor.
-
-The rest of P2ABCE code is unchanged and will work as if a real smart card was in use. \textit{IoTsmartcardio} implementation will transmit the APDUs through the TCP socket with the format mentioned.
-
-\hfil
-
-Our test consists on the tutorial available in P2ABCE's repository. From the Omega2's terminal we run \textit{curl} commands to send REST petitions. For example:
-
-\begin{verbatim}
-$ curl -X POST --header 'Content-Type: text/xml'
-  'http://localhost:9200/user/initIoTsmartcard/http%3A%2F%2
-  Fticketcompany%2FMyFavoriteSoccerTeam%2Fissuance%3Aidemix
-  ?host=192.168.3.1&port=8888'
-\end{verbatim}
-
-The IoT device will only manage XML as data files to exchange between the third party actor and the P2ABCE delegation service, without parsing them.
-
-After a REST call to the User Service, the P2ABC Engine will talk to the IoT smart card process.
-
-\paragraph{APDU transmission with BIOSC}
-
-% TODO : diagrama del software
-
-Before sending the REST message, the Omega2 must be listening in the specified port.
-At start up, BIOSC reads from a JSON file the status of the smart card (credentials in the BLOB, private keys, PIN and PUK codes, etc.), then opens the TCP socket and listens in a loop.
-
-When the delegation server sends an APDU Command following the simple protocol, BIOSC stores in a byte array the APDU Command bytes and calls the \textit{handle\_APDU} function in \textit{Core Smart Card}.
-
-The \textit{APDU\_handler} will parse the APDU bytes and check both CLA and INS bytes.
-For each possible INS the is a function that must always finish calling either \textit{mExit()} or \textit{output\_large\_data()}.
-
-The \textit{mExit} function is the reimplementation of MULTOS C-API exit functionality, that finishes the application execution returning to the MULTOS OS, that will send the APDU Response bytes to the terminal.
-This led in the original ABC4Trust Card's code to some tricky situations.
-Imagine a function that ends with an if-else expression.
-Many times we save a line writing something like
-
-\begin{verbatim}
-if(condition)
-    return a;
-return b;
-\end{verbatim}
-
-instead of the complete
-
-\begin{verbatim}
-if(condition)
-    return a;
-else
-    return b;
-\end{verbatim}
-
-This is a swift example of what the \textit{multosExit} function led to, because it is called in both the \textit{switch-case} processing the APDU instruction, and the \textit{Subroutines}.
-For example, the PIN code check was programmed in a way every reason to fail finished the application execution, with different error codes.
-Another example, checking if a credential is stored, if not, fail and exit inside the \textit{readBlob} subroutine.
-
-In our standard architecture machine, if we call a subroutine, it must end, and return the control to the handling function, and so the \textit{APDU\_handler} must return to the listening loop of BIOSC.
-
-Having all that in count, the \textit{mExit} function is not implemented as the documentation specifies, but only saves the smart card status serializing it in the JSON file again, and once it's saved, parses the APDU Response and sends it to the socket.
-
-The \textit{output\_large\_data} function is a tool used in MULTOS smart cards that don't support Extended APDUs.
-The APDU Command handler saves in a buffer the data to send, and its size.
-Then, \textit{output\_large\_data} will manage the buffer expecting future GET RESPONSE commands from the P2ABCE server.
-
-A future change to the project may include support for Extended APDUs, avoiding the use of multiple GET RESPONSE.
-To do this, the \textit{output\_large\_data} function and the \textit{IoTsmartcardio} implementation of \textit{javax.smartcardio} must support the feature.
-
-\section{Performance Evaluation}
-
-\subsection{Testbed description}
-
-% Hardware, network,
-
-\subsubsection{P2ABCE setting}
-
-To test the correct execution of the IoT smart card, we will use the ABC system from the tutorial in the P2ABCE Wiki \citep{p2abcurlwiki}.
-It is based on a soccer club which wishes to issue VIP-tickets to a soccer match.
-The VIP-member number in the ticket is inspectable for a lottery, ie.
-after the game, a random presentation token is inspected and the winning member is notified.
-
-First the various entities are \textbf{setup}, where several artifacts are generated and distributed. Then a ticket credential containing the following attributes is issued:
-
-\begin{verbatim}
-First name: John
-Last name: Dow
-Birthday: 1985-05-05Z
-Member number: 23784638726
-Matchday: 2013-08-07Z
-\end{verbatim}
-
-During \textbf{issuance}, a \textit{scope exclusive pseudonym} is established and the newly issued credential is bound to this pseudonym.
-This ensures that the ticket credential can not be used without the smart card.
-
-Then \textbf{presentation} is performed.
-The \textit{presentation policy} specifies that the member number is inspectable and a predicate ensures that the matchday is in fact $2013-08-07Z$.
-This last part ensures that a ticket issued for another match can not be used.
-
-The ticket holder was lucky and his presentation token was chosen in the lottery. The presentation token is therefore inspected.
-
-\hfil
-
-\subsubsection{Execution environment}
-
-First we will execute the test in our development machine (laptop).
-After asserting that the services work as expected, we then run the test in a Raspberry Pi 3, exactly like in the laptop.
-Finally, we will deploy the IoT smart card in a Omega2 and the delegation services in the Raspberry Pi 3.
-After every test, we checked that the issuing and proving were successful, in case a cryptographic error appeared in the implementation.
-
-Lets have a closer look at the hardware of each device:
-
-\paragraph{Development laptop}\marginpar{
-	\vskip0pt
-	\includegraphics[width=0.8\linewidth]{gfx/DELL}
-
-	DELL XPS 15
-}
-Our device is a DELL XPS 15, with a Core i7-6700HQ at 3.5GHz quad core processor and 32GB of DDR4 RAM, running Ubuntu 16.10.
-
-This is a powerful machine that can simulate the performance of many servers and clients that would implement P2ABCE in a real environment, giving a reference point for performance comparisons.
-
-\paragraph{Raspberry Pi 3}\marginpar{
-	\vskip0pt
-	\includegraphics[width=0.8\linewidth]{gfx/RPi3}
-
-	Raspberry Pi 3
-} A familiar environment, powerful enough to debug and hold the delegated P2ABCE Java services of P2ABCE with its 1GB of RAM, and with two network interfaces, perfect to work as the gateway for the IoT devices to the Internet.
-
-Only a microSD with enough space to burn the OS is needed to plug\&play with the Raspberry Pi.
-We use Raspbian, a stable Debian based distro, recommended by the Raspberry Pi designers, and ready to use with the P2ABCE compiled \textit{self-contained .jar} services.
-
-\begin{table}[h]
-	\myfloatalign
-	\begin{tabularx}{0.75\textwidth}{ll} \toprule
-		CPU & ARMv8 64bit quad-core @1.2GHz \\
-		RAM & 1GB \\
-		Storage & microSD \\
-		Firmware & Raspbian (Debian based distro) \\
-		Connectivity & Wifi n + Ethernet \\
-		Power & 5V 2A \\
-		\bottomrule
-	\end{tabularx}
-	\caption[Raspberry Pi 3 Specifications]{Raspberry Pi 3 Specifications.}
-	\label{tab:RPi3Specs}
-\end{table}
-
-\paragraph{Onion Omega2}\marginpar{
-	\vskip0pt
-	\includegraphics[width=\linewidth]{gfx/Omega2}
-
-	Onion Omega2
-} A device that falls inside the category of IoT, powerful enough to run a Linux environment, LEDE, where we can develop and debug the first PoC without troubling ourselves with problems not related to the project itself.
-
-\begin{table}[h]
-	\myfloatalign
-	\begin{tabularx}{0.75\textwidth}{ll} \toprule
-		MCU & Mediatek MT688 \citep{MT7688} \\
-		CPU & MIPS32 24KEc 580MHz \\
-		RAM & 64MB \\
-		Storage & 16MB \\
-		Firmware & LEDE (OpenWRT fork distro) \\
-		Connectivity & Wifi b/g/n \\
-		Power & 3.3V 300mA \\
-		\bottomrule
-	\end{tabularx}
-	\caption[Onion Omega 2 Specifications]{Onion Omega2 Specifications.}
-	\label{tab:Omega2Specs}
-\end{table}
-
-Nonetheless, the Omega2 needs fine tunning to start operating, and basic knowledge of electronics to make it work. The two main things to begin with Omega2 are:
-\begin{itemize}
-	\item A reliable 3.3V with a maximum of 800mA power supply, e.g.
-a USB2.0 with a step-down circuit, with quality soldering and wires to avoid unwanted resistances.
-The Omega2 will usually use up to 350mA, when the WiFi module is booting up.
-The mean consumption is about 200mA.
-	\item A Serial to USB adapter wired to the TX and RX UART pins to use the Serial Terminal, to avoid the use of SSH over WiFi.
-\end{itemize}
-
-Also, we will need a cross-compiler to generate the smart card binary.
-Because we will use GMP and OpenSSL as shared libraries, the best option is to use the LEDE SDK.
-The SDK manages the available packages and generates the GCC toolchain.
-Also, using CMake,
-%TODO
-
-\paragraph{The network} In our third scenario, the Raspberry Pi 3 and the Omega2 will talk to each other over TCP.
-This implies possible network delays depending on the quality of the connection.
-The Raspberry Pi 3 is connected over Ethernet to a switch with WiFi access point.
-The Omega2 is connected over WiFi n to said AP.
-To ensure the delay wasn't significant, we measured 6000 APDU messages, and the results show that the mean transmission time is less than half a millisecond per APDU.
-As we will see in the results section, this network time is negligible.
-
-\subsection{The test code}
-
-There are three pieces of software that conform the test: the P2ABCE services, the IoT smart card, and a shell script automatizing the REST calls to all the services.
-
-\paragraph{P2ABCE services} This is the common part to our three sets.
-The services are compiled in a self-contained Jetty web server, or in WAR format, ready to be deployed in a server like Tomcat.
-We use the same JAR files with the Jetty web server for the PC and Raspberry Pi.
-
-We modified the User Service code to measure the execution time for each REST method.
-The time stamp starts inside the method and ends before it's \textit{return} instruction.
-This way, we filter the time the server spends processing the HTTP protocol and deciding which Java method to call.
-
-\paragraph{IoT smart card} Our C implementation of the P2ABCE smart card, compiled for the Omega2, with BIOSC listening on port 8888 for the APDU messages.
-
-We tested the execution in two modes, a full logging where every step was printed in terminal, and another one with no logging.
-With the first mode, we can check a proper execution, every byte exchanged, and with the second one, we measure the execution without unnecessary I/O.
-
-\paragraph{Shell script} To orchestrate all the services we use a simple script that performs the REST calls using  \textit{curl}.
-Here we perform the mentioned steps: setup of the P2ABCE system, issuance of the credential and a prove for the presentation policy.
-
-In the setup, the system parameters are generated, indicating key sizes of 1024 bits, the ones currently supported by the ABC4Trust and IoT smart cards.
-Then the system parameters and public keys from the services (issuer, inspector, revocation authority) are also exchanged.
-
-In the issuance, a two step protocol is performed by the User.
-The Issuer and the User send each other the XML data with the cryptographic information, and finish the protocol with a credential issued inside a smart card, software (Java) or hardware (physical or IoT).
-
-Finally, a Verifier sends the User a \textit{presentation policy}. The User generates a \textit{presentation token} for the Verifier and Inspector.
-
-%%%%%
-There are two script versions, one where every REST call is done from the same machine, therefore avoiding the exchange of the XML files by other means (like \textit{scp}); the other one is ready to be executed on a real setup, two scripts, one for the Omega2, where the REST calls are the delegation on the Raspberry Pi 3, and other for the PC, running the Issuer and Verifier as third party entities.
-The XML files generated by each service can be sent with the same protocol the IoT device and the Verifier, for example, would communicate in the real world.
-This way it's clearer that the \textit{curl} commands are the delegation protocol for the IoT device, and then the IoT device can send the XML as it wishes, over the Internet, Bluetooth, NFC, etc.
-%%%%%
-
-\subsection{Results}

### 2017-05-18 Doc/classicthesis/Chapters/Draft.tex
objectives
structure

@@
 
+we decide that the best option is the use of P2ABCE, a framework defining how entities of the system must communicate and using as cryptographic engine Idemix or U-Prove
+
 % TODO
 % From Idemix To P2ABCE: reasons
-
-\section{Smart card APDU}
-
-To communicate the smart cards and the reader an standardized protocol is specified in ISO/IEC 7816-4 \citep{APDUISO}.
-
-The messages, also kown as \ac{APDU}, are divided in APDU Commands and APDU Responses.
-
-\textbf{APDU Commands} consist in 4 mandatory bytes (CLA, INS, P1, P2), and an optional payload.
-
-\begin{itemize}
-	\item CLA byte: Instruction class. Denotes if the command is interindustry standard or proprietary.
-	\item INS byte: Instruction code. Indicates the specific command.
-	\item P1, P2 bytes: Instruction parameters.
-	\item Lc, 0-3 bytes: Command data length.
-	\item Command data: Lc bytes of data.
-	\item Le, 0-3 bytes: Expected response data length.
-\end{itemize}
-
-This way, minimal number of bytes are needed to transmit commands to the smart card, allowing manufacturer's personalization of the smart card behavior and capabilities along with standard operations.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=0.55\linewidth]{gfx/APDUCommand}
-	\end{center}
-	\caption{APDU Command}
-	\label{fig:APDUCommand}
-\end{figure}
-
-\textbf{APDU Responses} are generated inside the smart card, always as an answer to an APDU Command. They consist on an optional payload and two mandatory status bytes.
-
-\begin{itemize}
-	\item Response data: At most Le bytes of data.
-	\item SW1-SW2 bytes: Status bytes. Encode the exit status of the instruction.
-\end{itemize}
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=0.55\linewidth]{gfx/APDUResponse}
-	\end{center}
-	\caption{APDU Response}
-	\label{fig:APDUResponse}
-\end{figure}
-
-\hfil
-
-The transmission protocol varies between different types of readers and smart cards (e.g.
-chip, contact-less), but what is common between every smart card interaction, is the \textit{APDU Command-Response Dialogue}.
-As long as the smart card has a power supply, it can maintain the dynamic memory in RAM between APDU Commands, what allows to do in two or more commmands complex operations, transmit more bytes than a single APDU can admit, etc.
-
-\hfil
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=0.75\linewidth]{gfx/APDUdialog}
-	\end{center}
-	\caption{APDU Command-Response Dialogue}
-	\label{fig:APDUdialog}
-\end{figure}
-
-Originally, the Lc and Le bytes had only 1 byte if present, restricting the payload data to be at most 256 bytes long.
-An extension to the protocol changed the meaning of a Lc or Le 0x00 byte (256 bytes long payload), so when the byte corresponding to Lc or Le started with 0x00, the next two bytes where the real length.
-With this, an Extended APDU lets up to 65536 bytes of data.
-
-The problem here is that not all readers or smart cards support extended APDUs.
-Originally, to send more than 256 bytes of data, a new APDU Command instruction was defined so the smart card stored the payload in a buffer, until other APDU Command made use of it.
-To send more data in an APDU Response, the status bytes were set to: SW1=0x61 and SW2 to the remaining bytes to send.
-Because a smart card can't send APDU Commands, the card terminal then sends a GET RESPONSE, a special APDU Command, with Le set to the number of bytes specified in SW2.
-Iterating this process, the smart card could send as many bytes as it wanted.
 
@@
 %TODO : hablar más de los servicios REST
-
-\section{MULTOS}
-
-MULTOS is a multi-application smart card operative system, which provides a custom developing environment, with rich documentation \citep{MultosTechLib}.
-MULTOS smart cards communicate like any other smart card following the standard, but internally offers a very specific architecture, affecting the way one must code applications for it.
-
-In this section we will present the main characteristics of a MULTOS smart card that shaped the ABC4Trust Card Lite code and that we had to be aware of when adapting it to IoT devices.
-
-\paragraph{MULTOS programming languages} A native assembly language called MEL, C and, to a lesser extent, Java, are the available languages to code for MULTOS.
-In our case, ABC4T Card Lite uses MEL and C.
-
-\paragraph{MULTOS Workflow}
-
-Most of the transmission and communication process is done by MULTOS core, and it then selects, based on the CLA byte of the APDU, the application to load.
-This application is what most developers will only worry about, and is where their $main()$ function will start.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/multosWorkflow}
-	\end{center}
-	\caption{MULTOS workflow}
-	\label{fig:multosWorkflow}
-\end{figure}
-
-The application uses then the \textit{multos.h} file that declares multiple global variables already loaded with the needed data, including the APDU Command bytes.
-
-Now the developer is in charge of checking what instruction was sent and if the APDU has the expected ISO Case.
-If everything is ok, code what needs to be run and write in specific data space the APDU Response bytes, call \textit{multosExit()} and MULTOS will be in charge to send the APDU Response.
-
-In summary, our application starts with all data loaded and exits without worrying how to send the answer.
-A very comfortable workflow that we must now implement for our IoT device if we would want ABC4T Card Lite code to work.
-
-\paragraph{MULTOS Memory Layout}
-
-Each application in MULTOS has access to a specific memory layout, divided in different categories:
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/multosMemLay}
-	\end{center}
-	\caption{MULTOS Memory Layout}
-	\label{fig:multosMemLay}
-\end{figure}
-
-The Code Space is where the application code is stored.
-The Data Space is divided in Static memory, Public memory and Dynamic memory.
-
-\textbf{Static memory} are the application variables declared after the specific \textit{\#pragma melstatic} compiler directive.
-These variables are stored in the non-volatile EEPROM, and any write is assured to be saved because they are not loaded into RAM.
-
-\textbf{Public memory} can be seen as the input/output buffer for applications and MULTOS system.
-The APDU header appears at the top of Public, and command data at the bottom.
-The application writes then the APDU Response bytes in Public, at specific position (see \autoref{fig:multosPubMem}).
-To declare variables in this data space, the \textit{\#pragma melpublic} directive is available.
-
-\textbf{Dynamic memory} works like usual program memory, with Session Data storing global variables and the Stack.
-The limited size of RAM in IoT devices and smart cards makes the use of dynamic memory not advisable.
-The compiler directive to use Session Data is \textit{\#pragma melsession}.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/multosPubMem}
-	\end{center}
-	\caption{MULTOS Public Memory Data Map}
-	\label{fig:multosPubMem}
-\end{figure}
-
-\hfil
-
-With regards to primitive types, to avoid confusion with their sizes, MULTOS defines and uses the following data types specified in \autoref{fig:multosDataTypes}.
-It's important to notice that MULTOS is Big Endian
-\marginpar{\includegraphics[width=\linewidth]{gfx/Big-Endian}\\Big-Endian, Wikipedia}
-and when storing structures there is no padding between defined variables, unlike modern compilers that perform data structure alignment \citep{dataStructAlign} for performance.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/multosDataTypes}
-	\end{center}
-	\caption{MULTOS Data Types}
-	\label{fig:multosDataTypes}
-\end{figure}
-
-\paragraph{MULTOS Standard C-API}
-
-A collection of more than a hundred functions are provided for arithmetic, cryptography, memory and smart card operations.
-The \textit{multos.h} interface provides access to these functions, that ultimately call their respective primitive instructions in assembly code.
-The primitive instructions are but a system call with an operation code, loading data in the needed registers.
-Therefore,  no implementation for these tools is available, nor in C, nor in assembly code.
-
-Nevertheless, the C-API documentation \citep{MultosTechLib} provides rich description for each function.
 
@@
 
+% TODO: esto para el diseño: fase de offloading
 \section{IoT and P2ABCE}
@@
 
+% TODO: for design
 \section{IoT Smart Card}

### 2017-05-18 Doc/classicthesis/Chapters/Introduction.tex
structure

%************************************************
\chapter{Introduction}\label{ch:introduction}
%************************************************

The \ac{IoT} is a term with a wide range of interpretations \citep{Atzori20102787}, but a breaf definition could be the set of devices, mainly resource constrained, that are interconnected between them in order to achieve a goal. This includes examples like lampposts with proximity sensors that talk to each other in order to light up part of the street when a passerby walks by, greenhouses with automated irrigation, or cars exchanging traffic data.

Thanks to organizations like \href{https://wikileaks.org/}{WikiLeaks}, people are aware of the implications of their data on the Internet, demanding more security and privacy for it. This includes not only the data shared with others, where one must trust they will keep it safe, but it's also the data collected about us and that we don't have direct control over it. The attack Sony suffered in 2011 to PSN \citep{PSN2011} is an example of the trust people had in Sony to store their billing information, and because the security of that information depended on both them and Sony, people found themselves compromised unable to do anything about it.

With the proliferation of IoT devices gathering as much information as they can with their sensors, the amount of data gathered about anyone can be immense. And IoT has proved to not address neither security nor privacy, with recent events like the Mirai botnet DDoS attack on October 2016, considered the biggest DDoS in history \citep{jeyanthi:2017}, or the multiple vulnerabilities affecting house devices, like baby monitors \citep{rapid7babycam}.

To address this problem of privacy in the Internet, a recent approach is the concept of \textit{strong anonymity}, that conceals our personal details while letting us continue to operate online as a clearly defined individuals \citep{stronganonymity}. One very promising way to achieve this is using \acp{ZKP}
\marginpar{To read more about ZKPs, refer to my Mathematics thesis \citep{tfgmates}.}
, cryptographic methods that allows to proof knowledge of data without disclosing it.

To understand how ZKPs work, in 1990 Guillou, Quisquater and Berson published in \textit{How to Explain Zero-Knowledge Protocols to Your Children} \citep{ZKPcave:story} a story about how Ali Baba proved he knew the magic words to open the cave, but without revealing the words to anyone. Here we present a brief version highlighting the properties of ZKP.

\hfil

\begin{quote}
	Imagine a cave, where the path forks in two passages, and at the end of each one, they join again, with the shape of a ring. In the point the paths meet, there is a magic door, that only opens when someones pronounces the magic work.

	\textbf{P}eggy knows the secret word and wants to \textbf{p}rove it to her friend, \textbf{V}ictor, but without revealing it.
	\marginpar{\includegraphics[width=1.\linewidth]{gfx/graficoJL_ZKP_1}\\The cave \citep{ZKPcave:fig}
		. Peggy takes randomly A or B. Victor awaits outside.}
	Peggy and Victor meet at the entrance of the cave, then Victor awaits while Peggy goes inside the cave, taking one of the passages, that we will name A and B. Victor can't see which way Peggy went.

	When Peggy arrives at the door, \textbf{V}ictor enters the cave, and when he arrive to the fork, stops and yells which path, A or B, he wants Peggy to come back, to \textbf{v}erify she knows how to open the door.

	If Peggy actually knows the secret, she always can take the requested path, opening the magic door if needed.
	\marginpar{\includegraphics[width=1.\linewidth]{gfx/graficoJL_ZKP_2}\\The cave. Victor chooses randomly the returning path for Peggy.}
	But if Peggy doesn't know the magic word, she had a chance of $50\%$ to guess correctly what passage Victor was going to ask. That means she had a chance to fool Victor.

	Victor then asks to repeat the experiment. With $20$ repetitions, the chances Peggy fools Victor in all of them is only  $2^{-20}$,
	\marginpar{\includegraphics[width=1.\linewidth]{gfx/graficoJL_ZKP_3}\\The cave. Peggy returns by the requested path.}

	\textbf{E}ve, curious about what Victor and Peggy were doing in the cave, \textbf{e}avesdrop Victor during the process. The problem is that Eve doesn't know if Peggy and Victor agreed on what paths to choose, because they wanted to prank her for being busybody. Only Victor is confident he is choosing the returning passage randomly.

	Later, Victor is convinced that the door can be opened and Peggy knows the word, but he can't prove it to Eve because he can't open the door. 
\end{quote}

Based on ZKP properties, IBM has developed the Identity Mixer, Idemix for short, protocol suite for privacy-preserving authentication and transfer of certified attributes \citep{idemixurl}. It allows user authentication without divulging any personal data. Users have a personal certificate with multiple attributes, but they can choose how many to disclose, or only give a proof of them, like being older than 18 years-old, living in a country without revealing the city, etc. Thus, no personal data is collected that needs to be protected, managed, and treated.

\begin{center}
	\textit{``If your personal data is never collected, it cannot be stolen.''}
\end{center}

The goal of this project is to integrate Idemix with the IoT. It will be done using the ABC4Trust's \ac{P2ABCE}, a framework that defines a common architecture, policy language and data artifacts for an attribute based ecosystem, cryptographically based on either IBM's Idemix or Microsoft's U-Prove \citep{p2abcurl}. This gives us a standardized language to exchange Idemix's messages between IoT devices and any other P2ABCE actor.

\hfil

\section{Outline of this thesis}

%TODO
TODO: organización, basada en los objetivos

### 2017-05-18 Doc/classicthesis/Chapters/Objectives.tex
structure

%************************************************
\chapter{Analysis and methodology}\label{ch:objectives}
%************************************************

% descripción del proyecto: objetivos, metodología: pasos del diseño y desarrollo, ver figura 3.1 (elección de sensores: computing offloading), contexto (partimos de ABC4T: recordar P2ABCE y que se basa en tarjetas), entorno de desarrollo (meter elección del Omega2, LEDE, teoría de APDUs, código heredado) el sw, los tests p2abce puro,

%TODO resumir organización del capítulo

In this chapter we describe the project objectives, and the methodology followed during its development. In section ... TODO

\section{Project description}

The purpose of this project is the integration of IBM's privacy preserving solution, Idemix, in the environment of the Internet of Things. The objective is to design a general solution for the existing IoT devices and systems, without compromising any feature of Idemix, and provide a working PoC in a real IoT environment, even though it isn't the most constrained scenario. The project is aimed to be used in privacy preserving environments, providing security for IoT, controlling what data is being disclosed and to whom. In a smart city project, citizens data can be privatized and continue to offer the benefits of the sensors around. Authorized personnel can disclose the information when required, like firefighters that could access a building sensors to know how many people and what conditions they are in, in case of a emergency; and keep such invasive data private to other non-critical services, for example, only giving a proof that there are more than a minimum number of people in the building, to activate or deactivate the air-conditioning system.

\hfil

We will divide the project in various \textbf{objectives}, starting from the principal goal, dividing our work in different categories.

\hfil

\paragraph{Idemix and the IoT}
We part from this main objective, integrating IBM's privacy preserving system in the IoT environment.

\paragraph{Analysis} Study the state of the art for Idemix and the IoT, analyzing related projects like ABC4Trust's \ac{P2ABCE}, papers with similar approaches, and our task is to consider the best fitting solutions from where to start.

\paragraph{Design and implementation} After studying the existing works, we must give a formal solution to be implemented. This includes the theoretical architecture, the steps to take, the software to implement, the hardware to use, and any simplifications to leave as future work.

Our \textbf{software} objectives include making it easily maintainable, structured and extensible, from the IoT and original project perspectives, that is, the IoT solution must be interoperable with other non-IoT solutions, now and in the future, with minimal effort, given new IoT systems or changes in the cryptographic protocols.

Our \textbf{hardware} objective is to use devices as constrained as possible. But considering our early situation in the project, we will use IoT devices that ease testing and development, having in mind the other devices, and what specific steps we should take for them in the future.

\paragraph{Validity and evaluation} Deploy a \ac{PoC} in a \textbf{real scenario}, without the need of simulators, checking it works as expected, and measuring its performance.

\section{Methodology}

Giving the vast range of IoT devices, a one-for-all solution must take in consideration multiple requirements and limitations. We will break down the original system, Idemix and P2ABCE, analyze every part of it and categorize them. We will consider what components are mandatory to be executed in the target device and which ones the devices would be able to execute.

Devices with equivalent processing power to smart phones are capable to execute the current Java implementations of P2ABCE and Idemix, but the most constrained ones can't handle the entire system, only the key cryptographic operations.

Using the technique known as \textit{Computation Offloading} we will design an IoT architecture where the most constrained targets can keep their private keys and certificates secure in the same device, and act as any other User in the P2ABCE system. Studying the P2ABCE architecture and implementations we will identify those mandatory operations to be executed in the constrained device, and how to communicate efficiently during the delegation.

After the technical design, we will implement the PoC, using software designs patterns known to many professionals, that will improve the maintainability of the project, for improvements and fixes. Taking advantage of this practices, we can document the immediate steps for future developers, how to reimplement certain interfaces when porting the application to a new system, or where the core logic lies for future protocol changes.

Finally, the PoC must be evaluated to assert we achieved our goals, and measure its performance, judging if it could be fitting for real IoT environments.

\section{Analysis of P2ABCE}

% TODO: explicar P2ABCE

In the \ac{P2ABCE} repository \citep{p2abcurl} is available the project's code, divided in two solutions: a complete P2ABCE implementation in Java and a Multos Smartcard implementation as companion for the project.

The Java code is managed by a Maven project, structured using various known design patterns, but not of our interest. The structure we are actually interested in are the REST Services and their use of the Components classes, in which the smartcards are included.

P2ABCE project is based on the concept of smartcards to store the credentials, logical or physical. An interface is defined to communicate with these smartcards, and then different implementations allow to use either \textit{Software Smartcards} or \textit{Hardware Smartcards}.

The \textit{SoftwareSmartcard} class implements the interface in Java, suitable for tests and self-stored smartcards that any application using P2ABCE may need.

The \textit{HardwareSmartcard} class uses the standard APDU messages [TODO:ref] to interact with smartcards. P2ABCE defines for every method in the mentioned interface, the necessary APDU instructions, and currently relies on \textit{javax.smartcardio} abstract classes (implemented by Oracle in their JRE) to communicate with the smartcard reader. This way, it doesn't matter what manufacturer issues the smartcard, or if it's an Android device, if they support the APDU API, P2ABCE will work with them.

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/p2abceBasicUML}
	\end{center}
	\caption{Basic P2ABCE structure}
	\label{fig:p2abceBasicUML}
\end{figure}

As a PoC the P2ABCE project includes the ABC4Trust Card Lite, an implementation for ML3-36K-R1 Multos Smartcards. The code is written in C, but is very dependent on the Multos framework, aside from numerous bugs and bad coding habits.

At this stage, we have two options to implement our IoT device compatible with P2ABCE:

\begin{itemize}
	\item Implement in C the \textit{Smartcard} interface used by P2ABCE architecture, and use some communication protocol to remotelly call the methods from the machine running the P2ABC Engine.
	\item Present the IoT device as a hardware smart card, using the APDU protocol (already defined, standard and with minimal overload). Providing a \textit{javax.smartcardio} ``IoT implementation'' to communicate with the IoT device through a transmission protocol, the already existing \textit{HardwareSmartcard} class can work with the new \textit{IoTSmartcard} in the IoT device.
\end{itemize}

\subsection{ABC4Trust Card Lite}

%TODO: as part of P2ABCE project, this MULTOS PoC

%TODO: check content

P2ABCE provides a smart card reference implementation, ABC4Trust Card Lite \citep{ABC4TCardLite}. It supports device-bound U-Prove and Idemix, and virtually any discrete logarithm based pABC system.

Version 1.2 is based on MULTOS ML3 cards, with approximately 64KB of EEPROM (non-volatile memory), 1KB of RAM and an Infineon SLE 78 microcontroller, a 16-bit based CPU aimed for chip cards.

The card stores the user's private key $x$ and any \ac{BLOB} that the P2ABCE may need (like user's credentials). Then P2ABCE delegates the cryptographic operations on the smart card, that operates with $x$.

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/ABC4TCardLite}
	\end{center}
	\caption{ABC4Trust Card Lite}
	\label{fig:ABC4TCardLite}
\end{figure}

The cryptographic operations performed by the smartcard are the modular exponentiation and addition that discrete logarithm ZKPs are based on.

\hfil

The code is available from the P2ABCE project and has some good and bad points to have in count:

The best asset of this code is that it's written in C aiming to a very constrained device, similar in computational power to many IoT devices, and very limited memory.

Some \textit{tricks} in the code include using \textit{union} data types for variables that will be stored on the same data location, but at different moments (e.g. depending on APDU Command INS byte), minimizing this way the use of RAM and making code readability better; or strong use of pointers and \textit{memcpy} calls to copy structures with multiple variables as arrays of bytes.

Among the many drawbacks, we could highlight the awful coding, the strong dependency on MULTOS framework and some bugs found.

The code is structured in two files, \textit{main.h} and \textit{main.c}, with 557 and 5157 lines of code respectively.

The file \textit{main.h} is mostly a reimplementation in assembly MEL of some MULTOS functionality already offered with latest \textit{multos.h}.

The \textit{main.c} consists on near 600 lines of variables and data structures declarations, followed by the \textit{main()} function, a 2635 lines long \textit{switch-case} with practically no comments, and to conclude, the implementation of thirty functions called \textit{Subroutines} at the end of the file.

This gives an idea of the problematic to maintain or even understand the code. But once one studies MULTOS framework in deep and applies many refactoring techniques to ABC4T Card's code, this becomes the best starting point for the IoT version.

\section{Development environment}

The development is divided between the \ac{IoT} device code and the \ac{P2ABCE} services. To ease the setup of new developing machines, we will use Docker to create compiling containers.

\hfil

The P2ABCE is already written in Java, and few modifications will be done to the code in comparison to the existing project size, so we will continue using Java with the P2ABCE part.

P2ABCE project needs some minor changes to work with the IoT architecture. It is written in Java using Maven to manage the project dependencies. Any text editor or Java IDE is suitable for the development, because the compilation is done through the terminal, with Maven commands.

We compile the project inside a Docker container, with OpenJDK 7, Maven and the Idemix Maven plugins installed, following the project \href{https://github.com/p2abcengine/p2abcengine/wiki/How-to-Build-the-ABC-Engine}{instructions} to use Idemix as the Engine for P2ABCE.

\hfil

We can assume all IoT devices have a C cross-compiler, some even a C++ cross-compiler. The worst case scenario is that one must write assembly code, and that code will be specific of that target, so we won't consider them.
If now we focus on the most constrained devices, we could find out that some can't compile C++, some may not have many common libraries, and that the memory limitations they face make practically impossible to use dynamic memory, if we want to avoid many execution malfunctions.

For that reason, the developed code for IoT devices should be written with standard C, without using dynamic memory or third party libraries.

\hfil

To manage the PoC code we chose CMake, providing many advantages over Makefiles:

\begin{itemize}
	\item Cross-platform. It works in many systems, and more specifically, in Linux it generates Makefiles.
	\item Simpler syntax. Adding a library, files to compile, set definitions, etc. can be done with one CMake command, with rich documentation on the project's \href{https://cmake.org/cmake/help/latest/}{website}.
	\item Cross-compilation. With only a \href{http://www.vtk.org/Wiki/CMake_Cross_Compiling#The_toolchain_file}{\small{CMAKE TOOLCHAIN}} file, CMake sets up automatically the cross-compilation with Makefiles and the C/C++ cross-compiler provided.
\end{itemize}

\hfil

Although the ideal final code is pure C, without external libraries or dynamic memory, the \ac{PoC} uses three major libraries:

\begin{itemize}
	\item OpenSSL: Provides reliable and tested AES128, SHA256 and random number generator implementations.
	\item LibGMP: Provides multiprecision integer modular arithmetic.
	\item cJSON: Provides a JSON parser to store and read the status of the device, in a human readable way.
\end{itemize}

These three libraries are used to implement different interfaces in the project, and C implementations of these interfaces should replace the external libraries in the future.

\hfil

Finally, we use Docker to deploy the compilation environment. Our container includes CMake and the LEDE SDK \citep{ledeproject}, configured for the Omega2 target, the device chose for the PoC.

The Dockerfiles can be found in the Appendix.
%TODO

### 2017-05-18 Doc/classicthesis/Chapters/StateOfTheArt.tex
structure

--- 2017-04-06 Doc/classicthesis/Chapters/Chapter02.tex
@@
 
-In this chapter we present the two dimensions of this project: the \ac{IoT} development state and an introduction to IBM's privacy-preserving solution, Idemix.
+In this chapter we present the two domains of this project: the Zero-Knowledge Proof based systems and the Internet of Things.
+
+\section{Zero-Knowledge Proof systems}
+% U -Prove
+% Persiano
+% Idemix
 
@@
 The development for Internet of Things depends heavily on each target device.
-We can differentiate two big groups: those with enough processing power to act like an usual computer, and those constrained devices that can't perform arbitrary tasks, sometimes called \textit{embedded}.
+We could categorize them in two classes, depending on their capabilities: those with enough processing power to act like an usual computer, and those constrained devices that can't perform arbitrary tasks, sometimes called \textit{embedded}.
 
@@
 \paragraph{TODO: Poner descripción + formal de Idemix, luego P2ABCE, y meter que en P2ABCE había una implementación de sc basada en C}
+
+%TODO: vanets, IoT, explicar Idemix, P2ABCE: ZKP systems: UProve, el extra de las vanets e Idemix

### 2017-05-18 Doc/classicthesis/FrontBackmatter/Declaration.tex
structure

%*******************************************************
% Declaration
%*******************************************************
\refstepcounter{dummy}
\pdfbookmark[0]{Declaración de originalidad}{Declaración de originalidad}
\chapter*{Declaración de originalidad}
\thispagestyle{empty}
Yo, \myName, autor del TFG \spacedlowsmallcaps{\myTitle}, bajo la tutela de los profesores \myProf y \myOtherProf, declaro que el trabajo que presento es original, en el sentido de que ha puesto el mayor empeño en citar debidamente todas las fuentes utilizadas.

\bigskip

\noindent\textit{\myLocation, \myTime}

\smallskip

\begin{flushright}
	\begin{tabular}{m{5cm}}
		\\ \hline
		\centering\myName \\
	\end{tabular}
\end{flushright}

### 2017-05-19 Doc/classicthesis/Chapters/Chapter0A.tex
fixes + organización

%********************************************************************
% Appendix
%*******************************************************
% If problems with the headers: get headings in appendix etc. right
%\markboth{\spacedlowsmallcaps{Appendix}}{\spacedlowsmallcaps{Appendix}}
\chapter{Test: APDU Commands exchanged}\label{ch:resultsdiagrams}

% TODO
In this appendix we provide 3 sequence diagrams highlighting the APDU Commands exchanged during our testbed. The first one is the setup of the IoT smart card, storing the system parameters needed to work in the deployed P2ABCE system. The second one is the issuance of the credential, divided in the three REST calls needed in the delegation. The last one is the proving for the Presentation Policy from the Verifier, using two REST calls.

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=0.9\linewidth]{gfx/APDUsInitIoTSC}
	\end{center}
	\caption{Init IoT Smart Card APDU Commands exchanged.}
	\label{fig:APDUsInitIoTSC}
\end{figure}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=0.78\linewidth]{gfx/IssuanceAPDUs}
	\end{center}
	\caption{Issuance APDU Commands.}
	\label{fig:IssuanceAPDUs}
\end{figure}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/APDUsProving}
	\end{center}
	\caption{Proving APDU Commands.}
	\label{fig:APDUsProving}
\end{figure}

\chapter{Guide on how to continue the project development}

### 2017-05-19 Doc/classicthesis/Chapters/Conclusions.tex
fixes + organización

%************************************************
\chapter{Conclusions and Future Work}\label{ch:conclusions}
%************************************************

To finish this document, we sum up some conclusions from the work done, and results 
obtained. We will also enumerate some future lines of research that could start from
the work done in this project.

\section{Conclusions}

The designed solution for the integration of Idemix and the Internet of Things provides new possibilities for the IoT security field, mainly in those scenarios where people's data is more vulnerable. However, there is a long path of research before we can see this in production. Many decisions depend on the specific deployment in course, and our solution tries to ease the best it all that future process.

With regards to the designed architecture, the flexibility of the Computation Offloading technique, identifying the key operations that can't be delegated, and those ones that can, allows us to define a general solution for the vast world of the Internet of Things. The IoT devices can operate as individual actors in the Idemix ecosystem. When in need of performing offloading, the delegation server also falls into the IoT class of devices, a great benefit for any real project, having as many options as possible, the solution delivered can be adapted to most of requirements. Nevertheless, being so open to any solution leaves us with a lot of work to do, researching what are the best options, comparing benefits and drawbacks.

Our PoC implementation demonstrates that this project is actually feasible, not by performing a simulation of an IoT device, like in \citep{vanet}. During its development, we had to investigate a lot of concepts related to IoT, smart cards, and even the insides of P2ABCE's code, to fix many existing bugs in the original project and minimize the amount of changes it had to undergo, in order to work with the IoT devices.

\section{Future work}

% Mejoras PoC
% Mejoras arquitectura: estudiar entornos típicos IoT y definir un despliegue realista usando esta cryptografía

### 2017-05-19 Doc/classicthesis/Chapters/Design.tex
fixes + organización

@@
 %TODO resumir organización del capítulo
-
-\section{Preliminaries about smart cards}
-
-\subsection{Smart Card Communication Protocol}
-
-To communicate the smart cards and the reader the ISO/IEC 7816-4 \citep{APDUISO} specifies a standardized protocol .
-
-The messages, also kown as \acp{APDU}, are divided in APDU Commands and APDU Responses.
-
-\textbf{APDU Commands} consist in 4 mandatory bytes (CLA, INS, P1, P2), and an optional payload.
-
-\begin{itemize}
-	\item CLA byte: Instruction class. Denotes if the command is interindustry standard or proprietary.
-	\item INS byte: Instruction code. Indicates the specific command.
-	\item P1, P2 bytes: Instruction parameters.
-	\item Lc, 0-3 bytes: Command data length.
-	\item Command data: Lc bytes of data.
-	\item Le, 0-3 bytes: Expected response data length.
-\end{itemize}
-
-This way, minimal number of bytes are needed to transmit commands to the smart card, allowing manufacturer's personalization of the smart card behavior and capabilities along with standard operations.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=0.55\linewidth]{gfx/APDUCommand}
-	\end{center}
-	\caption{APDU Command}
-	\label{fig:APDUCommand}
-\end{figure}
-
-\textbf{APDU Responses} are generated inside the smart card, always as an answer to an APDU Command. They consist on an optional payload and two mandatory status bytes.
-
-\begin{itemize}
-	\item Response data: At most Le bytes of data.
-	\item SW1-SW2 bytes: Status bytes. Encode the exit status of the instruction.
-\end{itemize}
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=0.55\linewidth]{gfx/APDUResponse}
-	\end{center}
-	\caption{APDU Response}
-	\label{fig:APDUResponse}
-\end{figure}
-
-\hfil
-
-The transmission protocol varies between different types of readers and smart cards (e.g.
-chip, contact-less), but what is common between every smart card interaction, is the \textit{APDU Command-Response Dialogue}.
-As long as the smart card has a power supply, it can maintain the memory in RAM between APDU Commands, what allows to do in two or more steps complex operations, transmit more bytes than a single APDU can admit, etc.
-
-\hfil
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=0.75\linewidth]{gfx/APDUdialog}
-	\end{center}
-	\caption{APDU Command-Response Dialogue}
-	\label{fig:APDUdialog}
-\end{figure}
-
-Originally, the Lc and Le bytes had only 1 byte, if present, restricting the payload data to be at most 256 bytes long.
-An extension to the protocol changed the meaning of a Lc or Le 0x00 byte (256 bytes long payload), so when the byte corresponding to Lc or Le started with 0x00, the next two bytes where the real length.
-With this, an Extended APDU lets up to $65536$ bytes of data.
-
-The problem here, is that not all readers or smart cards support extended APDUs.
-Originally, to send more than 256 bytes of data in an APDU Command, a \textit{Put Data} instruction is defined, so the smart card stores the payload in a buffer, until other APDU Command indicates how to use it.
-
-To send more data in an APDU Response, the status bytes are set to: SW1=0x61 and SW2 to the remaining bytes to send.
-Because a smart card can't send APDU Commands, the card terminal must send a \textit{GET RESPONSE}, a special APDU Command, with Le set to the number of bytes specified in SW2.
-Iterating this process, the smart card can send as many bytes as it wants as Response.
-
-With the introduction of Extended APDUs, this technique is no longer needed.
-
-\subsection{MULTOS}
-
-MULTOS is a multi-application smart card operative system, which provides a custom developing environment, with rich documentation \citep{MultosTechLib}.
-MULTOS smart cards communicate like any other smart card following the standard, but internally offers a very specific architecture, affecting the way one must code applications for it.
-
-In this section we will present the main characteristics of a MULTOS smart card that shaped the ABC4Trust Card Lite code and that we had to be aware of when adapting it to IoT devices.
-
-\paragraph{MULTOS programming languages} A native assembly language called MEL, C and, to a lesser extent, Java, are the available languages to code for MULTOS.
-In our case, ABC4T Card Lite uses MEL and C.
-
-\paragraph{MULTOS Workflow}
-
-Most of the transmission and communication process is done by MULTOS core, and it then selects, based on the CLA byte of the APDU, the application to load.
-This application is what most developers will only worry about, and is where their $main()$ function will start.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/multosWorkflow}
-	\end{center}
-	\caption{MULTOS workflow}
-	\label{fig:multosWorkflow}
-\end{figure}
-
-The application uses then the \textit{multos.h} file that declares multiple global variables already loaded with the needed data, including the APDU Command bytes.
-
-Now the developer is in charge of checking what instruction was sent and if the APDU has the expected ISO Case.
-If everything is ok, code what needs to be run and write in specific data space the APDU Response bytes, call \textit{multosExit()} and MULTOS will be in charge to send the APDU Response.
-
-In summary, our application starts with all data loaded and exits without worrying how to send the answer.
-A very comfortable workflow that we must now implement for our IoT device if we would want ABC4T Card Lite code to work.
-
-\paragraph{MULTOS Memory Layout}
-
-Each application in MULTOS has access to a specific memory layout, divided in different categories:
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/multosMemLay}
-	\end{center}
-	\caption{MULTOS Memory Layout}
-	\label{fig:multosMemLay}
-\end{figure}
-
-The Code Space is where the application code is stored.
-The Data Space is divided in Static memory, Public memory and Dynamic memory.
-
-\textbf{Static memory} are the application variables declared after the specific \textit{\#pragma melstatic} compiler directive.
-These variables are stored in the non-volatile EEPROM, and any write is assured to be saved because they are not loaded into RAM.
-
-\textbf{Public memory} can be seen as the input/output buffer for applications and MULTOS system.
-The APDU header appears at the top of Public, and command data at the bottom.
-The application writes then the APDU Response bytes in Public, at specific position (see \autoref{fig:multosPubMem}).
-To declare variables in this data space, the \textit{\#pragma melpublic} directive is available.
-
-\textbf{Dynamic memory} works like usual program memory, with Session Data storing global variables and the Stack.
-The limited size of RAM in IoT devices and smart cards makes the use of dynamic memory not advisable.
-The compiler directive to use Session Data is \textit{\#pragma melsession}.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/multosPubMem}
-	\end{center}
-	\caption{MULTOS Public Memory Data Map}
-	\label{fig:multosPubMem}
-\end{figure}
-
-\hfil
-
-With regards to primitive types, to avoid confusion with their sizes, MULTOS defines and uses the following data types specified in \autoref{fig:multosDataTypes}.
-It's important to notice that MULTOS is Big Endian
-\marginpar{\includegraphics[width=\linewidth]{gfx/Big-Endian}\\Big-Endian, Wikipedia}
-and when storing structures there is no padding between defined variables, unlike modern compilers that perform data structure alignment \citep{dataStructAlign} for performance.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/multosDataTypes}
-	\end{center}
-	\caption{MULTOS Data Types}
-	\label{fig:multosDataTypes}
-\end{figure}
-
-\paragraph{MULTOS Standard C-API}
-
-A collection of more than a hundred functions are provided for arithmetic, cryptography, memory and smart card operations.
-The \textit{multos.h} interface provides access to these functions, that ultimately call their respective primitive instructions in assembly code.
-The primitive instructions are but a system call with an operation code, loading data in the needed registers.
-Therefore,  no implementation for these tools is available, nor in C, nor in assembly code.
-
-Nevertheless, the C-API documentation \citep{MultosTechLib} provides rich description for each function.

### 2017-05-19 Doc/classicthesis/Chapters/Draft.tex
fixes + organización

@@
 % From Idemix To P2ABCE: reasons
-
-\section{P2ABCE}
-
-In the \ac{P2ABCE} repository \citep{p2abcurl} is available the project's code, divided in two solutions: a complete P2ABCE implementation in Java and a Multos Smartcard implementation as companion for the project.
-
-The Java code is managed by a Maven project, structured using various known design patterns, but not of our interest.
-The structure we are actually interested in are the REST Services and their use of the Components classes, in which the smartcards are included.
-
-P2ABCE project is based on the concept of smartcards to store the credentials, logical or physical.
-An interface is defined to communicate with these smartcards, and then different implementations allow to use either \textit{Software Smartcards} or \textit{Hardware Smartcards}.
-
-The \textit{SoftwareSmartcard} class implements the interface in Java, suitable for tests and self-stored smartcards that any application using P2ABCE may need.
-
-The \textit{HardwareSmartcard} class uses the standard APDU messages [TODO:ref] to interact with smartcards.
-P2ABCE defines for every method in the mentioned interface, the necessary APDU instructions, and currently relies on \textit{javax.smartcardio} abstract classes (implemented by Oracle in their JRE) to communicate with the smartcard reader.
-This way, it doesn't matter what manufacturer issues the smartcard, or if it's an Android device, if they support the APDU API, P2ABCE will work with them.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/p2abceBasicUML}
-	\end{center}
-	\caption{Basic P2ABCE structure}
-	\label{fig:p2abceBasicUML}
-\end{figure}
-
-As a PoC the P2ABCE project includes the ABC4Trust Card Lite, an implementation for ML3-36K-R1 Multos Smartcards.
-The code is written in C, but is very dependent on the Multos framework, aside from numerous bugs and bad coding habits.
-
-At this stage, we have two options to implement our IoT device compatible with P2ABCE:
-
-\begin{itemize}
-	\item Implement in C the \textit{Smartcard} interface used by P2ABCE architecture, and use some communication protocol to remotelly call the methods from the machine running the P2ABC Engine.
-	\item Present the IoT device as a hardware smart card, using the APDU protocol (already defined, standard and with minimal overload).
-Providing a \textit{javax.smartcardio} ``IoT implementation'' to communicate with the IoT device through a transmission protocol, the already existing \textit{HardwareSmartcard} class can work with the new \textit{IoTSmartcard} in the IoT device.
-\end{itemize}
-
-%TODO : hablar más de los servicios REST
-
-\section{ABC4Trust Card Lite}
-
-P2ABCE provides a smart card reference implementation, ABC4Trust Card Lite \citep{ABC4TCardLite}. It supports device-bound U-Prove and Idemix, and virtually any discrete logarithm based pABC system.
-
-Version 1.2 is based on MULTOS ML3 cards, with approximately 64KB of EEPROM (non-volatile memory), 1KB of RAM and an Infineon SLE 78 microcontroller, a 16-bit based CPU aimed for chip cards.
-
-The card stores the user's private key $x$ and any \ac{BLOB} that the P2ABCE may need (like user's credentials).
-Then P2ABCE delegates the cryptographic operations on the smart card, that operates with $x$.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/ABC4TCardLite}
-	\end{center}
-	\caption{ABC4Trust Card Lite}
-	\label{fig:ABC4TCardLite}
-\end{figure}
-
-The cryptographic operations performed by the smartcard are the modular exponentiation and addition that discrete logarithm ZKPs are based on.
-
-\hfil
-
-The code is available from the P2ABCE project and has some good and bad points to have in count:
-
-The best asset of this code is that it's written in C aiming to a very constrained device, similar in computational power to many IoT devices, and very limited memory.
-
-Some \textit{tricks} in the code include using \textit{union} data types for variables that will be stored on the same data location, but at different moments (e.g.
-depending on APDU Command INS byte), minimizing this way the use of RAM and making code readability better; or strong use of pointers and \textit{memcpy} calls to copy structures with multiple variables as arrays of bytes.
-
-Among the many drawbacks, we could highlight the awful coding, the strong dependency on MULTOS framework and some bugs found.
-
-The code is structured in two files, \textit{main.h} and \textit{main.c}, with 557 and 5157 lines of code respectively.
-
-The file \textit{main.h} is mostly a reimplementation in assembly MEL of some MULTOS functionality already offered with latest \textit{multos.h}.
-
-The \textit{main.c} consists on near 600 lines of variables and data structures declarations, followed by the \textit{main()} function, a 2635 lines long \textit{switch-case} with practically no comments, and to conclude, the implementation of thirty functions called \textit{Subroutines} at the end of the file.
-
-This gives an idea of the problematic to maintain or even understand the code.
-But once one studies MULTOS framework in deep and applies many refactoring techniques to ABC4T Card's code, this becomes the best starting point for the IoT version.
 
@@
 For example, Atmel's ATAES132A \citep{ATAES132A}
-\marginpar{\includegraphics[width=\linewidth]{gfx/atmel} \\ Atmel's cryptography chips.}
+\marginpar{\includegraphics[width=0.9\linewidth]{gfx/atmel} \\ Atmel's cryptography chips.}
 offers a serial chip for secure key storage, AES128 execution and random number generation.
@@
 
-\section{Performance Evaluation}
-
-\subsection{Testbed description}
-
-% Hardware, network,
-
-\subsubsection{P2ABCE setting}
-
-To test the correct execution of the IoT smart card, we will use the ABC system from the tutorial in the P2ABCE Wiki \citep{p2abcurlwiki}.
-It is based on a soccer club which wishes to issue VIP-tickets to a soccer match.
-The VIP-member number in the ticket is inspectable for a lottery, ie.
-after the game, a random presentation token is inspected and the winning member is notified.
-
-First the various entities are \textbf{setup}, where several artifacts are generated and distributed. Then a ticket credential containing the following attributes is issued:
-
-\begin{verbatim}
-First name: John
-Last name: Dow
-Birthday: 1985-05-05Z
-Member number: 23784638726
-Matchday: 2013-08-07Z
-\end{verbatim}
-
-During \textbf{issuance}, a \textit{scope exclusive pseudonym} is established and the newly issued credential is bound to this pseudonym.
-This ensures that the ticket credential can not be used without the smart card.
-
-Then \textbf{presentation} is performed.
-The \textit{presentation policy} specifies that the member number is inspectable and a predicate ensures that the matchday is in fact $2013-08-07Z$.
-This last part ensures that a ticket issued for another match can not be used.
-
-The ticket holder was lucky and his presentation token was chosen in the lottery. The presentation token is therefore inspected.
-
-\hfil
-
-\subsubsection{Execution environment}
-
-First we will execute the test in our development machine (laptop).
-After asserting that the services work as expected, we then run the test in a Raspberry Pi 3, exactly like in the laptop.
-Finally, we will deploy the IoT smart card in a Omega2 and the delegation services in the Raspberry Pi 3.
-After every test, we checked that the issuing and proving were successful, in case a cryptographic error appeared in the implementation.
-
-Lets have a closer look at the hardware of each device:
-
-\paragraph{Development laptop}\marginpar{
-	\vskip0pt
-	\includegraphics[width=0.8\linewidth]{gfx/DELL}
-
-	DELL XPS 15
-}
-Our device is a DELL XPS 15, with a Core i7-6700HQ at 3.5GHz quad core processor and 32GB of DDR4 RAM, running Ubuntu 16.10.
-
-This is a powerful machine that can simulate the performance of many servers and clients that would implement P2ABCE in a real environment, giving a reference point for performance comparisons.
-
-\paragraph{Raspberry Pi 3}\marginpar{
-	\vskip0pt
-	\includegraphics[width=0.8\linewidth]{gfx/RPi3}
-
-	Raspberry Pi 3
-} A familiar environment, powerful enough to debug and hold the delegated P2ABCE Java services of P2ABCE with its 1GB of RAM, and with two network interfaces, perfect to work as the gateway for the IoT devices to the Internet.
-
-Only a microSD with enough space to burn the OS is needed to plug\&play with the Raspberry Pi.
-We use Raspbian, a stable Debian based distro, recommended by the Raspberry Pi designers, and ready to use with the P2ABCE compiled \textit{self-contained .jar} services.
-
-\begin{table}[h]
-	\myfloatalign
-	\begin{tabularx}{0.75\textwidth}{ll} \toprule
-		CPU & ARMv8 64bit quad-core @1.2GHz \\
-		RAM & 1GB \\
-		Storage & microSD \\
-		Firmware & Raspbian (Debian based distro) \\
-		Connectivity & Wifi n + Ethernet \\
-		Power & 5V 2A \\
-		\bottomrule
-	\end{tabularx}
-	\caption[Raspberry Pi 3 Specifications]{Raspberry Pi 3 Specifications.}
-	\label{tab:RPi3Specs}
-\end{table}
-
-\paragraph{Onion Omega2}\marginpar{
-	\vskip0pt
-	\includegraphics[width=\linewidth]{gfx/Omega2}
-
-	Onion Omega2
-} A device that falls inside the category of IoT, powerful enough to run a Linux environment, LEDE, where we can develop and debug the first PoC without troubling ourselves with problems not related to the project itself.
-
-\begin{table}[h]
-	\myfloatalign
-	\begin{tabularx}{0.75\textwidth}{ll} \toprule
-		MCU & Mediatek MT688 \citep{MT7688} \\
-		CPU & MIPS32 24KEc 580MHz \\
-		RAM & 64MB \\
-		Storage & 16MB \\
-		Firmware & LEDE (OpenWRT fork distro) \\
-		Connectivity & Wifi b/g/n \\
-		Power & 3.3V 300mA \\
-		\bottomrule
-	\end{tabularx}
-	\caption[Onion Omega 2 Specifications]{Onion Omega2 Specifications.}
-	\label{tab:Omega2Specs}
-\end{table}
-
-Nonetheless, the Omega2 needs fine tunning to start operating, and basic knowledge of electronics to make it work. The two main things to begin with Omega2 are:
-\begin{itemize}
-	\item A reliable 3.3V with a maximum of 800mA power supply, e.g.
-a USB2.0 with a step-down circuit, with quality soldering and wires to avoid unwanted resistances.
-The Omega2 will usually use up to 350mA, when the WiFi module is booting up.
-The mean consumption is about 200mA.
-	\item A Serial to USB adapter wired to the TX and RX UART pins to use the Serial Terminal, to avoid the use of SSH over WiFi.
-\end{itemize}
-
-Also, we will need a cross-compiler to generate the smart card binary.
-Because we will use GMP and OpenSSL as shared libraries, the best option is to use the LEDE SDK.
-The SDK manages the available packages and generates the GCC toolchain.
-Also, using CMake,
-%TODO
-
-\paragraph{The network} In our third scenario, the Raspberry Pi 3 and the Omega2 will talk to each other over TCP.
-This implies possible network delays depending on the quality of the connection.
-The Raspberry Pi 3 is connected over Ethernet to a switch with WiFi access point.
-The Omega2 is connected over WiFi n to said AP.
-To ensure the delay wasn't significant, we measured 6000 APDU messages, and the results show that the mean transmission time is less than half a millisecond per APDU.
-As we will see in the results section, this network time is negligible.
-
-\paragraph{Future work in tests} The lack of a physical MULTOS smart card precludes us to load and test the ABC4Trust code and measure the time P2ABCE needed using the \textit{HardwareSmartcard} class.
-This would be really interesting because for a single method from the \textit{Smartcard} interface, \textit{HardwareSmartcard} needs to send various APDUs, but \textit{SoftwareSmartcard} can perform the operations immediately.
-
-\subsection{The test code}
-
-There are three pieces of software that conform the test: the P2ABCE services, the IoT smart card, and a shell script automatizing the REST calls to all the services.
-
-\paragraph{P2ABCE services} This is the common part to our three sets.
-The services are compiled in a self-contained Jetty web server, or in WAR format, ready to be deployed in a server like Tomcat.
-We use the same JAR files with the Jetty web server for the PC and Raspberry Pi.
-
-We modified the User Service code to measure the execution time for each REST method.
-The time stamp starts inside the method and ends before it's \textit{return} instruction.
-This way, we filter the time the server spends processing the HTTP protocol and deciding which Java method to call.
-
-\paragraph{IoT smart card} Our C implementation of the P2ABCE smart card, compiled for the Omega2, with BIOSC listening on port 8888 for the APDU messages.
-
-We tested the execution in two modes, a full logging where every step was printed in terminal, and another one with no logging.
-With the first mode, we can check a proper execution, every byte exchanged, and with the second one, we measure the execution without unnecessary I/O.
-
-\paragraph{Shell script} To orchestrate all the services we use a simple script that performs the REST calls using  \textit{curl}.
-Here we perform the mentioned steps: setup of the P2ABCE system, issuance of the credential and a prove for the presentation policy.
-
-In the setup, the system parameters are generated, indicating key sizes of 1024 bits, the ones currently supported by the ABC4Trust and IoT smart cards.
-Then the system parameters and public keys from the services (issuer, inspector, revocation authority) are also exchanged.
-
-In the issuance, a two step protocol is performed by the User.
-The Issuer and the User send each other the XML data with the cryptographic information, and finish the protocol with a credential issued inside a smart card, software (Java) or hardware (physical or IoT).
-
-Finally, a Verifier sends the User a \textit{presentation policy}. The User generates a \textit{presentation token} for the Verifier and Inspector.
-
-%%%%%
-There are two script versions, one where every REST call is done from the same machine, therefore avoiding the exchange of the XML files by other means (like \textit{scp}); the other one is ready to be executed on a real setup, two scripts, one for the Omega2, where the REST calls are the delegation on the Raspberry Pi 3, and other for the PC, running the Issuer and Verifier as third party entities.
-The XML files generated by each service can be sent with the same protocol the IoT device and the Verifier, for example, would communicate in the real world.
-This way it's clearer that the \textit{curl} commands are the delegation protocol for the IoT device, and then the IoT device can send the XML as it wishes, over the Internet, Bluetooth, NFC, etc.
-%%%%%
-
-\subsection{Results}
-
-After 15 executions for each scenario (laptop, RPi3, Omega2), we take the means and compare each step.
-
-It is worth noting first that during the test, the measured use of the CPU showed that P2ABCE does not benefit of multi core processors, therefore, it only uses one of the four cores in the Raspberry Pi 3.
-
-\paragraph{The network}
-To test the network we sent six thousand APDUs to the Omega2, but instead of calling the \textit{APDU handler}, the Omega2 responded with the same bytes back.
-This way, the Omega2 only performed the simple BIOSC protocol, reading and writing the TCP socket.
-The APDUs had multiple sizes, taken from the most common APDUs logged in a successful execution. The test showed that our test network speed was around $0.014$ ms per byte.
-
-\paragraph{The setup}
-Because the Omega2 doesn't intervene until the creation of the smart card, the times measured in the second and third scenarios are practically identical.
-
-\begin{figure}[bth]
-	\myfloatalign
-	\subfloat[Times and relative speedup]
-	{\includegraphics[width=\linewidth]{gfx/graphics/setuptable}} \quad
-	\subfloat[Comparison graph]
-	{\includegraphics[width=.8\linewidth]{gfx/graphics/setup}} \\
-	\caption{Setup times (milliseconds)}
-	\label{fig:setup:graph}
-\end{figure}
-
-As we can see, the laptop is about ten times faster than Raspberry Pi 3, but considering that the highest time is less than two and a half seconds, and that the setup is done only once, this isn't a worrisome problem.
-
-\paragraph{Creation of the smart card}
-Here we create a \textit{SoftwareSmartcard} or a \textit{HardwareSmartcard} object that the User service will use in the following REST calls.
-
-The REST method to create a \textit{SoftwareSmartcard} is $/createSmartcard$, and to create a \textit{HardwareSmartcard}, using the \textit{IoTsmartcardio} implementation, we use $/initIoTsmartcard$.
-
-\begin{figure}[bth]
-	\myfloatalign
-	\subfloat[Times and relative speedup]
-	{\includegraphics[width=0.5\linewidth]{gfx/graphics/createSCtable}} \quad
-	\subfloat[Comparison graph]
-	{\includegraphics[width=0.45\linewidth]{gfx/graphics/createSC}} \\
-	\caption{Create smart card times (seconds)}
-	\label{fig:createSmartCard:graph}
-\end{figure}
-
-From \autoref{fig:createSmartCard:graph} we see that the RPi3 is about 16 times slower than the laptop in the creation of the \textit{SoftwareSmartcard}, but almost 9 times faster than the setup of the smart card in the Omega2 using APDUs.
-This gives us that the laptop is 145 times faster than the combination of RPi3 and Omega2 in our IoT deployment.
-But looking at the times, this process lasts up to 20 seconds, making it something feasible.
-
-Again, this operation is done only once per device, and includes commands from the creation of the PIN and PUK of the smart card, to storing the system parameters of P2ABCE, equivalent to the previous setup step.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=0.8\linewidth]{gfx/APDUsInitIoTSC}
-	\end{center}
-	\caption{Init IoT Smart Card APDU Commands exchanged.}
-	\label{fig:APDUsInitIoTSC}
-\end{figure}
-
-This is the first interaction between the RPi3 and the Omega2 IoT smart card, \textbf{using 30 APDU Commands, and Responses}, as shown in \autoref{fig:APDUsInitIoTSC}, with a total of $1109$ bytes exchanged.
-From our network benchmark, using TCP sockets, the delay in the transmission is only around 15 and 20 ms, negligible as we said compared to the almost 20 seconds the operation lasts.
-
-\paragraph{Issuing of the credential}
-The issuance is done in three steps for the User service, shown in \autoref{fig:IssuanceInteraction} with a red note showing the start of each step for the User delegation, in green for the interactions between Issuer and User, and the darker red is the Identity service, choosing the first available identity to use.
-The arrows in the figure show the REST calls performed during the test, where the IoT device acts as User, the RPi3 hosts the P2ABCE delegation services and the laptop is the Issuer.
-
-The three delegation steps and the REST method called are:
-
-\begin{center}
-	\begin{tabular}{l|l}
-	First issuance protocol step & $/issuanceProtocolStep$ \\
-	Second issuance protocol step  & $/issuanceProtocolStepUi$ \\
-	\textit{(end of first step for the User)} & \\
-	Third issuance protocol step & $/issuanceProtocolStep$ \\
-	\textit{(second step for the User)}  & \\
-\end{tabular}
-\end{center}
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/IssuanceInteraction}
-	\end{center}
-	\caption{Issuance interaction.}
-	\label{fig:IssuanceInteraction}
-\end{figure}
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=0.7\linewidth]{gfx/IssuanceAPDUs}
-	\end{center}
-	\caption{Issuance APDU Commands.}
-	\label{fig:IssuanceAPDUs}
-\end{figure}
-
-As we can see, the three REST calls to the delegation User service involve communication with the smart card. We show in \autoref{fig:IssuanceAPDUs} the APDU Commands for each REST call.
-There are 45 APDU Commands in total, $3197$ bytes exchanged, that would introduce a latency of 45ms in the network, negligible.
-
-\begin{figure}[bth]
-	\myfloatalign
-	\subfloat[Times (ms) and relative speedup]
-	{\includegraphics[width=\linewidth]{gfx/graphics/issuancetable}} \quad
-	\subfloat[Comparison graph]
-	{\includegraphics[width=.8\linewidth]{gfx/graphics/issuance}} \\
-	\caption{Issuance times (milliseconds)}
-	\label{fig:issuance:graph}
-\end{figure}
-
-In \autoref{fig:issuance:graph} we have the times spent in each REST call.
-The laptop shows again to be many times faster than the other two scenarios, but the times are again feasible even for the IoT environment.
-
-Lets compare the Raspberry Pi 3 and the Omega2 executions.
-There is a correlation between the number of APDU Commands needed in each step with the increment in time when using the IoT smart card, that is, how much the delegation server.
-
-The first one only involved one APDU, with $33$ bytes total (Command and Response), and times are almost identical.
-The second call needed $34$ APDUs, with $1623$ bytes, and the increase in time is around tree times slower than the RPi3 on its own.
-The third call used $20$ APDUs, $1541$ bytes, and makes the IoT scenario almost 7 times slower.
-
-The analysis shows where there are more cryptographic operations involving the Omega2, and because the amount of data exchanged is minimal, the difference in processing power between Omega2 and Raspberry Pi 3 is clear.
-
-\paragraph{Presentation token}
-The final step of the test involves a Prove, or Presentation in P2ABCE, where the Verifier sends the User or Prover the Presentation Policy, and the User answers with the Presentation Token, without more steps.
-In \autoref{fig:ProvingInteraction}, using the same colors as in the Issuance interaction, we can see the delegation messages done by the Omega2.
-
-To ensure that all the process was successful, it's enough to check if the Verifier and the Inspector returned XML files, accepting the prove, or an error code.
-Of course, every execution measured in the test was successful.
-
-In \autoref{fig:APDUsProving} the APDU Commands for each step are shown, $28$ in total, $1939$ bytes, $27$ms approx. in the network.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/ProvingInteraction}
-	\end{center}
-	\caption{Proving interaction.}
-	\label{fig:ProvingInteraction}
-\end{figure}
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=0.7\linewidth]{gfx/APDUsProving}
-	\end{center}
-	\caption{Proving APDU Commands.}
-	\label{fig:APDUsProving}
-\end{figure}
-
-Again, as shown in \autoref{fig:proving:graph}, there is a correlation between the number of APDU Commands used, the work the IoT smart card must perform, and the time measured.
-The $20$ APDU Commands in the first call make the IoT deployment almost $8$ times slower than the Raspberry Pi 3; but with only $8$ APDU Commands, the second one is less than $1.5$ times slower.
-
-Nonetheless, it's significant the difference in performance between the laptop and the Raspberry Pi 3 in the last REST call, more than $40$ times slower, even using the \textit{SoftwareSmartcard}.
-
-\begin{figure}[bth]
-	\myfloatalign
-	\subfloat[Times (ms) and relative speedup]
-	{\includegraphics[width=.8\linewidth]{gfx/graphics/provingtable}} \quad
-	\subfloat[Comparison graph]
-	{\includegraphics[width=.7\linewidth]{gfx/graphics/proving}} \\
-	\caption{Proving times (milliseconds)}
-	\label{fig:proving:graph}
-\end{figure}
-
-Unlike the previous steps, the Presentation or Proving is done more than once, being the key feature of ZKP protocols.
-The laptop performs a prove in less than one second, the RPi3 needs $15$ seconds, but our P2ABCE IoT deployment needs $15$ seconds for the first step, and $18$s for the second step, $33$ seconds total to generate a Presentation Token.
-
-\hfil
-
-\paragraph{Memory usage on the Omega2}
-Using the tool \textit{time -v} we can get a lot of useful information about a program, once it finishes.
-In our case, the binary with BIOSC and the smart card logic starts as an empty smart card, goes through the described process, and then we can stop it, as the User Service won't use it anymore.
-
-After another round of tests, now using \textit{time -v}, the field named \textit{Maximum resident set size (kbytes)} shows the \textbf{maximum} size of RAM used by the process since its launch.
-In our case, this involves the use of static memory for the \textit{global variables} of the smart card logic, and the dynamic memory used by the third party libraries, like GMPlib, OpenSSL and cJSON.
-
-GMP and OpenSSL always allocate the data in their own ADT, what involves copying the arrays of bytes representing the big modular integers from the cryptographic operations.
-cJSON, used in the serialization of the smart card for storage, and debug being human readable, stores a copy of every saved variable in the JSON tree structure, then creates a string (array of char) with the JSON, that the user can write to a file.
-
-Understanding the many bad uses of memory done in this PoC is important for future improvements and ports.
-A custom modular library using the same array of bytes that the smart card logic, a binary serialization, and many improvements, are our future work.
-
-With all that said, the mean of the maximum memory usage measured is $6569.6$ kbytes.
-Compared to the $64$MB of RAM available in the Omega2, our PoC could be executed in more constrained devices, given the system is compatible.
-
 %%%%%%%%%%%%%%%%%%%%%%%

### 2017-05-19 Doc/classicthesis/Chapters/Objectives.tex
fixes + organización

@@
 
+%TODO : hablar más de los servicios REST
+
 \subsection{ABC4Trust Card Lite}
@@
 But once one studies MULTOS framework in deep and applies many refactoring techniques to ABC4T Card's code, this becomes the best starting point for the IoT version.
+
+\section{Preliminaries about smart cards}
+
+\subsection{Smart Card Communication Protocol}
+
+To communicate the smart cards and the reader the ISO/IEC 7816-4 \citep{APDUISO} specifies a standardized protocol .
+
+The messages, also kown as \acp{APDU}, are divided in APDU Commands and APDU Responses.
+
+\textbf{APDU Commands} consist in 4 mandatory bytes (CLA, INS, P1, P2), and an optional payload.
+
+\begin{itemize}
+	\item CLA byte: Instruction class. Denotes if the command is interindustry standard or proprietary.
+	\item INS byte: Instruction code. Indicates the specific command.
+	\item P1, P2 bytes: Instruction parameters.
+	\item Lc, 0-3 bytes: Command data length.
+	\item Command data: Lc bytes of data.
+	\item Le, 0-3 bytes: Expected response data length.
+\end{itemize}
+
+This way, minimal number of bytes are needed to transmit commands to the smart card, allowing manufacturer's personalization of the smart card behavior and capabilities along with standard operations.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=0.55\linewidth]{gfx/APDUCommand}
+	\end{center}
+	\caption{APDU Command}
+	\label{fig:APDUCommand}
+\end{figure}
+
+\textbf{APDU Responses} are generated inside the smart card, always as an answer to an APDU Command. They consist on an optional payload and two mandatory status bytes.
+
+\begin{itemize}
+	\item Response data: At most Le bytes of data.
+	\item SW1-SW2 bytes: Status bytes. Encode the exit status of the instruction.
+\end{itemize}
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=0.55\linewidth]{gfx/APDUResponse}
+	\end{center}
+	\caption{APDU Response}
+	\label{fig:APDUResponse}
+\end{figure}
+
+\hfil
+
+The transmission protocol varies between different types of readers and smart cards (e.g.
+chip, contact-less), but what is common between every smart card interaction, is the \textit{APDU Command-Response Dialogue}.
+As long as the smart card has a power supply, it can maintain the memory in RAM between APDU Commands, what allows to do in two or more steps complex operations, transmit more bytes than a single APDU can admit, etc.
+
+\hfil
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=0.75\linewidth]{gfx/APDUdialog}
+	\end{center}
+	\caption{APDU Command-Response Dialogue}
+	\label{fig:APDUdialog}
+\end{figure}
+
+Originally, the Lc and Le bytes had only 1 byte, if present, restricting the payload data to be at most 256 bytes long.
+An extension to the protocol changed the meaning of a Lc or Le 0x00 byte (256 bytes long payload), so when the byte corresponding to Lc or Le started with 0x00, the next two bytes where the real length.
+With this, an Extended APDU lets up to $65536$ bytes of data.
+
+The problem here, is that not all readers or smart cards support extended APDUs.
+Originally, to send more than 256 bytes of data in an APDU Command, a \textit{Put Data} instruction is defined, so the smart card stores the payload in a buffer, until other APDU Command indicates how to use it.
+
+To send more data in an APDU Response, the status bytes are set to: SW1=0x61 and SW2 to the remaining bytes to send.
+Because a smart card can't send APDU Commands, the card terminal must send a \textit{GET RESPONSE}, a special APDU Command, with Le set to the number of bytes specified in SW2.
+Iterating this process, the smart card can send as many bytes as it wants as Response.
+
+With the introduction of Extended APDUs, this technique is no longer needed.
+
+\subsection{MULTOS}
+
+MULTOS is a multi-application smart card operative system, which provides a custom developing environment, with rich documentation \citep{MultosTechLib}.
+MULTOS smart cards communicate like any other smart card following the standard, but internally offers a very specific architecture, affecting the way one must code applications for it.
+
+In this section we will present the main characteristics of a MULTOS smart card that shaped the ABC4Trust Card Lite code and that we had to be aware of when adapting it to IoT devices.
+
+\paragraph{MULTOS programming languages} A native assembly language called MEL, C and, to a lesser extent, Java, are the available languages to code for MULTOS.
+In our case, ABC4T Card Lite uses MEL and C.
+
+\paragraph{MULTOS Workflow}
+
+Most of the transmission and communication process is done by MULTOS core, and it then selects, based on the CLA byte of the APDU, the application to load.
+This application is what most developers will only worry about, and is where their $main()$ function will start.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=0.6\linewidth]{gfx/multosWorkflow}
+	\end{center}
+	\caption{MULTOS Workflow}
+	\label{fig:multosWorkflow}
+\end{figure}
+
+The application uses then the \textit{multos.h} file that declares multiple global variables already loaded with the needed data, including the APDU Command bytes.
+
+Now the developer is in charge of checking what instruction was sent and if the APDU has the expected ISO Case.
+If everything is ok, code what needs to be run and write in specific data space the APDU Response bytes, call \textit{multosExit()} and MULTOS will be in charge to send the APDU Response.
+
+In summary, our application starts with all data loaded and exits without worrying how to send the answer.
+A very comfortable workflow that we must now implement for our IoT device if we would want ABC4T Card Lite code to work.
+
+\paragraph{MULTOS Memory Layout}
+
+Each application in MULTOS has access to a specific memory layout, divided in different categories:
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=0.8\linewidth]{gfx/multosMemLay}
+	\end{center}
+	\caption{MULTOS Memory Layout}
+	\label{fig:multosMemLay}
+\end{figure}
+
+The Code Space is where the application code is stored.
+The Data Space is divided in Static memory, Public memory and Dynamic memory.
+
+\textbf{Static memory} are the application variables declared after the specific \textit{\#pragma melstatic} compiler directive.
+These variables are stored in the non-volatile EEPROM, and any write is assured to be saved because they are not loaded into RAM.
+
+\textbf{Public memory} can be seen as the input/output buffer for applications and MULTOS system.
+The APDU header appears at the top of Public, and command data at the bottom.
+The application writes then the APDU Response bytes in Public, at specific position (see \autoref{fig:multosPubMem}).
+To declare variables in this data space, the \textit{\#pragma melpublic} directive is available.
+
+\textbf{Dynamic memory} works like usual program memory, with Session Data storing global variables and the Stack.
+The limited size of RAM in IoT devices and smart cards makes the use of dynamic memory not advisable.
+The compiler directive to use Session Data is \textit{\#pragma melsession}.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/multosPubMem}
+	\end{center}
+	\caption{MULTOS Public Memory Data Map}
+	\label{fig:multosPubMem}
+\end{figure}
+
+\hfil
+
+With regards to primitive types, to avoid confusion with their sizes, MULTOS defines and uses the following data types specified in \autoref{fig:multosDataTypes}.
+It's important to notice that MULTOS is Big Endian
+\marginpar{\includegraphics[width=\linewidth]{gfx/Big-Endian}\\Big-Endian, Wikipedia}
+and when storing structures there is no padding between defined variables, unlike modern compilers that perform data structure alignment \citep{dataStructAlign} for performance.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/multosDataTypes}
+	\end{center}
+	\caption{MULTOS Data Types}
+	\label{fig:multosDataTypes}
+\end{figure}
+
+\paragraph{MULTOS Standard C-API}
+
+A collection of more than a hundred functions are provided for arithmetic, cryptography, memory and smart card operations.
+The \textit{multos.h} interface provides access to these functions, that ultimately call their respective primitive instructions in assembly code.
+The primitive instructions are but a system call with an operation code, loading data in the needed registers.
+Therefore,  no implementation for these tools is available, nor in C, nor in assembly code.
+
+Nevertheless, the C-API documentation \citep{MultosTechLib} provides rich description for each function.
 

### 2017-05-19 Doc/classicthesis/Chapters/Validation.tex
fixes + organización

%************************************************
\chapter{Validation and Performance Evaluation}\label{ch:validation}
%************************************************

In this chapter, we will describe the deployment of three testing scenarios: a laptop, a Raspberry Pi 3, and a Omega2 IoT device with the Raspberry Pi 3 as the delegation server. We will measure and compare the results to determine if the proposed solution is feasible or must be submitted to revision.

\section{Testbed description}

First, we shall describe the example Attribute Based Credential system in use. Then, the hardware we will use in our benchmarking.

\subsection{P2ABCE setting}

To test the correct execution of the \textit{IoT smart card}, we will use the ABC system from the tutorial in the P2ABCE Wiki \citep{p2abcurlwiki}. It is based on a soccer club, which wishes to issue VIP-tickets for a match. The VIP-member number in the ticket is inspectable for a lottery, ie. after the game, a random presentation token is inspected and the winning member is notified.

First the various entities are \textbf{setup}, where several artifacts are generated and distributed. Then a ticket credential containing the following attributes is issued:

\begin{verbatim}
First name: John
Last name: Dow
Birthday: 1985-05-05Z
Member number: 23784638726
Matchday: 2013-08-07Z
\end{verbatim}

During \textbf{issuance}, a \textit{scope exclusive pseudonym} is established and the newly issued credential is bound to this pseudonym. This ensures that the ticket credential can not be used without the smart card.

Then \textbf{presentation} is performed. The \textit{presentation policy} specifies that the member number is inspectable and a predicate ensures that the matchday is in fact $2013-08-07Z$. This last part ensures that a ticket issued for another match can not be used.

The ticket holder was lucky and his presentation token was chosen in the lottery. The presentation token is therefore inspected.

\hfil

\subsection{Execution environment}

First we will execute the test in our development machine (laptop). After asserting that the services work as expected, we then run the test in a Raspberry Pi 3, exactly like in the laptop. Finally, we will deploy the IoT smart card in a Omega2 and the delegation services in the Raspberry Pi 3. After every test, we checked that the issuing and proving were successful, in case a cryptographic error appeared in the implementation.

Lets have a closer look at the hardware of each device:

\paragraph{Development laptop}\marginpar{
	\vskip0pt
	\includegraphics[width=0.8\linewidth]{gfx/DELL}

	DELL XPS 15
}
Our device is a DELL XPS 15, with a Core i7-6700HQ at 3.5GHz quad core processor and 32GB of DDR4 RAM, running Ubuntu 16.10.

This is a powerful machine that can simulate the performance of many servers and clients that would implement P2ABCE in a real environment, giving a reference point for performance comparisons.

\paragraph{Raspberry Pi 3}\marginpar{
	\vskip0pt
	\includegraphics[width=0.8\linewidth]{gfx/RPi3}

	Raspberry Pi 3
} A familiar environment, powerful enough to debug and hold the delegated P2ABCE Java services of P2ABCE with its 1GB of RAM, and with two network interfaces, perfect to work as the gateway for the IoT devices to the Internet.

Only a microSD with enough space to burn the OS is needed to plug\&play with the Raspberry Pi. We use Raspbian, a stable Debian based distro, recommended by the Raspberry Pi designers, and ready to use with the P2ABCE compiled \textit{self-contained .jar} services.

\begin{table}[h]
	\myfloatalign
	\begin{tabularx}{0.75\textwidth}{ll} \toprule
		CPU & ARMv8 64bit quad-core @1.2GHz \\
		RAM & 1GB \\
		Storage & microSD \\
		Firmware & Raspbian (Debian based distro) \\
		Connectivity & Wifi n + Ethernet \\
		Power & 5V 2A \\
		\bottomrule
	\end{tabularx}
	\caption[Raspberry Pi 3 Specifications]{Raspberry Pi 3 Specifications.}
	\label{tab:RPi3Specs}
\end{table}

\paragraph{Onion Omega2}\marginpar{
	\vskip0pt
	\includegraphics[width=\linewidth]{gfx/Omega2}

	Onion Omega2
} A device that falls inside the category of IoT, powerful enough to run a Linux environment, LEDE, where we can develop and debug the first PoC without troubling ourselves with problems not related to the project itself.

\begin{table}[h]
	\myfloatalign
	\begin{tabularx}{0.75\textwidth}{ll} \toprule
		MCU & Mediatek MT688 \citep{MT7688} \\
		CPU & MIPS32 24KEc 580MHz \\
		RAM & 64MB \\
		Storage & 16MB \\
		Firmware & LEDE (OpenWRT fork distro) \\
		Connectivity & Wifi b/g/n \\
		Power & 3.3V 300mA \\
		\bottomrule
	\end{tabularx}
	\caption[Onion Omega 2 Specifications]{Onion Omega2 Specifications.}
	\label{tab:Omega2Specs}
\end{table}

Nonetheless, the Omega2 needs fine tunning to start operating, and basic knowledge of electronics to make it work. The two main things to begin with Omega2 are:
\begin{itemize}
	\item A reliable 3.3V with a maximum of 800mA power supply, e.g. a USB2.0 with a step-down circuit, with quality soldering and wires to avoid unwanted resistances. The Omega2 will usually use up to 350mA, when the WiFi module is booting up. The mean consumption is about 200mA.
	\item A Serial to USB adapter wired to the TX and RX UART pins to use the Serial Terminal, to avoid the use of SSH over WiFi.
\end{itemize}

Also, we will need a cross-compiler to generate the smart card binary. Because we will use GMP and OpenSSL as shared libraries, the best option is to use the LEDE SDK. The SDK manages the available packages and generates the GCC toolchain. Also, using CMake, 
%TODO

\paragraph{The network} In our third scenario, the Raspberry Pi 3 and the Omega2 will talk to each other over TCP. This implies possible network delays depending on the quality of the connection. The Raspberry Pi 3 is connected over Ethernet to a switch with WiFi access point. The Omega2 is connected over WiFi n to said AP. To ensure the delay wasn't significant, we measured 6000 APDU messages, and the results show that the mean transmission time is less than half a millisecond per APDU. As we will see in the results section, this network time is negligible.

\paragraph{Future work for tests} The lack of a physical MULTOS smart card precludes us to load and test the ABC4Trust Card Lite's code and measure the time P2ABCE would need when using the \textit{HardwareSmartcard} class. This would be really interesting because for a single method from the \textit{Smartcard} interface, \textit{HardwareSmartcard} implementation needs to send multiple APDU Commands, but \textit{SoftwareSmartcard} can perform the operations immediately, with the full computer's resources.

\section{The test code}

In this section we present the scripts and binaries used during the tests.

There are three pieces of software that conform the test: the P2ABCE services, the IoT smart card, and shell scripts automatizing the REST calls, from the terminal.

\paragraph{P2ABCE services}\hfil

This is a common part to our three sets. The services are compiled in a self-contained Jetty web server, or in WAR format, ready to be deployed in a server like Tomcat. We use the same JAR files with the embedded Jetty web server for the PC and Raspberry Pi.

We modified the User Service code to measure the execution elapsed time for each REST method. We don't measure the time the web server spends processing the HTTP protocol and deciding which Java method to call.

\paragraph{IoT smart card}\hfil

Our C implementation of the P2ABCE smart card, compiled for the Omega2, with BIOSC listening on port 8888 for the APDU messages.

We tested the execution in two modes, a full logging where every step was printed in terminal, and another one with no logging. With the first mode, we can check a proper execution, every byte exchanged, and with the second one, we measure the execution without unnecessary I/O.

\paragraph{Shell script} \hfil

To orchestrate all the services we use a simple script that performs the REST calls using  \textit{curl}. Here we perform the mentioned steps: setup of the P2ABCE system, issuance of the credential and a prove for the presentation policy.

In the setup, the system parameters are generated, indicating key sizes of 1024 bits, the ones currently supported by the ABC4Trust and IoT smart cards. Then the system parameters and public keys from the services (issuer, inspector, revocation authority) are also exchanged.

In the issuance, a two step protocol is performed by the User. The Issuer and the User send each other the XML data with the cryptographic information, and finish the protocol with a credential issued inside a smart card, software (Java) or hardware (physical or IoT).

Finally, a Verifier sends the User a \textit{presentation policy}. The User generates a \textit{presentation token} for the Verifier and Inspector.

%%%%%
There are two script versions, one where every REST call is done from the same machine, therefore avoiding the exchange of the XML files by other means (like \textit{scp}); the other one is ready to be executed on a real setup, two scripts, one for the Omega2, where the REST calls are the delegation on the Raspberry Pi 3, and other for the PC, running the Issuer and Verifier as third party entities. The XML files generated by each service can be sent with the same protocol the IoT device and the Verifier, for example, would communicate in the real world. This way it's clearer that the \textit{curl} commands are the delegation protocol for the IoT device, and then the IoT device can send the XML as it wishes, over the Internet, Bluetooth, NFC, etc.
%%%%%

\section{Results}

After 20 executions for each scenario (laptop, RPi3, Omega2+RPi3), we take the means and compare each step of the testbed.

It is worth noting that during the test, the measured use of the CPU showed that P2ABCE does not benefit of parallelization, therefore, it only uses one of the four cores in the laptop and Raspberry Pi 3.

To test the network, we sent six thousand APDUs to the Omega2, but instead of calling the \textit{APDU handler}, the Omega2 responded with the same bytes back. This way, the Omega2 only performed the simple BIOSC protocol, reading from and writing to the TCP socket.
The APDUs had multiple sizes, taken from the most common APDUs logged in during a successful execution. The test showed that our network speed was around $0.014$ ms per byte.

\paragraph{The setup}\hfil

The first step of our testbed. The Omega2 doesn't intervene until the creation of the smart card, therefore, the times measured in the second and third scenarios are practically identical.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[Times and relative speedup]
	{\includegraphics[width=\linewidth]{gfx/graphics/setuptable}} \quad
	\subfloat[Comparison graph]
	{\includegraphics[width=.8\linewidth]{gfx/graphics/setup}} \\
	\caption{Setup times (milliseconds)}
	\label{fig:setup:graph}
\end{figure}

As we can see in \autoref{fig:setup:graph}, the laptop is about ten times faster than Raspberry Pi 3, but considering that the highest time is less than two and a half seconds, and that the setup is done only once, this isn't a worrisome problem.

\paragraph{Creation of the smart card}\hfil

Here we create a \textit{SoftwareSmartcard} or a \textit{HardwareSmartcard} object that the User service will use in the following REST calls.

The REST method to create a \textit{SoftwareSmartcard} is $/createSmartcard$, and to create a \textit{HardwareSmartcard}, using the \textit{IoTsmartcardio} implementation, we use $/initIoTsmartcard$.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[Times and relative speedup]
	{\includegraphics[width=0.5\linewidth]{gfx/graphics/createSCtable}} \quad
	\subfloat[Comparison graph]
	{\includegraphics[width=0.45\linewidth]{gfx/graphics/createSC}} \\
	\caption{Create smart card times (seconds)}
	\label{fig:createSmartCard:graph}
\end{figure}

From \autoref{fig:createSmartCard:graph} we see that the RPi3 is about 16 times slower than the laptop in the creation of the \textit{SoftwareSmartcard}, but almost 9 times faster than the setup of the smart card in the Omega2 using APDUs. This gives us that the laptop is 145 times faster than the combination of RPi3 and Omega2 in our IoT deployment.
But looking at the times, this process lasts up to 20 seconds, making it something feasible.

Again, this operation is done only once per device, and includes commands from the creation of the PIN and PUK of the smart card, to storing the system parameters of P2ABCE, equivalent to the previous setup step.

This is the first interaction between the RPi3 and the IoT smart card running in the Omega2. To setup the smart card \textbf{30 APDU Commands}, and their respective Responses, are exchanged, as shown in \autoref{ch:resultsdiagrams}, \autoref{fig:APDUsInitIoTSC}, with a total of $1109$ bytes. From our network benchmark, using TCP sockets, the delay in the transmission is only around 15 and 20 ms, negligible, as we said, compared to the almost 20 seconds the operation lasts.

\paragraph{Issuing of the credential}\hfil

The issuance is done in three steps for the User service, shown in \autoref{fig:IssuanceInteraction} with a red note showing the start of each step for the User delegation, in green for the interactions between Issuer and User, and the darker red is the Identity service, choosing the first available identity to use. The arrows in the figure show the REST calls performed during the test, where the IoT device acts as User, the RPi3 hosts the P2ABCE delegation services and the laptop is the Issuer.

The three delegation steps and the REST method called are:

\begin{center}
	\begin{tabular}{l|l}
	First issuance protocol step & $/issuanceProtocolStep$ \\
	Second issuance protocol step  & $/issuanceProtocolStepUi$ \\
	\textit{(end of first step for the User)} & \\
	Third issuance protocol step & $/issuanceProtocolStep$ \\
	\textit{(second step for the User)}  & \\
\end{tabular}
\end{center}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/IssuanceInteraction}
	\end{center}
	\caption{Issuance interaction.}
	\label{fig:IssuanceInteraction}
\end{figure}

As we can see, the three REST calls to the delegation User service involve communication with the smart card. We show in \autoref{ch:resultsdiagrams}, \autoref{fig:IssuanceAPDUs}, the APDU Commands used for each REST call in the issuance.
There are 45 APDU Commands in total, $3197$ bytes exchanged, that would introduce a latency of 45ms in the network, negligible.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[Times (ms) and relative speedup]
	{\includegraphics[width=\linewidth]{gfx/graphics/issuancetable}} \quad
	\subfloat[Comparison graph]
	{\includegraphics[width=.8\linewidth]{gfx/graphics/issuance}} \\
	\caption{Issuance times (milliseconds)}
	\label{fig:issuance:graph}
\end{figure}

In \autoref{fig:issuance:graph} we have the times spent in each REST call. The laptop shows again to be many times faster than the other two scenarios, but the times are again feasible even for the IoT environment.

Lets compare the Raspberry Pi 3 and the Omega2 executions. There is a correlation between the number of APDU Commands needed in each step with the increment in time when using the IoT smart card, that is, how much the delegation server.

The first one only involved one APDU, with $33$ bytes total (Command and Response), and times are almost identical. The second call needed $34$ APDUs, with $1623$ bytes, and the increase in time is around tree times slower than the RPi3 on its own. The third call used $20$ APDUs, $1541$ bytes, and makes the IoT scenario almost 7 times slower.

The analysis shows where there are more cryptographic operations involving the Omega2, and because the amount of data exchanged is minimal, the difference in processing power between Omega2 and Raspberry Pi 3 is clear.

\paragraph{Presentation token}\hfil

The final step of the test involves a Prove, or Presentation in P2ABCE, where the Verifier sends the User or Prover the Presentation Policy, and the User answers with the Presentation Token, without more steps. In \autoref{fig:ProvingInteraction}, using the same colors as in the Issuance interaction, we can see the delegation messages done by the Omega2.

To ensure that all the process was successful, it's enough to check if the Verifier and the Inspector returned XML files, accepting the prove, or an error code. Of course, every execution measured in the test was successful.

In \autoref{ch:resultsdiagrams}, \autoref{fig:APDUsProving}, we provide the APDU Commands for each step, $28$ in total, with $1939$ bytes, giving us about $27$ms of delay in the network transmission.

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/ProvingInteraction}
	\end{center}
	\caption{Proving interaction.}
	\label{fig:ProvingInteraction}
\end{figure}

Again, as shown in \autoref{fig:proving:graph}, there is a correlation between the number of APDU Commands used, the work the IoT smart card must perform, and the time measured. The $20$ APDU Commands in the first call make the IoT deployment almost $8$ times slower than the Raspberry Pi 3; but with only $8$ APDU Commands, the second one is less than $1.5$ times slower.

Nonetheless, it's significant the difference in performance between the laptop and the Raspberry Pi 3 in the last REST call, more than $40$ times slower, even using the \textit{SoftwareSmartcard}.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[Times (ms) and relative speedup]
	{\includegraphics[width=.8\linewidth]{gfx/graphics/provingtable}} \quad
	\subfloat[Comparison graph]
	{\includegraphics[width=.7\linewidth]{gfx/graphics/proving}} \\
	\caption{Proving times (milliseconds)}
	\label{fig:proving:graph}
\end{figure}

Unlike the previous steps, the Presentation or Proving is done more than once, being the key feature of ZKP protocols. The laptop performs a prove in less than one second, the RPi3 needs $15$ seconds, but our P2ABCE IoT deployment needs $15$ seconds for the first step, and $18$s for the second step, $33$ seconds total to generate a Presentation Token.

\hfil

\paragraph{Memory usage on the Omega2}\hfil

Using the tool \textit{time -v} we can get a lot of useful information about a program, once it finishes. In our case, the binary with BIOSC and the smart card logic starts as an empty smart card, goes through the described process, and then we can stop it, as the User Service won't use it anymore.

After another round of tests, now using \textit{time -v}, the field named \textit{Maximum resident set size (kbytes)} shows the \textbf{maximum} size of RAM used by the process since its launch. In our case, this involves the use of static memory for the \textit{global variables} of the smart card logic, and the dynamic memory used by the third party libraries, like GMPlib, OpenSSL and cJSON.

GMP and OpenSSL always allocate the data in their own ADT, what involves copying the arrays of bytes representing the big modular integers from the cryptographic operations. cJSON, used in the serialization of the smart card for storage, and debug being human readable, stores a copy of every saved variable in the JSON tree structure, then creates a string (array of char) with the JSON, that the user can write to a file.

Understanding the many bad uses of memory done in this PoC is important for future improvements and ports. A custom modular library using the same array of bytes that the smart card logic, a binary serialization, and many improvements, are our future work.

With all that said, the mean of the maximum memory usage measured is $6569.6$ kbytes. Compared to the $64$MB of RAM available in the Omega2, our PoC could be executed in more constrained devices, given the system is compatible.

%%%%%%%%%%%%%%%%%%%%%%%

% Consideraciones futuras: implementar directamente Smartcard interface vs seguir usando APDUs:  minimizan uso de red, facilitan protocolo de comunicación de las APDUs, son estándar, mejoran el mantenimiento del código: P2ABCE puede cambiar las interfaces, pero mantendrá compatibilidad con las smartcards, si implementamos Smartcard.java por cada cambio en la interfaz deberemos actualizar cada IoT smart card.

### 2017-05-22 Doc/classicthesis/Chapters/Chapter0A.tex
objectives + org.

@@
 %\markboth{\spacedlowsmallcaps{Appendix}}{\spacedlowsmallcaps{Appendix}}
+
+\chapter{Dockerfiles}\label{ch:docker}
+
+\begin{lstlisting}[language=bash,frame=tb,caption={Dockerfile for P2ABCE},label=lst:dockerP2ABCE]
+FROM openjdk:7
+
+COPY idemix-3.0.36-binaries/ /idemix/
+
+RUN apt-get update && apt-get install -y maven && rm -rf /var/lib/apt/lists/*
+
+RUN	cd /idemix/com/ibm/zurich/idmx/com.ibm.zurich.idmx.3_x_x/3.0.36 && \
+	mvn install:install-file     \
+	-Dfile=com.ibm.zurich.idmx.3_x_x-3.0.36.jar   \
+	-DpomFile=com.ibm.zurich.idmx.3_x_x-3.0.36.pom   && \
+
+	cd /idemix/com/ibm/zurich/idmx/com.ibm.zurich.idmx.interfaces/3.0.36  && \
+	mvn install:install-file      \
+	-Dfile=com.ibm.zurich.idmx.interfaces-3.0.36.jar  \
+	-DpomFile=com.ibm.zurich.idmx.interfaces-3.0.36.pom  && \
+
+	cd /idemix/com/ibm/zurich/idmx/pom/3.0.36     && \
+	mvn install:install-file          \
+	-Dfile=pom-3.0.36.pom          \
+	-DpomFile=pom-3.0.36.pom
+
+# docker run -it --name p2abce_dev_env -v /p2abcengine/source/code/:/usr/src/mymaven -w /usr/src/mymaven/Code/core-abce p2abce_env bash
+# docker start p2abce_dev_env
+# docker attach p2abce_dev_env
+\end{lstlisting}
+
+\hfil
+
+\begin{lstlisting}[language=bash,frame=tb,caption={Dockerfile for Omega2 SDK},label=lst:dockerLEDE]
+
+# Linux dev environment for LEDE project
+# Based on borromeotlhs' dockerfile
+
+FROM ubuntu:14.04
+
+LABEL maintainer="joseluis.canovas@outlook.com"
+LABEL description="Dockerfile for Onion Omega2 SDK environment. It can be modified for Omega2+."
+
+RUN apt-get update && DEBIAN_FRONTEND=noninteractive apt-get install -y \
+subversion g++ zlib1g-dev build-essential git python libncurses5-dev gawk gettext unzip file libssl-dev wget \
+&& rm -rf /var/lib/apt/lists/*
+
+# Install CMake 3.7.2
+RUN mkdir ~/temp &&\
+cd ~/temp &&\
+wget https://cmake.org/files/v3.7/cmake-3.7.2.tar.gz  &&\
+tar xzvf cmake-3.7.2.tar.gz &&\
+cd cmake-3.7.2/ &&\
+./bootstrap &&\
+make  &&\
+make install &&\
+rm -rf ~/temp
+
+RUN git clone https://github.com/lede-project/source.git lede
+
+RUN adduser omega &&  echo 'omega:omega' | chpasswd   && chown -R omega:omega lede
+WORKDIR lede
+USER omega
+
+RUN ./scripts/feeds update -a  && ./scripts/feeds install -a
+
+# Set SDK environment for Omega2
+# For Omega2+ change the third echo line with: (notice the 'p' for plus)
+#  echo "CONFIG_TARGET_ramips_mt7688_DEVICE_omega2p=y" > .config && \
+
+RUN echo "CONFIG_TARGET_ramips=y" > .config  && \
+echo "CONFIG_TARGET_ramips_mt7688=y" >> .config  && \
+echo "CONFIG_TARGET_ramips_mt7688_DEVICE_omega2=y" >> .config && \
+make defconfig
+
+RUN make tools/install
+RUN make toolchain/install
+
+ENV PATH "$PATH:/lede/staging_dir/toolchain-mipsel_24kc_gcc-5.4.0_musl/bin"
+ENV STAGING_DIR "/lede/staging_dir"
+ENV CFLAGS=-I/lede/staging_dir/target-mipsel_24kc_musl/usr/include
+ENV LDFLAGS=-L/lede/staging_dir/target-mipsel_24kc_musl/usr/lib
+
+# Use this command to run with shared directory:
+# docker run -it --name omega2-sdk-app -v /my_host_dir:/remote omega2-sdk bash
+
+\end{lstlisting}
+
 \chapter{Test: APDU Commands exchanged}\label{ch:resultsdiagrams}

### 2017-05-22 Doc/classicthesis/Chapters/Objectives.tex
objectives + org.

@@
 
-% TODO: explicar P2ABCE
+In the P2ABCE repository \citep{p2abcurl} there is available the project's code, divided in two solutions: a complete P2ABCE implementation in Java and a MULTOS smart card implementation as PoC for the project.
 
-In the \ac{P2ABCE} repository \citep{p2abcurl} is available the project's code, divided in two solutions: a complete P2ABCE implementation in Java and a Multos Smartcard implementation as companion for the project.
+\subsection{P2ABCE system}
+%TODO: actores y sus papeles
 
+\subsection{P2ABCE Structure and Services}
 The Java code is managed by a Maven project, structured using various known design patterns, but not of our interest.
-The structure we are actually interested in are the REST Services and their use of the Components classes, in which the smartcards are included.
+The part we are actually interested in are the \textit{REST Services} and their use of the \textit{Components} classes, where the smart card's logic and use are defined.
 
-P2ABCE project is based on the concept of smartcards to store the credentials, logical or physical.
-An interface is defined to communicate with these smartcards, and then different implementations allow to use either \textit{Software Smartcards} or \textit{Hardware Smartcards}.
+P2ABCE project is based on the concept of smart cards, virtual or physical, to store the credentials.
+An interface is defined to communicate with these smart cards, and then different implementations allow to use either \textit{Software Smartcards} or \textit{Hardware Smartcards}.
 
-The \textit{SoftwareSmartcard} class implements the interface in Java, suitable for tests and self-stored smartcards that any application using P2ABCE may need.
+The \textit{SoftwareSmartcard} class implements the interface in Java, suitable for applications using P2ABCE self-storing digital smart cards, like a virtual wallet.
 
-The \textit{HardwareSmartcard} class uses the standard APDU messages [TODO:ref] to interact with smartcards.
-P2ABCE defines for every method in the mentioned interface, the necessary APDU instructions, and currently relies on \textit{javax.smartcardio} abstract classes (implemented by Oracle in their JRE) to communicate with the smartcard reader.
-This way, it doesn't matter what manufacturer issues the smartcard, or if it's an Android device, if they support the APDU API, P2ABCE will work with them.
+The \textit{HardwareSmartcard} class uses the standard APDU messages (\ref{subsec:APDU}) to interact with smartcards.
+P2ABCE defines the necessary APDU instructions for the smart card needed to implement each method of the interface.
+It relies on \textit{javax.smartcardio} abstract classes (implemented by Oracle in their JRE) to communicate the smart card reader and the smart card.
+This way, it doesn't matter what manufacturer issues the smartcard, or if it's an Android device with NFC, if they support the APDU instructions, P2ABCE will work with them.
 
@@
 
-As a PoC the P2ABCE project includes the ABC4Trust Card Lite, an implementation for ML3-36K-R1 Multos Smartcards.
-The code is written in C, but is very dependent on the Multos framework, aside from numerous bugs and bad coding habits.
+\hfil
 
-At this stage, we have two options to implement our IoT device compatible with P2ABCE:
+The project also provides a set of REST services to control each role of the P2ABCE systeam (User, Issuer, Verifier, Inspector, etc.).
+The methods implemented include the creation of \textit{Software} smart cards within the User Service, and store them in a data base for future REST calls that may need them.
 
-\begin{itemize}
-	\item Implement in C the \textit{Smartcard} interface used by P2ABCE architecture, and use some communication protocol to remotelly call the methods from the machine running the P2ABC Engine.
-	\item Present the IoT device as a hardware smart card, using the APDU protocol (already defined, standard and with minimal overload).
-Providing a \textit{javax.smartcardio} ``IoT implementation'' to communicate with the IoT device through a transmission protocol, the already existing \textit{HardwareSmartcard} class can work with the new \textit{IoTSmartcard} in the IoT device.
-\end{itemize}
-
-%TODO : hablar más de los servicios REST
+The services receive parameters like the length of the cryptographic keys, IDs, or XML files to parse.
+The REST API is not meant for a User Service to communicate with an Issuer Service, for example, but for the User actor to call the Engine through the User Service, to perform specific actions, and then the Service returns the response XML.
+Those XML files are the way to communicate two different actors.
+The transmission method to exchange the XML depends on the specific scenario.
 
@@
 
-%TODO: as part of P2ABCE project, this MULTOS PoC
+As a PoC the P2ABCE project includes a smart card reference implementation, the ABC4Trust Card Lite \citep{ABC4TCardLite}.
+It supports device-bound U-Prove and Idemix, and virtually any discrete logarithm based pABC system.
 
-%TODO: check content
-
-P2ABCE provides a smart card reference implementation, ABC4Trust Card Lite \citep{ABC4TCardLite}. It supports device-bound U-Prove and Idemix, and virtually any discrete logarithm based pABC system.
-
-Version 1.2 is based on MULTOS ML3 cards, with approximately 64KB of EEPROM (non-volatile memory), 1KB of RAM and an Infineon SLE 78 microcontroller, a 16-bit based CPU aimed for chip cards.
+Version 1.2 is written for \texttt{ML3-36K-R1} MULTOS smart cards, with approximately 64KB of EEPROM (non-volatile memory), 1KB of RAM and an Infineon SLE 78 microcontroller, a 16-bit based CPU aimed for chip cards.
 
@@
 
-The cryptographic operations performed by the smartcard are the modular exponentiation and addition that discrete logarithm ZKPs are based on.
+The cryptographic operations performed by the smartcard are the modular exponentiation and addition used by ZKPs based on the discrete logarithm problem.
 
@@
 
-The code is available from the P2ABCE project and has some good and bad points to have in count:
+The code is available from the P2ABCE's repository and has some good and bad points to have in count:
 
-The best asset of this code is that it's written in C aiming to a very constrained device, similar in computational power to many IoT devices, and very limited memory.
+The best asset of this code is that it's written in C aiming to a very constrained device, with very limited memory and similar computational power to many other IoT devices.
 
-Some \textit{tricks} in the code include using \textit{union} data types for variables that will be stored on the same data location, but at different moments (e.g.
-depending on APDU Command INS byte), minimizing this way the use of RAM and making code readability better; or strong use of pointers and \textit{memcpy} calls to copy structures with multiple variables as arrays of bytes.
+The code uses some good practices when programming for IoT devices, like using \textit{union} data types for variables that will be stored on the same data location, but at different times, minimizing this way the use of RAM and making code readability better; or like the strong use of pointers and \textit{memcpy} calls to copy structures with multiple variables as arrays of bytes.
 
-Among the many drawbacks, we could highlight the awful coding, the strong dependency on MULTOS framework and some bugs found.
+Although, among the many drawbacks, we could highlight the \textit{awful} coding style, the strong dependency on MULTOS framework and many bugs found.
 
-The code is structured in two files, \textit{main.h} and \textit{main.c}, with 557 and 5157 lines of code respectively.
+The code is structured in two files, \textit{main.h} and \textit{main.c}, with around $550$ and $5200$ lines of code, respectively.
 
-The file \textit{main.h} is mostly a reimplementation in assembly MEL of some MULTOS functionality already offered with latest \textit{multos.h}.
+The file \textit{main.h} is mostly a reimplementation in assembly MEL code of some MULTOS functionality already offered by their API.
 
-The \textit{main.c} consists on near 600 lines of variables and data structures declarations, followed by the \textit{main()} function, a 2635 lines long \textit{switch-case} with practically no comments, and to conclude, the implementation of thirty functions called \textit{Subroutines} at the end of the file.
+The \textit{main.c} consists on nearly $600$ lines of variables and data structures declarations; followed by the \textit{main()} function, a $2600$ lines long \textit{switch-case} expression, with practically no comments; and to conclude, the implementation of thirty functions called \textit{Subroutines} at the end of the file, around other $2000$ lines of code.
 
-This gives an idea of the problematic to maintain or even understand the code.
-But once one studies MULTOS framework in deep and applies many refactoring techniques to ABC4T Card's code, this becomes the best starting point for the IoT version.
+\hfil
+
+At this stage, we have two options to implement our IoT device compatible with P2ABCE:
+
+\begin{itemize}
+	\item Implement in C the \textit{Smartcard} interface used by P2ABCE architecture, and use some communication protocol to remotelly call the methods from the machine running the P2ABC Engine.
+	\item Present the IoT device as a hardware smart card, using the APDU protocol (already defined, standard and with minimal overload).
+Providing a \textit{javax.smartcardio} ``IoT implementation'' to communicate with the IoT device through a transmission protocol, making the already existing \textit{HardwareSmartcard} class compatible with the new \textit{IoTSmartcard} running in the IoT device.
+\end{itemize}
+
+Even with the problematic to maintain or even understand the code of ABC4Trust Card Lite, once one studies MULTOS framework in deep and applies many refactoring techniques to the code, it becomes the best starting point for the IoT version, making us opt for the second option.
 
@@
 
-\subsection{Smart Card Communication Protocol}
+In this section we will introduce a brief description of the APDU standard protocol, and the MULTOS framework used by the ABC4Trust Card Lite.
+
+\subsection{Smart Card Communication Protocol}\label{subsec:APDU}
 
@@
 
-The development is divided between the \ac{IoT} device code and the \ac{P2ABCE} services. To ease the setup of new developing machines, we will use Docker to create compiling containers.
+For almost every IoT device in the market, there exists a C compiler and many frameworks available to build firmware binaries, like Arduino Core, Contiki, Mongoose OS, ThreadX OS, OpenWrt, LEDE, proprietary SDKs, etc.
+Each firmware targets a specific range of devices, depending on processing power and memory limitations.
+For example, Arduino and Contiki aim for very constrained microcontrollers, like Atmel's ATmega or TI's MSP430, but can also be used in ESP8266, a more powerful device, with WiFi capabilities.
+\marginpar{\includegraphics[width=0.8\linewidth]{gfx/ATmega328P}\\$ATmega328P$}
+
+Starting a big project development for \ac{IoT}, aiming the most constrained devices may not be a good idea.
+The lack of usual OS tools, like POSIX, threads, or minimum I/O, like a terminal, can make debugging a tedious task.
+With good programming practices, one can start from the top and slowly end with more constrained devices and reliable code.
+
+For this reason, the current \ac{PoC} is developed on LEDE, Linux Embedded Development Environment \citep{ledeproject}, using the Onion Omega2 development board.
+Although the Omega2 uses LEDE, its microcontroller is also listed as compatible with ThreadX OS \citep{THREADX}, a Real-Time Operative System for embedded devices, therefore, the performance measured in our PoC can be relevant to real scenarios with similar hardware.
+
+The PoC will take advantage of the Linux system using mainly files and sockets like in any other Linux desktop distribution, so we can focus on the project itself rather than the specific platform APIs for storage and connectivity.
 
@@
 
-The P2ABCE is already written in Java, and few modifications will be done to the code in comparison to the existing project size, so we will continue using Java with the P2ABCE part.
+The project development is divided between the IoT device code and the P2ABCE services. To ease the setup of new developing machines, we will use Docker to deploy containers ready to compile the code.
 
-P2ABCE project needs some minor changes to work with the IoT architecture.
-It is written in Java using Maven to manage the project dependencies.
+\hfil
+
+P2ABCE is already written in Java and uses Maven to manage dependencies.
+The project needs some minor changes to work with our IoT architecture.
 Any text editor or Java IDE is suitable for the development, because the compilation is done through the terminal, with Maven commands.
 
-We compile the project inside a Docker container, with OpenJDK 7, Maven and the Idemix Maven plugins installed, following the project \href{https://github.com/p2abcengine/p2abcengine/wiki/How-to-Build-the-ABC-Engine}{instructions} to use Idemix as the Engine for P2ABCE.
+We compile the project inside a Docker container, with OpenJDK 7, Maven 3 and Idemix 3.0.36, following the project \href{https://github.com/p2abcengine/p2abcengine/wiki/How-to-Build-the-ABC-Engine}{instructions} to use Idemix as the Engine for P2ABCE.
 
@@
 The worst case scenario is that one must write assembly code, and that code will be specific of that target, so we won't consider them.
-If now we focus on the most constrained devices, we could find out that some can't compile C++, some may not have many common libraries, and that the memory limitations they face make practically impossible to use dynamic memory, if we want to avoid many execution malfunctions.
+If now we focus on the most constrained devices, we could find out that for some we can't use C++, some may not have many usual libraries, moreover, the memory limitations they face make practically impossible to use dynamic memory, if we want to avoid many execution malfunctions.
 
@@
 
-Although the ideal final code is pure C, without external libraries or dynamic memory, the \ac{PoC} uses three major libraries:
+Although the ideal final code is written in pure C, without external libraries or dynamic memory, the PoC uses three major libraries:
 
@@
 
-The Dockerfiles can be found in the Appendix.
-%TODO
+The Dockerfiles can be found in \autoref{ch:docker}.

### 2017-05-22 Doc/classicthesis/Chapters/StateOfTheArt.tex
fixes + organización
objectives + org.

@@
 
-\section{Zero-Knowledge Proof systems}
+\section{Systems based on Zero-Knowledge Proofs}
+% ZKP vs minimal disclosure como OAuth
+% Persiano
 % U -Prove
-% Persiano
 % Idemix
+% P2ABCE 
+% TODO: explicar P2ABCE
+%\ac{P2ABCE}
+
+%%%
+Other solutions are based on minimal disclosure.
+Standards like OAuth offer secure delegated access to the user information and when registering to a new service, the user can give a key to access only the data they want from another trusted service.
+This lets the service provider to work with the OAuth server, offering the same service as before without knowing as much data.
+
+But this is only minimizes how many services have our data. Our OAuth provider could be attacked, revealing all our data, or our service provider, revealing now less data.
+%%%
 
@@
 
-The development for Internet of Things depends heavily on each target device.
-We could categorize them in two classes, depending on their capabilities: those with enough processing power to act like an usual computer, and those constrained devices that can't perform arbitrary tasks, sometimes called \textit{embedded}.
+% Adaptar seguridad "clásica" de cifrado de mensajes con claves simétrica y asimétrica, RSA, curva elíptica, DH, 
+% Que el principio de las ZKP ya se usaba en las tarjetas, de manera básica, pero no como parte de un sistema mayor como Idemix
+% Vanets y otros intentos de meter sistemas previos en IoT
 
-We can consider in the first category powerful ARM devices like Raspberry Pi 3, with a 64-bit architecture, 4 CPU cores, 1GB of RAM, which can even compile its own binaries, run the \textit{Java Virtual Machine}, etc., working in practice like any other computer.
-These kind of devices do not present any major difficulty in terms of research.
-
-What we will consider to be a more \textit{pure} \ac{IoT} device will be the constrained ones, where it's not trivial to develop any algorithm and run it successfully.
-
-Very known devices fall into this category, like Arduino, powered by Atmel's AVR ATmega328 8-bit microprocessors, with 32KB of program flash memory, 2KB of SRAM, 16MHz of CPU \citep{ATmega328}.
-It seems clear that memory and computation power are a very big issue to deal with when developing to this devices.
-
-A step above in power we can find ESP8266, the most famous Espressif's microcontroller, with built-in WiFi antenna, a Tensilica 16bit RISC microcontroller at 80MHz, 50Kb of RAM, and 1MB flash memory \citep{ESP8266}.
-The possibility of direct WiFi connectivity is its best selling point, putting the \textit{Internet} in \acl{IoT}.
-
-In another level of power we have microcontrollers usually found in routers, but used in many other applications, like the On-Board Units (OBU) used in Vehicular ad hoc networks (VANET).
-Characteristics in this range vary around a single core 32-bit CPU, at some hundreds MHz, with tens or hundreds MB of RAM and flash memory, which places them near the first \ac{IoT} mentioned category.
-
-Although one can code in assembly language for these microcontrollers, there exist C compilers, and many frameworks to build firmware binaries: Arduino Core, Contiki, propietary SDKs, Mongoose OS, ThreadX OS (Real-Time OS), OpenWrt, LEDE, etc.
-Each firmware targets specific ranges of devices, depending on processing power and memory limitations.
-For example, Arduino and Contiki aim for microcontrollers like Atmel's ATmega and TI's MSP430, but can also be used in ESP8266, a more powerful microcontroller.
-
-In particular OpenWrt and LEDE (a fork of OpenWrt) are based on Linux, with optimized library binaries, providing many packages through \textit{opkg} \citep{opkg}.
-To compile C/C++ code, build the firmware or packages, a complete build system and cross-compiler toolchain can be installed in a x86 host, and using Makefiles select the target hardware \citep{openwrtbuildsystem}.
-
-Devices running OpenWRT and other Linux distros are in the limit between \ac{IoT} categories, but the need of a cross-compiler marks that they belong to the second category.
-
-%%%
-Starting a big project development for \ac{IoT} aiming the most constrained devices may not be a good idea.
-The lack of usual operative system tools (like POSIX), I/O, or even threads can make debugging a tedious task.
-With good programming practices one can start from the top and slowly end at the bottom with reliable code.
-
-For this reason the current \ac{PoC} is developed on LEDE, Linux Embedded Development Environment \citep{ledeproject}, using the Onion Omega2 development board, a Mediatek MT688 microcontroller \citep{MT7688} with a 32-bit MIPS 24KEc CPU at 580MHz, 64MB of RAM and 16MB of flash and built-in WiFi.
-This development board uses LEDE as its firmware, but its CPU is also listed as compatible with ThreadX OS \citep{THREADX}, a Real-Time Operative System for embedded devices.
-
-The \ac{PoC} will take advantage of the Linux system using files and sockets like in any other Linux desktop distribution, so we can focus on the project itself rather than the specific platform APIs for storage and connectivity.
-
-%%%
+Regarding the Internet of Things security,
 
@@
 %%%%
-The problem of Internet privacy has been approached by securing the transmission channel (e.g.
+The problem of Internet security has been approached by securing the transmission channel (e.g.
 SSL/TLS) and the data stored in both ends (strict access policies, local encryption, etc.).
@@
 %%%%
-
-Other solutions are based on minimal disclosure.
-Standards like OAuth offer secure delegated access to the user information and when registering to a new service, the user can give a key to access only the data they want from another trusted service.
-This lets the service provider to work with the OAuth server, offering the same service as before without knowing as much data.
-
-But this is only minimizes how many services have our data. Our OAuth provider could be attacked, revealing all our data, or our service provider, revealing now less data.
-
-%%%%
-\hfill
-
-IBM proposes a step forward using \aclp{ZKP}:
-
-%%%%
-
-\begin{center}
-	\textit{If your personal data is never collected, it cannot be stolen}.
-\end{center}
-
-\paragraph{TODO: Poner descripción + formal de Idemix, luego P2ABCE, y meter que en P2ABCE había una implementación de sc basada en C}
-
-%TODO: vanets, IoT, explicar Idemix, P2ABCE: ZKP systems: UProve, el extra de las vanets e Idemix

### 2017-05-23 Doc/classicthesis/Chapters/Chapter0A.tex
Analysis

@@
 
-\chapter{Dockerfiles}\label{ch:docker}
+\chapter{Docker and CMake files}\label{ch:docker}
+
+\begin{lstlisting}[language=bash,frame=tb,caption={Dockerfile for P2ABCE},label=lst:cmakeCrossFiles]
+SET(CMAKE_SYSTEM_NAME Linux)
+SET(CMAKE_SYSTEM_VERSION 1)
+
+# specify the cross compiler
+SET(CMAKE_C_COMPILER   /lede/staging_dir/toolchain-mipsel_24kc_gcc-5.4.0_musl/bin/mipsel-openwrt-linux-gcc)
+SET(CMAKE_CXX_COMPILER /lede/staging_dir/toolchain-mipsel_24kc_gcc-5.4.0_musl/bin/mipsel-openwrt-linux-g++)
+
+# where is the target environment 
+SET(CMAKE_FIND_ROOT_PATH  /lede/staging_dir/toolchain-mipsel_24kc_gcc-5.4.0_musl)
+
+# search for programs in the build host directories
+SET(CMAKE_FIND_ROOT_PATH_MODE_PROGRAM NEVER)
+# for libraries and headers in the target directories
+SET(CMAKE_FIND_ROOT_PATH_MODE_LIBRARY ONLY)
+SET(CMAKE_FIND_ROOT_PATH_MODE_INCLUDE ONLY)
+
+# Use the commands:
+# mkdir build
+# cd build
+# cmake -DCMAKE_TOOLCHAIN_FILE=Toolchain-omega2-mipsel.cmake ..
+# make
+\end{lstlisting}
+
+\hfil
 

### 2017-05-23 Doc/classicthesis/Chapters/Objectives.tex
Analysis

@@
 %TODO resumir organización del capítulo
+% Incluir bibliografía: vanet, wiki
 
@@
 
+A Privacy Preserving Attribute Based Credential system is composed by several actors, each one of them with different roles.
+One entity could act with more than one role, e.g., an Issuer could also act as a Verifier, but one can assume each actor acts with one role at a time.
+
+The roles are:
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/actors}
+	\end{center}
+	\caption{Entities in a P2ABC System}
+	\label{fig:actors}
+\end{figure}
+
+\begin{itemize}
+	\item \textbf{Issuer}\\
+	In the ABC infrastructure, the Issuer is a trusted entity responsible for issuing credentials, vouching for the correctness of the information contained.
+Each Issuer generates a secret key and publishes the Issuer parameters that include the corresponding public verification key.
+
+	\item \textbf{User}\\
+	Entities that collect credentials from various Issuers.
+They can decide between all their credentials which attributes and values to present when making assertions about their identity to service providers.
+
+	\item \textbf{Verifier}\\
+	A service provider that protects access to a resource by imposing restrictions on the credentials that users must own and the information from these credentials that users must present in order to access the service.
+
+	\item \textbf{Revocation Authority} (optional)\\
+	This entity is responsible for revoking issued credential, preventing their further usage to generate a presentation token.
+Each revocation authority generates and publishes its revocation parameters.
+
+	\item \textbf{Inspector} (optional)\\
+	The Inspector's duty is to de-anonymize the user's presentation token under specific circumstances (e.g.
+misuse or liability).
+At setup, each inspector generates a private decryption key and a corresponding public encryption key.
+Usually, the capability of inspection should be bonded to privacy protection laws.
+
+\end{itemize}
+
+To sum up, an attribute-based credential contains attribute-value pairs that are certified by a trusted Issuer.
+A credential can also specify one or more Revocation Authorities who are able to revoke the credential if necessary.
+Using her credentials, a User can form a presentation token that contains a subset of the certified attributes, provided that the corresponding credentials have not been revoked.
+Additionally, some of the attributes can be encoded in the presentation token so that they can only be retrieved by an Inspector.
+Receiving a presentation token from a User, a Verifier checks whether the presentation token is valid w.r.t.
+the relevant Issuers and Inspector's public keys and the latest revocation information.
+If the verification succeeds, the Verifier will be convinced that the attributes contained in the presentation token are vouched by the corresponding Issuers.
+
+\hfil
+
 In the P2ABCE repository \citep{p2abcurl} there is available the project's code, divided in two solutions: a complete P2ABCE implementation in Java and a MULTOS smart card implementation as PoC for the project.
 
-\subsection{P2ABCE system}
-%TODO: actores y sus papeles
+\subsection{P2ABCE Code Structure and REST Services}
 
-\subsection{P2ABCE Structure and Services}
 The Java code is managed by a Maven project, structured using various known design patterns, but not of our interest.
@@
 
+\hfil
+
 \section{Development environment}
@@
 
-The Dockerfiles can be found in \autoref{ch:docker}.
+The Dockerfiles and CMake cross-compilation toolchain can be found in \autoref{ch:docker}.

### 2017-05-25 Doc/classicthesis/Chapters/StateOfTheArt.tex
Analysis
state of the art: recopilation

@@
 
-\section{Systems based on Zero-Knowledge Proofs}
+\section{Anonymous Credentials Systems}
 % ZKP vs minimal disclosure como OAuth
@@
 %\ac{P2ABCE}
+% Credential: a certified container of attributes where an attribute has a type and a value
 
@@
 %%%
+
+%%%%
+
+A Quick Introduction to
+Anonymous Credentials
+Gregory Neven
+IBM Zürich Research Laboratory
+August 2008
+
+Intuitively, an anonymous credential can be thought of as a digital signature by the Issuer on a list of
+attribute-value pairs, e.g. the list
+(fname=”Alice”, lname=”Anderson”, bdate=”1977/05/10”, nation=”DE”).
+The most straightforward way for the User to convince a Verifier of her list of attributes would be to
+simply transmit her credential to the Verifier. This approach has a number of disadvantages, most
+notably
+· that the User has to reveal all of her attributes so that the Verifier can check the signature;
+· that the Verifier can reuse the credential to impersonate Alice wrt other Verifiers.
+With anonymous credentials, the User never transmits the credential itself, but rather uses it to convince
+the Verifier that her attributes satisfy certain properties – without leaking anything about the credential
+other than the shown properties. This has the obvious advantage that the Verifier can no longer reuse the
+credential to impersonate Alice. Another advantage is that anonymous credentials allow the User to
+reveal a selected subset of her attributes. In the example above, Alice could for example reveal her first
+name and last name, but keep her birth date hidden.
+Stronger even, apart from showing the exact value of an attribute, the User can even convince the
+Verifier that some complex predicate over the attributes holds. In particular, she can demonstrate any
+predicate consisting of the following operations:
+· Integer arithmetic: att+c , att+att , att·c : sum of attributes and/or constants, product of
+attributes with constants
+· Comparative: exp1 = exp2 , exp1 < exp2 , exp1 > exp2 , exp1  exp2 , exp1  exp2 : compare
+arithmetic expressions over attributes/constants
+· Logical: pred1 AND pred2 , pred1 OR pred2
+Note that revealing a subset of attributes comes down to the special of showing that the predicate
+att1 = val1 AND att2 = val2 AND …
+holds. In the example above, Alice could show that she’s an overage national of an EU country by
+showing that he has a credential satisfying
+bdate  today – 18y AND ( nation=”DE” OR nation=”FR” OR … ) .
+The Show protocol does not leak any information other than the truth of the statement. So in the
+example above, not only does Alice’s name remain hidden from the Verifier, but so do her exact birth
+date and nationality; the only information leaked is that her birth date is more than 18 years ago, and that
+her nationality is one of the EU countries.
+%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+
+@book{book:185217,
+	title =     {Financial Cryptography: 8th International Conference, FC 2004, Key West, FL, USA, February 9-12, 2004. Revised Papers},
+	author =    {Jack R. Selby (auth.), Ari Juels (eds.)},
+	publisher = {Springer-Verlag Berlin Heidelberg},
+	isbn =      {3540224203,9783540224204,9783540278092},
+	year =      {2004},
+	series =    {Lecture Notes in Computer Science 3110},
+	edition =   {1},
+	volume =    {},
+	pages = {196-211},
+	url =       {http://gen.lib.rus.ec/book/index.php?md5=24CD58AE88D5564A03C2E44B96732298}}
+
+Persiano and Visconti presented a non-transferable anonymous
+credential system that is multi-show and for which it is possible to prove
+properties (encoded by a linear Boolean formula) of the credentials. Unfortunately,
+their proof system is not efficient since the step in which a user proves
+possession of credentials (that needs a number of modular exponentiations that
+is linear in the number of credentials) must be repeated times (where is the
+security parameter) in order to obtain a satisfying soundness
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%
+@book{book:947508,
+	title =     {Security and Privacy in Communication Networks: 7th International ICST Conference, SecureComm 2011, London, UK, September 7-9, 2011, Revised Selected Papers},
+	author =    {Zhiyun Qian, Z. Morley Mao, Ammar Rayes, David Jaffe (auth.), Muttukrishnan Rajarajan, Fred Piper, Haining Wang, George Kesidis (eds.)},
+	publisher = {Springer-Verlag Berlin Heidelberg},
+	isbn =      {978-3-642-31908-2,978-3-642-31909-9},
+	year =      {2012},
+	series =    {Lecture Notes of the Institute for Computer Sciences, Social Informatics and Telecommunications Engineering 96},
+	edition =   {1},
+	volume =    {},
+	pages = {243-260},
+	url =       {http://gen.lib.rus.ec/book/index.php?md5=A7667DB3EE5B36AAA6E93AE07E888D8D}}
+
+U-Prove. Stefan Brands provided the first integral description of the U-Prove
+technology in his thesis [6] in 2000, after which he founded the company Credentica
+in 2002 to implement and sell this technology. Microsoft acquired Credentica
+in 2008 and published the U-Prove protocol specification [5] in 2010
+under the Open Specification Promise4 together with open source reference software
+development kits (SDKs) in C# and Java.
+The U-Prove technology is centred around a so-called U-Prove token. This
+token serves as a pseudonym for the prover. It contains a number of attributes
+which can be selectively disclosed to a verifier. Hence the prover decides which
+attributes to show and which to withhold (e.g. one can reveal the birth date,
+but not the residence address). Besides the attributes the token contains two
+information fields, one defined by the issuer, and one by the prover. These fields
+are always disclosed and can be used to provide some meta data such as a validity
+date of the token. Finally there is the token’s public-key, which aggregates all
+information in the token, and a signature from the issuer over this public-key to
+ensure the authenticity
+
+5. Brands, S., Paquin, C.: U-Prove cryptographic specification v1.0. Tech. rep., Microsoft
+Corporation (March 2010)
+6. Brands, S.A.: Rethinking Public Key Infrastructures and Digital Certificates:
+Building in Privacy. MIT Press (August 2000)
+%%%%%%%%%%%
+
+%% Idemix:
+
+En \citep{Camenisch:GroupSig} \citep{Camenisch:AnonCred} \citep{camenisch2002signature} describen los métodos criptográficos iniciales utilizando ZKPs que evolucionarán a Idemix.
+En \citep{idemixSpec} se encuentra la especificación actual del protocolo Identity Mixer de IBM.
 
@@
 %%%%
+
+%%%%%%%%%%%%%%
+
+Estudio teórico de cómo integrar Idemix en una smart card, según el reparto de cálculos y valores secretos \citep{luuk}.
+Implementación de Idemix en smart cards MULTOS \citep{vullers2013efficient}
+\citep{sucasas} Diseño de sistema anonymous credential system with sc support, y haciendo pruebas en simulador de smart card, sin embargo no utilizan Idemix ni dan toda su funcionalidad, pero sí aportan un estudio de cómo afrontar la implementación: smart cards.
+Integración en IoT device con P2ABCE en \citep{vanet}
+
+Why ABC4Trust Card Lite: soporte de las APDUs del proyecto P2ABCE, más grande, multi-engine, como dice su powerpoint, adaptable a cualquiera que use Discrete Log.

### 2017-05-31 Doc/classicthesis/Chapters/Introduction.tex
fixes + organización
estado del arte + limpieza

@@
 
-The \ac{IoT} is a term with a wide range of interpretations \citep{Atzori20102787}, but a breaf definition could be the set of devices, mainly resource constrained, that are interconnected between them in order to achieve a goal.
-This includes examples like lampposts with proximity sensors that talk to each other in order to light up part of the street when a passerby walks by, greenhouses with automated irrigation, or cars exchanging traffic data.
+The \ac{IoT} is a term with a wide range of interpretations \citep{Atzori20102787}, breafly, we can think of it as a set of devices, mainly resource constrained, that are interconnected between them in order to achieve a goal.
+For example, a network of lampposts with proximity sensors that talk to each other so they light up part of the street when a passerby walks by, but save energy when not; greenhouses with automated irrigation to balance costs and quality vegetables; or cars exchanging traffic data to reduce city pollution from traffic jams or avoid accidents.
 
-Thanks to organizations like \href{https://wikileaks.org/}{WikiLeaks}, people are aware of the implications of their data on the Internet, demanding more security and privacy for it.
+Many of this objectives require the use of a great amount of data, and thanks to organizations like \href{https://wikileaks.org/}{WikiLeaks}, people are aware of the implications of their data on the Internet, demanding more security and privacy for it.
 This includes not only the data shared with others, where one must trust they will keep it safe, but it's also the data collected about us and that we don't have direct control over it.
@@
 To address this problem of privacy in the Internet, a recent approach is the concept of \textit{strong anonymity}, that conceals our personal details while letting us continue to operate online as a clearly defined individuals \citep{stronganonymity}.
-One very promising way to achieve this is using \acp{ZKP}
+To achieve it, we must address a way to perform authentication and authorization in the most privacy-friendly approach.
+Attribute-based credentials and \textit{selective disclosure} allow to control what information we reveal, under a trusted environment.
+
+Intuitively, an attribute-based credential can be thought of as a digital signature by the Issuer on a list of attribute-value pairs, e.g.
+the list \texttt{\justify(fname=Alice, lname=Anderson, bdate=1977-05-10, nation=DE)} \citep{introCredIBM}.
+The most straightforward way for the User to convince a Verifier of her list of attributes would be to simply transmit her credential to the Verifier.
+With anonymous credentials, the User never transmits the credential itself, but rather uses it to convince the Verifier that her attributes satisfy certain properties – without leaking anything about the credential other than the shown properties.
+This has the obvious advantage that the Verifier can no longer reuse the credential to impersonate \texttt{Alice}.
+Another advantage is that anonymous credentials allow the User to reveal a selected subset of her attributes.
+Stronger even, apart from showing the exact value of an attribute, the User can even convince the
+Verifier that some complex predicate over the attributes holds, e.g. that her birth date was more than 18 years ago, without revealing the real date.
+
+With usual symmetric and asymmetric cryptography it seems impossible to create such credentials, without an explosion of signatures over every possible combination.
+Thats why current solutions rely on  \acp{ZKP}
 \marginpar{To read more about ZKPs, refer to my Mathematics thesis \citep{tfgmates}.}
-, cryptographic methods that allows to proof knowledge of data without disclosing it.
+, cryptographic methods that allow to proof knowledge of some data without disclosing it.
 
-To understand how ZKPs work, in 1990 Guillou, Quisquater and Berson published in \textit{How to Explain Zero-Knowledge Protocols to Your Children} \citep{ZKPcave:story} a story about how Ali Baba proved he knew the magic words to open the cave, but without revealing the words to anyone.
-Here we present a brief version highlighting the properties of ZKP.
+To understand how ZKPs work, in 1990 Guillou, Quisquater and Berson published in \textit{How to Explain Zero-Knowledge Protocols to Your Children} \citep{ZKPcave:story} a story about how Ali Baba proved that he knew the magic words to open the cave, but without revealing those words to anyone.
+Here we present a brief version highlighting the properties of ZKPs.
 
@@
 
-	\textbf{E}ve, curious about what Victor and Peggy were doing in the cave, \textbf{e}avesdrop Victor during the process.
+	\textbf{E}ve, curious about what Victor and Peggy were doing in the cave, \textbf{e}avesdrops Victor during the process.
 The problem is that Eve doesn't know if Peggy and Victor agreed on what paths to choose, because they wanted to prank her for being busybody.
@@
 
-Based on ZKP properties, IBM has developed the Identity Mixer, Idemix for short, protocol suite for privacy-preserving authentication and transfer of certified attributes \citep{idemixurl}.
+Based on ZKP properties, IBM has developed the Identity Mixer\footnote{Identity Mixer - \url{https://www.research.ibm.com/labs/zurich/idemix/}}, Idemix for short, protocol suite for privacy-preserving authentication and transfer of certified attributes.
 It allows user authentication without divulging any personal data.

### 2017-05-31 Doc/classicthesis/Chapters/StateOfTheArt.tex
estado del arte + limpieza

--- 2017-05-22 Doc/classicthesis/Chapters/StateOfTheArt.tex
@@
 
-In this chapter we present the two domains of this project: the Zero-Knowledge Proof based systems and the Internet of Things.
+In this project we will study Identity Mixer as an Attribute-Based Credentials (ABC) solution for privacy-preserving scenarios, but there exist other solutions competing to give the best performance and capabilities as possible.
+The two most known alternatives to Idemix are Microsoft's U-Prove and Persiano's ABC systems.
 
-\section{Systems based on Zero-Knowledge Proofs}
-% ZKP vs minimal disclosure como OAuth
+%minimal disclosure
 % Persiano
+% Verheul's self-blindable: only binary attributes [14]deIdemixSC
 % U -Prove
@@
 
-%%%
-Other solutions are based on minimal disclosure.
-Standards like OAuth offer secure delegated access to the user information and when registering to a new service, the user can give a key to access only the data they want from another trusted service.
-This lets the service provider to work with the OAuth server, offering the same service as before without knowing as much data.
+%%%%%%%%%%%%%%%%%%%%%%%%
 
-But this is only minimizes how many services have our data. Our OAuth provider could be attacked, revealing all our data, or our service provider, revealing now less data.
-%%%
+Persiano and Visconti presented a non-transferable anonymous
+credential system that is multi-show and for which it is possible to prove
+properties (encoded by a linear Boolean formula) of the credentials \citep{book:185217}. Unfortunately, their proof system is not efficient since the step in which a user proves
+possession of credentials (that needs a number of modular exponentiations that
+is linear in the number of credentials) must be repeated times (where is the
+security parameter) in order to obtain a satisfying soundness.
 
-\section{Internet of Things}
+%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Adaptar seguridad "clásica" de cifrado de mensajes con claves simétrica y asimétrica, RSA, curva elíptica, DH, 
-% Que el principio de las ZKP ya se usaba en las tarjetas, de manera básica, pero no como parte de un sistema mayor como Idemix
-% Vanets y otros intentos de meter sistemas previos en IoT
+%%%%%%%%%
 
-Regarding the Internet of Things security,
+Stefan Brands provided the first integral description of the U-Prove
+technology in his thesis \citep{uprove} in 2000, after which he founded the company Credentica
+in 2002 to implement and sell this technology. Microsoft acquired Credentica
+in 2008 and published the U-Prove protocol specification \citep{uprove2} in 2010
+under the Open Specification Promise4 together with open source reference software
+development kits (SDKs) in C\# and Java.
+The U-Prove technology is centered around a so-called U-Prove token. This
+token serves as a pseudonym for the prover. It contains a number of attributes
+which can be selectively disclosed to a verifier. Hence the prover decides which
+attributes to show and which to withhold. Finally there is the token’s public-key, which aggregates all information in the token, and a signature from the issuer
+over this public-key to ensure the authenticity \citep{book:947508}.
 
-\section{Idemix}
-
-\paragraph{TODO: Move to motivation}
-
-%%%%
-The problem of Internet security has been approached by securing the transmission channel (e.g.
-SSL/TLS) and the data stored in both ends (strict access policies, local encryption, etc.).
-In the end, the data exists in two entities, the owner of the data and the service provider.
-The owner is the most interested in securing his data, and can apply as many measures as he wants, but only on his side of the table.
-The service provider that stores the user data needs it to provide the service, and a successful attack would reveal many users data, aside from how many measures each one used to protect it.
-The case of PlayStation Network outage in 2011 \citep{PSN2011} affected 77 million accounts, with suspected credit card fraud, is an example of this kind of attacks.
+%%%%%%%%%%%
 
@@
 
+Jan Camenisch, Markus Stadler and Anna Lysyanskaya studied in \citep{Camenisch:GroupSig}, \citep{Camenisch:AnonCred} and \citep{camenisch2002signature} the cryptographic bases for signature schemes and anonymous credentials, that later became IBM's Identity Mixer protocol specification \citep{idemixSpec}.
+
+Luuk Danes in 2007 studied theoretically how Idemix's User role could be implemented using  
+smart cards \citep{luuk}, identifying what data and operations should be kept inside the device to perform different levels of security.
+The User role was divided between the smart card, holding secret keys, and the Idemix terminal, that commanded operations inside the smart card, or read the keys in it to perform the instructions itself.
+The studied sets were:
+\begin{itemize}
+	\item The smart card gives all information to the terminal.
+	\item The smart card only keeps the master key secret.
+	\item The smart card only gives the pseudonym with the verifier to the terminal.
+	\item The smart card keeps everything secret.
+\end{itemize}
 %%%%
+
+Later, in 2008 Víctor Sucasas also studied an anonymous credential system with smart card support \citep{sucasas}, equivalent to a basic version of Idemix, using a simulator to test the PoC and pointing out some crucial implementation details for performance.
+The researching tendency starts to show that smart cards are the best solution to hold safely the User's credentials.
+
+In 2009, some Java smart card PoC for Idemix were developed in \citep{javaIdemix1} and \citep{javaIdemix2}, but they weren't optimal and didn't include some Idemix's functionalities, like selective disclosure.
+
+Later, in 2013, Vullers and Alpar, implemented an efficient smart card for Idemix \citep{vullers2013efficient}, aiming to integrate it in the IRMA\footnote{The IRMA project has been recently included in the Privacy by Design Foundation: \url{https://privacybydesign.foundation/}} project, and comparing the performance with U-Prove's smart cards.
+This new implementation was written in C, under the MULTOS platform for smart cards, and describes many decisions made during the development to improve the performance on such constrained devices.
+The terminal application was written in Java and used an extension of the Idemix cryptographic library to take care of the smart card specifics.
+
+Extending the concept of smart cards, physical or logical, as holders of the credentials, the ABC4Trust's project, P2ABCE\footnote{\acl{P2ABCE} - \url{https://github.com/p2abcengine/p2abcengine}}, was created as a unified ABC system for different cryptographic engines, currently supporting U-Prove and Idemix.
+The Idemix library was updated to support P2ABCE and the last version is interoperable with U-Prove.
+Therefore, the smart card specification from the P2ABCE project could be considered the official version to work with.
+
+% Uso de P2ABCE en VANETs
+Related to the IoT, the P2ABCE project has been used to test in a VANET\footnote{Vehicular Ad-Hoc Network} scenario how an OBU (On Board Unit), with constrained hardware, could act as a User in an P2ABC system \citep{vanet}.
+However, after the theoretical analysis, the paper only simulates a computer with similar performance as an OBU, without adapting the existing Java implementation of P2ABCE to a real VANET system.
+In our project, we can consider ourselves as part of their \textit{future work}, because our PoC will execute on hardware actually used in VANET systems, although the OS may not be the same.
+
+%%%%%%%%%%%%%%

### 2017-05-31 Doc/classicthesis/Chapters/Objectives.tex
estado del arte + limpieza
Objectives + +

@@
 %************************************************
-\chapter{Analysis and methodology}\label{ch:objectives}
+\chapter{Analysis and objectives}\label{ch:objectives}
 %************************************************
@@
 The objective is to design a general solution for the existing IoT devices and systems, without compromising any feature of Idemix, and provide a working PoC in a real IoT environment, even though it isn't the most constrained scenario.
-The project is aimed to be used in privacy preserving environments, providing security for IoT, controlling what data is being disclosed and to whom.
-In a smart city project, citizens data can be privatized and continue to offer the benefits of the sensors around.
-Authorized personnel can disclose the information when required, like firefighters that could access a building sensors to know how many people and what conditions they are in, in case of a emergency; and keep such invasive data private to other non-critical services, for example, only giving a proof that there are more than a minimum number of people in the building, to activate or deactivate the air-conditioning system.
+The project is aimed to be used in privacy-preserving environments, providing security for IoT, controlling what data is being disclosed and to whom.
+
+In a smart city project, citizens' data can be privatized and at the same time continue to offer the benefits of the sensors around.
+Authorized personnel can disclose the information when required, like fire-fighters accessing a building's sensors to check how many people there are and what conditions they are in, in case of an emergency; but keep such invasive data private to other non-critical services, for example, only giving a proof that there are people in a floor of the building, to activate or deactivate the air-conditioning system.
 
@@
 \paragraph{Idemix and the IoT}
-We part from this main objective, integrating IBM's privacy preserving system in the IoT environment.
+This is our main objective, integrating IBM's privacy-preserving system in the IoT environment.
 
-\paragraph{Analysis} Study the state of the art for Idemix and the IoT, analyzing related projects like ABC4Trust's \ac{P2ABCE}, papers with similar approaches, and our task is to consider the best fitting solutions from where to start.
+\paragraph{Analysis} Study the state of the art of Idemix and the IoT, analysing related projects, papers with similar approaches, and consider the best fitting solutions from where to begin.
 
@@
 We will break down the original system, Idemix and P2ABCE, analyze every part of it and categorize them.
-We will consider what components are mandatory to be executed in the target device and which ones the devices would be able to execute.
+We will consider what components are mandatory to be executed in the target device that wants to act as User of Verifier, and which components the devices would actually be able to execute.
 
-Devices with equivalent processing power to smart phones are capable to execute the current Java implementations of P2ABCE and Idemix, but the most constrained ones can't handle the entire system, only the key cryptographic operations.
+Devices with equivalent processing power to smart phones are capable of running the current Java implementations of P2ABCE with Idemix, but the most constrained IoT devices can not handle the entire system, only the mandatory cryptographic operations.
 
-Using the technique known as \textit{Computation Offloading} we will design an IoT architecture where the most constrained targets can keep their private keys and certificates secure in the same device, and act as any other User in the P2ABCE system.
-Studying the P2ABCE architecture and implementations we will identify those mandatory operations to be executed in the constrained device, and how to communicate efficiently during the delegation.
+Using the technique known as \textit{Computation Offloading} we can design an IoT architecture where the most constrained targets can keep their private keys and certificates secure within the device, and act as any other actor in the P2ABCE system.
+Studying the original P2ABCE architecture and implementations we will identify the key operations to be executed in the constrained device, and how to communicate efficiently during the delegation.
 
-After the technical design, we will implement the PoC, using software designs patterns known to many professionals, that will improve the maintainability of the project, for improvements and fixes.
-Taking advantage of this practices, we can document the immediate steps for future developers, how to reimplement certain interfaces when porting the application to a new system, or where the core logic lies for future protocol changes.
+After the technical design, we will implement the PoC, using known software designs patterns, that will improve the maintainability of the project.
+Taking advantage of this practices, we can document the immediate steps for future developers, how to reimplement certain interfaces when porting the application to a new system, or where the core logic lies, to implement future protocol changes.
 
-Finally, the PoC must be evaluated to assert we achieved our goals, and measure its performance, judging if it could be fitting for real IoT environments.
+Finally, the PoC will be evaluated to assert that we achieved our goals, and measure its performance, judging if it can indeed be suitable for a IoT deployment.
 
@@
 
-A Privacy Preserving Attribute Based Credential system is composed by several actors, each one of them with different roles.
-One entity could act with more than one role, e.g., an Issuer could also act as a Verifier, but one can assume each actor acts with one role at a time.
+There are several cryptographic systems for dealing with attribute-based
+identities. Typically these systems distinguish credentials and attributes. Informally,
+a credential is a cryptographic container of attributes, where an attribute has a type and a value \citep{vullers2013efficient}.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=0.4\linewidth]{gfx/ABC}
+	\end{center}
+	\caption{A first look at an attribute-based credential with four attributes.}
+	\label{fig:abc}
+\end{figure}
+
+The Privacy Preserving Attribute Based Credentials system is composed by several actors, each one of them with different roles.
+One entity could act with more than one role, e.g., in a M2M (Machine To Machine) scenario, a device could act as both User and Verifier to other peers; but one can assume each actor acts with one role at a time.
 
@@
 \end{itemize}
+
+The cryptographic nature of the credential-as-container concept includes the
+following four security aspects \citep{vullers2013efficient}:
+\begin{itemize}
+	\item The issuer’s digital signature ensures \textbf{authenticity}: the credential originates
+	from the issuer, and this issuer asserts that the attributes hold for the person.
+	\item This signature also guarantees integrity: the attributes contained in the credential
+	have not been altered since they were issued.
+	\item A credential is \textbf{non-transferable} as it is bound to the secret key, only known by the credential owner.
+	\item A credential \textbf{hides} its content, so it does not reveal the attributes it contains.
+	Furthermore, a credential protects the privacy of its owner through the following
+	cryptographic properties.
+	\item Issuer \textbf{unlinkability} ensures that any information gathered during issuing
+	cannot be used to link a verification of the credential to its issuance.
+	\item \textbf{Multi-show unlinkability} guarantees that when a credential is verified multiple
+	times, these sessions cannot be linked.
+\end{itemize}
+
+Zero-knowledge proofs, used in Idemix, allow a user to prove ownership of a credential without
+revealing the credential itself. Since the verifier does not see the credential, verification
+instances are unlinkable and they also cannot be related to the issuing procedure.
+The privacy of users is protected by these unlinkability properties, even if the credential issuer and all verifiers collude.
 
@@
 
-The code uses some good practices when programming for IoT devices, like using \textit{union} data types for variables that will be stored on the same data location, but at different times, minimizing this way the use of RAM and making code readability better; or like the strong use of pointers and \textit{memcpy} calls to copy structures with multiple variables as arrays of bytes.
+The code uses some good practices when programming for constrained devices, also explained in the first implementation of Idemix in smart cards \citep{vullers2013efficient}.
+Aside the implementation in assembly code of some operations to accelerate the execution, most techniques aim to reduce the memory usage.
+They define \texttt{union} data types to join variables never used at the same time, so they can be stored on the same data location.
+Instead of function parameters, most variables are made global, reducing the space on the stack used when calling other functions.
+If parameters are needed, they are pointers to shared buffers with fixed sizes, avoiding the use of any dynamic memory allocation.
 
@@
 
-To communicate the smart cards and the reader the ISO/IEC 7816-4 \citep{APDUISO} specifies a standardized protocol .
+To communicate the smart cards and the reader the \texttt{ISO/IEC 7816-4} \citep{APDUISO} specifies a standardized protocol .
 
@@
 
-\paragraph{MULTOS Workflow}
+\paragraph{Execution Model}
+Applications on a MULTOS card are executed in a virtual machine, called the Application Abstract Machine (AAM). The AAM is a stack machine that interprets instructions from the MULTOS Executable
+Language (MEL).
 
-Most of the transmission and communication process is done by MULTOS core, and it then selects, based on the CLA byte of the APDU, the application to load.
-This application is what most developers will only worry about, and is where their $main()$ function will start.
+The transmission and communication process is done by MULTOS core, and it then selects, based on the CLA byte of the APDU, the application to load.
+This application is what most developers will only worry about, and is where their compiled \texttt{main()} function will start.
 
@@
 
-The application uses then the \textit{multos.h} file that declares multiple global variables already loaded with the needed data, including the APDU Command bytes.
+Now the developer is in charge of checking what instruction was sent, handle it with regard to his domain logic,  write in the specific data space the APDU Response bytes, and call \texttt{multosExit()}, a MULTOS API function that will be in charge to send the APDU Response.
+In summary, our application starts with all data loaded in memory and exits without worrying how the answer is sent back.
 
-Now the developer is in charge of checking what instruction was sent and if the APDU has the expected ISO Case.
-If everything is ok, code what needs to be run and write in specific data space the APDU Response bytes, call \textit{multosExit()} and MULTOS will be in charge to send the APDU Response.
-
-In summary, our application starts with all data loaded and exits without worrying how to send the answer.
-A very comfortable workflow that we must now implement for our IoT device if we would want ABC4T Card Lite code to work.
+As we can see, MULTOS is a comfortable framework to develop smart card applications, and now we must adapt and implement it for our IoT devices, if we want to port ABC4Trust Card Lite's code.
 
@@
 \textbf{Static memory} are the application variables declared after the specific \textit{\#pragma melstatic} compiler directive.
-These variables are stored in the non-volatile EEPROM, and any write is assured to be saved because they are not loaded into RAM.
+These variables are stored in the non-volatile secure EEPROM, and any write is assured to be saved because they are not loaded into RAM.
 

### 2017-06-01 Doc/classicthesis/Chapters/Design.tex
design init

@@
 %TODO resumir organización del capítulo
+
+% Estructura:
+% +Design
+%  -Architecture design: delegation, computation offloading, duality User/SC + Engine
+% 
+% +Implementation:
+%  -REST as current delegation protocol.
+%  -BIOSC as current APDU transmission protocol: no security -> no overload.
+%  -IoT Smart Card software design and implementation notes.
+%  -Sequence diagrams. TODO
+
+In this chapter we will describe the process for defining how a constrained IoT device may be integrated in a system like P2ABCE.
+We will also describe the PoC implementation carried out to test a realistic deployment.
+
+\section{Design}
+
+In this section we will define how an IoT device may be integrated in the P2ABCE architecture, being totally compatible with any other system using P2ABCE, addressing the power and memory constrains many IoT devices face.
+
+We decided to use P2ABCE with the Idemix as its Engine, because it is officially supported by the Idemix Library, it has the most up-to-date implementation, as we saw in the state of the art, and adds capabilities to Idemix like the Presentation Policies, or interoperability with U-Prove, not available without the P2ABCE project.
+
+Our main goal is to make an IoT device capable to act as a User or Verifier in the P2ABCE architecture.
+For this, the device should be able to \textbf{communicate} with the Verifier or Prover with which it is interacting, manage the P2ABCE complex \textbf{XML schemas} transmitted, and perform the \textbf{cryptographic operations} required.
+
+The communication between actors depends on each IoT scenario, it can be achieved with many existing standard solutions, e.g. an IP network, a Bluetooth M2M connection, RF communication, etc.
+
+Our real concerns are, on one side, parsing the XML data, based on the P2ABCE's XML schema, that specifies the data artifacts created and exchanged during the issuance, presentation, revocation and inspection of pABCs; and on the other side, the cryptographic operations, that involve the use of secret keys, stored privately in the IoT device.
+
+After the analysis done in the previous sections to the P2ABCE architecture, emphasizing that the logic of smart cards gathers the cryptographic operations independently from how the data is exchanged between P2ABCE actors.
+
+Using the \textit{computation offloading} technique to our scenario, our design consists on implementing the smart card logic inside the IoT device, keeping secure our master key and credentials, and for the rest of the P2ABCE system, if the device can not run the complete Engine, it may delegate to a server running it, indicating how to send APDU Commands to the \textit{IoT smart card}.
+
+Even in the case we were to implement all P2ABCE inside an IoT device, we would have to implement the support for software smart cards, to keep the secret inside the IoT device.
+Therefore, we can begin implementing the smart card logic inside the IoT device, and later, if the device resources admit it, other components of the P2ABCE project.
+
+\hfil
+
+Computation offloading is not new to IoT deployments.
+For example, IPv6 involves managing 128 bits per address and other headers, and many IoT scenes only need to communicate inside a private network, making only the last 64 bits in an address relevant.
+To reduce that overhead, instead of IPv6 they use 6LoWPAN to compress packets and use smaller address sizes.
+To communicate a 6LoWPAN with the Internet or other networks, the IoT devices delegate the networking workload on a proxy that can manage the 6LoWPAN and IPv6 stacks.
+In the scope of consumer devices, smart bands or watches can install applications, bit many of them delegate on the user's phone to accomplish their task.
+
+\subsection{System architecture}
+
+% TODO : system architecture:
+% Definir cada etapa: inicio conversación con otro actor, delegación indicando cómo accede a la iot sc, transmisión de APDUs, de los dos últimos la seguridad y escenarios posibles en un último párrafo.
+%describir fig:P2ABCE-IoT
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/P2ABCE-IoT-bw}
+	\end{center}
+	\caption{IoT in P2ABCE Architecture.}
+	\label{fig:P2ABCE-IoT}
+\end{figure}
+
+%\begin{figure}[bth]
+%	\begin{center}
+%		\includegraphics[width=\linewidth]{gfx/P2ABCE-IoT-color}
+%	\end{center}
+%	\caption{IoT in P2ABCE deployment diagram.}
+%	\label{fig:P2ABCE-IoT-color}
+%\end{figure}
+
+\hfil
+
+Therefore, the IoT device now has a \textbf{duality} in its functions, because it is the User that starts any interaction with other actors, and it's also the smart card that a P2ABCE server must ask for cryptographic operations.
+It can also be seen as a \textbf{double delegation}.
+The IoT device delegates on the external P2ABCE server to manage the protocol, and the P2ABCE server delegates on the IoT, acting now as a smart card, for the cryptography.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/DelegationProving}
+	\end{center}
+	\caption{IoT Delegation in P2ABCE for Proving.}
+	\label{fig:DelegationProving}
+\end{figure}
+
+\paragraph{PoC Delegation to the P2ABCE Server} \hfil
+
+We now will address each type of delegation: from the IoT device as a P2ABCE actor to the P2ABCE delegation server, and from the server to the smart card in the IoT device.
+
+\hfil
+
+Currently P2ABCE offers multiple REST web services to run different roles in P2ABCE system: User Service, Issuer Service, Verification Service, etc.
+An application that integrates P2ABCE can make use of this services in the same machine or implement the functionality using the core components written in Java, the same ones the REST services use.
+Our PoC machine, the Omega2, can make REST calls easily, but other devices may use \ac{CoAP}, and in that case, the P2ABCE REST services should be rewritten to offer CoAP support.
+The commands needed to delegate to the P2ABCE delegation server will be the same to operate with the REST services.
+This way, the first issue is solved.
+
+The transmission of the messages will depend on the specific use cases, capabilities and resources available.
+If the delegation server is connected, for example, through RS-232 serial with the IoT device, and physically inaccessible, in the same way an IoT device on its own would be protected, the communication is simple, and not far away from the Arduino Yun idea of combining two devices, one more powerful but to use only when needed.
+But if the IoT device and the delegation server are apart, or more than one IoT device delegates to it, then the transmission must be secured.
+They could use 6LoWPAN to talk to each other (the delegation service could be deployed in the proxy) and then secure communications with existing solutions, like with pre-shared symmetric keys, certificates for authentication and authorization, etc., it depends on each particular deployment.
+
+At the end of the day, this is all about usual security in IoT. Many other studies focus on this matter, so we will assume it can be done, and focus on what's new, P2ABCE in IoT.
+
+\hfil
+
+% TODO: imagen con servicios desplegados/actores en cada máquina
+
+To sum up, our IoT device will act as User (Prover or Verifier) keeping its secrets in a software smart card.
+When it starts an interaction with other actor of the P2ABCE system (Issuer, Verifier, etc.), the IoT device will delegate with a remote call (using REST in our PoC) to a P2ABCE delegation server, attaching the XML file and the necessary information for the server to send the APDUs to the software smart card (in our PoC using TCP sockets, giving the IP and listening port).
+
+\hfil
+
+This simple design keeps the benefit of a 100\% compatible P2ABCE deployment, and the integration of IoT devices to the P2ABCE ecosystem.
+
+In the future, more functionality currently delegated in the P2ABCE server can be implemented in the IoT device, if its resources allow it.
+For example, in a M2M environment, where an IoT device can act as User and Verifier, the verification consists on sending a Presentation Policy, and verifying the Presentation Token, which implies less logic than generating it as the User.
+Therefore, the implementation of the Verifier functionality would reduce significantly the need of a delegation server, but as we said, managing complex XML squemas is not something many IoT devices could do.
+
+\section{Implementation}
+
+\section{IoT Smart Card}
+
+After many design decisions in the process to adapt the original ABC4T Card Lite code to pure C, working over a more usual architecture machine, in this section, we present the current \ac{PoC} code, most important decisions, workflow execution, and future work.
+
+First, let's define what a \textit{more usual architecture} is.
+If we remember the MULTOS section, the framework gives an application a very specific memory layout and entry and output points of execution, that could be seen as a single process execution machine.
+Many IoT devices work like a computer, with multiple processes or threads, without pre-loaded data on startup (like the APDU MULTOS loads for the application), a non-volatile memory for data and code, maybe a basic file system in this memory, and RAM with the program's stack, heap, data and code.
+
+Our PoC is tested on a Linux system, and we will give instructions on how to adapt each part to work with other typical IoT systems.
+For example, other IoT devices may work like MULTOS and let access variables in non-volatile memory during execution, and in that case, the port should be changed according to these particularities.
+
+\hfil
+
+\subsection{Code structure}
+
+We divide the project in three different sections with the objective of enhancing maintainability, improving future changes, ports, fixes, etc.
+
+The first section is what could be called as the core of the smart card, the second one the interface for the tools the core need and may depend on the platform, and finally third party libraries, that in may be empty if the interfaces implementation doesn't need any.
+
+In our PoC we used CMake to manage the project, due to the cross-compilation tools, integration with multiple IDEs and tests.
+
+%\begin{figure}[bth]
+%	\begin{center}
+%		\includegraphics[width=\linewidth]{gfx/IoTCScomponents-color}
+%	\end{center}
+%	\caption{IoT Smart Card Code Structure.}
+%	\label{fig:IoTCScomponents-color}
+%\end{figure}
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/IoTCScomponents-bw}
+	\end{center}
+	\caption{IoT Smart Card Code Structure.}
+	\label{fig:IoTCScomponents-bw}
+\end{figure}
+
+\hfil
+
+\paragraph{Core smart card}
+
+The smart card logic lies in this section, the concepts of APDU Commands, what instructions are defined in P2ABCE smart cards and how to process them and generate proper APDU Responses.
+
+Changes in the APDU protocol for P2ABCE must be done here, independently of the target platform.
+
+After refactoring the original ABC4Trust Card's code, most of it fell in what we will call the core of the smart card.
+
+All types and variable definitions and the APDU handling is done in this code.
+However, the ABC4Trust's code depended on the MULTOS C-API for the input/output of data, modular arithmetic, and even AES128 and SHA256 cryptography.
+
+A characteristic of MULTOS C-API is that every function name starts with \textit{multos}, but as we said, the \textit{main.h} file implemented equivalent functions to some available in \textit{multos.h}.
+Our first step was to replace the \textit{main.h} functions for the standard ones in the C-API.
+Then, we implemented, following the C-API documentation, the functions from \textit{multos.h} (only the used ones) changing their names from \textit{multosFoo()} to \textit{mFoo()} for readability and emphasize that they were no longer from MULTOS.
+
+Future changes in the code may refactor it so there's no longer need for the MULTOS framework functions.
+
+\paragraph{Interfaces}
+
+To implement MULTOS functions, we needed to use some libraries, so we defined a facade to isolate the implementation of the core smart card from our different options, that could vary depending on the hardware or the system used by the IoT device.
+
+The use of a facade lets us, for example, change the implementation of modular arithmetic with a hardware optimized version, or a future more lightweight library, or our very own software implementation using the same data types that the core uses, minimizing the data transformations needed.
+
+Taking a step forward, we make the core smart card totally independent of any library, only on our interfaces.
+This means that typical C libraries, like the standard \textit{stdlib.h}, or  \textit{string.h} are also behind the facade, in case some IoT system doesn't support them.
+The main goal we go after with this decision is that future developers adapting the code to a specific platform need to make no change to the \textit{core smart card}'s code, only to the interfaces implementation.
+
+\paragraph{External utilities}
+
+If the IoT system offers well tested libraries that could aid in the interfaces implementation, or we simply found a pure C implementation for the task, these third party libraries belong to this section.
+
+In our PoC, we use two ANSI C libraries, for base64 and JSON, and two shared libraries available in as packages in LEDE, GMPLib and OpenSSL.
+The last two libraries offer more functionality than we need, hence, it's desired in a production code to implement \textit{Modular Arithmetic} and \textit{Cryptography} interfaces with more lightweight alternatives.
+
+For example, Atmel's ATAES132A \citep{ATAES132A}
+\marginpar{\includegraphics[width=0.9\linewidth]{gfx/atmel} \\ Atmel's cryptography chips.}
+offers a serial chip for secure key storage, AES128 execution and random number generation.
+Another serial chip like ESP8266 offers WiFi connectivity, typically used with Arduino, and can also perform AES encryption.
+For random number generation, a technique used with Contiki devices is to read from sensors aleatory data and use it as seed.
+All these alternatives depend on the target device, but are all valid.
+The \textit{interfaces} and \textit{external utilities} sections allow for a clean and fast port of the code.
+
+\subsection{PoC Workflow}
+
+% PoC BIOSC : TCP sockets, JSON serialization
+
+% PoC IoT javax.smartcardio : protocolo de 2 bytes + APDU
+
+% 1. Boot: deserialize sc status
+% 2. Listening: open tcp socket
+% 3. The IoT device delegates on the P2ABCE for some task and sends the IP and Port of the IoT Smart Card.
+% 4. An APDU Command arrives : read in 2 bytes the length of the apdu and
+
+In a real deployment, we would offer a full library with an API to other processes to delegate on the P2ABC Engine, that automatizes the listening and security, that we presented in the deployment diagram \textbf{TODO:ref-img} as the \textit{P2ABCE IoT Toolkit}.
+
+But to test the IoT smart card, we use the \textit{curl} command for REST delegation calls and the BIOSC (Basic Input Output Smart Card) for APDU communication.
+
+To transmit the APDU messages in our PoC we use a simple protocol, consisting in one first byte for the instruction: receive an APDU Command or close the connection.
+In the first case, then we read two header bytes with the length of the APDU Command or Response to receive, followed by those APDU bytes.
+The message is sent over TCP for a reliable transmission (concept of session, packet retransmission, reordering, etc.).
+
+We lack of any security (authentication and authorization) that a real system should implement.
+It is vital to authenticate the delegation service, to authorize it to make APDU Commands, and the same with the IoT device, to prevent attacks.
+This belongs to \textit{usual} security, as we already said, and for that reason it's not in the PoC.
+
+\paragraph{P2ABCE REST Delegation}
+
+The only P2ABCE Service that needed to be modified was the User Service. We added the REST call
+
+\begin{center}
+	\textit{/initIoTsmartcard/{issuerParametersUid}?host=\&port=}
+\end{center}
+
+where we communicate the P2ABCE server that a IoT Smart Card is accessible via \textit{host} and \textit{port}.
+Then a new \textit{HardwareSmartcard} object is stored in the P2ABC Engine, but instead of the \textit{javax.smartcardio} Oracle's \textit{CardTerminal} implementation, we use our \textit{IoTsmartcardio} implementation for the \textit{HardwareSmartcard} constructor.
+
+The rest of P2ABCE code is unchanged and will work as if a real smart card was in use. \textit{IoTsmartcardio} implementation will transmit the APDUs through the TCP socket with the format mentioned.
+
+\hfil
+
+Our test consists on the tutorial available in P2ABCE's repository. From the Omega2's terminal we run \textit{curl} commands to send REST petitions. For example:
+
+\begin{verbatim}
+$ curl -X POST --header 'Content-Type: text/xml'
+'http://localhost:9200/user/initIoTsmartcard/http%3A%2F%2
+Fticketcompany%2FMyFavoriteSoccerTeam%2Fissuance%3Aidemix
+?host=192.168.3.1&port=8888'
+\end{verbatim}
+
+The IoT device will only manage XML as data files to exchange between the third party actor and the P2ABCE delegation service, without parsing them.
+
+After a REST call to the User Service, the P2ABC Engine will talk to the IoT smart card process.
+
+\paragraph{APDU transmission with BIOSC}
+
+% TODO : diagrama del software
+
+Before sending the REST message, the Omega2 must be listening in the specified port.
+At start up, BIOSC reads from a JSON file the status of the smart card (credentials in the BLOB, private keys, PIN and PUK codes, etc.), then opens the TCP socket and listens in a loop.
+
+When the delegation server sends an APDU Command following the simple protocol, BIOSC stores in a byte array the APDU Command bytes and calls the \textit{handle\_APDU} function in \textit{Core Smart Card}.
+
+The \textit{APDU\_handler} will parse the APDU bytes and check both CLA and INS bytes.
+For each possible INS the is a function that must always finish calling either \textit{mExit()} or \textit{output\_large\_data()}.
+
+The \textit{mExit} function is the reimplementation of MULTOS C-API exit functionality, that finishes the application execution returning to the MULTOS OS, that will send the APDU Response bytes to the terminal.
+This led in the original ABC4Trust Card's code to some tricky situations.
+Imagine a function that ends with an if-else expression.
+Many times we save a line writing something like
+
+\begin{verbatim}
+if(condition)
+return a;
+return b;
+\end{verbatim}
+
+instead of the complete
+
+\begin{verbatim}
+if(condition)
+return a;
+else
+return b;
+\end{verbatim}
+
+This is a swift example of what the \textit{multosExit} function led to, because it is called in both the \textit{switch-case} processing the APDU instruction, and the \textit{Subroutines}.
+For example, the PIN code check was programmed in a way every reason to fail finished the application execution, with different error codes.
+Another example, checking if a credential is stored, if not, fail and exit inside the \textit{readBlob} subroutine.
+
+In our standard architecture machine, if we call a subroutine, it must end, and return the control to the handling function, and so the \textit{APDU\_handler} must return to the listening loop of BIOSC.
+
+Having all that in count, the \textit{mExit} function is not implemented as the documentation specifies, but only saves the smart card status serializing it in the JSON file again, and once it's saved, parses the APDU Response and sends it to the socket.
+
+The \textit{output\_large\_data} function is a tool used in MULTOS smart cards that don't support Extended APDUs.
+The APDU Command handler saves in a buffer the data to send, and its size.
+Then, \textit{output\_large\_data} will manage the buffer expecting future GET RESPONSE commands from the P2ABCE server.
+
+A future change to the project may include support for Extended APDUs, avoiding the use of multiple GET RESPONSE.
+To do this, the \textit{output\_large\_data} function and the \textit{IoTsmartcardio} implementation of \textit{javax.smartcardio} must support the feature.
+
+% TODO -Sequence diagrams.

### 2017-06-02 Doc/classicthesis/Chapters/Design.tex
design arch

@@
 
+Therefore, the IoT device now has a \textbf{duality} in its functions, because it is the User that starts any interaction with other actors, and it's also the smart card that a P2ABCE server must ask for cryptographic operations.
+It can also be seen as a \textbf{double delegation}.
+The IoT device delegates on the external P2ABCE server to manage the protocol, and the P2ABCE server delegates on the IoT, acting now as a smart card, for the cryptography.
+
 \subsection{System architecture}
-
-% TODO : system architecture:
-% Definir cada etapa: inicio conversación con otro actor, delegación indicando cómo accede a la iot sc, transmisión de APDUs, de los dos últimos la seguridad y escenarios posibles en un último párrafo.
-%describir fig:P2ABCE-IoT
 
@@
 
-Therefore, the IoT device now has a \textbf{duality} in its functions, because it is the User that starts any interaction with other actors, and it's also the smart card that a P2ABCE server must ask for cryptographic operations.
-It can also be seen as a \textbf{double delegation}.
-The IoT device delegates on the external P2ABCE server to manage the protocol, and the P2ABCE server delegates on the IoT, acting now as a smart card, for the cryptography.
+The system will be compounded by the IoT device, the P2ABCE delegation server and the third party P2ABCE actors.
+
+\begin{itemize}
+
+	\item \textbf{IoT device}
+
+	In \autoref{fig:P2ABCE-IoT} the IoT device is represented with two interfaces, physical or virtual.
+One allows external communications to other machines, including other P2ABCE actors, that could be on the Internet, a corporate LAN, a M2M overlay network, etc.
+Through this interface, the P2ABCE XML messages are exchanged as in any other deployment.
+This allows an IoT device to interact with other actors without special adaptations to the protocol.
+The other interface allows a secure communication with the delegation server.
+Both the delegation messages and the APDU Dialogue are transmitted over this interface, making it a point of attack to the system, and we will talk about its security in the delegation process.
+
+	The scheme also shows the \textit{P2ABCE IoT Toolkit}. This piece of software includes the IoT Smart Card, and the API for other processes that want to use the P2ABCE system.
+
+	The IoT Smart Card is the implementation of a software smart card, listens for APDU Commands from the secure interface and stores securely the credentials and private keys within the device's memory.
+
+	The P2ABCE API is an interface for other processes that wish to use the private-preserving environment of P2ABCE.
+It provides access to every operation available, hiding the delegation process.
+In the future, if for example the Verification Service is implemented for the IoT device, i.e., there's no need to delegate to other machine to act as a Verifier, then any program using the API won't need to change anything, the toolkit conceals the transition from delegating to native execution.
+
+	\item \textbf{P2ABCE actors}
+
+	If we recall from \autoref{analysisP2ABCE}, the possible roles in the system were the Issuer, the User, the Verifier, the Revocation Authority and the Inspector.
+All of them use the P2ABCE XML schema in the specification to communicate to each other.
+Any third party actor will be unaware of the fact that the device is a constrained IoT device that delegates on the P2ABCE server.
+
+	\item \textbf{P2ABCE Delegation Server}
+
+	The machine in charge of receiving authorized IoT devices' commands to parse the XML files exchanged and orchestrate the cryptographic operations the IoT smart card must perform.
+
+\end{itemize}
+
+\hfil
+
+\begin{flushleft}
+	\textbf{Delegation process}
+\end{flushleft}
+
+Here we describe the computation offloading carried out by the IoT device.
+In \autoref{fig:DelegationProving} we show an example of the IoT acting as a User, Proving a Presentation Policy to a third party Verifier.
+
+\begin{enumerate}
+	\item Communication with P2ABCE actor.
+
+	The IoT device starts an interaction with another actor, e.g. an Issuer or Verifier, receiving a P2ABCE XML file.
+
+	\item Delegation to the P2ABCE Server.
+
+	Depending on what role the IoT device is acting as, it will delegate in the corresponding service, e.g.
+User Service.
+The delegation message must include the XML file, and any parameter required to accomplish the task, like the information on how to communicate with the IoT smart card (listening port, security challenge, etc.).
+
+	\item APDU Dialogue (if necessary).
+
+	The server may need to send APDU Commands to the IoT smart card to read the credential information or perform cryptographic operations involving private keys.
+
+	\item Server response.
+
+	The server may return a status code or a XML file if the first one required an answer from the IoT device, in which case, it will send as response to the third party actor, resuming the communication.
+
+\end{enumerate}
 
@@
 
+Transmissions over the \textit{Server-IoT} channel must be secured in order to avoid attacks like: impersonate the P2ABCE delegation server, having access to the IoT smart card sending the APDU Commands the attacker wishes; delegate as a device on the server but giving the parameters of another device, making the delegation server send the APDU Commands to a victim IoT smart card.
+
+We could use a corporative PKI to issue certificates to the server and devices and configure policies for access control; design a challenge-response system combined with the smart card PIN, like a password and TOTP\footnote{Time-based One-time Password} in a 2FA\footnote{Two-factor authentication} login; we also could connect physically the delegation service through RS-232 serial to the IoT device, securing both physically as we would do with the IoT device on its own, isolating the delegation system from any network attack.
+
+As we can see, there are many state of the art solutions for all this threads, therefore, we can assume a secure channel without mentioning a specific solution, providing freedom to choose the most fitting one in a real deployment.
+
+\paragraph{Modifications for more constrained devices}
+
+Our architecture is designed for devices that could in a future run a reimplemented version of P2ABCE, that means, the devices could perform more tasks than only running the smart card software and their main purpose process, e.g.
+recollecting sensor data.
+But if our target devices are so constrained that can barely run the smart card, they may not be able to handle the XML files because of memory restrictions, like a MSP430\footnote{\url{http://www.ti.com/lsds/ti/microcontrollers-16-bit-32-bit/msp/overview.page}} running Contiki-OS, the microcontroller has between hundred of bytes to tens of kilobytes of memory, making impossible to store multiple XML files in the size range of tens of kilobytes.
+
+In these cases, the delegation in the server goes a step forward, making the server a proxy to communicate with other P2ABCE actors, and the IoT device only acts as a smart card.
+The IoT device would still act as the User, or any other P2ABCE, role because it orchestrates when and how an interaction with other actor is executed, but the communications would be between the proxy and the third party actor.
+
+\section{Proof of Concept Implementation}
+
 \paragraph{PoC Delegation to the P2ABCE Server} \hfil
@@
 An application that integrates P2ABCE can make use of this services in the same machine or implement the functionality using the core components written in Java, the same ones the REST services use.
-Our PoC machine, the Omega2, can make REST calls easily, but other devices may use \ac{CoAP}, and in that case, the P2ABCE REST services should be rewritten to offer CoAP support.
+Our PoC machine, the Omega2, can make REST calls easily, but other devices may use \ac{CoAP}, and in that case, the P2ABCE REST services should be adapted to offer CoAP support.
 The commands needed to delegate to the P2ABCE delegation server will be the same to operate with the REST services.
@@
 The transmission of the messages will depend on the specific use cases, capabilities and resources available.
-If the delegation server is connected, for example, through RS-232 serial with the IoT device, and physically inaccessible, in the same way an IoT device on its own would be protected, the communication is simple, and not far away from the Arduino Yun idea of combining two devices, one more powerful but to use only when needed.
+If the delegation server is connected, for example, through RS-232 serial to the IoT device, it benefits from the same level of security as the physical device,  reducing the security considerations of the delegation channel to that of the devices security, the communication is simple, and not far away from the Arduino Yun idea of combining two devices, one more powerful but to use only when needed.
 But if the IoT device and the delegation server are apart, or more than one IoT device delegates to it, then the transmission must be secured.
@@
 \hfil
-
-% TODO: imagen con servicios desplegados/actores en cada máquina
 
@@
 Therefore, the implementation of the Verifier functionality would reduce significantly the need of a delegation server, but as we said, managing complex XML squemas is not something many IoT devices could do.
-
-\section{Implementation}
 

### 2017-06-03 Doc/classicthesis/Chapters/Objectives.tex
design init
design arch
UML sequence diagram

@@
 \hfil
+
+% TODO: numerar objetivos: nombrarlos luego por el número cuando se consiguen, empiezan, en conclusiones...
+% TODO: no tan obvio que se han definido objetivos tras el proyecto.
 
@@
 
-\section{Analysis of P2ABCE}
+\section{Analysis of P2ABCE}\label{analysisP2ABCE}
 
@@
 P2ABCE defines the necessary APDU instructions for the smart card needed to implement each method of the interface.
-It relies on \textit{javax.smartcardio} abstract classes (implemented by Oracle in their JRE) to communicate the smart card reader and the smart card.
+It relies on \texttt{javax.smartcardio} abstract classes (implemented by Oracle in their JRE) to communicate the smart card reader and the smart card.
 This way, it doesn't matter what manufacturer issues the smartcard, or if it's an Android device with NFC, if they support the APDU instructions, P2ABCE will work with them.
@@
 	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/p2abceBasicUML}
+		\includegraphics[width=\linewidth]{gfx/UML/p2abceBasicUML}
 	\end{center}

### 2017-06-05 Doc/classicthesis/Chapters/Design.tex
UML sequence diagram
impl avance

@@
 %
-% +Implementation:
-%  -REST as current delegation protocol.
-%  -BIOSC as current APDU transmission protocol: no security -> no overload.
-%  -IoT Smart Card software design and implementation notes.
-%  -Sequence diagrams. TODO
 
@@
 	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/DelegationProving}
+		\includegraphics[width=\linewidth]{gfx/UML/DelegationProving}
 	\end{center}
@@
 
-We could use a corporative PKI to issue certificates to the server and devices and configure policies for access control; design a challenge-response system combined with the smart card PIN, like a password and TOTP\footnote{Time-based One-time Password} in a 2FA\footnote{Two-factor authentication} login; we also could connect physically the delegation service through RS-232 serial to the IoT device, securing both physically as we would do with the IoT device on its own, isolating the delegation system from any network attack.
+We could use a corporative PKI to issue certificates to the server and devices and configure policies for access control; design a challenge-response system combined with the smart card PIN, like a password and TOTP\footnote{Time-based One-time Password} in a 2FA\footnote{Two-factor authentication} login.
+We also could connect physically the delegation service through RS-232 serial to the IoT device, securing both physically as we would do with the IoT device on its own, isolating the delegation system from any network attack.
+This last idea is an approximation to the Arduino Yún\footnote{\url{https://www.arduino.cc/en/Main/ArduinoBoardYun}}, a development board that integrates two microcontrollers, one a typical Arduino with very low resources, and another one running a fork of OpenWrt.
+The Arduino microcontroller can control the terminal of the more powerful one, using the serial pins as commented before.
 
@@
 
-\paragraph{Modifications for more constrained devices}
+\paragraph{Notes for more constrained devices}
 
@@
 \section{Proof of Concept Implementation}
+% +Implementation:
+% *System:
+% 	-Raspbian OS: distro de debian para Raspberry Pi, que se describirá en los tests
+% 	-LEDE: ash terminal (shell command interpreter), POSIX, typical hardware, libraries with no hardware aceleration vs smart card MULTOS,
+%
+% *Delegation
+%  -REST as current delegation protocol: curl script bash
+%  -BIOSC as current APDU transmission protocol: no security -> no overload.
+% *Smart Card
+%  
+%  -IoT Smart Card software design and implementation notes.
+%  -Sequence diagrams. TODO
 
-\paragraph{PoC Delegation to the P2ABCE Server} \hfil
+In this section we present the first PoC implementation, introducing the IoT system where we are going to work, then we will describe the delegation protocols, one for the computation offloading of the IoT device on the P2ABCE server, and another for the transmission of APDU Commands, and finally, we will describe the IoT smart card implementation.
 
-We now will address each type of delegation: from the IoT device as a P2ABCE actor to the P2ABCE delegation server, and from the server to the smart card in the IoT device.
+\subsection{IoT system}
+
+We will develop our PoC in Linux based systems in order to avoid complications with firmware specific issues that we are not familiar with.
+Even so, the linux systems used are aimed for the IoT environment and serve as a starting point for future implementations in more constrained devices or different systems.
+
+In our delegation server we will run Raspbian OS, a distribution based on Debian for the Raspberry Pi.
+We will talk more about the hardware specifications in the benchmark chapter.
+We chose this system because it has a great package repository support, including the Java Runtime Environment, needed to run the P2ABCE Services.
+
+For our IoT device, we will use LEDE, Linux Embedded Development Environment, a distribution born from OpenWrt, aimed for routers and embedded chips with low resources requirements.
+It offers the \texttt{ash} command interpreter, or \texttt{shell}, and the \texttt{opkg} package manager, that allows an easy installation of some libraries and tools needed during the early development.
+
+\subsection{PoC Delegation}
+
+As we explained in the design section, the delegation has two steps, the IoT device calling the P2ABCE server to offload the parsing of the XML data, and the P2ABCE server sending APDU Commands to the IoT smart card in the device.
+
+\subsubsection{PoC Delegation to the P2ABCE Server}
+
+Currently P2ABCE offers multiple REST web services to run different roles in P2ABCE system: User Service, Issuer Service, Verification Service, etc.
+Any third party application that integrates P2ABCE with their system can make use of these services or implement the functionality using the library written in Java, the tool that the REST services actually use.
+
+Our PoC machine, the Omega2, can make REST calls easily with the \texttt{curl} command, but other devices may use \ac{CoAP}, but in that case, the P2ABCE REST services should be adapted to offer CoAP support.
+The commands needed to delegate to the P2ABCE server would be the same that those defined now to operate with the REST services, but with the modifications needed to pass the parameters on how to communicate with the IoT smart card.
+
+In this PoC the only P2ABCE Service that needed to be modified was the User Service. We added the following REST call
+
+\begin{center}
+	\textit{/initIoTsmartcard/{issuerParametersUid}?host=\&port=}
+\end{center}
+
+where we communicate the P2ABCE server that an IoT Smart Card is accessible via \textit{host} and \textit{port}.
+Then a new \textit{HardwareSmartcard} object is stored in the P2ABC Engine, but instead of the \textit{javax.smartcardio} Oracle's \textit{CardTerminal} implementation, we use our own \textit{IoTsmartcardio} implementation for the \textit{HardwareSmartcard} constructor, that we will discuss in the following section.
 
@@
 
-Currently P2ABCE offers multiple REST web services to run different roles in P2ABCE system: User Service, Issuer Service, Verification Service, etc.
-An application that integrates P2ABCE can make use of this services in the same machine or implement the functionality using the core components written in Java, the same ones the REST services use.
-Our PoC machine, the Omega2, can make REST calls easily, but other devices may use \ac{CoAP}, and in that case, the P2ABCE REST services should be adapted to offer CoAP support.
-The commands needed to delegate to the P2ABCE delegation server will be the same to operate with the REST services.
-This way, the first issue is solved.
+In a real deployment, we would offer a full library with an API for other processes that want to use the P2ABC Engine, which automatizes the boot of the IoT smart card and handles the security policies, like the one we presented in the deployment diagram \ref{fig:P2ABCE-IoT} of the \textit{P2ABCE IoT Toolkit}.
 
-The transmission of the messages will depend on the specific use cases, capabilities and resources available.
-If the delegation server is connected, for example, through RS-232 serial to the IoT device, it benefits from the same level of security as the physical device,  reducing the security considerations of the delegation channel to that of the devices security, the communication is simple, and not far away from the Arduino Yun idea of combining two devices, one more powerful but to use only when needed.
-But if the IoT device and the delegation server are apart, or more than one IoT device delegates to it, then the transmission must be secured.
-They could use 6LoWPAN to talk to each other (the delegation service could be deployed in the proxy) and then secure communications with existing solutions, like with pre-shared symmetric keys, certificates for authentication and authorization, etc., it depends on each particular deployment.
+Now we show some examples of the PoC \texttt{curl} commands, run in the device's terminal, or from a shell script for automation:
 
-At the end of the day, this is all about usual security in IoT. Many other studies focus on this matter, so we will assume it can be done, and focus on what's new, P2ABCE in IoT.
+\begin{lstlisting}[language=bash]
+$ curl -X POST --header 'Content-Type: text/xml' "http://$DelegationServerIP:9200/user/initIoTsmartcard/http%3A%2F%2Fticketcompany%2FMyFavoriteSoccerTeam%2Fissuance%3Aidemix?host=192.168.3.1&port=8888"
+
+$ curl -X POST --header 'Content-Type: text/xml' -d @firstIssuanceMessage.xml "http://$DelegationServerIP:9200/user/issuanceProtocolStep/" > issuanceReturn.xml
+\end{lstlisting}
+
+The IoT device will only manage XML as data files to exchange between the third party actor and the P2ABCE delegation service, attaching them in the body of the REST call.
+In the parameters of the first call we can see the IP and port where the IoT smart card will be listening.
+
+\subsubsection{APDU Dialogue Transmission}
+
+To transmit the APDU messages in our PoC we use a simple protocol, that we will refer as BIOSC (Basic Input Output Smart Card), consisting in one first byte for the instruction:
 
@@
 
-To sum up, our IoT device will act as User (Prover or Verifier) keeping its secrets in a software smart card.
-When it starts an interaction with other actor of the P2ABCE system (Issuer, Verifier, etc.), the IoT device will delegate with a remote call (using REST in our PoC) to a P2ABCE delegation server, attaching the XML file and the necessary information for the server to send the APDUs to the software smart card (in our PoC using TCP sockets, giving the IP and listening port).
+\begin{tabular}{|c|l|}
+	\hline
+	\texttt{0x01} & APDU Command or Response: Read a 2 bytes for the\\& length and then as many bytes as said length. \\
+	\hline
+	\texttt{0xff} & Finish connection: close the current TCP socket and\\& end the IoT smart card execution. \\
+	\hline
+\end{tabular}
 
@@
 
-This simple design keeps the benefit of a 100\% compatible P2ABCE deployment, and the integration of IoT devices to the P2ABCE ecosystem.
+In the first instruction, we read two header bytes with the length of the APDU Command or Response to receive, followed by said APDU bytes.
+The message is sent over TCP for a reliable transmission (concept of session, packet retransmission, reordering, etc.).
+The second instruction serves for closing the TCP socket in the IoT device when the smart card is no longer needed.
 
-In the future, more functionality currently delegated in the P2ABCE server can be implemented in the IoT device, if its resources allow it.
-For example, in a M2M environment, where an IoT device can act as User and Verifier, the verification consists on sending a Presentation Policy, and verifying the Presentation Token, which implies less logic than generating it as the User.
-Therefore, the implementation of the Verifier functionality would reduce significantly the need of a delegation server, but as we said, managing complex XML squemas is not something many IoT devices could do.
-
-\section{IoT Smart Card}
-
-After many design decisions in the process to adapt the original ABC4T Card Lite code to pure C, working over a more usual architecture machine, in this section, we present the current \ac{PoC} code, most important decisions, workflow execution, and future work.
-
-First, let's define what a \textit{more usual architecture} is.
-If we remember the MULTOS section, the framework gives an application a very specific memory layout and entry and output points of execution, that could be seen as a single process execution machine.
-Many IoT devices work like a computer, with multiple processes or threads, without pre-loaded data on startup (like the APDU MULTOS loads for the application), a non-volatile memory for data and code, maybe a basic file system in this memory, and RAM with the program's stack, heap, data and code.
-
-Our PoC is tested on a Linux system, and we will give instructions on how to adapt each part to work with other typical IoT systems.
-For example, other IoT devices may work like MULTOS and let access variables in non-volatile memory during execution, and in that case, the port should be changed according to these particularities.
+We lack any security (authentication or authorization) that a real system should implement.
+It is vital to authenticate the delegation service, to authorize it to make APDU Commands, and the same with the IoT device, to prevent attacks.
+But using this method for the transmission of the APDU messages, with only 3 bytes of overhead, helps us in the  benchmarks to measure the real performance of the system.
 
 \hfil
+
+The implementation of this simple protocol is done in Java and C, one for the P2ABCE server and the other one for the IoT smart card.
+
+% TODO : IoTsmartcardio
+% TODO : BIOSC.c
+%TODO : optional: diagrama secuencia de P2ABCE, pero si eso solo de la reimplementación de javax.smartcardio y su uso por HardwareSmartcard
+
+\subsection{IoT Smart Card Implementation}
+
+After many design decisions in the process to adapt the original ABC4Trust Card Lite code to pure C, working on a MIPS architecture, in this section, we present the current \ac{PoC} code, most important decisions and execution workflow in the form of sequence diagrams.
+
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/UML/sequenceBIOSC}
+	\end{center}
+	\caption{IoT Smart Card Sequence Diagram.}
+	\label{fig:sequenceBIOSC}
+\end{figure}
+
+\hfil
+
+%%%%%%%%%%%%%%%%%%%%%%%
+
+%% DRAFT %%
 
@@
 
-The first section is what could be called as the core of the smart card, the second one the interface for the tools the core need and may depend on the platform, and finally third party libraries, that in may be empty if the interfaces implementation doesn't need any.
-
-In our PoC we used CMake to manage the project, due to the cross-compilation tools, integration with multiple IDEs and tests.
+The first section is what could be called as the core of the smart card, the second one the interface for those tools the core needs and may depend on the target's platform, and finally, the third party libraries, that may be empty if the interfaces implementation doesn't need any.
 
@@
 
-The smart card logic lies in this section, the concepts of APDU Commands, what instructions are defined in P2ABCE smart cards and how to process them and generate proper APDU Responses.
+The smart card logic lies in this section, the concepts of APDU Commands, what instructions are defined in P2ABCE smart cards, and how to process them and generate proper APDU Responses.
 
-Changes in the APDU protocol for P2ABCE must be done here, independently of the target platform.
+If the APDU protocol defined for P2ABCE smart cards changed, with the consequence of having to update all smart cards in use, the changes to adapt the IoT smart card should be applied here.
+This also implies that if the P2ABCE Java code changes their \texttt{Smartcard} interface, as long as it is backwards compatible with existing smart cards, the IoT smart card will still work as intended.
 
-After refactoring the original ABC4Trust Card's code, most of it fell in what we will call the core of the smart card.
+Most of the original ABC4Trust Card's code has been reused in this section, like all the variables and custom data type definitions, or the APDU Commands handling.
 
-All types and variable definitions and the APDU handling is done in this code.
-However, the ABC4Trust's code depended on the MULTOS C-API for the input/output of data, modular arithmetic, and even AES128 and SHA256 cryptography.
+However, the ABC4Trust's code heavily depended on the MULTOS C-API for the input/output of data, modular arithmetic, and AES128 and SHA256 cryptographic operations.
 
-A characteristic of MULTOS C-API is that every function name starts with \textit{multos}, but as we said, the \textit{main.h} file implemented equivalent functions to some available in \textit{multos.h}.
-Our first step was to replace the \textit{main.h} functions for the standard ones in the C-API.
-Then, we implemented, following the C-API documentation, the functions from \textit{multos.h} (only the used ones) changing their names from \textit{multosFoo()} to \textit{mFoo()} for readability and emphasize that they were no longer from MULTOS.
+A characteristic of MULTOS C-API is that every function name starts with ``\texttt{multos}'', so we implemented, following the C-API specification, the functions used, but changing their names from \texttt{multosFoo()} to \texttt{mFoo()} for readability and to emphasize that they were no longer MULTOS functions.
+
+The \texttt{main.h} file, from the original project, also implemented equivalent functions to some available in \texttt{multos.h}.
+We replaced them for the standard functions in the C-API, checking that the MEL code did the same as the documentation specified.
+This last detail is very important, because in \citep{vullers2013efficient} they noted that \texttt{ModularExponentiation} function dos not accept exponents larger than the modulus size, and they added \texttt{SpecialModularExponentiation}, dividing the exponentiation in two that MULTOS could perform.
+After this analysis, when those particularities appeared, we decided to adapt the reimplementation of the C-API.
 
@@
 \paragraph{External utilities}
+% Podría haber implementaciones de librerías criptográficas en ensamblador de la plataforma, se incluyen aquí y la interfaz lo usa al implementar
 
@@
 
-\subsection{PoC Workflow}
-
 % PoC BIOSC : TCP sockets, JSON serialization
@@
 % PoC IoT javax.smartcardio : protocolo de 2 bytes + APDU
+
+\subsection{PoC Workflow}
 
@@
 
-\paragraph{P2ABCE REST Delegation}
-
-The only P2ABCE Service that needed to be modified was the User Service. We added the REST call
-
-\begin{center}
-	\textit{/initIoTsmartcard/{issuerParametersUid}?host=\&port=}
-\end{center}
-
-where we communicate the P2ABCE server that a IoT Smart Card is accessible via \textit{host} and \textit{port}.
-Then a new \textit{HardwareSmartcard} object is stored in the P2ABC Engine, but instead of the \textit{javax.smartcardio} Oracle's \textit{CardTerminal} implementation, we use our \textit{IoTsmartcardio} implementation for the \textit{HardwareSmartcard} constructor.
-
-The rest of P2ABCE code is unchanged and will work as if a real smart card was in use. \textit{IoTsmartcardio} implementation will transmit the APDUs through the TCP socket with the format mentioned.
-
-\hfil
-
-Our test consists on the tutorial available in P2ABCE's repository. From the Omega2's terminal we run \textit{curl} commands to send REST petitions. For example:
-
-\begin{verbatim}
-$ curl -X POST --header 'Content-Type: text/xml'
-'http://localhost:9200/user/initIoTsmartcard/http%3A%2F%2
-Fticketcompany%2FMyFavoriteSoccerTeam%2Fissuance%3Aidemix
-?host=192.168.3.1&port=8888'
-\end{verbatim}
-
-The IoT device will only manage XML as data files to exchange between the third party actor and the P2ABCE delegation service, without parsing them.
-
-After a REST call to the User Service, the P2ABC Engine will talk to the IoT smart card process.
-
 \paragraph{APDU transmission with BIOSC}
-
-% TODO : diagrama del software
 
@@
 To do this, the \textit{output\_large\_data} function and the \textit{IoTsmartcardio} implementation of \textit{javax.smartcardio} must support the feature.
-
-% TODO -Sequence diagrams.

### 2017-06-05 Doc/classicthesis/Chapters/Objectives.tex
impl avance

@@
 
-The cryptographic nature of the credential-as-container concept includes the
-following four security aspects \citep{vullers2013efficient}:
-\begin{itemize}
-	\item The issuer’s digital signature ensures \textbf{authenticity}: the credential originates
-	from the issuer, and this issuer asserts that the attributes hold for the person.
-	\item This signature also guarantees integrity: the attributes contained in the credential
-	have not been altered since they were issued.
-	\item A credential is \textbf{non-transferable} as it is bound to the secret key, only known by the credential owner.
-	\item A credential \textbf{hides} its content, so it does not reveal the attributes it contains.
-	Furthermore, a credential protects the privacy of its owner through the following
-	cryptographic properties.
-	\item Issuer \textbf{unlinkability} ensures that any information gathered during issuing
-	cannot be used to link a verification of the credential to its issuance.
-	\item \textbf{Multi-show unlinkability} guarantees that when a credential is verified multiple
-	times, these sessions cannot be linked.
-\end{itemize}
-
-Zero-knowledge proofs, used in Idemix, allow a user to prove ownership of a credential without
+Zero-knowledge proofs used in Idemix allow a user to prove ownership of a credential without
 revealing the credential itself. Since the verifier does not see the credential, verification
@@
 The privacy of users is protected by these unlinkability properties, even if the credential issuer and all verifiers collude.
+We can identify the following key features of the Idemix system \citep{vanet}:
+
+\begin{itemize}
+
+	\item \textbf{Issuance unlinkability}: the issuer cannot link an issued credential to the presentation of such credential.
+
+	\item \textbf{Multi-show unlinkability}: a credential can be used multiple times without
+	the resulting evidence becoming linkable.
+
+	\item \textbf{Selective disclosure of attributes}: allows users to prove only a subset of
+	attributes to a verifier.
+
+	\item \textbf{Predicate proof}: it consists of statements that allow to prove a property
+	of an attribute without disclosing its actual value. Example of these statements
+	are the logical operators > or <.
+
+	\item \textbf{Proof of holdership}: a cryptographic evidence for proving ownership or
+	possession of a credential without disclosing the attributes contained in
+	that credential.
+
+	\item \textbf{Non-transferability}: key binding can be used to bind one or more credentials
+	of a user to the same secret.
+
+	\item \textbf{Scope-exclusive pseudonyms}: a certified pseudonym unique for a specific
+	scope and secret key, i.e. a single pseudonym can be created for each credential.
+
+	\item \textbf{Carry-over attributes}: it relies on the assumption that the user already
+	possesses a credential, from which a given attribute can be carried over
+	into the new credential without disclosing the attribute value to the Issuer.
+
+	\item \textbf{Cross-credential proofs}: it allows users to prove relations between attributes
+	from two or more credentials without revealing them to the verifier. For
+	instance, that the name contained on a credit card and on a passport
+	match.
+
+	\item \textbf{De-anonymization}: it is an optional feature that allows an Inspector, either
+	alone or in cooperation with other entities, to reveal the identity of users
+	in cases of accountability and non-repudiation.
+
+	\item \textbf{Revocation}: in case of misuse, it allows the revocation of issued credentials
+	to (misbehaving) users. Thus, revoked credentials can no longer be used to
+	generate Presentation Tokens.
+
+\end{itemize}
 

### 2017-06-05 Doc/classicthesis/Chapters/Validation.tex
UML sequence diagram
impl avance

@@
 
-Also, we will need a cross-compiler to generate the smart card binary.
-Because we will use GMP and OpenSSL as shared libraries, the best option is to use the LEDE SDK.
-The SDK manages the available packages and generates the GCC toolchain.
-Also, using CMake,
-%TODO
+A downside of using LEDE and the Omega2 is the lack of hardware acceleration for cryptographic operations, unlike the MULTOS applications, because the smart cards hardware and MULTOS API include support for such common operations in smart cards.
 
@@
 	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/IssuanceInteraction}
+		\includegraphics[width=\linewidth]{gfx/UML/IssuanceInteraction}
 	\end{center}
@@
 	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/ProvingInteraction}
+		\includegraphics[width=\linewidth]{gfx/UML/ProvingInteraction}
 	\end{center}

### 2017-06-06 Doc/classicthesis/Chapters/Design.tex
finishing touches

@@
 %************************************************
-
-%TODO resumir organización del capítulo
 
@@
 	\hline
-	\texttt{0x01} & APDU Command or Response: Read a 2 bytes for the\\& length and then as many bytes as said length. \\
+	\texttt{0x01} & APDU Command or Response: Read 2 bytes for the\\& length and then as many bytes as said length. \\
 	\hline
@@
 
-The implementation of this simple protocol is done in Java and C, one for the P2ABCE server and the other one for the IoT smart card.
+The implementation of this simple protocol is done in both Java and C, one for the P2ABCE server and the other one for the IoT smart card.
 
-% TODO : IoTsmartcardio
-% TODO : BIOSC.c
+\hfil
+
+As we mentioned in the analysis of P2ABCE's code, the \texttt{HardwareSmartcard} class implements the \texttt{Smartcard} interface using the package of abstract classes \texttt{javax.smartcardio} to communicate with the physical smart cards.
+The Oracle JRE implements these package for the majority of smart card manufacturers.
+For our PoC, we implement the \texttt{javax.smartcardio} package so it transmits the APDUs with BIOSC, making the use of a physical or IoT smart card totally transparent to the \texttt{HardwareSmartcard} class, enhancing maintainability, and following the \textit{expert pattern} from the known GRASP guidelines.
+
 %TODO : optional: diagrama secuencia de P2ABCE, pero si eso solo de la reimplementación de javax.smartcardio y su uso por HardwareSmartcard
+
+\textbf{TODO} :  diagrama secuencia de la reimplementación de javax.smartcardio y su uso por HardwareSmartcard
+
+\hfil
+
+In the PoC IoT device, we implement in \texttt{BIOS.c} the \texttt{main()} function, that reads from a Json file the previous status of the smart card, and then opens the TCP socket in the $8888$ port, and loops listening for BIOSC transmissions until an $0xff$ is received.
+When an APDU Command is sent, it calls the APDU Handler, that we will talk about in the next section.
 
@@
 
-After many design decisions in the process to adapt the original ABC4Trust Card Lite code to pure C, working on a MIPS architecture, in this section, we present the current \ac{PoC} code, most important decisions and execution workflow in the form of sequence diagrams.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/UML/sequenceBIOSC}
-	\end{center}
-	\caption{IoT Smart Card Sequence Diagram.}
-	\label{fig:sequenceBIOSC}
-\end{figure}
+After many design decisions in the process to adapt the original ABC4Trust Card Lite code to pure C, working on a MIPS architecture, in this section, we present the current \ac{PoC} code, most important decisions taken and the execution workflow in the form of sequence diagrams.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%
-
-%% DRAFT %%
-
-\subsection{Code structure}
+\subsubsection{Code structure}
 
@@
 
-The first section is what could be called as the core of the smart card, the second one the interface for those tools the core needs and may depend on the target's platform, and finally, the third party libraries, that may be empty if the interfaces implementation doesn't need any.
+In \autoref{fig:IoTCScomponents-bw} we show the project's structure.
+The first section is what could be called as the core of the smart card, the second one the interface for those tools the core needs and that may depend on the target's platform, and finally, the third party libraries, that may be empty if the interfaces implementation doesn't require any.
 
@@
 \paragraph{Core smart card}
+\hfil
 
-The smart card logic lies in this section, the concepts of APDU Commands, what instructions are defined in P2ABCE smart cards, and how to process them and generate proper APDU Responses.
+The smart card logic lies in this section, the concepts of APDU Commands, what instructions are defined for P2ABCE smart cards, and how to process them and generate proper APDU Responses.
 
-If the APDU protocol defined for P2ABCE smart cards changed, with the consequence of having to update all smart cards in use, the changes to adapt the IoT smart card should be applied here.
-This also implies that if the P2ABCE Java code changes their \texttt{Smartcard} interface, as long as it is backwards compatible with existing smart cards, the IoT smart card will still work as intended.
+If the APDU protocol defined for P2ABCE smart cards changed, with the consequence of having to update all smart cards in use, the changes to adapt the IoT smart card should be applied in this part of the code.
+This also implies that if the P2ABCE Java code changes their \texttt{Smartcard} interface, something more probable, the IoT smart card will still work as intended.
 
-Most of the original ABC4Trust Card's code has been reused in this section, like all the variables and custom data type definitions, or the APDU Commands handling.
+Most of the original ABC4Trust Card's code has been reused in this section, e.g., all the global variables and custom data type definitions, the APDU Command handling and auxiliary subroutines.
 
-However, the ABC4Trust's code heavily depended on the MULTOS C-API for the input/output of data, modular arithmetic, and AES128 and SHA256 cryptographic operations.
+However, the ABC4Trust's code heavily depended on the MULTOS platform for the input and output of data, modular arithmetic, memory management, and AES128 and SHA256 cryptographic operations.
+Therefore, we reimplement the used MULTOS functions following the documentation, adapting it in some cases.
 
-A characteristic of MULTOS C-API is that every function name starts with ``\texttt{multos}'', so we implemented, following the C-API specification, the functions used, but changing their names from \texttt{multosFoo()} to \texttt{mFoo()} for readability and to emphasize that they were no longer MULTOS functions.
+A characteristic of MULTOS C-API is that every function name starts with ``\texttt{multos}'', so we changing their names from \texttt{multosFoo()} to \texttt{mFoo()} for readability and to emphasize that they were no longer MULTOS functions.
 
-The \texttt{main.h} file, from the original project, also implemented equivalent functions to some available in \texttt{multos.h}.
+The \texttt{main.h} file, from the original ABC4Trust project, also implemented some functions that were equivalent to the ones available in \texttt{multos.h}.
 We replaced them for the standard functions in the C-API, checking that the MEL code did the same as the documentation specified.
-This last detail is very important, because in \citep{vullers2013efficient} they noted that \texttt{ModularExponentiation} function dos not accept exponents larger than the modulus size, and they added \texttt{SpecialModularExponentiation}, dividing the exponentiation in two that MULTOS could perform.
-After this analysis, when those particularities appeared, we decided to adapt the reimplementation of the C-API.
+This last detail is very important, because in \citep{vullers2013efficient} they noted that MULTOS' \texttt{ModularExponentiation} function does not accept exponents larger than the modulus size, so they implemented \texttt{SpecialModularExponentiation}, dividing the exponentiation in two that MULTOS could perform.
+After this analysis, when those particularities appeared, our reimplementation of the MULTOS API accepts the \textit{expanded functionality}.
 
-Future changes in the code may refactor it so there's no longer need for the MULTOS framework functions.
+\hfil
 
-\paragraph{Interfaces}
+Because of the inherited code, many particularities of the internals of the MULTOS platform had to be addressed.
 
-To implement MULTOS functions, we needed to use some libraries, so we defined a facade to isolate the implementation of the core smart card from our different options, that could vary depending on the hardware or the system used by the IoT device.
+One was the different memory zones, where the variables in \texttt{static} were read and written directly from the secure EEPROM, defining the status of the smart card between executions.
+We keep those variables in RAM, but save them in a Json file, deserializing them at the smart card boot up, and serializing always before an APDU Response is sent.
+From the MULTOS documentation, the \texttt{static} variables are read and written atomically, and if power source is lost during any process, memory won't get corrupted.
+This also means that if a MULTOS smart card sends an APDU Response, every variable that had to change, is saved in the EEPROM.
+Our PoC uses Json for human readability during debugging, but a real deployment should apply secure reads and writes on the file, preventing data corruption and undesired access, for example, ciphering the file with a key derived from the smart card PIN.
+It is also desirable another data serialization method, reducing the size of the file to store in memory constrained devices.
 
-The use of a facade lets us, for example, change the implementation of modular arithmetic with a hardware optimized version, or a future more lightweight library, or our very own software implementation using the same data types that the core uses, minimizing the data transformations needed.
+Another of the mentioned particularities is that the MULTOS compiler does not apply padding between variables in the data structures.
+For example, in an X86 machine, where a \texttt{short} is stored in 2 bytes, in the following code\footnote{From \url{https://en.wikipedia.org/wiki/Data_structure_alignment}}
 
-Taking a step forward, we make the core smart card totally independent of any library, only on our interfaces.
+\begin{lstlisting}[language=C]
+	struct MyData
+	{
+		short Data1;
+		short Data2;
+		short Data3;
+	};
+\end{lstlisting}
+
+each member of the data structure would be 2-byte aligned.
+Data1 would be at offset 0, Data2 at offset 2, and Data3 at offset 4.
+The size of this structure would be 6 bytes.
+But if we defined the following structure
+
+\begin{lstlisting}[language=C]
+	struct MixedData
+	{
+		char Data1;
+		short Data2;
+		int Data3;
+		char Data4;
+	};
+\end{lstlisting}
+
+where a \texttt{char} is 1 byte, and an \texttt{int} uses 4 bytes, although the total bytes used are 8 bytes, the compiler will align them in memory adding padding, like if we defined the following:
+
+\begin{lstlisting}[language=C]
+struct MixedData  /* After compilation in 32-bit x86 machine */
+{
+	char Data1; /* 1 byte */
+	char Padding1[1]; /* 1 byte for the following 'short' to be aligned on a 2 byte boundary
+	assuming that the address where structure begins is an even number */
+	short Data2; /* 2 bytes */
+	int Data3;  /* 4 bytes - largest structure member */
+	char Data4; /* 1 byte */
+	char Padding2[3]; /* 3 bytes to make total size of the structure 12 bytes */
+};
+\end{lstlisting}
+
+\hfil
+
+In MULTOS, to reduce the memory usage, there is no padding, and this affects the inherited ABC4Trust's code because of the use of \texttt{memcpy} to copy zones of memory from one address to another.
+The problem is the code copies multiple variables in one \texttt{memcpy} call, because, for example, the APDU Command payload includes multiple data, and instead of copying one variable at a time, the \texttt{struct} is defined with the same order and copies everything at once.
+
+An example taken from the parsing of instruction \texttt{SET\_PROVER}:
+
+\begin{lstlisting}[language=C,frame=tblr]
+typedef struct
+{
+	BYTE prover_id;
+	unsigned int ksize;
+	unsigned int csize;
+	BYTE kx[MAX_SMALLINT_SIZE];
+	BYTE c[HASH_SIZE];
+	BYTE proofsession[PROOFSESSION_SIZE];
+	BYTE proofstatus;
+	BYTE cred_ids[NUM_CREDS];
+	BYTE cred_ids_size; // also called 't' in the documentation
+	BYTE exists;
+} PROVER;
+/*****/
+PROVER provers[NUM_PROVERS];
+
+/*****/
+case INS_SET_PROVER:
+  // [...]
+  memcpy(&(provers[temp_prover_id-1].prover_id), &(apdu_data.set_prover_in.prover_id), 5); // under the hood, this also initializes ksize and csize
+\end{lstlisting}
+This is the only case where a comment showcases that it is copying more than one byte.
+
+The temporal solution is to use \texttt{struct \_\_attribute\_\_((\_\_packed\_\_))} to ask a GCC compiler to not use padding in the structs, but this is not standard, neither a good practice.
+A deeper refactorization of the code is needed where the hidden copies of variables are made explicit, letting the compiler manage the memory layout.
+
+\hfil
+
+An internal alarm should always warn us when we see raw memory copy from serialized data, that is, the APDU Payload.
+We mentioned in the MULTOS introduction that the platform is Big Endian.
+The serialized data is also typically represented in Big Endian, and our APDU Payloads do so.
+In our code we added a new Subroutine to check whether the device is Big or Little Endian, and to rectify the Endianness from those variables, with 2 or more bytes (the only ones affected by the Endianness), copied from or written to APDU Payloads.
+
+\hfil
+
+Finally, we must talk about the \texttt{multosExit} function.
+In a MULTOS application, to finish the execution and send the APDU Response, the programmer can call in any moment to \texttt{multosExit}, and the MULTOS OS will continue the execution, not returning to the application again.
+
+Think about how we all have, at least once, used an \texttt{if} and \texttt{return} to avoid the \texttt{else}, like in this example:
+
+\begin{lstlisting}[language=C]
+if(condition)
+	return a;
+return b;
+\end{lstlisting}
+
+The use of \texttt{multosExit} in ABC4Trust code is similar, but used in almost every function.
+This leads to a sequence execution with many possible interruptions, leaving a full stack behind with \textit{un-returned} function calls, i.e., parameters stored in memory after a function call that never returns.
+In MULTOS, this memory won't affect future executions, because with the next APDU Command, the stack will be loaded with the clean application, and the Session memory stays untouched between APDU Commands, during an APDU Dialogue.
+
+To adapt the code, the reimplemented \texttt{mExit} function can not end the program execution, because we would lose the variables in RAM needed during the APDU Dialogue, as well as that the TCP socket would close.
+The solution is to refactor every use of \texttt{multosExit} in a way that, if it is called from a Subroutine, the function returns a success or error code, and the only functions that can call \texttt{mExit} are the ones handling an specific instruction.
+
+Nevertheless, there is one exception, the \texttt{output\_large\_data()} Subroutine, which is a tool to send large data payloads without Extended APDU support.
+We talked about this in \ref{subsec:APDU}, mentioning the special instruction \texttt{GET RESPONSE}.
+
+Summarizing, every Subroutine must return from its execution, as well as every instruction handler, which must call either \texttt{mExit} or \texttt{output\_large\_data()} before returning.
+After the instruction has been handled, the execution returns to the listening loop of BIOSC.
+
+\paragraph{Interfaces}\hfil
+
+To reimplement some of the MULTOS functions, we needed to use some libraries, so we defined a facade to isolate the implementation of the core smart card from our different options, that could vary depending on the hardware or the system used by the IoT device.
+
+The use of a facade lets us, for example, change the implementation of modular arithmetic with a hardware optimized version, or a future more lightweight library, or our very own software implementation using the same data types that the core uses, minimizing the data usage.
+
+Taking a step forward, we make the core smart card totally independent of any library, only dependant of our interfaces.
 This means that typical C libraries, like the standard \textit{stdlib.h}, or  \textit{string.h} are also behind the facade, in case some IoT system doesn't support them.
@@
 
-\paragraph{External utilities}
-% Podría haber implementaciones de librerías criptográficas en ensamblador de la plataforma, se incluyen aquí y la interfaz lo usa al implementar
+The interfaces defined can be organized in 5 groups (see \autoref{fig:IoTCScomponents-bw}), depending on their purpose: Modular Arithmetic, Cryptography, Memory Management, Serialization, APDU Parsing.
 
-If the IoT system offers well tested libraries that could aid in the interfaces implementation, or we simply found a pure C implementation for the task, these third party libraries belong to this section.
+\paragraph{External utilities}\hfil
+
+If the IoT system offers well tested libraries that could aid in the interfaces implementation, for example, a assembly optimized code for the AES128 and SHA256 operations, these third party libraries belong to this section.
 
 In our PoC, we use two ANSI C libraries, for base64 and JSON, and two shared libraries available in as packages in LEDE, GMPLib and OpenSSL.
-The last two libraries offer more functionality than we need, hence, it's desired in a production code to implement \textit{Modular Arithmetic} and \textit{Cryptography} interfaces with more lightweight alternatives.
+These libraries use dynamic memory and offer more functionality than we need, although they are a good tool in this PoC, future versions should use more lightweight solutions.
 
@@
 For random number generation, a technique used with Contiki devices is to read from sensors aleatory data and use it as seed.
-All these alternatives depend on the target device, but are all valid.
-The \textit{interfaces} and \textit{external utilities} sections allow for a clean and fast port of the code.
 
-% PoC BIOSC : TCP sockets, JSON serialization
+All these alternatives depend on the target device, but are all valid, and would substitute OpenSSL with only a Serial communication library, so we can talk to the dedicated chips.
 
-% PoC IoT javax.smartcardio : protocolo de 2 bytes + APDU
+\hfil
 
-\subsection{PoC Workflow}
+The \textit{interfaces} and \textit{external utilities} sections  allow that the project is easily ported to specific targets without modifying the smart card logic.
 
-% 1. Boot: deserialize sc status
-% 2. Listening: open tcp socket
-% 3. The IoT device delegates on the P2ABCE for some task and sends the IP and Port of the IoT Smart Card.
-% 4. An APDU Command arrives : read in 2 bytes the length of the apdu and
+\subsubsection{Execution workflow}
 
-In a real deployment, we would offer a full library with an API to other processes to delegate on the P2ABC Engine, that automatizes the listening and security, that we presented in the deployment diagram \textbf{TODO:ref-img} as the \textit{P2ABCE IoT Toolkit}.
+The sequence diagram from \autoref{fig:sequenceBIOSC} shows the execution of the PoC IoT smart card.
 
-But to test the IoT smart card, we use the \textit{curl} command for REST delegation calls and the BIOSC (Basic Input Output Smart Card) for APDU communication.
+\begin{figure}[bth]
+	\begin{center}
+		\includegraphics[width=\linewidth]{gfx/UML/sequenceBIOSC}
+	\end{center}
+	\caption{IoT Smart Card Sequence Diagram.}
+	\label{fig:sequenceBIOSC}
+\end{figure}
 
-To transmit the APDU messages in our PoC we use a simple protocol, consisting in one first byte for the instruction: receive an APDU Command or close the connection.
-In the first case, then we read two header bytes with the length of the APDU Command or Response to receive, followed by those APDU bytes.
-The message is sent over TCP for a reliable transmission (concept of session, packet retransmission, reordering, etc.).
+The program starts with the \texttt{main} function in BIOSC, that deserializes the status from the Json file, and listens on a loop for APDU Commands from the delegation server.
 
-We lack of any security (authentication and authorization) that a real system should implement.
-It is vital to authenticate the delegation service, to authorize it to make APDU Commands, and the same with the IoT device, to prevent attacks.
-This belongs to \textit{usual} security, as we already said, and for that reason it's not in the PoC.
+Every time an APDU Command arrives, it calls the function \texttt{handle\_APDU()} with the raw APDU bytes.
+The Handler calls the APDU I/O interface to parse the bytes, storing in global variables the APDU structure.
+Using a \texttt{switch-case} expression on the \texttt{INS} byte, the Handler calls an \textit{Instruction handler} function.
 
-\paragraph{APDU transmission with BIOSC}
+Inside this function, it may call multiple functions from the Subroutines, that may call MULTOS C-API functions, Which in turn may use an interface to perform its functionality.
 
-Before sending the REST message, the Omega2 must be listening in the specified port.
-At start up, BIOSC reads from a JSON file the status of the smart card (credentials in the BLOB, private keys, PIN and PUK codes, etc.), then opens the TCP socket and listens in a loop.
+As we mentioned, every instruction handler must end, before the \texttt{return;} expression, with \texttt{mExit} or  \texttt{output\_large\_data()}, that will use \texttt{mExit}.
+This reimplemented MULTOS function will save the current status of the smart card, with the help of the serialization interface, and once the file is saved, it outputs the APDU Response back to the delegation server.
 
-When the delegation server sends an APDU Command following the simple protocol, BIOSC stores in a byte array the APDU Command bytes and calls the \textit{handle\_APDU} function in \textit{Core Smart Card}.
-
-The \textit{APDU\_handler} will parse the APDU bytes and check both CLA and INS bytes.
-For each possible INS the is a function that must always finish calling either \textit{mExit()} or \textit{output\_large\_data()}.
-
-The \textit{mExit} function is the reimplementation of MULTOS C-API exit functionality, that finishes the application execution returning to the MULTOS OS, that will send the APDU Response bytes to the terminal.
-This led in the original ABC4Trust Card's code to some tricky situations.
-Imagine a function that ends with an if-else expression.
-Many times we save a line writing something like
-
-\begin{verbatim}
-if(condition)
-return a;
-return b;
-\end{verbatim}
-
-instead of the complete
-
-\begin{verbatim}
-if(condition)
-return a;
-else
-return b;
-\end{verbatim}
-
-This is a swift example of what the \textit{multosExit} function led to, because it is called in both the \textit{switch-case} processing the APDU instruction, and the \textit{Subroutines}.
-For example, the PIN code check was programmed in a way every reason to fail finished the application execution, with different error codes.
-Another example, checking if a credential is stored, if not, fail and exit inside the \textit{readBlob} subroutine.
-
-In our standard architecture machine, if we call a subroutine, it must end, and return the control to the handling function, and so the \textit{APDU\_handler} must return to the listening loop of BIOSC.
-
-Having all that in count, the \textit{mExit} function is not implemented as the documentation specifies, but only saves the smart card status serializing it in the JSON file again, and once it's saved, parses the APDU Response and sends it to the socket.
-
-The \textit{output\_large\_data} function is a tool used in MULTOS smart cards that don't support Extended APDUs.
-The APDU Command handler saves in a buffer the data to send, and its size.
-Then, \textit{output\_large\_data} will manage the buffer expecting future GET RESPONSE commands from the P2ABCE server.
-
-A future change to the project may include support for Extended APDUs, avoiding the use of multiple GET RESPONSE.
-To do this, the \textit{output\_large\_data} function and the \textit{IoTsmartcardio} implementation of \textit{javax.smartcardio} must support the feature.
+After returning from \texttt{mExit}, the instruction handler and the APDU handler functions, the program listens again from the socket, until an $0xff$ byte is received.

### 2017-06-06 Doc/classicthesis/Chapters/Introduction.tex
finishing touches

@@
 
-%TODO
-TODO: organización, basada en los objetivos
+This document is structured as follows: In \autoref{ch:stateoftheart} we show a state of the art analysis through the history of Idemix and related works, analysing what is of the most interest for the IoT perspective; then, in \autoref{ch:objectives} we outline the project's objectives and analyse in depth the existing solutions that we are going to deal with; in \autoref{ch:design} we describe the formal design of the IoT and Idemix solution, and describe the PoC implementation developed; after an implementation, it is a must to validate it, as showed during the performance tests in \autoref{ch:validation}; finally, our conclusions and lines for future work are described in \autoref{ch:conclusions}.

### 2017-06-06 Doc/classicthesis/Chapters/Objectives.tex
finishing touches

@@
 %************************************************
-\chapter{Analysis and objectives}\label{ch:objectives}
+\chapter{Objectives and Analysis}\label{ch:objectives}
 %************************************************
 
-% descripción del proyecto: objetivos, metodología: pasos del diseño y desarrollo, ver figura 3.1 (elección de sensores: computing offloading), contexto (partimos de ABC4T: recordar P2ABCE y que se basa en tarjetas), entorno de desarrollo (meter elección del Omega2, LEDE, teoría de APDUs, código heredado) el sw, los tests p2abce puro,
-
-%TODO resumir organización del capítulo
-% Incluir bibliografía: vanet, wiki
-
-In this chapter we describe the project objectives, and the methodology followed during its development. In section ... TODO
+In this chapter we describe the project objectives, and the methodology followed during its development.
+We continue then describing the privacy-preserving attribute based credentials and the ecosystem surrounding it, from the P2ABCE's perspective, as it is the starting point of our project.
+We finish with an introduction to all the technologies involved in our development, from P2ABCE's code, to how smart cards work.
 
@@
 
-% TODO: numerar objetivos: nombrarlos luego por el número cuando se consiguen, empiezan, en conclusiones...
-% TODO: no tan obvio que se han definido objetivos tras el proyecto.
+We will divide the project in various objectives:
 
-We will divide the project in various \textbf{objectives}, starting from the principal goal, dividing our work in different categories.
+\begin{enumerate}
+	\item Idemix and the IoT
 
-\hfil
-
-\paragraph{Idemix and the IoT}
 	This is our main objective, integrating IBM's privacy-preserving system in the IoT environment.
 
-\paragraph{Analysis} Study the state of the art of Idemix and the IoT, analysing related projects, papers with similar approaches, and consider the best fitting solutions from where to begin.
+	\item Study of the Idemix ecosystem
 
-\paragraph{Design and implementation} After studying the existing works, we must give a formal solution to be implemented.
-This includes the theoretical architecture, the steps to take, the software to implement, the hardware to use, and any simplifications to leave as future work.
+	Study the state of the art of Idemix and the IoT, analysing the related projects and papers with similar approaches.
+Once we know as much as we can from the existing ecosystem, we will be able to deliver the most fitting solution to our project.
 
-Our \textbf{software} objectives include making it easily maintainable, structured and extensible, from the IoT and original project perspectives, that is, the IoT solution must be interoperable with other non-IoT solutions, now and in the future, with minimal effort, given new IoT systems or changes in the cryptographic protocols.
+	\item Design of the IoT+Idemix solution
 
-Our \textbf{hardware} objective is to use devices as constrained as possible.
-But considering our early situation in the project, we will use IoT devices that ease testing and development, having in mind the other devices, and what specific steps we should take for them in the future.
+	A formal description of the IoT architecture that will use Idemix.
 
-\paragraph{Validity and evaluation} Deploy a \ac{PoC} in a \textbf{real scenario}, without the need of simulators, checking it works as expected, and measuring its performance.
+	\item Implementation of a functional PoC
+
+	\subitem Software objectives
+
+	A maintainable, structured and extensible solution, from the IoT and original project perspectives, that is, the IoT solution must be interoperable with other non-IoT solutions, now and in the future, given new IoT systems or changes in the cryptographic protocols.
+
+	\subitem Hardware objectives
+
+	Be compatible with devices as constrained as possible, considering the multiple IoT platforms existing.
+
+	\item Validation of the PoC
+
+	Deploy a \ac{PoC} in a real scenario, without the need of simulators, checking that it works as expected.
+
+	\item Evaluation of performance
+
+	Measure and compare the solution's performance against the existing ones.
+
+	\item Integrate the IoT+Idemix as a security solution for other projects
+
+	The IoT+Idemix project is a solution for the privacy threads in the IoT environment, and does not work on its own, but as part of other projects.
+
+\end{enumerate}
 
@@
 To send more data in an APDU Response, the status bytes are set to: SW1=0x61 and SW2 to the remaining bytes to send.
-Because a smart card can't send APDU Commands, the card terminal must send a \textit{GET RESPONSE}, a special APDU Command, with Le set to the number of bytes specified in SW2.
+Because a smart card can't send APDU Commands, the card terminal must send a \texttt{GET RESPONSE}, a special APDU Command, with Le set to the number of bytes specified in SW2.
 Iterating this process, the smart card can send as many bytes as it wants as Response.
@@
 
-\textbf{Static memory} are the application variables declared after the specific \textit{\#pragma melstatic} compiler directive.
+\textbf{Static memory} are the application variables declared after the specific \texttt{\#pragma melstatic} compiler directive.
 These variables are stored in the non-volatile secure EEPROM, and any write is assured to be saved because they are not loaded into RAM.
@@
 The application writes then the APDU Response bytes in Public, at specific position (see \autoref{fig:multosPubMem}).
-To declare variables in this data space, the \textit{\#pragma melpublic} directive is available.
+To declare variables in this data space, the \texttt{\#pragma melpublic} directive is available.
 
@@
 The limited size of RAM in IoT devices and smart cards makes the use of dynamic memory not advisable.
-The compiler directive to use Session Data is \textit{\#pragma melsession}.
+The compiler directive to use Session Data is \texttt{\#pragma melsession}.
 
@@
 Finally, we use Docker to deploy the compilation environment. Our container includes CMake and the LEDE SDK \citep{ledeproject}, configured for the Omega2 target, the device chose for the PoC.
-
-The Dockerfiles and CMake cross-compilation toolchain can be found in \autoref{ch:docker}.

### 2017-06-06 Doc/classicthesis/Chapters/Validation.tex
finishing touches

@@
 
-To test the correct execution of the \textit{IoT smart card}, we will use the ABC system from the tutorial in the P2ABCE Wiki \citep{p2abcurlwiki}.
+To test the correct execution of the \textit{IoT smart card}, we will use the ABC system from the tutorial in the P2ABCE Wiki\footnote{\url{https://github.com/p2abcengine/p2abcengine/wiki/}}.
 It is based on a soccer club, which wishes to issue VIP-tickets for a match.
@@
 This would be really interesting because for a single method from the \textit{Smartcard} interface, \textit{HardwareSmartcard} implementation needs to send multiple APDU Commands, but \textit{SoftwareSmartcard} can perform the operations immediately, with the full computer's resources.
+Also, physical smart cards benefit from hardware acceleration in most of the cryptographic operations, unlike our software implementations in the PoC.
 
@@
 Compared to the $64$MB of RAM available in the Omega2, our PoC could be executed in more constrained devices, given the system is compatible.
-
-%%%%%%%%%%%%%%%%%%%%%%%
-
-% Consideraciones futuras: implementar directamente Smartcard interface vs seguir usando APDUs:  minimizan uso de red, facilitan protocolo de comunicación de las APDUs, son estándar, mejoran el mantenimiento del código: P2ABCE puede cambiar las interfaces, pero mantendrá compatibilidad con las smartcards, si implementamos Smartcard.java por cada cambio en la interfaz deberemos actualizar cada IoT smart card.

### 2017-06-07 Doc/classicthesis/Chapters/Conclusions.tex
Analysis
state of the art: recopilation
design init
design arch
impl avance
finishing touches
First draft

@@
 To finish this document, we sum up some conclusions from the work done, and results 
-obtained. We will also enumerate some future lines of research that could start from
-the work done in this project.
+obtained. We will also enumerate some future lines of research.
 
@@
 
-The designed solution for the integration of Idemix and the Internet of Things provides new possibilities for the IoT security field, mainly in those scenarios where people's data is more vulnerable.
-However, there is a long path of research before we can see this in production.
-Many decisions depend on the specific deployment in course, and our solution tries to ease the best it all that future process.
+In the memory of this project we try to show the work done from the beginning, but we only showed our right decisions, and the information that is significant for the final result.
+The truth is that, aside from the information included in this paper, we have worked with other systems that ended up discarded.
+This isn't a negative aspect, because if we didn't, for example, study the Contiki OS, Cooja simulator and the hardware used, we could not be sure the development of the first PoC for that system would be impossible in the time given.
 
-With regards to the designed architecture, the flexibility of the Computation Offloading technique, identifying the key operations that can't be delegated, and those ones that can, allows us to define a general solution for the vast world of the Internet of Things.
-The IoT devices can operate as individual actors in the Idemix ecosystem.
-When in need of performing offloading, the delegation server also falls into the IoT class of devices, a great benefit for any real project, having as many options as possible, the solution delivered can be adapted to most of requirements.
-Nevertheless, being so open to any solution leaves us with a lot of work to do, researching what are the best options, comparing benefits and drawbacks.
+With regard to the work presented, the flexibility of the computation offloading technique, identifying the key operations that can be delegated, and those ones that can't, has allowed us to define a general solution for the vast world of the Internet of Things.
+The IoT devices can operate as individual actors in the P2ABCE ecosystem, and when in need of performing computation offloading, the delegation server can also be a device considered into the IoT class.
+
+% Resultados son válidos: feseability, problema de tiempo en sistemas de tiempo real
+% Se han conseguido objetivos planteados
+% Experiencia: qué era más tedioso, dificultades durante diseño y desarrollo
+% Primera aproximación de este tipo ~~: más bien somos el future work de...
+% Aplicación de procedimientos aprendidos/aplicados durante el TFG/carrera
+% ^ Novelty
+
+During the development, we had to investigate a lot of concepts related to IoT, smart cards, and even the insides of P2ABCE's code, to fix many existing bugs in the original project and minimize the amount of changes it had to undergo, in order to work with the IoT devices.
+In the implementation chapter we give guidelines to port MULTOS applications, considering the particularities we encountered, that's why we can't consider it an \textit{instruction manual to port MULTOS apps}, but an interesting reading on how to confront a similar project.
+For example, if we wanted the Idemix implementation from \citep{vullers2013efficient}, previous to P2ABCE, in a IoT device, we could apply almost the same steps, obtaining a core functionality of Idemix for constrained devices in C.
 
 Our PoC implementation demonstrates that this project is actually feasible, not by performing a simulation of an IoT device, like in \citep{vanet}.
-During its development, we had to investigate a lot of concepts related to IoT, smart cards, and even the insides of P2ABCE's code, to fix many existing bugs in the original project and minimize the amount of changes it had to undergo, in order to work with the IoT devices.
+However, the use of third party libraries and no hardware acceleration support, makes the PoC too slow for certain cases, e.g., a Real-Time system requires almost immediate operations, like a car warning that another one is approaching too fast.
+There is a long path of research before we can see this design in production, as many decisions depend on the specific deployment in consideration.
+
+Recalling the objectives listed in \autoref{objectives:section}, we think we accomplish them, except the implementation of a PoC in the most constrained device possible, where we used LEDE to ease our development, and the last objective, where we could have given a use to the PoC, and we leave as part of the future work.
 
@@
 
-% Mejoras PoC
-% Mejoras arquitectura: estudiar entornos típicos IoT y definir un despliegue realista usando esta cryptografía
+Due to the nature of the project, there exist many options to continue researching in this area. We list below a collection of the tasks we couldn't accomplish but would have been our next steps:
+
+\begin{itemize}
+	\item Second PoC for Arduino systems
+
+	We have already developed a working PoC for Linux based systems, and from our objective of using the most constrained devices available, we can follow an iterative process, building for devices a little more constrained than the last one.
+The next natural step, after building with LEDE/OpenWrt, is to aim for Arduino.
+
+	The critical changes to the actual PoC are: 
+	\subitem The transmission channel, if the device has no TCP/IP stack, it could use a Serial connection to transmit BIOSC and delegation messages.
+	\subitem The use of a Json file to store the smart card status depends on whether the device has enough space or a file system.
+The current Json parser also uses too much dynamic memory.
+At this stage the serialization should be performed in a binary format, in a file or writing variables directly to the EEPROM memory.
+	\subitem GMPLib and OpenSSL libraries must be replaced by others with similar capabilities, or special implementations for the particular device in use.
+
+	All these changes affect only the \textit{external utilities} and \textit{interfaces} sections of the code, leaving the \textit{core} untouched, because the smart card logic doesn't change.
+
+	\item Implement more P2ABCE functionality in the device
+
+	If instead of aiming for more constrained devices, we work with similar devices to the Omega2, the computational power of the device could assume all the functionality of P2ABCE, at least for some roles.
+In a M2M (machine to machine) environment, the devices could act as Provers and Verifiers in the same interaction, and the verifying logic could be implemented entirely in the device (generate a Presentation Policy and checking a prove in a Presentation Token is valid), reducing the need of a delegation server.
+
+	\item Improve the \textit{core smart card} code
+
+	From the simple refactoring of removing the MULTOS reimplemented API for direct calls to the \textit{interfaces}, to identify all the \texttt{memcpy} copies of multiple variables with one call, removing the compiler directive for not using memory padding in the defined structures, letting the compiler manage the memory automatically.
+
+	\item Support Extended APDUs
+
+	Currently, the APDU instruction handlers use the GET RESPONSE command to send large APDU Response Payloads, but we can reduce the number of APDUs exchanged supporting the Extended APDU format.
+Because the P2ABCE already sent some Extended APDUs, the APDU I/O interface in the PoC has support for Extended APDU Commands, but not for Extended Responses.
+
+	\item Integrate the IoT+Idemix solution in other system
+
+	Our seventh objective, that we couldn't get to design in depth.
+The original idea was to integrate in a FIWARE\footnote{FIWARE \url{https://www.fiware.org/}} environment the Fiware Keyrock and Idemix Issuer, so both, users and IoT devices' identities are managed in the IdM\footnote{Identity Management - KeyRock \url{https://catalogue.fiware.org/enablers/identity-management-keyrock}} system.
+The Issuer generates credentials based on the user’s attributes held in the Keyrock, and IoT devices can authenticate against Fi-ware services (using Idemix), or authenticate against other IoT devices (M2M).
+
+	\begin{figure}[bth]
+		\begin{center}
+			\includegraphics[width=0.8\linewidth]{gfx/fiware}
+		\end{center}
+		\caption{IoT+Idemix Fi-Ware integration.}
+		\label{fig:fiware}
+	\end{figure}
+
+\end{itemize}
+
+% Último objetivo: FIWARE

### 2017-06-07 Doc/classicthesis/Chapters/Introduction.tex
First draft

@@
 This includes not only the data shared with others, where one must trust they will keep it safe, but it's also the data collected about us and that we don't have direct control over it.
-The attack Sony suffered in 2011 to PSN \citep{PSN2011} is an example of the trust people had in Sony to store their billing information, and because the security of that information depended on both them and Sony, people found themselves compromised unable to do anything about it.
+The attack Sony suffered in 2011 to PSN\footnote{2011 PlayStation Network outage -
+\url{https://en.wikipedia.org/wiki/2011_PlayStation_Network_outage}} is an example of the trust people had in Sony to store their billing information, and because the security of that information depended on both them and Sony, people found themselves compromised unable to do anything about it.
 
@@
 The goal of this project is to integrate Idemix with the IoT.
-It will be done using the ABC4Trust's \ac{P2ABCE}, a framework that defines a common architecture, policy language and data artifacts for an attribute based ecosystem, cryptographically based on either IBM's Idemix or Microsoft's U-Prove \citep{p2abcurl}.
+It will be done using the ABC4Trust's \ac{P2ABCE}, a framework that defines a common architecture, policy language and data artifacts for an attribute based ecosystem, cryptographically based on either IBM's Idemix or Microsoft's U-Prove \footnote{P2ABCEngine \url{https://github.com/p2abcengine/p2abcengine}}.
 This gives us a standardized language to exchange Idemix's messages between IoT devices and any other P2ABCE actor.

### 2017-06-07 Doc/classicthesis/Chapters/Objectives.tex
First draft

@@
 
-\section{Project description}
+\section{Project description}\label{objectives:section}
 
@@
 
-In the P2ABCE repository \citep{p2abcurl} there is available the project's code, divided in two solutions: a complete P2ABCE implementation in Java and a MULTOS smart card implementation as PoC for the project.
+In the P2ABCE repository\footnote{\url{https://github.com/p2abcengine/p2abcengine}} there is available the project's code, divided in two solutions: a complete P2ABCE implementation in Java and a MULTOS smart card implementation as PoC for the project.
 
@@
 
-MULTOS is a multi-application smart card operative system, which provides a custom developing environment, with rich documentation \citep{MultosTechLib}.
+MULTOS is a multi-application smart card operative system, which provides a custom developing environment, with rich documentation\footnote{MULTOS technical library - \url{http://www.multos.com/developer_centre/technical_library/}}.
 MULTOS smart cards communicate like any other smart card following the standard, but internally offers a very specific architecture, affecting the way one must code applications for it.
@@
 \marginpar{\includegraphics[width=\linewidth]{gfx/Big-Endian}\\Big-Endian, Wikipedia}
-and when storing structures there is no padding between defined variables, unlike modern compilers that perform data structure alignment \citep{dataStructAlign} for performance.
+and when storing structures there is no padding between defined variables, unlike modern compilers that perform data structure alignment \footnote{\url{https://en.wikipedia.org/wiki/Data_structure_alignment}} for performance.
 
@@
 
-Nevertheless, the C-API documentation \citep{MultosTechLib} provides rich description for each function.
+Nevertheless, the C-API documentation\footnote{\url{http://www.multos.com/developer_centre/technical_library/}} provides rich description for each function.
 
@@
 
-For this reason, the current \ac{PoC} is developed on LEDE, Linux Embedded Development Environment \citep{ledeproject}, using the Onion Omega2 development board.
-Although the Omega2 uses LEDE, its microcontroller is also listed as compatible with ThreadX OS \citep{THREADX}, a Real-Time Operative System for embedded devices, therefore, the performance measured in our PoC can be relevant to real scenarios with similar hardware.
+For this reason, the current \ac{PoC} is developed on LEDE\footnote{Linux Embedded Development Environment - \url{https://lede-project.org/}} using the Onion Omega2 development board.
+Although the Omega2 uses LEDE, its microcontroller is also listed as compatible with ThreadX\footnote{THREADX Real-Time Operative System for embedded devices - \url{http://rtos.com/products/threadx/}}, therefore, the performance measured in our PoC can be relevant to real scenarios with similar hardware.
 

### 2017-06-07 Doc/classicthesis/Chapters/Validation.tex
First draft

@@
 	\begin{tabularx}{0.75\textwidth}{ll} \toprule
-		MCU & Mediatek MT688 \citep{MT7688} \\
+		MCU & Mediatek MT688\footnote{MediaTek MT7688 Datasheet - \url{https://goo.gl/kqD2gF}} \\
 		CPU & MIPS32 24KEc 580MHz \\

### 2017-06-17 Doc/classicthesis/Chapters/Chapter0A.tex
UML sequence diagram
First draft
segundo borrador

--- 2017-05-19 Doc/classicthesis/Chapters/Chapter0A.tex
@@
 %\markboth{\spacedlowsmallcaps{Appendix}}{\spacedlowsmallcaps{Appendix}}
+
 \chapter{Test: APDU Commands exchanged}\label{ch:resultsdiagrams}
 
-% TODO
 In this appendix we provide 3 sequence diagrams highlighting the APDU Commands exchanged during our testbed.
+
 The first one is the setup of the IoT smart card, storing the system parameters needed to work in the deployed P2ABCE system.
-The second one is the issuance of the credential, divided in the three REST calls needed in the delegation.
-The last one is the proving for the Presentation Policy from the Verifier, using two REST calls.
+The first Command is called \texttt{isAndroid} because the \texttt{HardwareSmartcard} class distinguishes Android phones acting as smart cards from MULTOS smart cards, in order to use Extended APDUs.
+Then configures the smart card, setting the PIN (``\texttt{1234}'' by default) and PUK.
+Finally, it copies the cryptographic system parameters to the smart card with multiple \texttt{SET} instructions.
+
+The second image shows the issuance of the credential, divided in the three REST calls needed during the delegation.
+If we recall Idemix's issuance protocol, the User performed, during his first step, an exponentiation with his secret key and a ZKP, which can be identified by the \texttt{COMMITMENT} and \texttt{ISSUANCE RESPONSE} commands.
+During the second step, the User only has to verify the Issuer's ZKP and store the credential if the signature is valid.
+Because the P2ABC Engine can perform Verification of ZKP without the need of secret keys stored in the smart card, the APDU Commands in this phase are only for storing the credential information inside the IoT smart card's BLOB\footnote{(\textbf{B}inary \textbf{L}arge \textbf{Ob}jects} database, with the \texttt{STORE BLOB} instruction.
+
+The last diagram represents the proving for the Presentation Policy received from a Verifier, using two REST calls.
+During the first REST call, the Service stores data for the proving in the smart card, but before starting the proving itself, the device must choose an identity for the proving.
+This is done with the Identity Service, that in the PoC chooses the first identity available.
+After this, the second REST call starts the proving, creating the \textit{commitment} and \textit{responses} that integrate any ZKP, and depends on the device's secret keys\footnote{We, again, refer the reader who wants to know more about ZKPs to \citep{tfgmates}.}.
 
@@
 	\begin{center}
-		\includegraphics[width=0.9\linewidth]{gfx/APDUsInitIoTSC}
+		\includegraphics[width=0.9\linewidth]{gfx/UML/APDUsInitIoTSC}
 	\end{center}
@@
 	\begin{center}
-		\includegraphics[width=0.78\linewidth]{gfx/IssuanceAPDUs}
+		\includegraphics[width=0.78\linewidth]{gfx/UML/IssuanceAPDUs}
 	\end{center}
@@
 	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/APDUsProving}
+		\includegraphics[width=\linewidth]{gfx/UML/APDUsProving}
 	\end{center}
@@
 
-\chapter{Guide on how to continue the project development}
+\chapter{Attached code structure}\label{ch:code}
+
+Given the size of the project, we identify four main directory structures to arrange the code.
+
+First, the Dockerfiles, scripts that automatize the creation of Docker images.
+These images are like virtual machine snapshots, although Docker is not a virtualization system, and we use the images to create containers, that are like new virtual machines started from the snapshots.
+We use two Dockerfiles, one for the P2ABCE compilation environment, with Java and Maven, and another one for building the Omega2 binaries, using the LEDE SDK.
+
+\dirtree{%
+	.1 Docker/. 
+	.2 P2ABCE/. 
+	.3 Dockerfile. 
+	.3 idemix-3.0.36-binaries/. 
+	.2 LEDE SDK/. 
+	.3 Dockerfile. 
+}
+
+\hfil
+
+The original P2ABCE project is of considerable size, but we list here the most interesting directories that we had to modify to make the project compatible with IoT smart cards.
+
+\hfil
+
+\dirtree{%
+	.1 P2ABCE/. 
+	.2 Code/. 
+	.3 core-abce/. 
+	.4 abce-components/. 
+	.5 src/main/java/eu/abc4trust/smartcard/. 
+	.6 IoTsmartcardio/.
+	.7 util/. 
+	.8 IoTsmartcardConnection.java.  
+	.7 IoTCard.java. 
+	.7 IoTCardChannel.java. 
+	.7 IoTCardTerminal.java. 
+	.6 HardwareSmartcard.java. 
+	.6 SoftwareSmartcard.java. 
+	.4 abce-services/. 
+	.5 src/main/java/eu/abc4trust/services/. 
+	.6 UserService.java. 
+	.6 VerificationService.java. 
+	.6 IssuanceService.java. 
+}
+
+\hfil
+
+The IoT smart card toolkit is the core of this project.
+Managed with CMake, we see the \texttt{CMakeLists.txt} files, including the \texttt{Toolchain-omega2-mipsel.cmake} file with the cross-compiler information.
+We divide the project between the \textit{source} C files, and the header C files.
+Inside these directories, the organization is as stated in the implementation chapter: \textit{core} or \textit{common}, \textit{util interfaces} and \textit{external utilities}.
+
+\hfil
+
+\dirtree{%
+	.1 p2abce\_iot\_toolkit/. 
+	.2 CMakeLists.txt. 
+	.2 Toolchain-omega2-mipsel.cmake. 
+	.2 util\_sc\_tests/. 
+	.2 util\_sc/. 
+	.3 CMakeLists.txt. 
+	.3 BIOSC.c. 
+	.3 p2abc\_iot\_toolkit\_src/. 
+	.4 smartcard\_common/. 
+	.5 APDU\_handler.c. 
+	.5 global\_vars.c. 
+	.5 m\_adapted\_API.c. 
+	.5 subroutines.c. 
+	.4 smartcard\_external\_utilities/. 
+	.5 cJSON.c. 
+	.4 smartcard\_utils\_interface/. 
+	.5 APDU\_IO\_util.c. 
+	.5 arithmetic\_util.c. 
+	.5 crypto\_util.c. 
+	.5 serialize\_util.c. 
+	.5 system\_funcs.c. 
+	.3 p2abc\_iot\_toolkit\_include/. 
+	.4 error\_codes.h. 
+	.4 macrologger.h. 
+	.4 smartcard\_common/. 
+	.5 APDU\_handler.h. 
+	.5 APDU\_types.h. 
+	.5 abc4T\_types.h. 
+	.5 defs\_consts.h. 
+	.5 defs\_errs.h. 
+	.5 defs\_ins.h. 
+	.5 defs\_types.h. 
+	.5 global\_vars.h. 
+	.5 m\_adapted\_API.h. 
+	.5 subroutines.h. 
+	.4 smartcard\_external\_utilities/. 
+	.5 base64.h. 
+	.5 cJSON.h. 
+	.4 smartcard\_utils\_interface/. 
+	.5 APDU\_IO\_util.h. 
+	.5 arithmetic\_util.h. 
+	.5 crypto\_util.h. 
+	.5 serialize\_util.h. 
+	.5 system\_funcs.h. 
+}
+
+\hfil
+
+Finally, the test scripts, including a Python file to test the BIOSC protocol, a directory with the testbed scripts (one file to orchestrate every device from one terminal), and a more didactic version with a script per device.
+
+\hfil
+
+\dirtree{%
+	.1 scripts/. 
+	.2 testBIOSC.py. 
+	.2 tests/\DTcomment{One test script}. 
+	.3 setupTest.sh. 
+	.3 testNoIoT.sh. 
+	.3 IoTtest.sh. 
+	.2 distributed-PoC/\DTcomment{Distributed test}. 
+	.3 Omega2/. 
+	.3 RaspberryPi3/. 
+	.3 ThirdPartyServer/. 
+}

### 2017-06-17 Doc/classicthesis/Chapters/Conclusions.tex
segundo borrador

@@
 Recalling the objectives listed in \autoref{objectives:section}, we think we accomplish them, except the implementation of a PoC in the most constrained device possible, where we used LEDE to ease our development, and the last objective, where we could have given a use to the PoC, and we leave as part of the future work.
+In spite of this negative auto-critic, we have to acknowledge that this is the first privacy-preserving ABC implementation for IoT, designed for extensibility, interoperability and maintenance.
+
+Finally, we are very thankful to IBM's grant for the \textit{Privacy Preserving Identity Management applied to IoT} project, which allowed us to get this far.
+Also, this paper will be presented for publication in a JCR magazine.
 
@@
 
+	\item Use cryptographic hardware
+
+	Atmel offers many solutions for secure memory and cryptographic operations chips, with serial interfaces for the IoT devices.
+We could benefit from this chips by storing the secret keys of the credentials inside the secure memory, and speed up the SHA256 and AES128 calculations, currently software implemented and slow compared to any hardware implementation.
+
 	\item Support Extended APDUs
@@
 
-	\item Integrate the IoT+Idemix solution in other system
+	\item Integrate the IoT+Idemix solution in FIWARE
 
-	Our seventh objective, that we couldn't get to design in depth.
+	Our sixth objective, that we couldn't get to design in depth.
 The original idea was to integrate in a FIWARE\footnote{FIWARE \url{https://www.fiware.org/}} environment the Fiware Keyrock and Idemix Issuer, so both, users and IoT devices' identities are managed in the IdM\footnote{Identity Management - KeyRock \url{https://catalogue.fiware.org/enablers/identity-management-keyrock}} system.
@@
 
+	\item Research P2ABCE policies for IoT applications
+
+	The next step, after integrating Idemix with an Identity Management system, should be to design more complex applications to IoT scenarios.
+We talked about a smart building in the introduction, but the straightforward proofs we can present depend on the signed values in the credentials.
+The Zero-Knowledge Proofs cryptography is already implemented, we only have to extend the protocols to benefit from them.
+
+	For example, a combination of authentication and arbitrary proof could be achieved by a combined proof of knowledge of a signed credential (current authentication system) and said arbitrary proof, although the values used in it aren't actually signed by the Issuer, the first proof gives credibility.
+
+	Another solution for the same problem, similar to a PKI scheme, the IoT device receives a signed credential including an attribute which is his own public key.
+The IoT device can issue new credentials with the parameters it needs for the arbitrary proofs.
+The Verifier then certifies the new credential is signed by the IoT device's public key.
+
 \end{itemize}
-
-% Último objetivo: FIWARE

### 2017-06-17 Doc/classicthesis/Chapters/Design.tex
First draft
segundo borrador

@@
 
-In this chapter we will describe the process for defining how a constrained IoT device may be integrated in a system like P2ABCE.
-We will also describe the PoC implementation carried out to test a realistic deployment.
+In this chapter we describe the design of our proposal to integrate IoT constrained devices as part of the privacy preserving system P2ABCE.
+The ultimate goal is to enable constrained IoT devices to play the Idemix User role, interacting autonomously in order to authenticate and demonstrate their credential attributes in a privacy-preserving fashion.
 
@@
 	\end{center}
-	\caption{IoT in P2ABCE Architecture.}
+	\caption{Proposed high level Architecture for integrating IoT devices in P2ABCE.}
 	\label{fig:P2ABCE-IoT}
@@
 
-	In \autoref{fig:P2ABCE-IoT} the IoT device is represented with two interfaces, physical or virtual.
+	In \autoref{fig:P2ABCE-IoT} shows our proposed architecture, in which the IoT device is represented with two interfaces, physical or virtual.
 One allows external communications to other machines, including other P2ABCE actors, that could be on the Internet, a corporate LAN, a M2M overlay network, etc.
@@
 
-	The IoT device starts an interaction with another actor, e.g. an Issuer or Verifier, receiving a P2ABCE XML file.
+	The IoT device, acting as a User, starts an interaction with another actor, e.g., against an Issuer to obtain a signed credential, or against a Verifier to demonstrate certain property of its attributes in a privacy-preserving way.
 
@@
 
-	The server may need to send APDU Commands to the IoT smart card to read the credential information or perform cryptographic operations involving private keys.
+	The server may need to send APDU Commands to the IoT smart card to read the credential information or perform cryptographic operations involving private keys, necessarily stored inside the IoT device.
+
+	These APDU Commands include a list of 70 different instructions, identified by the APDU INS byte.
+Some of them manage the smart card, like changing the PIN code, and many are P2ABCE specific, like storing the system cryptographic parameters.
+Some examples:
+
+	\begin{tabular}{|c|l|}
+		\hline
+		INS & Description \\
+		\hline
+		\texttt{0x52} & \texttt{SET\_CREDENTIAL} \\ & Sets a new credential inside the smart card.\\
+		\texttt{0x5A} & \texttt{GET\_CREDENTIAL\_PUBLIC\_KEY} \\ & Returns a stored credential's public key.\\
+		\texttt{0x60} & \texttt{GET\_PRESENTATION\_COMMITMENT} \\ & First cryptographic operation during Proving.\\
+		\texttt{0x62} & \texttt{GET\_PRESENTATION\_RESPONSE} \\ & Second cryptographic operation during Proving.\\
+		\hline		
+	\end{tabular}
+
+	\hfil
 
@@
 	\end{center}
-	\caption{IoT Delegation in P2ABCE for Proving.}
+	\caption{Designed interactions exchanged during the IoT delegation for P2ABCE Proving operation.}
 	\label{fig:DelegationProving}
@@
 
-In this PoC the only P2ABCE Service that needed to be modified was the User Service. We added the following REST call
-
+In this PoC, only the P2ABCE's User Service needed to be modified, because this is the role the IoT device plays. We added the following REST call:
 \begin{center}
@@
 \end{center}
-
-where we communicate the P2ABCE server that an IoT Smart Card is accessible via \textit{host} and \textit{port}.
+where we communicate the P2ABCE server that an IoT Smart Card is accessible via the IP address \textit{host} and TCP port \textit{port}.
 Then a new \textit{HardwareSmartcard} object is stored in the P2ABC Engine, but instead of the \textit{javax.smartcardio} Oracle's \textit{CardTerminal} implementation, we use our own \textit{IoTsmartcardio} implementation for the \textit{HardwareSmartcard} constructor, that we will discuss in the following section.
@@
 
-The implementation of this simple protocol is done in both Java and C, one for the P2ABCE server and the other one for the IoT smart card.
+The implementation of this simple protocol is done in Java for the P2abce delegation server and in C for the IoT smart card device.
 
@@
 %TODO : optional: diagrama secuencia de P2ABCE, pero si eso solo de la reimplementación de javax.smartcardio y su uso por HardwareSmartcard
-
-\textbf{TODO} :  diagrama secuencia de la reimplementación de javax.smartcardio y su uso por HardwareSmartcard
 
@@
 After this analysis, when those particularities appeared, our reimplementation of the MULTOS API accepts the \textit{expanded functionality}.
+In total, we reduced to eighteen the number of MULTOS functions to reimplement, and with future refactoring this number may decrease.
 
@@
 
-For example, Atmel's ATAES132A \citep{ATAES132A}
+For example, Atmel's ATAES132A\footnote{ATAES132A 32K AES Serial EEPROM  Datasheet - \url{http://ww1.microchip.com/downloads/en/DeviceDoc/Atmel-8914-CryptoAuth-ATAES132A-Datasheet.pdf}}
 \marginpar{\includegraphics[width=0.9\linewidth]{gfx/atmel} \\ Atmel's cryptography chips.}

### 2017-06-17 Doc/classicthesis/Chapters/Introduccion.tex
segundo borrador

%************************************************
\chapter*{Introducción}\label{ch:introduccion}
%************************************************

El término Internet de las Cosas, o \textit{Internet of Things} (IoT) en inglés, tiene un amplio rango de interpretaciones \citep{Atzori20102787}, pero podemos resumirlo en miles de millones de dispositivos, en su mayoría de capacidad computacional limitada, que están interconectados entre sí e Internet, a fin de realizar algún objetivo común. Por ejemplo, una red de farolas en una calle de la ciudad, donde cada farola tiene un sensor de proximidad, y se comunica con las siguientes farolas para iluminar antes la calle, mejorando la seguridad y ahorrando energía; invernaderos con riego automatizado en base a los sensores de humedad repartidos por la tierra, mejorando la calidad de la plantación, y ahorrando agua; coches comunicándose entre sí para dar aviso de accidentes a los que vienen detrás, o indicando rutas alternativas para minimizar los atascos y la contaminación dentro de la ciudad.

Muchos de estos objetivos requieren recopilar una gran cantidad de datos por medio de sensores, y gracias a organizaciones como \href{https://wikileaks.org/}{WikiLeaks}, los usuarios son cada vez más conscientes de la implicación que suponen sus datos en Internet, y demandan más seguridad y privacidad para ellos. Esto incluye tanto los datos que los usuarios comparten, sino también los datos recopilados sobre nosotros y sobre los que no tenemos un control directo. El ataque informático sufrido por Sony en 2011\footnote{2011 PlayStation Network outage - 
\url{https://en.wikipedia.org/wiki/2011_PlayStation_Network_outage}} afectó a su red de jugadores de consola, cuya información de facturación estaba guardada en sus servidores, y debido a que dicha información dependía tanto de los usuarios como de Sony para estar segura, aunque los usuarios la protegieran, se vieron comprometidos igualmente.

Con la proliferación de dispositivos IoT recopilando tanta información como son capaces con sus sensores, la cantidad de datos recopilados sobre cualquier persona puede ser enorme. Y el IoT ha demostrado no ser de confianza para la seguridad ni la privacidad, como en el reciente ataque DDoS de la \textit{botnet} Mirai, en octubre de 2016, considerado el ataque DDoS más grande de la historia\citep{jeyanthi:2017}, o las múltiples vulnerabilidades descubiertas en aparatos del hogar, como monitores de bebé\citep{rapid7babycam}.

Para solucionar este problema de privacidad en Internet, aparece el concepto de \textit{anonimidad fuerte}, que oculta nuestros datos personales, pero que nos permite poder seguir operando en Internet como individuos independientes\citep{stronganonymity}. Para conseguir esto, debemos realizar los procesos de autenticación y autorización de la manera más privada posible. Las credenciales basadas en atributos y la \textit{selección selectiva} de los mismos, permite controlar qué información revelamos a terceros, bajo un sistema criptográfico de confianza para ambas partes.

Una credencial basada en atributos puede entenderse como una firma digital, realizada por un Emisor, sobre una lista de pares atributo-valor, por ej., la lista \texttt{\justify{(nombre=Alice, apellido=Anderson, fechaNac=1977-05-10, nacionalidad=DE)}}\citep{introCredIBM}.

El método más directo, como Usuario, para convencer a un Verificador de que se poseen los valores firmados en el certificado, sería simplemente enviar la credencial en claro, mostrando todos los atributos. Con credenciales anónimas, el Usuario nunca transmite la credencial en sí, sino que usa una criptografía especial para convencer al Verificador de que sus atributos cumplen alguna propiedad, sin mostrarlos directamente, por ejemplo, que la diferencia entre la fecha actual y la fecha de nacimiento en el certificado es mayor a 18 años. Una ventaja directa es que el Verificador no puede reutilizar esa prueba con terceras partes, pues no posee el certificado, lo que evita el robo de información.

Con la criptografía simétrica o asimétrica usual utilizada en los certificados más populares, como los TLS/SSL para las webs de Internet, no es posible llevar a cabo esas \textit{pruebas} sin una explosión combinatoria de firmas de certificados de todas las posibles pruebas a presentar. Las soluciones para la privacidad actuales utilizan las llamadas Pruebas de Conocimiento Cero (ZKP), cuya teoría estudio más en profundidad en mi TFG de Matemáticas\citep{tfgmates}, y en resumen, permiten demostrar conocimiento de algún dato secreto, sin revelar nada más que el hecho de conocer dicho dato, de ahí la parte de \textit{conocimiento cero}.

Para explicar brevemente cómo las ZKP funcionan, en 1990 Guillou, Quisquater y Berson publicaron \textit{How to Explain Zero-Knowledge Protocols to Your Children} \citep{ZKPcave:story}, una historia sobre cómo Alí Babá demostró que sabía las palabras mágicas para abrir la cueva, pero sin decirle a nadie dichas palabras. A continuación presentamos una versión reducida, pero que subraya las propiedades básicas de toda ZKP.

To understand how ZKPs work, in 1990 Guillou, Quisquater and Berson published in \textit{How to Explain Zero-Knowledge Protocols to Your Children} \citep{ZKPcave:story} a story about how Ali Baba proved that he knew the magic words to open the cave, but without revealing those words to anyone. Here we present a brief version highlighting the properties of ZKPs.

\hfil

\begin{quote}
	Imaginemos una cueva, donde el camino principal se bifurca y al final de cada pasillo los caminos se vuelven a encontrar, formando una especie de anillo. En el punto en que se unen, dentro de la cueva, hay una puerta mágica con una palabra secreta, la cual permite abrirla y cruzar al otro lado.

	\textbf{P}aula conoce la clave secreta y quiere \textbf{p}robarlo a su amigo Víctor, pero sin tener que revelársela.
	\marginpar{\includegraphics[width=1.\linewidth]{gfx/graficoJL_ZKP_1}\\La cueva \citep{ZKPcave:fig}
		. Paula entra por A o B al azar. Víctor espera fuera.}
	Paula y Víctor quedan en la entrada de la cueva con unos \textit{walkie-talkies}, de modo que Víctor esperará fuera y Paula entrará a la cueva y tomará uno de los pasillos, que llamaremos A y B, sin decirle cuál a Víctor.

	%\begin{figure}[bth]
	%	\begin{center}
	%		\includegraphics[width=.45\linewidth]{gfx/graficoJL_ZKP_1}
	%	\end{center}
	%	\caption{La cueva \citep{ZKPcave:fig}. Paula entra por A o B al azar, Víctor espera fuera.}
	%	\label{fig:ZKPcave1}
	%\end{figure}

	Al llegar a la puerta, Paula avisa a Víctor para que entre a la cueva y espere en la bifurcación, donde \textbf{V}íctor, para intentar \textbf{v}erificar que Paula conoce la palabra secreta, le indicará por qué pasillo quiere que vuelva, el A o el B.

	%\begin{figure}[bth]
	%	\begin{center}
	%		\includegraphics[width=.45\linewidth]{gfx/graficoJL_ZKP_2}
	%	\end{center}
	%	\caption{La cueva. Víctor elige al azar por dónde quiere que regrese Paula.}
	%	\label{fig:ZKPcave2}
	%\end{figure}

	Si Paula realmente conoce el secreto, podrá volver a la bifurcación por el pasillo solicitado, abriendo, si es preciso, la puerta.\marginpar{\includegraphics[width=1.\linewidth]{gfx/graficoJL_ZKP_2}\\La cueva. Víctor elige al azar por dónde quiere que regrese Paula.}
	Pero en caso de no conocer la clave, al entrar por uno de los pasillos, tenía una probabilidad del $50\%$ de adivinar cuál pediría Víctor.

	%\begin{figure}[bth]
	%	\begin{center}
	%		\includegraphics[width=.45\linewidth]{gfx/graficoJL_ZKP_3}
	%	\end{center}
	%	\caption{La cueva. Paula vuelve por el camino pedido.}
	%	\label{fig:ZKPcave3}
	%\end{figure}

	Víctor no se queda contento con una sola prueba, pues Paula podría haber tenido suerte, así que la repiten hasta que se convence. Con unas 20 repeticiones, Paula tendría solo una probabilidad de $2^{-20}$, prácticamente nula, de acertar todas las veces y engañar así a Víctor.
	\marginpar{\includegraphics[width=1.\linewidth]{gfx/graficoJL_ZKP_3}\\La cueva. Paula vuelve por el camino pedido.}

	\textbf{E}va, curiosa de qué hacían Víctor y Paula en la cueva, \textbf{e}spía a Víctor durante todo el proceso. Eva no sabe si Paula y Víctor han acordado previamente qué pasillo pedir por el \textit{walkie-talkie}, y sólo Víctor está seguro de que él los estaba eligiendo al azar y sin previo acuerdo.%, por eso, Eva no puede estar segura de si Paula conoce la clave secreta, o bien estaban \textbf{s}imulando todo para engañarla por cotilla.

	Más tarde Eva habla con Víctor, que está seguro de que Paula conoce la clave, y éste querría convencer también a Eva, pero como él no conoce la clave, no puede repetir la prueba a Eva, sólo Paula puede realizarla con éxito.
\end{quote}

\hfil

Utilizando las ZKP, IBM ha desarrollado el protocolo Identity Mixer\footnote{Identity Mixer - \url{https://www.research.ibm.com/labs/zurich/idemix/}}, Idemix para acortar, para autenticación y transferencia de atributos certificados preservando la privacidad. Idemix permite a un usuario autenticarse sin divulgar ningún dato personal. Cada usuario recibe un certificado con múltiples atributos, pero a la hora de presentarlo, pueden decidir qué atributos revelar o de cuáles dar prueba de alguna propiedad, como la de la mayoría de edad, pertenencia a un país de la Unión Europea, etc. De este modo, los datos como su fecha de nacimiento, o país específico no son recopilados y no necesitan ser protegidos, ni procesados por terceros.

\begin{center}
	\textit{``Si tus datos personales nunca son recopilados, no pueden ser robados.''}
\end{center}

Hasta ahora, Idemix o credenciales privadas basadas en atributos han conseguido lidiar con los escenarios típicos de Internet, donde los usuarios pueden autenticarse frente a un proveedor de servicios, demostrando que poseen un credencial válido con permisos, pero sin dar más información. Veremos en el capítulo del estado de arte que las implementaciones actuales están realizadas en Java, lo cual requiere unos recursos mínimos para su ejecución, y, hasta donde llega nuestro conocimiento, este trabajo es la primera propuesta de implementar una solución para la privacidad en IoT, basada en sistemas de credenciales anónimas como Idemix.

\hfil

El objetivo de este proyecto es integrar Idemix con el Internet de las Cosas. Para ello nos apoyaremos en el proyecto europeo de ABC4Trust, \ac{P2ABCE}\footnote{P2ABCEngine \url{https://github.com/p2abcengine/p2abcengine}}, que define una arquitectura común, lenguaje y artefactos para unificar distintas soluciones como Idemix de IBM, o U-Prove de Microsoft. Con esto todo dispositivo IoT que lo implemente será compatible con el ecosistema existente.

Una vez los dispositivos pueden ejecutar Idemix, el siguiente paso es utilizar esta herramienta en otros proyectos. Por ejemplo, en un edificio inteligente, distintas políticas de privacidad podrían combinar la protección de los datos recogidos con la seguridad, como el número de personas en el edificio, al termostato sólo le debe interesar si hay más de un mínimo para encender más aparatos de aire acondicionado, pero en caso de incendio, el cuerpo de bomberos puede tener acceso a los datos ocultos por los sensores, para conocer el paradero de las personas atrapadas. Para llegar a soluciones de ese tipo, el primer paso será integrar Idemix en sistemas de gestión de identidad de IoT existentes, como el proyecto FIWARE.

\hfil

\subsection*{Distribución del trabajo}

Este documento se estructura como sigue: En el capítulo \ref{ch:stateoftheart} mostramos el estado del arte mediante la historia de Idemix y trabajos relacionados, analizando los aspectos más interesantes de cara al IoT; luego, en el capítulo \ref{ch:objectives} destacamos los objetivos del proyecto y analizamos en detalle las tecnologías existentes con las que trataremos; en el capítulo \ref{ch:design} describiremos el diseño formal de nuestra solución para IoT e Idemix, y describiremos la implementación de la prueba de concepto; tras esto, en el capítulo \ref{ch:validation} validaremos la implementación y realizaremos tests para comprobar su viabilidad; finalmente, describimos nuestras conclusiones y líneas de trabajo futuro en el capítulo \ref{ch:conclusions}.

### 2017-06-17 Doc/classicthesis/Chapters/Introduction.tex
segundo borrador

@@
 
-The \ac{IoT} is a term with a wide range of interpretations \citep{Atzori20102787}, breafly, we can think of it as a set of devices, mainly resource constrained, that are interconnected between them in order to achieve a goal.
+The \ac{IoT} is a term with a wide range of interpretations \citep{Atzori20102787}, breafly, we can think of it as billions of devices, mainly resource constrained, that are interconnected between them, and the Internet, in order to achieve a goal.
 For example, a network of lampposts with proximity sensors that talk to each other so they light up part of the street when a passerby walks by, but save energy when not; greenhouses with automated irrigation to balance costs and quality vegetables; or cars exchanging traffic data to reduce city pollution from traffic jams or avoid accidents.
@@
 With usual symmetric and asymmetric cryptography it seems impossible to create such credentials, without an explosion of signatures over every possible combination.
-Thats why current solutions rely on  \acp{ZKP}
-\marginpar{To read more about ZKPs, refer to my Mathematics thesis \citep{tfgmates}.}
+For this reason, current solutions rely on  \acp{ZKP}
+\marginpar{To read more about ZKPs, refer to my Mathematics thesis \citep{tfgmates}}
 , cryptographic methods that allow to proof knowledge of some data without disclosing it.
@@
 
+%%%
+So far, Idemix or privacy-ABCs have been successfully applied to deal with traditional Internet scenarios, in which users can authenticate and prove their attributes against Service provider.
+However, due to the reduced computational capabilities of certain IoT devices, it has not been yet considered for IoT scenarios.
+As we study in the state of the art chapter, current implementations are based on Java, which requires high computational and memory resources to be executed, and to the best of our knowledge, this is the first proposal that tries to apply an IoT solution for privacy-preserving authentication and authorization, based on Anonymous credential Systems, like Idemix.
+%%5
+
 The goal of this project is to integrate Idemix with the IoT.
-It will be done using the ABC4Trust's \ac{P2ABCE}, a framework that defines a common architecture, policy language and data artifacts for an attribute based ecosystem, cryptographically based on either IBM's Idemix or Microsoft's U-Prove \footnote{P2ABCEngine \url{https://github.com/p2abcengine/p2abcengine}}.
+It will be done using the ABC4Trust's \acf{P2ABCE}, a framework that defines a common architecture, policy language and data artifacts for an attribute based ecosystem, cryptographically based on either IBM's Idemix or Microsoft's U-Prove \footnote{P2ABCEngine \url{https://github.com/p2abcengine/p2abcengine}}.
 This gives us a standardized language to exchange Idemix's messages between IoT devices and any other P2ABCE actor.
+
+Once the IoT devices can execute Idemix, the new step is to take advantage of it in other deployments.
+In a smart building, different privacy policies could protect sensitive data, like how many people there are, and their identities, where a thermostat would only need to know if the amount of people is between some boundaries, but in case of emergency, the police department could request all the information available.
+We think that the first approach to achieve these ideas is first to integrate Idemix in known IoT identity systems, like FIWARE project.
 

### 2017-06-17 Doc/classicthesis/Chapters/Objectives.tex
segundo borrador

@@
 
+The main objective of this thesis is to integrate IBM’s privacy-preserving Attribute Based System in the IoT environment.
+The idea is to enable IoT constrained devices to act as a autonomous entities that can authenticate and prove their attributes against a verifier entity using the Idemix protocol, thereby, allowing a privacy-preserving IdM solution for IoT.
+
+This general objective, can be split in 6 main subobjectives:
+
 \begin{enumerate}
-	\item Idemix and the IoT
-
-	This is our main objective, integrating IBM's privacy-preserving system in the IoT environment.
-
 	\item Study of the Idemix ecosystem
@@
 
-	A formal description of the IoT architecture that will use Idemix.
+	A formal description of the IoT architecture that will that will allow to integrate Idemix in IoT scenarios.
 
@@
 
-	Be compatible with devices as constrained as possible, considering the multiple IoT platforms existing.
+	The implemented software needs to be optimized to be deployable in IoT devices as constrained as possible, and consider the multiple IoT platforms in existence.
 
@@
 
-	Measure and compare the solution's performance against the existing ones.
+	This evaluation will allow to validate the feasibility of our solution and its implementation.
 
@@
 Giving the vast range of IoT devices, a one-for-all solution must take in consideration multiple requirements and limitations.
-We will break down the original system, Idemix and P2ABCE, analyze every part of it and categorize them.
-We will consider what components are mandatory to be executed in the target device that wants to act as User of Verifier, and which components the devices would actually be able to execute.
+We will break down the original system, Idemix and P2ABCE, analyse every part of it and categorize them.
+We will consider what components are mandatory to be executed in the target device, and which components the devices would actually be able to execute.
+Given the mathematical base of the Idemix protocol and the complexity of the P2ABCE architecture and code, this may be the most challenging step, because it determines how the rest of the project will be developed.
 
-Devices with equivalent processing power to smart phones are capable of running the current Java implementations of P2ABCE with Idemix, but the most constrained IoT devices can not handle the entire system, only the mandatory cryptographic operations.
+Devices with equivalent processing power to smart phones are currently capable of running the Java implementations of P2ABCE with Idemix, but the most constrained IoT devices can not handle the entire system, only the mandatory cryptographic calculations, that is, the operations required to keep secured the privacy information in the device, their keys and certificates.
 
@@
 
-Zero-knowledge proofs used in Idemix allow a user to prove ownership of a credential without
+In our work, we aim at making IoT constrained devices capable to act as a User or Verifier in the P2ABCE architecture, given, for example, a M2M (Machine to Machine) setup, a device could identify itself to others (User), and ask for the respective cryptographic evidence to its peer (Verifier).
+
+The key operations of Idemix are the Issuance and Proving, but there are also others like the revocation of credentials, update of attribute values or inspection of a prove.
+We consider that the cryptographic details of Idemix are out of the scope of this thesis, and can be consulted in \citep{idemixSpec}, but we will give a brief introduction of the Issuance and Proving interactions.
+Also, for those who want a little more information about the mathematical operations performed, without the deep explanations from the Idemix specification, we recommend to read the section 3.2 from \citep{vullers2013efficient}.
+
+\hfil
+
+The \textbf{Issuance} protocol consists on an interactive Camenisch - Lysyanskaya signature, that is, unlike usual signatures, where the Issuer alone hashes the credential (implying she knows all the values in it) and ciphers the digest, in this signature, the signing is performed between both actors.
+
+This is possible given the nature of the CL signature, where the values of the parameters act as exponents of some public bases, and the use of ZKPs assure each party performs the correct calculations, avoiding malicious users.
+The User will perform the exponentiation with her secret key and some hidden attributes (the Issuer won't know their true value, only that they have some property, through a ZKP).
+The User then sends this new value and a ZKP that she indeed knows the exponents (the hidden values in the credential) for the public bases, without revealing said exponents.
+The Verifier performs the same exponentiation with the public attributes (not hidden by the User), and using her secret key, finishes the signature.
+The User then receives a valid CL signature for the credential, and no one except her knows the secret key and hidden values, thanks to the ZKPs.
+
+The \textbf{Proving} protocol is simpler than the Issuance.
+It is usually performed in a three step protocol.
+First, the Verifier generates a nonce for the current interaction.
+Then, the User performs a non-interactive Zero-Knowledge Proof of any predicate she wants to proof.
+This is performed doing various ZKPs in parallel, using the same nonce from the Verifier.
+After the non-interactive ZKP is performed, the User sends it to the Verifier, which needs less calculations to verify the predicate, and learns nothing more than that the predicate itself is true, that is the nature of Zero-Knowledge Proofs.
+
+Although we see both protocols are interactive, with at least 3 messages each, we mention \textit{non-interactive} ZKPs.
+If we recall the cave story, Victor needed to repeat the experiment many times before believing that Peggy knew the secret.
+Here, using the Fiat-Shamir heuristic and the Verifier nonce, the Verifier challenges (in the story, the path to return from) are substituted by a Hash function, e.g.
+SHA256, performed over some values.
+The User can't anticipate the digest result, so the Verifier trusts the Zero-Knowledge Proof, even when she doesn't choose the challenge.
+The Fiat-Shamir heuristic significantly increases the performance of these protocols, reducing the number of messages exchanged.
+
+\hfil
+
+As we have seen, Zero-Knowledge Proofs used in Idemix allow a user to prove ownership of a credential without
 revealing the credential itself. Since the verifier does not see the credential, verification
@@
 
+\subsection{P2ABCE Code Structure and REST Services}
+
 In the P2ABCE repository\footnote{\url{https://github.com/p2abcengine/p2abcengine}} there is available the project's code, divided in two solutions: a complete P2ABCE implementation in Java and a MULTOS smart card implementation as PoC for the project.
 
-\subsection{P2ABCE Code Structure and REST Services}
-
-The Java code is managed by a Maven project, structured using various known design patterns, but not of our interest.
+The Java code is managed by a Maven project, structured using various known design patterns, but this code is not our main target, as it won't be executed in the device.
 The part we are actually interested in are the \textit{REST Services} and their use of the \textit{Components} classes, where the smart card's logic and use are defined.
 
-P2ABCE project is based on the concept of smart cards, virtual or physical, to store the credentials.
+P2ABCE project is based on the concept of smart cards, virtual or physical, to store the Idemix credentials and keys.
 An interface is defined to communicate with these smart cards, and then different implementations allow to use either \textit{Software Smartcards} or \textit{Hardware Smartcards}.
@@
 The \textit{SoftwareSmartcard} class implements the interface in Java, suitable for applications using P2ABCE self-storing digital smart cards, like a virtual wallet.
+This implementation is useful, for instance, to be deployed in more capable IoT devices (that can run Java) such as smartphones or smartwatches.
 
@@
 
-The project also provides a set of REST services to control each role of the P2ABCE systeam (User, Issuer, Verifier, Inspector, etc.).
+The project also provides a set of REST services to control each role of the P2ABCE project (User, Issuer, Verifier, Inspector, etc.).
 The methods implemented include the creation of \textit{Software} smart cards within the User Service, and store them in a data base for future REST calls that may need them.
@@
 The services receive parameters like the length of the cryptographic keys, IDs, or XML files to parse.
-The REST API is not meant for a User Service to communicate with an Issuer Service, for example, but for the User actor to call the Engine through the User Service, to perform specific actions, and then the Service returns the response XML.
-Those XML files are the way to communicate two different actors.
+The REST API is not meant for a User Service to communicate with an Issuer Service or Verifier, but for the User actor to call the P2ABC Engine through the User Service, to perform specific actions (execute the User side of the Idemix issuance and proving protocols), and then this Service returns the Response XML.
+Those XML files are the way to communicate two different actors (e.g.
+User and Verifier).
 The transmission method to exchange the XML depends on the specific scenario.
+
+This Services as \textit{tools} for the actors are only present in the P2ABCE project, not part of the Idemix or U-Prove engines.
+They are useful in a desktop environment, where we have a User Service in the background, and many programs, browser plugins, etc., make use of the Engine through the REST interface, reducing the number of P2ABCE entities in execution.
 
@@
 	\end{center}
-	\caption{ABC4Trust Card Lite}
+	\caption{ABC4Trust Card Lite. \textit{Image from ABC4Trust EU Project.}}
 	\label{fig:ABC4TCardLite}
@@
 
-The code is available from the P2ABCE's repository and has some good and bad points to have in count:
+The code is available from the P2ABCE's repository and has some good and bad points to take into account:
 
-The best asset of this code is that it's written in C aiming to a very constrained device, with very limited memory and similar computational power to many other IoT devices.
+The best asset of this code is that it is written in C aiming to a very constrained device, with very limited memory and similar computational power to many other IoT devices.
 
@@
 Instead of function parameters, most variables are made global, reducing the space on the stack used when calling other functions.
-If parameters are needed, they are pointers to shared buffers with fixed sizes, avoiding the use of any dynamic memory allocation.
+If parameters are needed, there are pointers to shared buffers with fixed sizes, avoiding the use of any dynamic memory allocation.
 
-Although, among the many drawbacks, we could highlight the \textit{awful} coding style, the strong dependency on MULTOS framework and many bugs found.
+On the other side, among the many \textbf{drawbacks}, we could highlight the \textit{awful} coding style, the strong dependency on MULTOS framework and the fact that many bugs were found during our work.
+
+The MULTOS C subroutines are an interface to hardware implementations for the smart cards, and it implies the ported code must be rewritten to an interoperable C code, not dependant on MULTOS specific functions.
+
+The code is also affected by the MULTOS architecture, breaking many execution sequences, that is, the running application can stop at any point, not returning from a function.
+The MULTOS OS will read the public memory to generate a Response APDU and then clean the stack memory, but our standard C application must return from every function, freeing the used memory, minimizing possible errors from bad memory management.
+As we will see in the next chapter, the sequence diagram shows how any execution should perform, returning to the caller, not finishing the process when an error happens.
+
+Although we mentioned the use of pointers as a way to reduce memory usage, we all agree that pointers are \textit{double–edged swords}, and a code with such an intense use of them is prone to undetectable errors before runtime.
 
 The code is structured in two files, \textit{main.h} and \textit{main.c}, with around $550$ and $5200$ lines of code, respectively.
-
 The file \textit{main.h} is mostly a reimplementation in assembly MEL code of some MULTOS functionality already offered by their API.
 
-The \textit{main.c} consists on nearly $600$ lines of variables and data structures declarations; followed by the \textit{main()} function, a $2600$ lines long \textit{switch-case} expression, with practically no comments; and to conclude, the implementation of thirty functions called \textit{Subroutines} at the end of the file, around other $2000$ lines of code.
+The \textit{main.c} consists on nearly $600$ lines of variables and data structures declarations; followed by the \textit{main()} function, a $2600$ lines long \textit{switch-case} expression, that handles every APDU instruction, with practically no comments; and to conclude, the implementation of thirty functions called \textit{Subroutines} at the end of the file, with around other $2000$ lines of code.
 
@@
 \begin{itemize}
-	\item Implement in C the \textit{Smartcard} interface used by P2ABCE architecture, and use some communication protocol to remotelly call the methods from the machine running the P2ABC Engine.
+	\item Implement in C the \texttt{Smartcard} interface used by P2ABCE architecture, like the \texttt{SoftwareSmartcard} class, and use some communication protocol to remotely call the methods from the machine running the P2ABC Engine.
 	\item Present the IoT device as a hardware smart card, using the APDU protocol (already defined, standard and with minimal overload).
@@
 
-Even with the problematic to maintain or even understand the code of ABC4Trust Card Lite, once one studies MULTOS framework in deep and applies many refactoring techniques to the code, it becomes the best starting point for the IoT version, making us opt for the second option.
+Even with the problematic to maintain or even understand the code of ABC4Trust Card Lite, once one studies MULTOS framework in deep, applies many refactoring techniques to the code, consider the applied optimizations for constrained devices, and the APDU protocol compatibility to work with \texttt{HardwareSmartcard} as any other smart card, it becomes the best starting point for the IoT version, making us opt for the second option.
 
@@
 
-In this section we will introduce a brief description of the APDU standard protocol, and the MULTOS framework used by the ABC4Trust Card Lite.
+We have decided to extend this chapter in order to explain the basis of smart card technologies, instead of only referencing the corresponding documentation, involving the reader to make the effort of understanding a system that is more complex than what we actually need to know at this point.
+
+This section is a summary of the smart card's communication protocol, the MULTOS ecosystem, and, therefore, the inherited ABC4Trust Card Lite code, and how our IoT solution will work in the inside, acting as a smart card, a port from a (poorly coded) MULTOS application.
 
@@
 
-As we can see, MULTOS is a comfortable framework to develop smart card applications, and now we must adapt and implement it for our IoT devices, if we want to port ABC4Trust Card Lite's code.
+As it can be sees, MULTOS is a comfortable framework to develop smart card applications, and now we must adapt and implement it for our IoT devices, if we want to port ABC4Trust Card Lite's code.
 
@@
 The \textit{multos.h} interface provides access to these functions, that ultimately call their respective primitive instructions in assembly code.
-The primitive instructions are but a system call with an operation code, loading data in the needed registers.
+The primitive instructions are but a system call with an operation code, loading data in the needed registers, and hardware implemented functionality.
 Therefore,  no implementation for these tools is available, nor in C, nor in assembly code.
@@
 Nevertheless, the C-API documentation\footnote{\url{http://www.multos.com/developer_centre/technical_library/}} provides rich description for each function.
+
+Following this documentation, we will have to reimplement in C the same or extended functionality, trying to be as efficient as the hardware implementations in MULTOS smart cards.
+For example, the cryptographic operations are common in smart cards and therefore hardware supported, but a constrained IoT device does not usually include a cryptographic chip or specialized processor instructions.
 
@@
 
-The project development is divided between the IoT device code and the P2ABCE services. To ease the setup of new developing machines, we will use Docker to deploy containers ready to compile the code.
+The project development is divided between the IoT device code and the P2ABCE services.
+To ease the setup of new developing machines, we will use Docker\footnote{\url{https://www.docker.com/what-docker}} to deploy two containers, one ready to cross-compile our IoT project's C code, and the other container will compile the P2ABCE's Java code.
 

### 2017-06-17 Doc/classicthesis/Chapters/StateOfTheArt.tex
finishing touches
segundo borrador

@@
 In this project we will study Identity Mixer as an Attribute-Based Credentials (ABC) solution for privacy-preserving scenarios, but there exist other solutions competing to give the best performance and capabilities as possible.
-The two most known alternatives to Idemix are Microsoft's U-Prove and Persiano's ABC systems.
-
-%minimal disclosure
-% Persiano
-% Verheul's self-blindable: only binary attributes [14]deIdemixSC
-% U -Prove
-% Idemix
-% P2ABCE 
-% TODO: explicar P2ABCE
-%\ac{P2ABCE}
+The two most notable alternatives to Idemix are Microsoft's U-Prove and Persiano's ABC systems.
 
@@
 However, after the theoretical analysis, the paper only simulates a computer with similar performance as an OBU, without adapting the existing Java implementation of P2ABCE to a real VANET system.
-In our project, we can consider ourselves as part of their \textit{future work}, because our PoC will execute on hardware actually used in VANET systems, although the OS may not be the same.
+In our project, we can consider ourselves as part of their \textit{future work}, because our PoC will run on hardware actually used in VANET systems, and has been implemented in C instead of Java, which is more realistic and efficient for its deployment in an OBU.
 

### 2017-06-17 Doc/classicthesis/Chapters/Validation.tex
segundo borrador

@@
 	Onion Omega2
-} A device that falls inside the category of IoT, powerful enough to run a Linux environment, LEDE, where we can develop and debug the first PoC without troubling ourselves with problems not related to the project itself.
+} A device that falls inside the category of IoT, powerful enough to run a Linux embedded environment, such as LEDE, where we can develop and debug the first PoC without troubling ourselves with problems not related to the project itself.
 
@@
 
-\paragraph{Shell script} \hfil
+\paragraph{Shell scripts} \hfil
 
-To orchestrate all the services we use a simple script that performs the REST calls using  \textit{curl}.
-Here we perform the mentioned steps: setup of the P2ABCE system, issuance of the credential and a prove for the presentation policy.
+To orchestrate all the services we use shell scripts that execute the REST calls using  \texttt{curl}.
+Here we perform the mentioned steps: setup of the P2ABCE system, issuance of the credential and proving for the presentation policy.
 
 In the setup, the system parameters are generated, indicating key sizes of 1024 bits, the ones currently supported by the ABC4Trust and IoT smart cards.
-Then the system parameters and public keys from the services (issuer, inspector, revocation authority) are also exchanged.
+Then the system parameters and public keys from the services (issuer, inspector, revocation authority) are also exchanged and stored in each Service.
 
-In the issuance, a two step protocol is performed by the User.
-The Issuer and the User send each other the XML data with the cryptographic information, and finish the protocol with a credential issued inside a smart card, software (Java) or hardware (physical or IoT).
+During the issuance and proving, the User and the Issuer and Verifier exchange multiple XML files, with the cryptographic information of each step of the Idemix protocol.
 
-Finally, a Verifier sends the User a \textit{presentation policy}. The User generates a \textit{presentation token} for the Verifier and Inspector.
+We offer two versions for the scripts.
+The first one is a single shell script file, to be executed from a single terminal.
+The advantage of this version is that, because every XML file is stored in the machine running the script, we don't have to transfer these files by other means, e.g.
+\texttt{scp}.
+An excerpt of this script showcases how we receive an XML file from the User Service, e.g.
+\texttt{secondIssuanceMessage.xml}, and send it to the Issuer Service, referencing the file stored in the script's working directory:
 
-%%%%%
-There are two script versions, one where every REST call is done from the same machine, therefore avoiding the exchange of the XML files by other means (like \textit{scp}); the other one is ready to be executed on a real setup, two scripts, one for the Omega2, where the REST calls are the delegation on the Raspberry Pi 3, and other for the PC, running the Issuer and Verifier as third party entities.
-The XML files generated by each service can be sent with the same protocol the IoT device and the Verifier, for example, would communicate in the real world.
-This way it's clearer that the \textit{curl} commands are the delegation protocol for the IoT device, and then the IoT device can send the XML as it wishes, over the Internet, Bluetooth, NFC, etc.
-%%%%%
+\begin{lstlisting}[language=bash]
+[...]
+# First issuance protocol step - UI (first step for the user).
+echo "Second issuance protocol step (first step for the user)"
+curl -X POST --header 'Content-Type: text/xml' -d @uiIssuanceReturn.xml 'http://localhost:9200/user/issuanceProtocolStepUi/' > secondIssuanceMessage.xml
+
+# Second issuance protocol step (second step for the issuer).
+echo "Second issuance protocol step (second step for the issuer)"
+curl -X POST --header 'Content-Type: text/xml' -d @secondIssuanceMessage.xml 'http://localhost:9100/issuer/issuanceProtocolStep/' > thirdIssuanceMessageAndBoolean.xml
+[...]
+\end{lstlisting}
+
+The second script version is intended for a more realistic execution.
+In a real scenario, the User and the Issuer, for example, would call the REST methods of their respective Services, but not the other actor's Service, exchanging the XML files through other means.
+Instead of one script running in a single machine, orchestrating the Omega2, Raspberry Pi 3 and laptop, we offer three scripts, one per machine, with \textit{pauses} where one device must send an XML file to another machine.
+An excerpt from the Omega2's script shows the same step as above, but instead of calling the Issuer's REST Service, waits until we send the corresponding files.
+
+\begin{lstlisting}[language=bash]	
+# First issuance protocol step - UI (first step for the user).
+echo "Second issuance protocol step (first step for the user)"
+curl -X POST --header 'Content-Type: text/xml' -d @uiIssuanceReturn.xml "http://$Rpi3IP:9200/user/issuanceProtocolStepUi/" > secondIssuanceMessage.xml
+
+read -p "Send => secondIssuanceMessage.xml <= back to the Issuer. Then press any key to continue... " -n1 -s
+
+read -p "Receive => thirdIssuanceMessageAndBoolean.xml <= from the Issuer. Then press any key to continue... " -n1 -s
+\end{lstlisting}
+
+\hfil
+
+The first version is recommended for tests, and the second version only for instructive purposes.
 
@@
 The REST method to create a \textit{SoftwareSmartcard} is $/createSmartcard$, and to create a \textit{HardwareSmartcard}, using the \textit{IoTsmartcardio} implementation, we use $/initIoTsmartcard$.
+This operation is done only once per device, and includes commands from the creation of the PIN and PUK of the smart card, to storing the system parameters of P2ABCE, equivalent to the previous setup step.
 
@@
 
-Again, this operation is done only once per device, and includes commands from the creation of the PIN and PUK of the smart card, to storing the system parameters of P2ABCE, equivalent to the previous setup step.
-
 This is the first interaction between the RPi3 and the IoT smart card running in the Omega2.
@@
 
-\paragraph{Issuing of the credential}\hfil
+\paragraph{Issuance of the credential}\hfil
+
+Our credential will have 5 attributes and key sizes of 1024 bits, as specified during the setup process.
 

### 2017-06-17 Doc/classicthesis/FrontBackmatter/Abstract.tex
segundo borrador

@@
 \chapter*{Abstract}
-Short summary of the contents in English\dots a great guide by 
-Kent Beck how to write good abstracts can be found here:  
-\begin{center}
-\url{https://plg.uwaterloo.ca/~migod/research/beckOOPSLA.html}
-\end{center}
+
+As part of IBM's academic grant for \textit{Privacy Preserving Identity Management applied to IoT}, we designed a solution to integrate IBM's Identity Mixer into the IoT ecosystem, in a privacy-preserving fashion thanks to the use of Zero-Knowledge Proof cryptography.
+To validate our design, we implemented a functional Proof of Concept written in C for IoT devices.
+The applications of such solution promise to improve both the security and privacy of IoT, which in recent years have proved to be compromised.
 

### 2017-06-19 Doc/classicthesis/Chapters/Chapter0A.tex
ieee

@@
 During the second step, the User only has to verify the Issuer's ZKP and store the credential if the signature is valid.
-Because the P2ABC Engine can perform Verification of ZKP without the need of secret keys stored in the smart card, the APDU Commands in this phase are only for storing the credential information inside the IoT smart card's BLOB\footnote{(\textbf{B}inary \textbf{L}arge \textbf{Ob}jects} database, with the \texttt{STORE BLOB} instruction.
+Because the P2ABC Engine can perform Verification of ZKP without the need of secret keys stored in the smart card, the APDU Commands in this phase are only for storing the credential information inside the IoT smart card's BLOB\footnote{\textbf{B}inary \textbf{L}arge \textbf{Ob}jects} database, with the \texttt{STORE BLOB} instruction.
 
@@
 
-\chapter{Attached code structure}\label{ch:code}
+\chapter{Paper proposal for JCR Magazine}\label{ch:JCR}
+
+a
+
+\chapter{Project's code structure}\label{ch:code}
 

### 2017-06-19 Doc/classicthesis/Chapters/Design.tex
ieee

@@
 After returning from \texttt{mExit}, the instruction handler and the APDU handler functions, the program listens again from the socket, until an $0xff$ byte is received.
+
+\section{Design and implementation conclusions}
+
+During this design and implementation process we had to study multiple technologies, including the ones mentioned in the state of the art, regarding Idemix, and those related to the IoT, we studied many systems (e.g.
+Contiki, Arduino) and hardware (e.g.
+MSP430, ESP8266), and after this, we had to take many decisions during development.
+
+The design is open in those aspects that depend on each specific project's requisites, and closed on the aspects involving the P2ABCE's core logic.
+
+However, the PoC had to make more concessions when implementing the cryptographic and modular arithmetic logic.
+But those are parts of the code that are more prone to be changed when developing for a new system.
+The core logic of the PoC is preserved as it is in every system and device.
+
+Even with the many points of future work left, we think we achieved the most extensible solution for Idemix+IoT at the time, our main objective of the project.

### 2017-06-19 Doc/classicthesis/Chapters/Introduction.tex
ieee

@@
 
-The goal of this project is to integrate Idemix with the IoT.
-It will be done using the ABC4Trust's \acf{P2ABCE}, a framework that defines a common architecture, policy language and data artifacts for an attribute based ecosystem, cryptographically based on either IBM's Idemix or Microsoft's U-Prove \footnote{P2ABCEngine \url{https://github.com/p2abcengine/p2abcengine}}.
+As part of IBM's academic grant for \textit{Privacy Preserving Identity Management applied to IoT}, the goal of this project is to integrate Idemix with the IoT.
+It will be done using ABC4Trust's \acf{P2ABCE}, a framework that defines a common architecture, policy language and data artifacts for an attribute based ecosystem, cryptographically based on either IBM's Idemix or Microsoft's U-Prove \footnote{P2ABCEngine \url{https://github.com/p2abcengine/p2abcengine}}.
 This gives us a standardized language to exchange Idemix's messages between IoT devices and any other P2ABCE actor.

### 2017-06-19 Doc/classicthesis/Chapters/Validation.tex
ieee

@@
 Compared to the $64$MB of RAM available in the Omega2, our PoC could be executed in more constrained devices, given the system is compatible.
+
+\section{Validation conclusions}
+
+Below \autoref{totaltime} sums up the time in seconds used in each step of the test for the Omega2 and Raspberry Pi 3 setup.
+
+The first step, \textit{System Setup} is done only once when the system is being deployed, and the \textit{IoT Smart Card Setup} only once per device.
+
+Because a device can have more than one credential, the Issuance step is significant when we issue multiple credentials over the device's lifetime.
+We recall that our tests used a credential with five attributes and key sizes of 1024 bits.
+
+Finally, the Proving step is expected to be the most performed calculation by the IoT device.
+The fact that it lasts over half a minute implies that we should not use this PoC for \textit{real-time} applications yet, usually used in critical secure systems.
+Nevertheless, for many other IoT applications, the fact this operation can be performed multiple times per hour, presents an useful tool for privacy, e.g.
+the thermostat system mentioned in the smart building, which can pool data from the sensors every few minutes.
+
+\begin{table}[!ht]
+	\begin{center}
+		\begin{tabular}{|r|r|r|r|}
+		\hline
+		System & IoT Smart  & Issue  & Prove Presen-\\
+		Setup & Card Setup & credential & tation Policy\\ \hline
+		3.77 s & 19.19 s & 36.48 s & 32.84 s \\ \hline
+	\end{tabular}
+	\end{center}
+\caption{Total time spent for each step in the Omega2+RPi3 setup.}
+\label{totaltime}
+\end{table}

### 2017-06-19 Doc/classicthesis/FrontBackmatter/Abstract.tex
ieee

@@
 
-As part of IBM's academic grant for \textit{Privacy Preserving Identity Management applied to IoT}, we designed a solution to integrate IBM's Identity Mixer into the IoT ecosystem, in a privacy-preserving fashion thanks to the use of Zero-Knowledge Proof cryptography.
+In this paper we design a solution to integrate IBM's Identity Mixer into the IoT ecosystem, integrating its privacy-preserving capabilities thanks to the use of Zero-Knowledge Proof cryptography.
 To validate our design, we implemented a functional Proof of Concept written in C for IoT devices.

### 2017-06-19 Doc/ieee/bare_jrnl.tex
ieee

%% bare_jrnl.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%\addbibresource{Bibliography.bib}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Integration of Idemix in IoT environments}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Jose Luis Canovas Sanchez, Antonio Skarmeta Gomez, Jorge Benal Bernabe}% <-this % stops a space
%\thanks{M. Shell was with the Department
%of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
%GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
%\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
%{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
In this paper we design a solution to integrate IBM's Identity Mixer into the IoT ecosystem, integrating its privacy-preserving capabilities thanks to the use of Zero-Knowledge Proof cryptography. To validate our design, we implemented a functional Proof of Concept written in C for IoT devices. The applications of such solution promise to improve both the security and privacy of IoT, which in recent years have proved to be compromised.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
%\begin{IEEEkeywords}
%IEEE, IEEEtran, journal, \LaTeX, paper, template.
%\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}

The IoT is a term with a wide range of interpretations \cite{Atzori20102787}, breafly, we can think of it as billions of devices, mainly resource constrained, that are interconnected between them, and the Internet, in order to achieve a goal. For example, a network of lampposts with proximity sensors that talk to each other so they light up part of the street when a passerby walks by, but save energy when not; greenhouses with automated irrigation to balance costs and quality vegetables; or cars exchanging traffic data to reduce city pollution from traffic jams or avoid accidents.

Many of this objectives require the use of a great amount of data, and thanks to organizations like WikiLeaks, people are aware of the implications of their data on the Internet, demanding more security and privacy for it. This includes not only the data shared with others, where one must trust they will keep it safe, but it's also the data collected about us and that we don't have direct control over it. The attack Sony suffered in 2011 to PSN\footnote{2011 PlayStation Network outage - 
	\url{https://en.wikipedia.org/wiki/2011_PlayStation_Network_outage}} is an example of the trust people had in Sony to store their billing information, and because the security of that information depended on both them and Sony, people found themselves compromised unable to do anything about it.

With the proliferation of IoT devices gathering as much information as they can with their sensors, the amount of data gathered about anyone can be immense. And IoT has proved to not address neither security nor privacy, with recent events like the Mirai botnet DDoS attack on October 2016, considered the biggest DDoS in history \cite{jeyanthi:2017}, or the multiple vulnerabilities affecting house devices, like baby monitors \cite{rapid7babycam}.

To address this problem of privacy in the Internet, a recent approach is the concept of \textit{strong anonymity}, that conceals our personal details while letting us continue to operate online as a clearly defined individuals \cite{stronganonymity}. To achieve it, we must address a way to perform authentication and authorization in the most privacy-friendly approach. Attribute-based credentials and \textit{selective disclosure} allow to control what information we reveal, under a trusted environment.

Intuitively, an attribute-based credential can be thought of as a digital signature by the Issuer on a list of attribute-value pairs, e.g. the list \texttt{(fname=Alice, lname=Anderson, bdate=1977-05-10, nation=DE)} \cite{introCredIBM}.
The most straightforward way for the User to convince a Verifier of her list of attributes would be to simply transmit her credential to the Verifier.
With anonymous credentials, the User never transmits the credential itself, but rather uses it to convince the Verifier that her attributes satisfy certain properties – without leaking anything about the credential other than the shown properties. This has the obvious advantage that the Verifier can no longer reuse the credential to impersonate \texttt{Alice}. Another advantage is that anonymous credentials allow the User to reveal a selected subset of her attributes.
Stronger even, apart from showing the exact value of an attribute, the User can even convince the
Verifier that some complex predicate over the attributes holds, e.g. that her birth date was more than 18 years ago, without revealing the real date.

With usual symmetric and asymmetric cryptography it seems impossible to create such credentials, without an explosion of signatures over every possible combination. For this reason, current solutions rely on  ZKPs, cryptographic methods that allow to proof knowledge of some data without disclosing it.

To understand how ZKPs work, in 1990 Guillou, Quisquater and Berson published in \textit{How to Explain Zero-Knowledge Protocols to Your Children} \cite{ZKPcave:story} a story about how Ali Baba proved that he knew the magic words to open the cave, but without revealing those words to anyone.

Based on ZKP properties, IBM has developed the Identity Mixer\footnote{Identity Mixer - \url{https://www.research.ibm.com/labs/zurich/idemix/}}, Idemix for short, protocol suite for privacy-preserving authentication and transfer of certified attributes. It allows user authentication without divulging any personal data. Users have a personal certificate with multiple attributes, but they can choose how many to disclose, or only give a proof of them, like being older than 18 years-old, living in a country without revealing the city, etc. Thus, no personal data is collected that needs to be protected, managed, and treated.

\begin{center}
	\textit{``If your personal data is never collected, it cannot be stolen.''}
\end{center}

%%%
So far, Idemix or privacy-ABCs have been successfully applied to deal with traditional Internet scenarios, in which users can authenticate and prove their attributes against Service provider. However, due to the reduced computational capabilities of certain IoT devices, it has not been yet considered for IoT scenarios. As we study in the state of the art chapter, current implementations are based on Java, which requires high computational and memory resources to be executed, and to the best of our knowledge, this is the first proposal that tries to apply an IoT solution for privacy-preserving authentication and authorization, based on Anonymous credential Systems, like Idemix.
%%5

As part of IBM's academic grant for \textit{Privacy Preserving Identity Management applied to IoT}, the goal of this project is to integrate Idemix with the IoT. It will be done using ABC4Trust's P2ABCE, a framework that defines a common architecture, policy language and data artifacts for an attribute based ecosystem, cryptographically based on either IBM's Idemix or Microsoft's U-Prove \footnote{P2ABCEngine \url{https://github.com/p2abcengine/p2abcengine}}. This gives us a standardized language to exchange Idemix's messages between IoT devices and any other P2ABCE actor.

Once the IoT devices can execute Idemix, the new step is to take advantage of it in other deployments. In a smart building, different privacy policies could protect sensitive data, like how many people there are, and their identities, where a thermostat would only need to know if the amount of people is between some boundaries, but in case of emergency, the police department could request all the information available. We think that the first approach to achieve these ideas is first to integrate Idemix in known IoT identity systems, like FIWARE project.

\hfil

\subsection{Outline of this thesis}

%This document is structured as follows: In \autoref{ch:stateoftheart} we show a state of the art analysis through the history of Idemix and related works, analysing what is of the most interest for the IoT perspective; then, in \autoref{ch:objectives} we outline the project's objectives and analyse in depth the existing solutions that we are going to deal with; in \autoref{ch:design} we describe the formal design of the IoT and Idemix solution, and describe the PoC implementation developed; after an implementation, it is a must to validate it, as showed during the performance tests in \autoref{ch:validation}; finally, our conclusions and lines for future work are described in \autoref{ch:conclusions}.

\section{State of the art}

In this project we will study Identity Mixer as an Attribute-Based Credentials (ABC) solution for privacy-preserving scenarios, but there exist other solutions competing to give the best performance and capabilities as possible. The two most notable alternatives to Idemix are Microsoft's U-Prove and Persiano's ABC systems.

%%%%%%%%%%%%%%%%%%%%%%%%

Persiano and Visconti presented a non-transferable anonymous
credential system that is multi-show and for which it is possible to prove
properties (encoded by a linear Boolean formula) of the credentials \cite{book:185217}. Unfortunately, their proof system is not efficient since the step in which a user proves
possession of credentials (that needs a number of modular exponentiations that
is linear in the number of credentials) must be repeated times (where is the
security parameter) in order to obtain a satisfying soundness.

%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%

Stefan Brands provided the first integral description of the U-Prove
technology in his thesis \cite{uprove} in 2000, after which he founded the company Credentica
in 2002 to implement and sell this technology. Microsoft acquired Credentica
in 2008 and published the U-Prove protocol specification \cite{uprove2} in 2010
under the Open Specification Promise4 together with open source reference software
development kits (SDKs) in C\# and Java.
The U-Prove technology is centered around a so-called U-Prove token. This
token serves as a pseudonym for the prover. It contains a number of attributes
which can be selectively disclosed to a verifier. Hence the prover decides which
attributes to show and which to withhold. Finally there is the token’s public-key, which aggregates all information in the token, and a signature from the issuer
over this public-key to ensure the authenticity \cite{book:947508}.

%%%%%%%%%%%

\hfil

Jan Camenisch, Markus Stadler and Anna Lysyanskaya studied in \cite{Camenisch:GroupSig}, \cite{Camenisch:AnonCred} and \cite{camenisch2002signature} the cryptographic bases for signature schemes and anonymous credentials, that later became IBM's Identity Mixer protocol specification \cite{idemixSpec}.

Luuk Danes in 2007 studied theoretically how Idemix's User role could be implemented using  
smart cards \cite{luuk}, identifying what data and operations should be kept inside the device to perform different levels of security. The User role was divided between the smart card, holding secret keys, and the Idemix terminal, that commanded operations inside the smart card, or read the keys in it to perform the instructions itself. The studied sets were:
\begin{itemize}
	\item The smart card gives all information to the terminal.
	\item The smart card only keeps the master key secret.
	\item The smart card only gives the pseudonym with the verifier to the terminal.
	\item The smart card keeps everything secret.
\end{itemize}
%%%%

Later, in 2008 Víctor Sucasas also studied an anonymous credential system with smart card support \cite{sucasas}, equivalent to a basic version of Idemix, using a simulator to test the PoC and pointing out some crucial implementation details for performance. The researching tendency starts to show that smart cards are the best solution to hold safely the User's credentials.

In 2009, some Java smart card PoC for Idemix were developed in \cite{javaIdemix1} and \cite{javaIdemix2}, but they weren't optimal and didn't include some Idemix's functionalities, like selective disclosure.

Later, in 2013, Vullers and Alpar, implemented an efficient smart card for Idemix \cite{vullers2013efficient}, aiming to integrate it in the IRMA\footnote{The IRMA project has been recently included in the Privacy by Design Foundation: \url{https://privacybydesign.foundation/}} project, and comparing the performance with U-Prove's smart cards. This new implementation was written in C, under the MULTOS platform for smart cards, and describes many decisions made during the development to improve the performance on such constrained devices. The terminal application was written in Java and used an extension of the Idemix cryptographic library to take care of the smart card specifics.

Extending the concept of smart cards, physical or logical, as holders of the credentials, the ABC4Trust's project, P2ABCE\footnote{\url{https://github.com/p2abcengine/p2abcengine}}, was created as a unified ABC system for different cryptographic engines, currently supporting U-Prove and Idemix. The Idemix library was updated to support P2ABCE and the last version is interoperable with U-Prove. Therefore, the smart card specification from the P2ABCE project could be considered the official version to work with.

% Uso de P2ABCE en VANETs
Related to the IoT, the P2ABCE project has been used to test in a VANET\footnote{Vehicular Ad-Hoc Network} scenario how an OBU (On Board Unit), with constrained hardware, could act as a User in an P2ABC system \cite{vanet}. However, after the theoretical analysis, the paper only simulates a computer with similar performance as an OBU, without adapting the existing Java implementation of P2ABCE to a real VANET system. In our project, we can consider ourselves as part of their \textit{future work}, because our PoC will run on hardware actually used in VANET systems, and has been implemented in C instead of Java, which is more realistic and efficient for its deployment in an OBU.

%%%%%%%%%%%%%%

\section{Design and implementation}

We now proceed describe the design of our proposal to integrate IoT constrained devices as part of the privacy preserving system P2ABCE. The ultimate goal is to enable constrained IoT devices to play the Idemix User role, interacting autonomously in order to authenticate and demonstrate their credential attributes in a privacy-preserving fashion.

\subsection{Design}

In this section we will define how an IoT device may be integrated in the P2ABCE architecture, being totally compatible with any other system using P2ABCE, addressing the power and memory constrains many IoT devices face.

We decided to use P2ABCE with the Idemix as its Engine, because it is officially supported by the Idemix Library, it has the most up-to-date implementation, as we saw in the state of the art, and adds capabilities to Idemix like the Presentation Policies, or interoperability with U-Prove, not available without the P2ABCE project.

Our main goal is to make an IoT device capable to act as a User or Verifier in the P2ABCE architecture. For this, the device should be able to \textbf{communicate} with the Verifier or Prover with which it is interacting, manage the P2ABCE complex \textbf{XML schemas} transmitted, and perform the \textbf{cryptographic operations} required.

The communication between actors depends on each IoT scenario, it can be achieved with many existing standard solutions, e.g. an IP network, a Bluetooth M2M connection, RF communication, etc.

Our real concerns are, on one side, parsing the XML data, based on the P2ABCE's XML schema, that specifies the data artifacts created and exchanged during the issuance, presentation, revocation and inspection of pABCs; and on the other side, the cryptographic operations, that involve the use of secret keys, stored privately in the IoT device.

After the analysis done in the previous sections to the P2ABCE architecture, emphasizing that the logic of smart cards gathers the cryptographic operations independently from how the data is exchanged between P2ABCE actors.

Using the \textit{computation offloading} technique to our scenario, our design consists on implementing the smart card logic inside the IoT device, keeping secure our master key and credentials, and for the rest of the P2ABCE system, if the device can not run the complete Engine, it may delegate to a server running it, indicating how to send APDU Commands to the \textit{IoT smart card}.

Even in the case we were to implement all P2ABCE inside an IoT device, we would have to implement the support for software smart cards, to keep the secret inside the IoT device. Therefore, we can begin implementing the smart card logic inside the IoT device, and later, if the device resources admit it, other components of the P2ABCE project.

\hfil

Computation offloading is not new to IoT deployments. For example, IPv6 involves managing 128 bits per address and other headers, and many IoT scenes only need to communicate inside a private network, making only the last 64 bits in an address relevant. To reduce that overhead, instead of IPv6 they use 6LoWPAN to compress packets and use smaller address sizes. To communicate a 6LoWPAN with the Internet or other networks, the IoT devices delegate the networking workload on a proxy that can manage the 6LoWPAN and IPv6 stacks. In the scope of consumer devices, smart bands or watches can install applications, bit many of them delegate on the user's phone to accomplish their task.

Therefore, the IoT device now has a \textbf{duality} in its functions, because it is the User that starts any interaction with other actors, and it's also the smart card that a P2ABCE server must ask for cryptographic operations. It can also be seen as a \textbf{double delegation}. The IoT device delegates on the external P2ABCE server to manage the protocol, and the P2ABCE server delegates on the IoT, acting now as a smart card, for the cryptography.

\subsubsection{System architecture}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/P2ABCE-IoT-bw}
	\end{center}
	\caption{Proposed high level Architecture for integrating IoT devices in P2ABCE.}
	\label{fig:P2ABCE-IoT}
\end{figure}

%\begin{figure}[bth]
%	\begin{center}
%		\includegraphics[width=\linewidth]{gfx/P2ABCE-IoT-color}
%	\end{center}
%	\caption{IoT in P2ABCE deployment diagram.}
%	\label{fig:P2ABCE-IoT-color}
%\end{figure}

\hfil

The system will be compounded by the IoT device, the P2ABCE delegation server and the third party P2ABCE actors.

\begin{itemize}

	\item \textbf{IoT device}

	In \autoref{fig:P2ABCE-IoT} shows our proposed architecture, in which the IoT device is represented with two interfaces, physical or virtual. One allows external communications to other machines, including other P2ABCE actors, that could be on the Internet, a corporate LAN, a M2M overlay network, etc. Through this interface, the P2ABCE XML messages are exchanged as in any other deployment. This allows an IoT device to interact with other actors without special adaptations to the protocol. The other interface allows a secure communication with the delegation server. Both the delegation messages and the APDU Dialogue are transmitted over this interface, making it a point of attack to the system, and we will talk about its security in the delegation process.

	The scheme also shows the \textit{P2ABCE IoT Toolkit}. This piece of software includes the IoT Smart Card, and the API for other processes that want to use the P2ABCE system.

	The IoT Smart Card is the implementation of a software smart card, listens for APDU Commands from the secure interface and stores securely the credentials and private keys within the device's memory.

	The P2ABCE API is an interface for other processes that wish to use the private-preserving environment of P2ABCE. It provides access to every operation available, hiding the delegation process. In the future, if for example the Verification Service is implemented for the IoT device, i.e., there's no need to delegate to other machine to act as a Verifier, then any program using the API won't need to change anything, the toolkit conceals the transition from delegating to native execution.

	\item \textbf{P2ABCE actors}

	If we recall from \autoref{analysisP2ABCE}, the possible roles in the system were the Issuer, the User, the Verifier, the Revocation Authority and the Inspector. All of them use the P2ABCE XML schema in the specification to communicate to each other. Any third party actor will be unaware of the fact that the device is a constrained IoT device that delegates on the P2ABCE server.

	\item \textbf{P2ABCE Delegation Server}

	The machine in charge of receiving authorized IoT devices' commands to parse the XML files exchanged and orchestrate the cryptographic operations the IoT smart card must perform.

\end{itemize}

\hfil

\begin{flushleft}
	\textbf{Delegation process}
\end{flushleft}

Here we describe the computation offloading carried out by the IoT device.  In \autoref{fig:DelegationProving} we show an example of the IoT acting as a User, Proving a Presentation Policy to a third party Verifier.

\begin{enumerate}
	\item Communication with P2ABCE actor.

	The IoT device, acting as a User, starts an interaction with another actor, e.g., against an Issuer to obtain a signed credential, or against a Verifier to demonstrate certain property of its attributes in a privacy-preserving way.

	\item Delegation to the P2ABCE Server.

	Depending on what role the IoT device is acting as, it will delegate in the corresponding service, e.g. User Service. The delegation message must include the XML file, and any parameter required to accomplish the task, like the information on how to communicate with the IoT smart card (listening port, security challenge, etc.).

	\item APDU Dialogue (if necessary).

	The server may need to send APDU Commands to the IoT smart card to read the credential information or perform cryptographic operations involving private keys, necessarily stored inside the IoT device.

	These APDU Commands include a list of 70 different instructions, identified by the APDU INS byte. Some of them manage the smart card, like changing the PIN code, and many are P2ABCE specific, like storing the system cryptographic parameters. Some examples:

	\begin{tabular}{|c|l|}
		\hline
		INS & Description \\
		\hline
		\texttt{0x52} & \texttt{SET\_CREDENTIAL} \\ & Sets a new credential inside the smart card.\\
		\texttt{0x5A} & \texttt{GET\_CREDENTIAL\_PUBLIC\_KEY} \\ & Returns a stored credential's public key.\\
		\texttt{0x60} & \texttt{GET\_PRESENTATION\_COMMITMENT} \\ & First cryptographic operation during Proving.\\
		\texttt{0x62} & \texttt{GET\_PRESENTATION\_RESPONSE} \\ & Second cryptographic operation during Proving.\\
		\hline		
	\end{tabular}

	\hfil

	\item Server response.

	The server may return a status code or a XML file if the first one required an answer from the IoT device, in which case, it will send as response to the third party actor, resuming the communication.

\end{enumerate}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/UML/DelegationProving}
	\end{center}
	\caption{Designed interactions exchanged during the IoT delegation for P2ABCE Proving operation.}
	\label{fig:DelegationProving}
\end{figure}

Transmissions over the \textit{Server-IoT} channel must be secured in order to avoid attacks like: impersonate the P2ABCE delegation server, having access to the IoT smart card sending the APDU Commands the attacker wishes; delegate as a device on the server but giving the parameters of another device, making the delegation server send the APDU Commands to a victim IoT smart card.

We could use a corporative PKI to issue certificates to the server and devices and configure policies for access control; design a challenge-response system combined with the smart card PIN, like a password and TOTP\footnote{Time-based One-time Password} in a 2FA\footnote{Two-factor authentication} login. We also could connect physically the delegation service through RS-232 serial to the IoT device, securing both physically as we would do with the IoT device on its own, isolating the delegation system from any network attack. This last idea is an approximation to the Arduino Yún\footnote{\url{https://www.arduino.cc/en/Main/ArduinoBoardYun}}, a development board that integrates two microcontrollers, one a typical Arduino with very low resources, and another one running a fork of OpenWrt. The Arduino microcontroller can control the terminal of the more powerful one, using the serial pins as commented before.

As we can see, there are many state of the art solutions for all this threads, therefore, we can assume a secure channel without mentioning a specific solution, providing freedom to choose the most fitting one in a real deployment.

\paragraph{Notes for more constrained devices}

Our architecture is designed for devices that could in a future run a reimplemented version of P2ABCE, that means, the devices could perform more tasks than only running the smart card software and their main purpose process, e.g. recollecting sensor data. But if our target devices are so constrained that can barely run the smart card, they may not be able to handle the XML files because of memory restrictions, like a MSP430\footnote{\url{http://www.ti.com/lsds/ti/microcontrollers-16-bit-32-bit/msp/overview.page}} running Contiki-OS, the microcontroller has between hundred of bytes to tens of kilobytes of memory, making impossible to store multiple XML files in the size range of tens of kilobytes.

In these cases, the delegation in the server goes a step forward, making the server a proxy to communicate with other P2ABCE actors, and the IoT device only acts as a smart card. The IoT device would still act as the User, or any other P2ABCE, role because it orchestrates when and how an interaction with other actor is executed, but the communications would be between the proxy and the third party actor.

\subsection{Proof of Concept Implementation}
% +Implementation:
% *System:
% 	-Raspbian OS: distro de debian para Raspberry Pi, que se describirá en los tests
% 	-LEDE: ash terminal (shell command interpreter), POSIX, typical hardware, libraries with no hardware aceleration vs smart card MULTOS,
%
% *Delegation
%  -REST as current delegation protocol: curl script bash
%  -BIOSC as current APDU transmission protocol: no security -> no overload.
% *Smart Card
%  
%  -IoT Smart Card software design and implementation notes.
%  -Sequence diagrams. TODO

In this section we present the first PoC implementation, introducing the IoT system where we are going to work, then we will describe the delegation protocols, one for the computation offloading of the IoT device on the P2ABCE server, and another for the transmission of APDU Commands, and finally, we will describe the IoT smart card implementation.

\subsection{IoT system}

We will develop our PoC in Linux based systems in order to avoid complications with firmware specific issues that we are not familiar with. Even so, the linux systems used are aimed for the IoT environment and serve as a starting point for future implementations in more constrained devices or different systems.

In our delegation server we will run Raspbian OS, a distribution based on Debian for the Raspberry Pi. We will talk more about the hardware specifications in the benchmark chapter. We chose this system because it has a great package repository support, including the Java Runtime Environment, needed to run the P2ABCE Services.

For our IoT device, we will use LEDE, Linux Embedded Development Environment, a distribution born from OpenWrt, aimed for routers and embedded chips with low resources requirements. It offers the \texttt{ash} command interpreter, or \texttt{shell}, and the \texttt{opkg} package manager, that allows an easy installation of some libraries and tools needed during the early development.

\subsection{PoC Delegation}

As we explained in the design section, the delegation has two steps, the IoT device calling the P2ABCE server to offload the parsing of the XML data, and the P2ABCE server sending APDU Commands to the IoT smart card in the device.

\subsubsection{PoC Delegation to the P2ABCE Server}

Currently P2ABCE offers multiple REST web services to run different roles in P2ABCE system: User Service, Issuer Service, Verification Service, etc. Any third party application that integrates P2ABCE with their system can make use of these services or implement the functionality using the library written in Java, the tool that the REST services actually use.

Our PoC machine, the Omega2, can make REST calls easily with the \texttt{curl} command, but other devices may use \ac{CoAP}, but in that case, the P2ABCE REST services should be adapted to offer CoAP support. The commands needed to delegate to the P2ABCE server would be the same that those defined now to operate with the REST services, but with the modifications needed to pass the parameters on how to communicate with the IoT smart card.

In this PoC, only the P2ABCE's User Service needed to be modified, because this is the role the IoT device plays. We added the following REST call:
\begin{center}
	\textit{/initIoTsmartcard/{issuerParametersUid}?host=\&port=}
\end{center}
where we communicate the P2ABCE server that an IoT Smart Card is accessible via the IP address \textit{host} and TCP port \textit{port}. Then a new \textit{HardwareSmartcard} object is stored in the P2ABC Engine, but instead of the \textit{javax.smartcardio} Oracle's \textit{CardTerminal} implementation, we use our own \textit{IoTsmartcardio} implementation for the \textit{HardwareSmartcard} constructor, that we will discuss in the following section.

\hfil

In a real deployment, we would offer a full library with an API for other processes that want to use the P2ABC Engine, which automatizes the boot of the IoT smart card and handles the security policies, like the one we presented in the deployment diagram \ref{fig:P2ABCE-IoT} of the \textit{P2ABCE IoT Toolkit}.

Now we show some examples of the PoC \texttt{curl} commands, run in the device's terminal, or from a shell script for automation:

\begin{lstlisting}[language=bash]
$ curl -X POST --header 'Content-Type: text/xml' "http://$DelegationServerIP:9200/user/initIoTsmartcard/http%3A%2F%2Fticketcompany%2FMyFavoriteSoccerTeam%2Fissuance%3Aidemix?host=192.168.3.1&port=8888"

$ curl -X POST --header 'Content-Type: text/xml' -d @firstIssuanceMessage.xml "http://$DelegationServerIP:9200/user/issuanceProtocolStep/" > issuanceReturn.xml
\end{lstlisting}

The IoT device will only manage XML as data files to exchange between the third party actor and the P2ABCE delegation service, attaching them in the body of the REST call. In the parameters of the first call we can see the IP and port where the IoT smart card will be listening.

\subsubsection{APDU Dialogue Transmission}

To transmit the APDU messages in our PoC we use a simple protocol, that we will refer as BIOSC (Basic Input Output Smart Card), consisting in one first byte for the instruction:

\hfil

\begin{tabular}{|c|l|}
	\hline
	\texttt{0x01} & APDU Command or Response: Read 2 bytes for the\\& length and then as many bytes as said length. \\
	\hline
	\texttt{0xff} & Finish connection: close the current TCP socket and\\& end the IoT smart card execution. \\
	\hline
\end{tabular}

\hfil

In the first instruction, we read two header bytes with the length of the APDU Command or Response to receive, followed by said APDU bytes. The message is sent over TCP for a reliable transmission (concept of session, packet retransmission, reordering, etc.). The second instruction serves for closing the TCP socket in the IoT device when the smart card is no longer needed.

We lack any security (authentication or authorization) that a real system should implement. It is vital to authenticate the delegation service, to authorize it to make APDU Commands, and the same with the IoT device, to prevent attacks. But using this method for the transmission of the APDU messages, with only 3 bytes of overhead, helps us in the  benchmarks to measure the real performance of the system.

\hfil

The implementation of this simple protocol is done in Java for the P2abce delegation server and in C for the IoT smart card device.

\hfil

As we mentioned in the analysis of P2ABCE's code, the \texttt{HardwareSmartcard} class implements the \texttt{Smartcard} interface using the package of abstract classes \texttt{javax.smartcardio} to communicate with the physical smart cards. The Oracle JRE implements these package for the majority of smart card manufacturers. For our PoC, we implement the \texttt{javax.smartcardio} package so it transmits the APDUs with BIOSC, making the use of a physical or IoT smart card totally transparent to the \texttt{HardwareSmartcard} class, enhancing maintainability, and following the \textit{expert pattern} from the known GRASP guidelines.

%TODO : optional: diagrama secuencia de P2ABCE, pero si eso solo de la reimplementación de javax.smartcardio y su uso por HardwareSmartcard

\hfil

In the PoC IoT device, we implement in \texttt{BIOS.c} the \texttt{main()} function, that reads from a Json file the previous status of the smart card, and then opens the TCP socket in the $8888$ port, and loops listening for BIOSC transmissions until an $0xff$ is received. When an APDU Command is sent, it calls the APDU Handler, that we will talk about in the next section.

\subsection{IoT Smart Card Implementation}

After many design decisions in the process to adapt the original ABC4Trust Card Lite code to pure C, working on a MIPS architecture, in this section, we present the current \ac{PoC} code, most important decisions taken and the execution workflow in the form of sequence diagrams.

\hfil

\subsubsection{Code structure}

We divide the project in three different sections with the objective of enhancing maintainability, improving future changes, ports, fixes, etc.

In \autoref{fig:IoTCScomponents-bw} we show the project's structure. The first section is what could be called as the core of the smart card, the second one the interface for those tools the core needs and that may depend on the target's platform, and finally, the third party libraries, that may be empty if the interfaces implementation doesn't require any.

%\begin{figure}[bth]
%	\begin{center}
%		\includegraphics[width=\linewidth]{gfx/IoTCScomponents-color}
%	\end{center}
%	\caption{IoT Smart Card Code Structure.}
%	\label{fig:IoTCScomponents-color}
%\end{figure}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/IoTCScomponents-bw}
	\end{center}
	\caption{IoT Smart Card Code Structure.}
	\label{fig:IoTCScomponents-bw}
\end{figure}

\hfil

\paragraph{Core smart card}
\hfil

The smart card logic lies in this section, the concepts of APDU Commands, what instructions are defined for P2ABCE smart cards, and how to process them and generate proper APDU Responses.

If the APDU protocol defined for P2ABCE smart cards changed, with the consequence of having to update all smart cards in use, the changes to adapt the IoT smart card should be applied in this part of the code. This also implies that if the P2ABCE Java code changes their \texttt{Smartcard} interface, something more probable, the IoT smart card will still work as intended.

Most of the original ABC4Trust Card's code has been reused in this section, e.g., all the global variables and custom data type definitions, the APDU Command handling and auxiliary subroutines.

However, the ABC4Trust's code heavily depended on the MULTOS platform for the input and output of data, modular arithmetic, memory management, and AES128 and SHA256 cryptographic operations. Therefore, we reimplement the used MULTOS functions following the documentation, adapting it in some cases.

A characteristic of MULTOS C-API is that every function name starts with ``\texttt{multos}'', so we changing their names from \texttt{multosFoo()} to \texttt{mFoo()} for readability and to emphasize that they were no longer MULTOS functions.

The \texttt{main.h} file, from the original ABC4Trust project, also implemented some functions that were equivalent to the ones available in \texttt{multos.h}. We replaced them for the standard functions in the C-API, checking that the MEL code did the same as the documentation specified. This last detail is very important, because in \citep{vullers2013efficient} they noted that MULTOS' \texttt{ModularExponentiation} function does not accept exponents larger than the modulus size, so they implemented \texttt{SpecialModularExponentiation}, dividing the exponentiation in two that MULTOS could perform. After this analysis, when those particularities appeared, our reimplementation of the MULTOS API accepts the \textit{expanded functionality}. In total, we reduced to eighteen the number of MULTOS functions to reimplement, and with future refactoring this number may decrease.

\hfil

Because of the inherited code, many particularities of the internals of the MULTOS platform had to be addressed.

One was the different memory zones, where the variables in \texttt{static} were read and written directly from the secure EEPROM, defining the status of the smart card between executions. We keep those variables in RAM, but save them in a Json file, deserializing them at the smart card boot up, and serializing always before an APDU Response is sent. From the MULTOS documentation, the \texttt{static} variables are read and written atomically, and if power source is lost during any process, memory won't get corrupted. This also means that if a MULTOS smart card sends an APDU Response, every variable that had to change, is saved in the EEPROM. Our PoC uses Json for human readability during debugging, but a real deployment should apply secure reads and writes on the file, preventing data corruption and undesired access, for example, ciphering the file with a key derived from the smart card PIN. It is also desirable another data serialization method, reducing the size of the file to store in memory constrained devices.

Another of the mentioned particularities is that the MULTOS compiler does not apply padding between variables in the data structures. For example, in an X86 machine, where a \texttt{short} is stored in 2 bytes, in the following code\footnote{From \url{https://en.wikipedia.org/wiki/Data_structure_alignment}}

\begin{lstlisting}[language=C]
struct MyData
{
short Data1;
short Data2;
short Data3;
};
\end{lstlisting}

each member of the data structure would be 2-byte aligned. Data1 would be at offset 0, Data2 at offset 2, and Data3 at offset 4. The size of this structure would be 6 bytes. But if we defined the following structure

\begin{lstlisting}[language=C]
struct MixedData
{
char Data1;
short Data2;
int Data3;
char Data4;
};
\end{lstlisting}

where a \texttt{char} is 1 byte, and an \texttt{int} uses 4 bytes, although the total bytes used are 8 bytes, the compiler will align them in memory adding padding, like if we defined the following:

\begin{lstlisting}[language=C]
struct MixedData  /* After compilation in 32-bit x86 machine */
{
char Data1; /* 1 byte */
char Padding1[1]; /* 1 byte for the following 'short' to be aligned on a 2 byte boundary
assuming that the address where structure begins is an even number */
short Data2; /* 2 bytes */
int Data3;  /* 4 bytes - largest structure member */
char Data4; /* 1 byte */
char Padding2[3]; /* 3 bytes to make total size of the structure 12 bytes */
};
\end{lstlisting}

\hfil

In MULTOS, to reduce the memory usage, there is no padding, and this affects the inherited ABC4Trust's code because of the use of \texttt{memcpy} to copy zones of memory from one address to another. The problem is the code copies multiple variables in one \texttt{memcpy} call, because, for example, the APDU Command payload includes multiple data, and instead of copying one variable at a time, the \texttt{struct} is defined with the same order and copies everything at once.

An example taken from the parsing of instruction \texttt{SET\_PROVER}:

\begin{lstlisting}[language=C,frame=tblr]
typedef struct
{
BYTE prover_id;
unsigned int ksize;
unsigned int csize;
BYTE kx[MAX_SMALLINT_SIZE];
BYTE c[HASH_SIZE];
BYTE proofsession[PROOFSESSION_SIZE];
BYTE proofstatus;
BYTE cred_ids[NUM_CREDS];
BYTE cred_ids_size; // also called 't' in the documentation
BYTE exists;
} PROVER;
/*****/
PROVER provers[NUM_PROVERS];

/*****/
case INS_SET_PROVER:
// [...]
memcpy(&(provers[temp_prover_id-1].prover_id), &(apdu_data.set_prover_in.prover_id), 5); // under the hood, this also initializes ksize and csize
\end{lstlisting}
This is the only case where a comment showcases that it is copying more than one byte.

The temporal solution is to use \texttt{struct \_\_attribute\_\_((\_\_packed\_\_))} to ask a GCC compiler to not use padding in the structs, but this is not standard, neither a good practice. A deeper refactorization of the code is needed where the hidden copies of variables are made explicit, letting the compiler manage the memory layout.

\hfil

An internal alarm should always warn us when we see raw memory copy from serialized data, that is, the APDU Payload. We mentioned in the MULTOS introduction that the platform is Big Endian. The serialized data is also typically represented in Big Endian, and our APDU Payloads do so. In our code we added a new Subroutine to check whether the device is Big or Little Endian, and to rectify the Endianness from those variables, with 2 or more bytes (the only ones affected by the Endianness), copied from or written to APDU Payloads.

\hfil

Finally, we must talk about the \texttt{multosExit} function. In a MULTOS application, to finish the execution and send the APDU Response, the programmer can call in any moment to \texttt{multosExit}, and the MULTOS OS will continue the execution, not returning to the application again.

Think about how we all have, at least once, used an \texttt{if} and \texttt{return} to avoid the \texttt{else}, like in this example:

\begin{lstlisting}[language=C]
if(condition)
return a;
return b;
\end{lstlisting}

The use of \texttt{multosExit} in ABC4Trust code is similar, but used in almost every function. This leads to a sequence execution with many possible interruptions, leaving a full stack behind with \textit{un-returned} function calls, i.e., parameters stored in memory after a function call that never returns. In MULTOS, this memory won't affect future executions, because with the next APDU Command, the stack will be loaded with the clean application, and the Session memory stays untouched between APDU Commands, during an APDU Dialogue.

To adapt the code, the reimplemented \texttt{mExit} function can not end the program execution, because we would lose the variables in RAM needed during the APDU Dialogue, as well as that the TCP socket would close. The solution is to refactor every use of \texttt{multosExit} in a way that, if it is called from a Subroutine, the function returns a success or error code, and the only functions that can call \texttt{mExit} are the ones handling an specific instruction.

Nevertheless, there is one exception, the \texttt{output\_large\_data()} Subroutine, which is a tool to send large data payloads without Extended APDU support. We talked about this in \ref{subsec:APDU}, mentioning the special instruction \texttt{GET RESPONSE}.

Summarizing, every Subroutine must return from its execution, as well as every instruction handler, which must call either \texttt{mExit} or \texttt{output\_large\_data()} before returning. After the instruction has been handled, the execution returns to the listening loop of BIOSC.

\paragraph{Interfaces}\hfil

To reimplement some of the MULTOS functions, we needed to use some libraries, so we defined a facade to isolate the implementation of the core smart card from our different options, that could vary depending on the hardware or the system used by the IoT device.

The use of a facade lets us, for example, change the implementation of modular arithmetic with a hardware optimized version, or a future more lightweight library, or our very own software implementation using the same data types that the core uses, minimizing the data usage.

Taking a step forward, we make the core smart card totally independent of any library, only dependant of our interfaces. This means that typical C libraries, like the standard \textit{stdlib.h}, or  \textit{string.h} are also behind the facade, in case some IoT system doesn't support them. The main goal we go after with this decision is that future developers adapting the code to a specific platform need to make no change to the \textit{core smart card}'s code, only to the interfaces implementation.

The interfaces defined can be organized in 5 groups (see \autoref{fig:IoTCScomponents-bw}), depending on their purpose: Modular Arithmetic, Cryptography, Memory Management, Serialization, APDU Parsing.

\paragraph{External utilities}\hfil

If the IoT system offers well tested libraries that could aid in the interfaces implementation, for example, a assembly optimized code for the AES128 and SHA256 operations, these third party libraries belong to this section.

In our PoC, we use two ANSI C libraries, for base64 and JSON, and two shared libraries available in as packages in LEDE, GMPLib and OpenSSL. These libraries use dynamic memory and offer more functionality than we need, although they are a good tool in this PoC, future versions should use more lightweight solutions.

For example, Atmel's ATAES132A\footnote{ATAES132A 32K AES Serial EEPROM  Datasheet - \url{http://ww1.microchip.com/downloads/en/DeviceDoc/Atmel-8914-CryptoAuth-ATAES132A-Datasheet.pdf}}
\marginpar{\includegraphics[width=0.9\linewidth]{gfx/atmel} \\ Atmel's cryptography chips.}
offers a serial chip for secure key storage, AES128 execution and random number generation. Another serial chip like ESP8266 offers WiFi connectivity, typically used with Arduino, and can also perform AES encryption. For random number generation, a technique used with Contiki devices is to read from sensors aleatory data and use it as seed.

All these alternatives depend on the target device, but are all valid, and would substitute OpenSSL with only a Serial communication library, so we can talk to the dedicated chips.

\hfil

The \textit{interfaces} and \textit{external utilities} sections  allow that the project is easily ported to specific targets without modifying the smart card logic.

\subsubsection{Execution workflow}

The sequence diagram from \autoref{fig:sequenceBIOSC} shows the execution of the PoC IoT smart card.

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/UML/sequenceBIOSC}
	\end{center}
	\caption{IoT Smart Card Sequence Diagram.}
	\label{fig:sequenceBIOSC}
\end{figure}

The program starts with the \texttt{main} function in BIOSC, that deserializes the status from the Json file, and listens on a loop for APDU Commands from the delegation server.

Every time an APDU Command arrives, it calls the function \texttt{handle\_APDU()} with the raw APDU bytes. The Handler calls the APDU I/O interface to parse the bytes, storing in global variables the APDU structure. Using a \texttt{switch-case} expression on the \texttt{INS} byte, the Handler calls an \textit{Instruction handler} function.

Inside this function, it may call multiple functions from the Subroutines, that may call MULTOS C-API functions, Which in turn may use an interface to perform its functionality.

As we mentioned, every instruction handler must end, before the \texttt{return;} expression, with \texttt{mExit} or  \texttt{output\_large\_data()}, that will use \texttt{mExit}. This reimplemented MULTOS function will save the current status of the smart card, with the help of the serialization interface, and once the file is saved, it outputs the APDU Response back to the delegation server.

After returning from \texttt{mExit}, the instruction handler and the APDU handler functions, the program listens again from the socket, until an $0xff$ byte is received.

\subsection{Design and implementation conclusions}

During this design and implementation process we had to study multiple technologies, including the ones mentioned in the state of the art, regarding Idemix, and those related to the IoT, we studied many systems (e.g. Contiki, Arduino) and hardware (e.g. MSP430, ESP8266), and after this, we had to take many decisions during development.

The design is open in those aspects that depend on each specific project's requisites, and closed on the aspects involving the P2ABCE's core logic.

However, the PoC had to make more concessions when implementing the cryptographic and modular arithmetic logic. But those are parts of the code that are more prone to be changed when developing for a new system. The core logic of the PoC is preserved as it is in every system and device.

Even with the many points of future work left, we think we achieved the most extensible solution for Idemix+IoT at the time, our main objective of the project.

%************************************************
\section{Validation and Performance Evaluation}\label{ch:validation}
%************************************************

In this chapter, we will describe the deployment of three testing scenarios: a laptop, a Raspberry Pi 3, and a Omega2 IoT device with the Raspberry Pi 3 as the delegation server. We will measure and compare the results to determine if the proposed solution is feasible or must be submitted to revision.

\subsection{Testbed description}

First, we shall describe the example Attribute Based Credential system in use. Then, the hardware we will use in our benchmarking.

\subsubsection{P2ABCE setting}

To test the correct execution of the \textit{IoT smart card}, we will use the ABC system from the tutorial in the P2ABCE Wiki\footnote{\url{https://github.com/p2abcengine/p2abcengine/wiki/}}. It is based on a soccer club, which wishes to issue VIP-tickets for a match. The VIP-member number in the ticket is inspectable for a lottery, ie. after the game, a random presentation token is inspected and the winning member is notified.

First the various entities are \textbf{setup}, where several artifacts are generated and distributed. Then a ticket credential containing the following attributes is issued:

\begin{verbatim}
First name: John
Last name: Dow
Birthday: 1985-05-05Z
Member number: 23784638726
Matchday: 2013-08-07Z
\end{verbatim}

During \textbf{issuance}, a \textit{scope exclusive pseudonym} is established and the newly issued credential is bound to this pseudonym. This ensures that the ticket credential can not be used without the smart card.

Then \textbf{presentation} is performed. The \textit{presentation policy} specifies that the member number is inspectable and a predicate ensures that the matchday is in fact $2013-08-07Z$. This last part ensures that a ticket issued for another match can not be used.

The ticket holder was lucky and his presentation token was chosen in the lottery. The presentation token is therefore inspected.

\hfil

\subsubsection{Execution environment}

First we will execute the test in our development machine (laptop). After asserting that the services work as expected, we then run the test in a Raspberry Pi 3, exactly like in the laptop. Finally, we will deploy the IoT smart card in a Omega2 and the delegation services in the Raspberry Pi 3. After every test, we checked that the issuing and proving were successful, in case a cryptographic error appeared in the implementation.

Lets have a closer look at the hardware of each device:

\paragraph{Development laptop}
Our device is a DELL XPS 15, with a Core i7-6700HQ at 3.5GHz quad core processor and 32GB of DDR4 RAM, running Ubuntu 16.10.

This is a powerful machine that can simulate the performance of many servers and clients that would implement P2ABCE in a real environment, giving a reference point for performance comparisons.

\paragraph{Raspberry Pi 3} A familiar environment, powerful enough to debug and hold the delegated P2ABCE Java services of P2ABCE with its 1GB of RAM, and with two network interfaces, perfect to work as the gateway for the IoT devices to the Internet.

Only a microSD with enough space to burn the OS is needed to plug\&play with the Raspberry Pi. We use Raspbian, a stable Debian based distro, recommended by the Raspberry Pi designers, and ready to use with the P2ABCE compiled \textit{self-contained .jar} services.

\begin{table}[h]
	\myfloatalign
	\begin{tabular}{ll} \toprule
		CPU & ARMv8 64bit quad-core @1.2GHz \\
		RAM & 1GB \\
		Storage & microSD \\
		Firmware & Raspbian (Debian based distro) \\
		Connectivity & Wifi n + Ethernet \\
		Power & 5V 2A \\
		\bottomrule
	\end{tabular}
	\caption[Raspberry Pi 3 Specifications]{Raspberry Pi 3 Specifications.}
	\label{tab:RPi3Specs}
\end{table}

\paragraph{Onion Omega2} A device that falls inside the category of IoT, powerful enough to run a Linux embedded environment, such as LEDE, where we can develop and debug the first PoC without troubling ourselves with problems not related to the project itself.

\begin{table}[h]
	\myfloatalign
	\begin{tabular}{ll} \toprule
		MCU & Mediatek MT688\footnote{MediaTek MT7688 Datasheet - \url{https://goo.gl/kqD2gF}} \\
		CPU & MIPS32 24KEc 580MHz \\
		RAM & 64MB \\
		Storage & 16MB \\
		Firmware & LEDE (OpenWRT fork distro) \\
		Connectivity & Wifi b/g/n \\
		Power & 3.3V 300mA \\
		\bottomrule
	\end{tabular}
	\caption[Onion Omega 2 Specifications]{Onion Omega2 Specifications.}
	\label{tab:Omega2Specs}
\end{table}

Nonetheless, the Omega2 needs fine tunning to start operating, and basic knowledge of electronics to make it work. The two main things to begin with Omega2 are:
\begin{itemize}
	\item A reliable 3.3V with a maximum of 800mA power supply, e.g. a USB2.0 with a step-down circuit, with quality soldering and wires to avoid unwanted resistances. The Omega2 will usually use up to 350mA, when the WiFi module is booting up. The mean consumption is about 200mA.
	\item A Serial to USB adapter wired to the TX and RX UART pins to use the Serial Terminal, to avoid the use of SSH over WiFi.
\end{itemize}

A downside of using LEDE and the Omega2 is the lack of hardware acceleration for cryptographic operations, unlike the MULTOS applications, because the smart cards hardware and MULTOS API include support for such common operations in smart cards.

\paragraph{The network} In our third scenario, the Raspberry Pi 3 and the Omega2 will talk to each other over TCP. This implies possible network delays depending on the quality of the connection. The Raspberry Pi 3 is connected over Ethernet to a switch with WiFi access point. The Omega2 is connected over WiFi n to said AP. To ensure the delay wasn't significant, we measured 6000 APDU messages, and the results show that the mean transmission time is less than half a millisecond per APDU. As we will see in the results section, this network time is negligible.

\paragraph{Future work for tests} The lack of a physical MULTOS smart card precludes us to load and test the ABC4Trust Card Lite's code and measure the time P2ABCE would need when using the \textit{HardwareSmartcard} class. This would be really interesting because for a single method from the \textit{Smartcard} interface, \textit{HardwareSmartcard} implementation needs to send multiple APDU Commands, but \textit{SoftwareSmartcard} can perform the operations immediately, with the full computer's resources. Also, physical smart cards benefit from hardware acceleration in most of the cryptographic operations, unlike our software implementations in the PoC.

\subsection{The test code}

In this section we present the scripts and binaries used during the tests.

There are three pieces of software that conform the test: the P2ABCE services, the IoT smart card, and shell scripts automatizing the REST calls, from the terminal.

\paragraph{P2ABCE services}\hfil

This is a common part to our three sets. The services are compiled in a self-contained Jetty web server, or in WAR format, ready to be deployed in a server like Tomcat. We use the same JAR files with the embedded Jetty web server for the PC and Raspberry Pi.

We modified the User Service code to measure the execution elapsed time for each REST method. We don't measure the time the web server spends processing the HTTP protocol and deciding which Java method to call.

\paragraph{IoT smart card}\hfil

Our C implementation of the P2ABCE smart card, compiled for the Omega2, with BIOSC listening on port 8888 for the APDU messages.

We tested the execution in two modes, a full logging where every step was printed in terminal, and another one with no logging. With the first mode, we can check a proper execution, every byte exchanged, and with the second one, we measure the execution without unnecessary I/O.

\paragraph{Shell scripts} \hfil

To orchestrate all the services we use shell scripts that execute the REST calls using  \texttt{curl}. Here we perform the mentioned steps: setup of the P2ABCE system, issuance of the credential and proving for the presentation policy.

In the setup, the system parameters are generated, indicating key sizes of 1024 bits, the ones currently supported by the ABC4Trust and IoT smart cards. Then the system parameters and public keys from the services (issuer, inspector, revocation authority) are also exchanged and stored in each Service.

During the issuance and proving, the User and the Issuer and Verifier exchange multiple XML files, with the cryptographic information of each step of the Idemix protocol.

We offer two versions for the scripts. The first one is a single shell script file, to be executed from a single terminal. The advantage of this version is that, because every XML file is stored in the machine running the script, we don't have to transfer these files by other means, e.g. \texttt{scp}. An excerpt of this script showcases how we receive an XML file from the User Service, e.g. \texttt{secondIssuanceMessage.xml}, and send it to the Issuer Service, referencing the file stored in the script's working directory:

\begin{lstlisting}[language=bash]
[...]
# First issuance protocol step - UI (first step for the user).
echo "Second issuance protocol step (first step for the user)"
curl -X POST --header 'Content-Type: text/xml' -d @uiIssuanceReturn.xml 'http://localhost:9200/user/issuanceProtocolStepUi/' > secondIssuanceMessage.xml

# Second issuance protocol step (second step for the issuer).
echo "Second issuance protocol step (second step for the issuer)"
curl -X POST --header 'Content-Type: text/xml' -d @secondIssuanceMessage.xml 'http://localhost:9100/issuer/issuanceProtocolStep/' > thirdIssuanceMessageAndBoolean.xml
[...]
\end{lstlisting}

The second script version is intended for a more realistic execution. In a real scenario, the User and the Issuer, for example, would call the REST methods of their respective Services, but not the other actor's Service, exchanging the XML files through other means. Instead of one script running in a single machine, orchestrating the Omega2, Raspberry Pi 3 and laptop, we offer three scripts, one per machine, with \textit{pauses} where one device must send an XML file to another machine. An excerpt from the Omega2's script shows the same step as above, but instead of calling the Issuer's REST Service, waits until we send the corresponding files.

\begin{lstlisting}[language=bash]	
# First issuance protocol step - UI (first step for the user).
echo "Second issuance protocol step (first step for the user)"
curl -X POST --header 'Content-Type: text/xml' -d @uiIssuanceReturn.xml "http://$Rpi3IP:9200/user/issuanceProtocolStepUi/" > secondIssuanceMessage.xml

read -p "Send => secondIssuanceMessage.xml <= back to the Issuer. Then press any key to continue... " -n1 -s

read -p "Receive => thirdIssuanceMessageAndBoolean.xml <= from the Issuer. Then press any key to continue... " -n1 -s
\end{lstlisting}

\hfil

The first version is recommended for tests, and the second version only for instructive purposes.

\subsection{Results}

After 20 executions for each scenario (laptop, RPi3, Omega2+RPi3), we take the means and compare each step of the testbed.

It is worth noting that during the test, the measured use of the CPU showed that P2ABCE does not benefit of parallelization, therefore, it only uses one of the four cores in the laptop and Raspberry Pi 3.

To test the network, we sent six thousand APDUs to the Omega2, but instead of calling the \textit{APDU handler}, the Omega2 responded with the same bytes back. This way, the Omega2 only performed the simple BIOSC protocol, reading from and writing to the TCP socket.
The APDUs had multiple sizes, taken from the most common APDUs logged in during a successful execution. The test showed that our network speed was around $0.014$ ms per byte.

\paragraph{The setup}\hfil

The first step of our testbed. The Omega2 doesn't intervene until the creation of the smart card, therefore, the times measured in the second and third scenarios are practically identical.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[Times and relative speedup]
	{\includegraphics[width=\linewidth]{gfx/graphics/setuptable}} \quad
	\subfloat[Comparison graph]
	{\includegraphics[width=.8\linewidth]{gfx/graphics/setup}} \\
	\caption{Setup times (milliseconds)}
	\label{fig:setup:graph}
\end{figure}

As we can see in \autoref{fig:setup:graph}, the laptop is about ten times faster than Raspberry Pi 3, but considering that the highest time is less than two and a half seconds, and that the setup is done only once, this isn't a worrisome problem.

\paragraph{Creation of the smart card}\hfil

Here we create a \textit{SoftwareSmartcard} or a \textit{HardwareSmartcard} object that the User service will use in the following REST calls.

The REST method to create a \textit{SoftwareSmartcard} is $/createSmartcard$, and to create a \textit{HardwareSmartcard}, using the \textit{IoTsmartcardio} implementation, we use $/initIoTsmartcard$. This operation is done only once per device, and includes commands from the creation of the PIN and PUK of the smart card, to storing the system parameters of P2ABCE, equivalent to the previous setup step.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[Times and relative speedup]
	{\includegraphics[width=0.5\linewidth]{gfx/graphics/createSCtable}} \quad
	\subfloat[Comparison graph]
	{\includegraphics[width=0.45\linewidth]{gfx/graphics/createSC}} \\
	\caption{Create smart card times (seconds)}
	\label{fig:createSmartCard:graph}
\end{figure}

From \autoref{fig:createSmartCard:graph} we see that the RPi3 is about 16 times slower than the laptop in the creation of the \textit{SoftwareSmartcard}, but almost 9 times faster than the setup of the smart card in the Omega2 using APDUs. This gives us that the laptop is 145 times faster than the combination of RPi3 and Omega2 in our IoT deployment.
But looking at the times, this process lasts up to 20 seconds, making it something feasible.

This is the first interaction between the RPi3 and the IoT smart card running in the Omega2. To setup the smart card \textbf{30 APDU Commands}, and their respective Responses, are exchanged, as shown in \autoref{ch:resultsdiagrams}, \autoref{fig:APDUsInitIoTSC}, with a total of $1109$ bytes. From our network benchmark, using TCP sockets, the delay in the transmission is only around 15 and 20 ms, negligible, as we said, compared to the almost 20 seconds the operation lasts.

\paragraph{Issuance of the credential}\hfil

Our credential will have 5 attributes and key sizes of 1024 bits, as specified during the setup process.

The issuance is done in three steps for the User service, shown in \autoref{fig:IssuanceInteraction} with a red note showing the start of each step for the User delegation, in green for the interactions between Issuer and User, and the darker red is the Identity service, choosing the first available identity to use. The arrows in the figure show the REST calls performed during the test, where the IoT device acts as User, the RPi3 hosts the P2ABCE delegation services and the laptop is the Issuer.

The three delegation steps and the REST method called are:

\begin{center}
	\begin{tabular}{l|l}
		First issuance protocol step & $/issuanceProtocolStep$ \\
		Second issuance protocol step  & $/issuanceProtocolStepUi$ \\
		\textit{(end of first step for the User)} & \\
		Third issuance protocol step & $/issuanceProtocolStep$ \\
		\textit{(second step for the User)}  & \\
	\end{tabular}
\end{center}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/UML/IssuanceInteraction}
	\end{center}
	\caption{Issuance interaction.}
	\label{fig:IssuanceInteraction}
\end{figure}

As we can see, the three REST calls to the delegation User service involve communication with the smart card. We show in \autoref{ch:resultsdiagrams}, \autoref{fig:IssuanceAPDUs}, the APDU Commands used for each REST call in the issuance.
There are 45 APDU Commands in total, $3197$ bytes exchanged, that would introduce a latency of 45ms in the network, negligible.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[Times (ms) and relative speedup]
	{\includegraphics[width=\linewidth]{gfx/graphics/issuancetable}} \quad
	\subfloat[Comparison graph]
	{\includegraphics[width=.8\linewidth]{gfx/graphics/issuance}} \\
	\caption{Issuance times (milliseconds)}
	\label{fig:issuance:graph}
\end{figure}

In \autoref{fig:issuance:graph} we have the times spent in each REST call. The laptop shows again to be many times faster than the other two scenarios, but the times are again feasible even for the IoT environment.

Lets compare the Raspberry Pi 3 and the Omega2 executions. There is a correlation between the number of APDU Commands needed in each step with the increment in time when using the IoT smart card, that is, how much the delegation server.

The first one only involved one APDU, with $33$ bytes total (Command and Response), and times are almost identical. The second call needed $34$ APDUs, with $1623$ bytes, and the increase in time is around tree times slower than the RPi3 on its own. The third call used $20$ APDUs, $1541$ bytes, and makes the IoT scenario almost 7 times slower.

The analysis shows where there are more cryptographic operations involving the Omega2, and because the amount of data exchanged is minimal, the difference in processing power between Omega2 and Raspberry Pi 3 is clear.

\paragraph{Presentation token}\hfil

The final step of the test involves a Prove, or Presentation in P2ABCE, where the Verifier sends the User or Prover the Presentation Policy, and the User answers with the Presentation Token, without more steps. In \autoref{fig:ProvingInteraction}, using the same colors as in the Issuance interaction, we can see the delegation messages done by the Omega2.

To ensure that all the process was successful, it's enough to check if the Verifier and the Inspector returned XML files, accepting the prove, or an error code. Of course, every execution measured in the test was successful.

In \autoref{ch:resultsdiagrams}, \autoref{fig:APDUsProving}, we provide the APDU Commands for each step, $28$ in total, with $1939$ bytes, giving us about $27$ms of delay in the network transmission.

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/UML/ProvingInteraction}
	\end{center}
	\caption{Proving interaction.}
	\label{fig:ProvingInteraction}
\end{figure}

Again, as shown in \autoref{fig:proving:graph}, there is a correlation between the number of APDU Commands used, the work the IoT smart card must perform, and the time measured. The $20$ APDU Commands in the first call make the IoT deployment almost $8$ times slower than the Raspberry Pi 3; but with only $8$ APDU Commands, the second one is less than $1.5$ times slower.

Nonetheless, it's significant the difference in performance between the laptop and the Raspberry Pi 3 in the last REST call, more than $40$ times slower, even using the \textit{SoftwareSmartcard}.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[Times (ms) and relative speedup]
	{\includegraphics[width=.8\linewidth]{gfx/graphics/provingtable}} \quad
	\subfloat[Comparison graph]
	{\includegraphics[width=.7\linewidth]{gfx/graphics/proving}} \\
	\caption{Proving times (milliseconds)}
	\label{fig:proving:graph}
\end{figure}

Unlike the previous steps, the Presentation or Proving is done more than once, being the key feature of ZKP protocols. The laptop performs a prove in less than one second, the RPi3 needs $15$ seconds, but our P2ABCE IoT deployment needs $15$ seconds for the first step, and $18$s for the second step, $33$ seconds total to generate a Presentation Token.

\hfil

\paragraph{Memory usage on the Omega2}\hfil

Using the tool \textit{time -v} we can get a lot of useful information about a program, once it finishes. In our case, the binary with BIOSC and the smart card logic starts as an empty smart card, goes through the described process, and then we can stop it, as the User Service won't use it anymore.

After another round of tests, now using \textit{time -v}, the field named \textit{Maximum resident set size (kbytes)} shows the \textbf{maximum} size of RAM used by the process since its launch. In our case, this involves the use of static memory for the \textit{global variables} of the smart card logic, and the dynamic memory used by the third party libraries, like GMPlib, OpenSSL and cJSON.

GMP and OpenSSL always allocate the data in their own ADT, what involves copying the arrays of bytes representing the big modular integers from the cryptographic operations. cJSON, used in the serialization of the smart card for storage, and debug being human readable, stores a copy of every saved variable in the JSON tree structure, then creates a string (array of char) with the JSON, that the user can write to a file.

Understanding the many bad uses of memory done in this PoC is important for future improvements and ports. A custom modular library using the same array of bytes that the smart card logic, a binary serialization, and many improvements, are our future work.

With all that said, the mean of the maximum memory usage measured is $6569.6$ kbytes. Compared to the $64$MB of RAM available in the Omega2, our PoC could be executed in more constrained devices, given the system is compatible.

\subsection{Validation conclusions}

Below \autoref{totaltime} sums up the time in seconds used in each step of the test for the Omega2 and Raspberry Pi 3 setup.

The first step, \textit{System Setup} is done only once when the system is being deployed, and the \textit{IoT Smart Card Setup} only once per device.

Because a device can have more than one credential, the Issuance step is significant when we issue multiple credentials over the device's lifetime. We recall that our tests used a credential with five attributes and key sizes of 1024 bits.

Finally, the Proving step is expected to be the most performed calculation by the IoT device. The fact that it lasts over half a minute implies that we should not use this PoC for \textit{real-time} applications yet, usually used in critical secure systems. Nevertheless, for many other IoT applications, the fact this operation can be performed multiple times per hour, presents an useful tool for privacy, e.g. the thermostat system mentioned in the smart building, which can pool data from the sensors every few minutes.

\begin{table}[!ht]
	\begin{center}
		\begin{tabular}{|r|r|r|r|}
			\hline
			System & IoT Smart  & Issue  & Prove Presen-\\
			Setup & Card Setup & credential & tation Policy\\ \hline
			3.77 s & 19.19 s & 36.48 s & 32.84 s \\ \hline
		\end{tabular}
	\end{center}
	\caption{Total time spent for each step in the Omega2+RPi3 setup.}
	\label{totaltime}
\end{table}

%************************************************
\section{Conclusions and Future Work}\label{ch:conclusions}
%************************************************

To finish this document, we sum up some conclusions from the work done, and results 
obtained. We will also enumerate some future lines of research.

\subsection{Conclusions}

In the memory of this project we try to show the work done from the beginning, but we only showed our right decisions, and the information that is significant for the final result. The truth is that, aside from the information included in this paper, we have worked with other systems that ended up discarded. This isn't a negative aspect, because if we didn't, for example, study the Contiki OS, Cooja simulator and the hardware used, we could not be sure the development of the first PoC for that system would be impossible in the time given.

With regard to the work presented, the flexibility of the computation offloading technique, identifying the key operations that can be delegated, and those ones that can't, has allowed us to define a general solution for the vast world of the Internet of Things. The IoT devices can operate as individual actors in the P2ABCE ecosystem, and when in need of performing computation offloading, the delegation server can also be a device considered into the IoT class.

% Resultados son válidos: feseability, problema de tiempo en sistemas de tiempo real
% Se han conseguido objetivos planteados
% Experiencia: qué era más tedioso, dificultades durante diseño y desarrollo
% Primera aproximación de este tipo ~~: más bien somos el future work de...
% Aplicación de procedimientos aprendidos/aplicados durante el TFG/carrera
% ^ Novelty

During the development, we had to investigate a lot of concepts related to IoT, smart cards, and even the insides of P2ABCE's code, to fix many existing bugs in the original project and minimize the amount of changes it had to undergo, in order to work with the IoT devices. In the implementation chapter we give guidelines to port MULTOS applications, considering the particularities we encountered, that's why we can't consider it an \textit{instruction manual to port MULTOS apps}, but an interesting reading on how to confront a similar project. For example, if we wanted the Idemix implementation from \citep{vullers2013efficient}, previous to P2ABCE, in a IoT device, we could apply almost the same steps, obtaining a core functionality of Idemix for constrained devices in C.

Our PoC implementation demonstrates that this project is actually feasible, not by performing a simulation of an IoT device, like in \citep{vanet}. However, the use of third party libraries and no hardware acceleration support, makes the PoC too slow for certain cases, e.g., a Real-Time system requires almost immediate operations, like a car warning that another one is approaching too fast.
There is a long path of research before we can see this design in production, as many decisions depend on the specific deployment in consideration.

Recalling the objectives listed in \autoref{objectives:section}, we think we accomplish them, except the implementation of a PoC in the most constrained device possible, where we used LEDE to ease our development, and the last objective, where we could have given a use to the PoC, and we leave as part of the future work. In spite of this negative auto-critic, we have to acknowledge that this is the first privacy-preserving ABC implementation for IoT, designed for extensibility, interoperability and maintenance.

Finally, we are very thankful to IBM's grant for the \textit{Privacy Preserving Identity Management applied to IoT} project, which allowed us to get this far. Also, in \autoref{ch:JCR} we show the version for publication in a JCR magazine.

\subsection{Future work}

Due to the nature of the project, there exist many options to continue researching in this area. We list below a collection of the tasks we couldn't accomplish but would have been our next steps:

\begin{itemize}
	\item Second PoC for Arduino systems

	We have already developed a working PoC for Linux based systems, and from our objective of using the most constrained devices available, we can follow an iterative process, building for devices a little more constrained than the last one. The next natural step, after building with LEDE/OpenWrt, is to aim for Arduino.

	The critical changes to the actual PoC are: 
	\subitem The transmission channel, if the device has no TCP/IP stack, it could use a Serial connection to transmit BIOSC and delegation messages.
	\subitem The use of a Json file to store the smart card status depends on whether the device has enough space or a file system. The current Json parser also uses too much dynamic memory. At this stage the serialization should be performed in a binary format, in a file or writing variables directly to the EEPROM memory.
	\subitem GMPLib and OpenSSL libraries must be replaced by others with similar capabilities, or special implementations for the particular device in use.

	All these changes affect only the \textit{external utilities} and \textit{interfaces} sections of the code, leaving the \textit{core} untouched, because the smart card logic doesn't change.

	\item Implement more P2ABCE functionality in the device

	If instead of aiming for more constrained devices, we work with similar devices to the Omega2, the computational power of the device could assume all the functionality of P2ABCE, at least for some roles. In a M2M (machine to machine) environment, the devices could act as Provers and Verifiers in the same interaction, and the verifying logic could be implemented entirely in the device (generate a Presentation Policy and checking a prove in a Presentation Token is valid), reducing the need of a delegation server.

	\item Improve the \textit{core smart card} code

	From the simple refactoring of removing the MULTOS reimplemented API for direct calls to the \textit{interfaces}, to identify all the \texttt{memcpy} copies of multiple variables with one call, removing the compiler directive for not using memory padding in the defined structures, letting the compiler manage the memory automatically.

	\item Use cryptographic hardware

	Atmel offers many solutions for secure memory and cryptographic operations chips, with serial interfaces for the IoT devices. We could benefit from this chips by storing the secret keys of the credentials inside the secure memory, and speed up the SHA256 and AES128 calculations, currently software implemented and slow compared to any hardware implementation.

	\item Support Extended APDUs

	Currently, the APDU instruction handlers use the GET RESPONSE command to send large APDU Response Payloads, but we can reduce the number of APDUs exchanged supporting the Extended APDU format. Because the P2ABCE already sent some Extended APDUs, the APDU I/O interface in the PoC has support for Extended APDU Commands, but not for Extended Responses.

	\item Integrate the IoT+Idemix solution in FIWARE

	Our sixth objective, that we couldn't get to design in depth. The original idea was to integrate in a FIWARE\footnote{FIWARE \url{https://www.fiware.org/}} environment the Fiware Keyrock and Idemix Issuer, so both, users and IoT devices' identities are managed in the IdM\footnote{Identity Management - KeyRock \url{https://catalogue.fiware.org/enablers/identity-management-keyrock}} system. The Issuer generates credentials based on the user’s attributes held in the Keyrock, and IoT devices can authenticate against Fi-ware services (using Idemix), or authenticate against other IoT devices (M2M).

	\begin{figure}[bth]
		\begin{center}
			\includegraphics[width=0.8\linewidth]{gfx/fiware}
		\end{center}
		\caption{IoT+Idemix Fi-Ware integration.}
		\label{fig:fiware}
	\end{figure}

	\item Research P2ABCE policies for IoT applications

	The next step, after integrating Idemix with an Identity Management system, should be to design more complex applications to IoT scenarios. We talked about a smart building in the introduction, but the straightforward proofs we can present depend on the signed values in the credentials. The Zero-Knowledge Proofs cryptography is already implemented, we only have to extend the protocols to benefit from them.

	For example, a combination of authentication and arbitrary proof could be achieved by a combined proof of knowledge of a signed credential (current authentication system) and said arbitrary proof, although the values used in it aren't actually signed by the Issuer, the first proof gives credibility.

	Another solution for the same problem, similar to a PKI scheme, the IoT device receives a signed credential including an attribute which is his own public key. The IoT device can issue new credentials with the parameters it needs for the arbitrary proofs. The Verifier then certifies the new credential is signed by the IoT device's public key.

\end{itemize}

\appendices

\section{Test: APDU Commands exchanged}\label{ch:resultsdiagrams}

In this appendix we provide 3 sequence diagrams highlighting the APDU Commands exchanged during our testbed.

The first one is the setup of the IoT smart card, storing the system parameters needed to work in the deployed P2ABCE system. The first Command is called \texttt{isAndroid} because the \texttt{HardwareSmartcard} class distinguishes Android phones acting as smart cards from MULTOS smart cards, in order to use Extended APDUs. Then configures the smart card, setting the PIN (``\texttt{1234}'' by default) and PUK. Finally, it copies the cryptographic system parameters to the smart card with multiple \texttt{SET} instructions.

The second image shows the issuance of the credential, divided in the three REST calls needed during the delegation. If we recall Idemix's issuance protocol, the User performed, during his first step, an exponentiation with his secret key and a ZKP, which can be identified by the \texttt{COMMITMENT} and \texttt{ISSUANCE RESPONSE} commands. During the second step, the User only has to verify the Issuer's ZKP and store the credential if the signature is valid. Because the P2ABC Engine can perform Verification of ZKP without the need of secret keys stored in the smart card, the APDU Commands in this phase are only for storing the credential information inside the IoT smart card's BLOB\footnote{\textbf{B}inary \textbf{L}arge \textbf{Ob}jects} database, with the \texttt{STORE BLOB} instruction.

The last diagram represents the proving for the Presentation Policy received from a Verifier, using two REST calls. During the first REST call, the Service stores data for the proving in the smart card, but before starting the proving itself, the device must choose an identity for the proving. This is done with the Identity Service, that in the PoC chooses the first identity available. After this, the second REST call starts the proving, creating the \textit{commitment} and \textit{responses} that integrate any ZKP, and depends on the device's secret keys.

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=0.9\linewidth]{gfx/UML/APDUsInitIoTSC}
	\end{center}
	\caption{Init IoT Smart Card APDU Commands exchanged.}
	\label{fig:APDUsInitIoTSC}
\end{figure}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=0.78\linewidth]{gfx/UML/IssuanceAPDUs}
	\end{center}
	\caption{Issuance APDU Commands.}
	\label{fig:IssuanceAPDUs}
\end{figure}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/UML/APDUsProving}
	\end{center}
	\caption{Proving APDU Commands.}
	\label{fig:APDUsProving}
\end{figure}

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

%\begin{IEEEbiography}{Michael Shell}
%Biography text here.
%\end{IEEEbiography}

% if you will not have a photo at all:
%\begin{IEEEbiographynophoto}{John Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

%\begin{IEEEbiographynophoto}{Jane Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2017-06-20 Doc/classicthesis/Chapters/Conclusions.tex
ieee
tercer borrador

@@
 Finally, we are very thankful to IBM's grant for the \textit{Privacy Preserving Identity Management applied to IoT} project, which allowed us to get this far.
-Also, this paper will be presented for publication in a JCR magazine.
+Also, in \autoref{ch:JCR} we show the version for publication in a JCR magazine.
 
@@
 	Our sixth objective, that we couldn't get to design in depth.
-The original idea was to integrate in a FIWARE\footnote{FIWARE \url{https://www.fiware.org/}} environment the Fiware Keyrock and Idemix Issuer, so both, users and IoT devices' identities are managed in the IdM\footnote{Identity Management - KeyRock \url{https://catalogue.fiware.org/enablers/identity-management-keyrock}} system.
+The original idea was to integrate in a FIWARE\footnote{FIWARE \url{https://www.fiware.org/}} environment the Fiware Keyrock and Idemix Issuer, so both users and IoT's identities are managed in the IdM\footnote{Identity Management - KeyRock \url{https://catalogue.fiware.org/enablers/identity-management-keyrock}} system.
 The Issuer generates credentials based on the user’s attributes held in the Keyrock, and IoT devices can authenticate against Fi-ware services (using Idemix), or authenticate against other IoT devices (M2M).

### 2017-06-20 Doc/classicthesis/Chapters/Introduccion.tex
ieee
tercer borrador

@@
 \url{https://en.wikipedia.org/wiki/2011_PlayStation_Network_outage}} afectó a su red de jugadores de consola, cuya información de facturación estaba guardada en sus servidores, y debido a que dicha información dependía tanto de los usuarios como de Sony para estar segura, aunque los usuarios la protegieran, se vieron comprometidos igualmente.
+
+In traditional M2M (Machine to Machine) environments the issues about security and privacy have already been treated deeply, but in the IoT world, due to it's recent and fast growth, lacks of those tools to solve autonomously these problems.
+
+En entornos tradicionales, los problemas de seguridad y privacidad ya han sido tratados en profundidad, pero en el mundo de IoT, debido a su reciente y rápido crecimiento, carece de esas herramientas para resolver estos problemas.
 
@@
 
-El objetivo de este proyecto es integrar Idemix con el Internet de las Cosas.
+Como parte de la beca académica de IBM para \textit{Administración de Identidad Preservando la Privacidad aplicada a Internet de las Cosas}, el objetivo de este proyecto es integrar Idemix con el Internet de las Cosas.
 Para ello nos apoyaremos en el proyecto europeo de ABC4Trust, \ac{P2ABCE}\footnote{P2ABCEngine \url{https://github.com/p2abcengine/p2abcengine}}, que define una arquitectura común, lenguaje y artefactos para unificar distintas soluciones como Idemix de IBM, o U-Prove de Microsoft.

### 2017-06-20 Doc/ieee/bare_jrnl.tex
tercer borrador

@@
 %
-%\usepackage{cite}
+\usepackage{cite}
 % cite.sty was written by Donald Arseneau
@@
 \ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
+  \usepackage[pdftex]{graphicx}
   % declare the path(s) where your graphic files are
@@
 %
-%\usepackage{amsmath}
+\usepackage{amsmath}
 % A popular package from the American Mathematical Society that provides
@@
 % http://www.ctan.org/pkg/algorithmicx
+
+\usepackage{listings}
+\lstset{
+	basicstyle=\ttfamily,
+	showstringspaces=false,
+	breaklines=true,
+	frame=single,
+	frameround=tttt,
+}
 
@@
 
+\usepackage[hidelinks]{hyperref}
+
 % *** Do not adjust lengths that control margins, column widths, etc. ***
@@
 
-\author{Jose Luis Canovas Sanchez, Antonio Skarmeta Gomez, Jorge Benal Bernabe}% <-this % stops a space
+%\author{Jose Luis Canovas Sanchez, Antonio Skarmeta Gomez, Jorge Benal Bernabe}% <-this % stops a space
+
+\author{\IEEEauthorblockN{Jose Luis Canovas Sanchez, Jorge Bernal Bernabe, Antonio F. Skarmeta}\\
+	\IEEEauthorblockA{Department of Information and Communications Engineering\\
+		Computer Science Faculty, 
+		University of Murcia, Spain\\
+		\{joseluiscanovas, jorgebernal, skarmeta\}@um.es\\
+	}
+}%
+
 %\thanks{M. Shell was with the Department
@@
 The IoT is a term with a wide range of interpretations \cite{Atzori20102787}, breafly, we can think of it as billions of devices, mainly resource constrained, that are interconnected between them, and the Internet, in order to achieve a goal.
-For example, a network of lampposts with proximity sensors that talk to each other so they light up part of the street when a passerby walks by, but save energy when not; greenhouses with automated irrigation to balance costs and quality vegetables; or cars exchanging traffic data to reduce city pollution from traffic jams or avoid accidents.
 
@@
 This includes not only the data shared with others, where one must trust they will keep it safe, but it's also the data collected about us and that we don't have direct control over it.
-The attack Sony suffered in 2011 to PSN\footnote{2011 PlayStation Network outage -
-	\url{https://en.wikipedia.org/wiki/2011_PlayStation_Network_outage}} is an example of the trust people had in Sony to store their billing information, and because the security of that information depended on both them and Sony, people found themselves compromised unable to do anything about it.
-
-With the proliferation of IoT devices gathering as much information as they can with their sensors, the amount of data gathered about anyone can be immense.
+
+In traditional M2M (Machine to Machine) environments the issues about security and privacy have already been treated deeply, but in the IoT world, due to it's recent and fast growth, lacks of those tools to solve autonomously these problems.
+
+With the proliferation of IoT devices gathering as much information as they can with their sensors, the amount of data stored about anyone can be immense.
 And IoT has proved to not address neither security nor privacy, with recent events like the Mirai botnet DDoS attack on October 2016, considered the biggest DDoS in history \cite{jeyanthi:2017}, or the multiple vulnerabilities affecting house devices, like baby monitors \cite{rapid7babycam}.
@@
 
-With usual symmetric and asymmetric cryptography it seems impossible to create such credentials, without an explosion of signatures over every possible combination.
-For this reason, current solutions rely on  ZKPs, cryptographic methods that allow to proof knowledge of some data without disclosing it.
+With usual symmetric and asymmetric cryptography it seems rather impossible to create such credentials, without an explosion of signatures over every possible combination.
+For this reason, current solutions rely on  Zero-Knowledge Proofs (ZKP), cryptographic methods that allow to proof knowledge of some data without disclosing it.
 
@@
 
-\subsection{Outline of this thesis}
-
-%This document is structured as follows: In \autoref{ch:stateoftheart} we show a state of the art analysis through the history of Idemix and related works, analysing what is of the most interest for the IoT perspective; then, in \autoref{ch:objectives} we outline the project's objectives and analyse in depth the existing solutions that we are going to deal with; in \autoref{ch:design} we describe the formal design of the IoT and Idemix solution, and describe the PoC implementation developed; after an implementation, it is a must to validate it, as showed during the performance tests in \autoref{ch:validation}; finally, our conclusions and lines for future work are described in \autoref{ch:conclusions}.
-
-\section{State of the art}
+This document is structured as follows: In \autoref{ch:stateoftheart} we show a state of the art analysis through the history of Idemix and related works, analysing what is of the most interest for the IoT perspective; in \autoref{ch:design} we describe the formal design of the IoT and Idemix solution, and describe the PoC implementation developed; after an implementation, it is a must to validate it, as showed during the performance tests in \autoref{ch:validation}; finally, our conclusions and lines for future work are described in \autoref{ch:conclusions}.
+
+\section{State of the art}\label{ch:stateoftheart}
 
@@
 
-\section{Design and implementation}
+\section{Design and implementation}\label{ch:design}
 
@@
 Some of them manage the smart card, like changing the PIN code, and many are P2ABCE specific, like storing the system cryptographic parameters.
-Some examples:
-
-	\begin{tabular}{|c|l|}
-		\hline
-		INS & Description \\
-		\hline
-		\texttt{0x52} & \texttt{SET\_CREDENTIAL} \\ & Sets a new credential inside the smart card.\\
-		\texttt{0x5A} & \texttt{GET\_CREDENTIAL\_PUBLIC\_KEY} \\ & Returns a stored credential's public key.\\
-		\texttt{0x60} & \texttt{GET\_PRESENTATION\_COMMITMENT} \\ & First cryptographic operation during Proving.\\
-		\texttt{0x62} & \texttt{GET\_PRESENTATION\_RESPONSE} \\ & Second cryptographic operation during Proving.\\
-		\hline		
-	\end{tabular}
 
@@
 In our delegation server we will run Raspbian OS, a distribution based on Debian for the Raspberry Pi.
-We will talk more about the hardware specifications in the benchmark chapter.
-We chose this system because it has a great package repository support, including the Java Runtime Environment, needed to run the P2ABCE Services.
-
-For our IoT device, we will use LEDE, Linux Embedded Development Environment, a distribution born from OpenWrt, aimed for routers and embedded chips with low resources requirements.
-It offers the \texttt{ash} command interpreter, or \texttt{shell}, and the \texttt{opkg} package manager, that allows an easy installation of some libraries and tools needed during the early development.
+For our IoT device, we will use LEDE, Linux Embedded Development Environment, a distribution forked from OpenWrt, aimed for routers and embedded chips with low resources requirements.
 
@@
 Any third party application that integrates P2ABCE with their system can make use of these services or implement the functionality using the library written in Java, the tool that the REST services actually use.
-
-Our PoC machine, the Omega2, can make REST calls easily with the \texttt{curl} command, but other devices may use \ac{CoAP}, but in that case, the P2ABCE REST services should be adapted to offer CoAP support.
-The commands needed to delegate to the P2ABCE server would be the same that those defined now to operate with the REST services, but with the modifications needed to pass the parameters on how to communicate with the IoT smart card.
 
@@
 
-In a real deployment, we would offer a full library with an API for other processes that want to use the P2ABC Engine, which automatizes the boot of the IoT smart card and handles the security policies, like the one we presented in the deployment diagram \ref{fig:P2ABCE-IoT} of the \textit{P2ABCE IoT Toolkit}.
-
-Now we show some examples of the PoC \texttt{curl} commands, run in the device's terminal, or from a shell script for automation:
-
-\begin{lstlisting}[language=bash]
-$ curl -X POST --header 'Content-Type: text/xml' "http://$DelegationServerIP:9200/user/initIoTsmartcard/http%3A%2F%2Fticketcompany%2FMyFavoriteSoccerTeam%2Fissuance%3Aidemix?host=192.168.3.1&port=8888"
-
-$ curl -X POST --header 'Content-Type: text/xml' -d @firstIssuanceMessage.xml "http://$DelegationServerIP:9200/user/issuanceProtocolStep/" > issuanceReturn.xml
-\end{lstlisting}
-
-The IoT device will only manage XML as data files to exchange between the third party actor and the P2ABCE delegation service, attaching them in the body of the REST call.
+%In a real deployment, we would offer a full library with an API for other processes that want to use the P2ABC Engine, which automatizes the boot of the IoT smart card and handles the security policies, like the one we presented in the deployment diagram \ref{fig:P2ABCE-IoT} of the \textit{P2ABCE IoT Toolkit}.
+
+%Now we show some examples of the PoC \texttt{curl} commands, run in the device's terminal, or from a shell script for automation:
+%
+%\begin{lstlisting}[language=bash]
+%$ curl -X POST --header 'Content-Type: text/xml' "http://$DelegationServerIP:9200/user/initIoTsmartcard/http%3A%2F%2Fticketcompany%2FMyFavoriteSoccerTeam%2Fissuance%3Aidemix?host=192.168.3.1&port=8888"
+%
+%$ curl -X POST --header 'Content-Type: text/xml' -d @firstIssuanceMessage.xml "http://$DelegationServerIP:9200/user/issuanceProtocolStep/" > issuanceReturn.xml
+%\end{lstlisting}
+
+%The IoT device will only manage XML as data files to exchange between the third party actor and the P2ABCE delegation service, attaching them in the body of the REST call.
 In the parameters of the first call we can see the IP and port where the IoT smart card will be listening.
@@
 
-To transmit the APDU messages in our PoC we use a simple protocol, that we will refer as BIOSC (Basic Input Output Smart Card), consisting in one first byte for the instruction:
-
-\hfil
-
-\begin{tabular}{|c|l|}
-	\hline
-	\texttt{0x01} & APDU Command or Response: Read 2 bytes for the\\& length and then as many bytes as said length. \\
-	\hline
-	\texttt{0xff} & Finish connection: close the current TCP socket and\\& end the IoT smart card execution. \\
-	\hline
-\end{tabular}
-
-\hfil
-
+To transmit the APDU messages in our PoC we use a simple protocol, that we will refer as BIOSC (Basic Input Output Smart Card), consisting in one first byte for the instruction.
 In the first instruction, we read two header bytes with the length of the APDU Command or Response to receive, followed by said APDU bytes.
@@
 
-As we mentioned in the analysis of P2ABCE's code, the \texttt{HardwareSmartcard} class implements the \texttt{Smartcard} interface using the package of abstract classes \texttt{javax.smartcardio} to communicate with the physical smart cards.
+After an analysis of P2ABCE's code, the \texttt{HardwareSmartcard} class implements the \texttt{Smartcard} interface using the package of abstract classes \texttt{javax.smartcardio} to communicate with the physical smart cards.
 The Oracle JRE implements these package for the majority of smart card manufacturers.
@@
 
-In the PoC IoT device, we implement in \texttt{BIOS.c} the \texttt{main()} function, that reads from a Json file the previous status of the smart card, and then opens the TCP socket in the $8888$ port, and loops listening for BIOSC transmissions until an $0xff$ is received.
-When an APDU Command is sent, it calls the APDU Handler, that we will talk about in the next section.
+In our PoC, we implement in \texttt{BIOS.c} the \texttt{main()} function, which reads from a Json file the status of the smart card (keys, credentials, ...), and then opens the TCP socket in the $8888$ port, and loops listening for BIOSC transmissions until an $0xff$ is received.
+When an APDU Command is sent, it calls the APDU Handler, that we will talk about below.
 
@@
 
-After many design decisions in the process to adapt the original ABC4Trust Card Lite code to pure C, working on a MIPS architecture, in this section, we present the current \ac{PoC} code, most important decisions taken and the execution workflow in the form of sequence diagrams.
+After many design decisions in the process to adapt the original ABC4Trust Card Lite code to pure C, working on a MIPS architecture, in this section, we present the current PoC code, most important decisions taken and the execution workflow in the form of sequence diagrams.
 
@@
 
-\paragraph{Core smart card}
+\textbf{Core smart card}
 \hfil
@@
 
-If the APDU protocol defined for P2ABCE smart cards changed, with the consequence of having to update all smart cards in use, the changes to adapt the IoT smart card should be applied in this part of the code.
-This also implies that if the P2ABCE Java code changes their \texttt{Smartcard} interface, something more probable, the IoT smart card will still work as intended.
-
-Most of the original ABC4Trust Card's code has been reused in this section, e.g., all the global variables and custom data type definitions, the APDU Command handling and auxiliary subroutines.
+Most of the original ABC4Trust Card Lite's code has been reused in this section, e.g., all the global variables and custom data type definitions, the APDU Command handling and auxiliary subroutines.
 
 However, the ABC4Trust's code heavily depended on the MULTOS platform for the input and output of data, modular arithmetic, memory management, and AES128 and SHA256 cryptographic operations.
-Therefore, we reimplement the used MULTOS functions following the documentation, adapting it in some cases.
-
-A characteristic of MULTOS C-API is that every function name starts with ``\texttt{multos}'', so we changing their names from \texttt{multosFoo()} to \texttt{mFoo()} for readability and to emphasize that they were no longer MULTOS functions.
-
-The \texttt{main.h} file, from the original ABC4Trust project, also implemented some functions that were equivalent to the ones available in \texttt{multos.h}.
-We replaced them for the standard functions in the C-API, checking that the MEL code did the same as the documentation specified.
-This last detail is very important, because in \citep{vullers2013efficient} they noted that MULTOS' \texttt{ModularExponentiation} function does not accept exponents larger than the modulus size, so they implemented \texttt{SpecialModularExponentiation}, dividing the exponentiation in two that MULTOS could perform.
-After this analysis, when those particularities appeared, our reimplementation of the MULTOS API accepts the \textit{expanded functionality}.
-In total, we reduced to eighteen the number of MULTOS functions to reimplement, and with future refactoring this number may decrease.
-
-\hfil
-
-Because of the inherited code, many particularities of the internals of the MULTOS platform had to be addressed.
-
-One was the different memory zones, where the variables in \texttt{static} were read and written directly from the secure EEPROM, defining the status of the smart card between executions.
-We keep those variables in RAM, but save them in a Json file, deserializing them at the smart card boot up, and serializing always before an APDU Response is sent.
-From the MULTOS documentation, the \texttt{static} variables are read and written atomically, and if power source is lost during any process, memory won't get corrupted.
-This also means that if a MULTOS smart card sends an APDU Response, every variable that had to change, is saved in the EEPROM.
-Our PoC uses Json for human readability during debugging, but a real deployment should apply secure reads and writes on the file, preventing data corruption and undesired access, for example, ciphering the file with a key derived from the smart card PIN.
-It is also desirable another data serialization method, reducing the size of the file to store in memory constrained devices.
-
-Another of the mentioned particularities is that the MULTOS compiler does not apply padding between variables in the data structures.
-For example, in an X86 machine, where a \texttt{short} is stored in 2 bytes, in the following code\footnote{From \url{https://en.wikipedia.org/wiki/Data_structure_alignment}}
-
-\begin{lstlisting}[language=C]
-struct MyData
-{
-short Data1;
-short Data2;
-short Data3;
-};
-\end{lstlisting}
-
-each member of the data structure would be 2-byte aligned.
-Data1 would be at offset 0, Data2 at offset 2, and Data3 at offset 4.
-The size of this structure would be 6 bytes.
-But if we defined the following structure
-
-\begin{lstlisting}[language=C]
-struct MixedData
-{
-char Data1;
-short Data2;
-int Data3;
-char Data4;
-};
-\end{lstlisting}
-
-where a \texttt{char} is 1 byte, and an \texttt{int} uses 4 bytes, although the total bytes used are 8 bytes, the compiler will align them in memory adding padding, like if we defined the following:
-
-\begin{lstlisting}[language=C]
-struct MixedData  /* After compilation in 32-bit x86 machine */
-{
-char Data1; /* 1 byte */
-char Padding1[1]; /* 1 byte for the following 'short' to be aligned on a 2 byte boundary
-assuming that the address where structure begins is an even number */
-short Data2; /* 2 bytes */
-int Data3;  /* 4 bytes - largest structure member */
-char Data4; /* 1 byte */
-char Padding2[3]; /* 3 bytes to make total size of the structure 12 bytes */
-};
-\end{lstlisting}
-
-\hfil
-
-In MULTOS, to reduce the memory usage, there is no padding, and this affects the inherited ABC4Trust's code because of the use of \texttt{memcpy} to copy zones of memory from one address to another.
+Therefore, we reimplement the used MULTOS functions following the documentation, adapting it in some cases for \textit{expanded functionality} when needed, like in \cite{vullers2013efficient} when they noted that MULTOS' \texttt{ModularExponentiation}  function does not accept exponents larger than the modulus size, so they implemented \texttt{SpecialModularExponentiation}.
+
+In MULTOS, to reduce the memory usage, the MULTOS compiler does not apply padding between variables in the data structures, i.e., data structure alignment\footnote{\url{https://en.wikipedia.org/wiki/Data_structure_alignment}}.
+This affects the inherited ABC4Trust's code because of the use of \texttt{memcpy} to copy zones of memory from one address to another.
 The problem is the code copies multiple variables in one \texttt{memcpy} call, because, for example, the APDU Command payload includes multiple data, and instead of copying one variable at a time, the \texttt{struct} is defined with the same order and copies everything at once.
-
-An example taken from the parsing of instruction \texttt{SET\_PROVER}:
-
-\begin{lstlisting}[language=C,frame=tblr]
-typedef struct
-{
-BYTE prover_id;
-unsigned int ksize;
-unsigned int csize;
-BYTE kx[MAX_SMALLINT_SIZE];
-BYTE c[HASH_SIZE];
-BYTE proofsession[PROOFSESSION_SIZE];
-BYTE proofstatus;
-BYTE cred_ids[NUM_CREDS];
-BYTE cred_ids_size; // also called 't' in the documentation
-BYTE exists;
-} PROVER;
-/*****/
-PROVER provers[NUM_PROVERS];
-
-/*****/
-case INS_SET_PROVER:
-// [...]
-memcpy(&(provers[temp_prover_id-1].prover_id), &(apdu_data.set_prover_in.prover_id), 5); // under the hood, this also initializes ksize and csize
-\end{lstlisting}
-This is the only case where a comment showcases that it is copying more than one byte.
 
@@
 
-\hfil
-
-An internal alarm should always warn us when we see raw memory copy from serialized data, that is, the APDU Payload.
-We mentioned in the MULTOS introduction that the platform is Big Endian.
-The serialized data is also typically represented in Big Endian, and our APDU Payloads do so.
-In our code we added a new Subroutine to check whether the device is Big or Little Endian, and to rectify the Endianness from those variables, with 2 or more bytes (the only ones affected by the Endianness), copied from or written to APDU Payloads.
-
-\hfil
-
-Finally, we must talk about the \texttt{multosExit} function.
-In a MULTOS application, to finish the execution and send the APDU Response, the programmer can call in any moment to \texttt{multosExit}, and the MULTOS OS will continue the execution, not returning to the application again.
-
-Think about how we all have, at least once, used an \texttt{if} and \texttt{return} to avoid the \texttt{else}, like in this example:
-
-\begin{lstlisting}[language=C]
-if(condition)
-return a;
-return b;
-\end{lstlisting}
-
-The use of \texttt{multosExit} in ABC4Trust code is similar, but used in almost every function.
-This leads to a sequence execution with many possible interruptions, leaving a full stack behind with \textit{un-returned} function calls, i.e., parameters stored in memory after a function call that never returns.
-In MULTOS, this memory won't affect future executions, because with the next APDU Command, the stack will be loaded with the clean application, and the Session memory stays untouched between APDU Commands, during an APDU Dialogue.
-
-To adapt the code, the reimplemented \texttt{mExit} function can not end the program execution, because we would lose the variables in RAM needed during the APDU Dialogue, as well as that the TCP socket would close.
-The solution is to refactor every use of \texttt{multosExit} in a way that, if it is called from a Subroutine, the function returns a success or error code, and the only functions that can call \texttt{mExit} are the ones handling an specific instruction.
-
-Nevertheless, there is one exception, the \texttt{output\_large\_data()} Subroutine, which is a tool to send large data payloads without Extended APDU support.
-We talked about this in \ref{subsec:APDU}, mentioning the special instruction \texttt{GET RESPONSE}.
-
-Summarizing, every Subroutine must return from its execution, as well as every instruction handler, which must call either \texttt{mExit} or \texttt{output\_large\_data()} before returning.
-After the instruction has been handled, the execution returns to the listening loop of BIOSC.
-
-\paragraph{Interfaces}\hfil
+\textbf{Interfaces}\hfil
 
@@
 
-Taking a step forward, we make the core smart card totally independent of any library, only dependant of our interfaces.
-This means that typical C libraries, like the standard \textit{stdlib.h}, or  \textit{string.h} are also behind the facade, in case some IoT system doesn't support them.
-The main goal we go after with this decision is that future developers adapting the code to a specific platform need to make no change to the \textit{core smart card}'s code, only to the interfaces implementation.
-
 The interfaces defined can be organized in 5 groups (see \autoref{fig:IoTCScomponents-bw}), depending on their purpose: Modular Arithmetic, Cryptography, Memory Management, Serialization, APDU Parsing.
 
-\paragraph{External utilities}\hfil
-
-If the IoT system offers well tested libraries that could aid in the interfaces implementation, for example, a assembly optimized code for the AES128 and SHA256 operations, these third party libraries belong to this section.
+\textbf{External utilities}\hfil
 
 In our PoC, we use two ANSI C libraries, for base64 and JSON, and two shared libraries available in as packages in LEDE, GMPLib and OpenSSL.
-These libraries use dynamic memory and offer more functionality than we need, although they are a good tool in this PoC, future versions should use more lightweight solutions.
-
-For example, Atmel's ATAES132A\footnote{ATAES132A 32K AES Serial EEPROM  Datasheet - \url{http://ww1.microchip.com/downloads/en/DeviceDoc/Atmel-8914-CryptoAuth-ATAES132A-Datasheet.pdf}}
-\marginpar{\includegraphics[width=0.9\linewidth]{gfx/atmel} \\ Atmel's cryptography chips.}
-offers a serial chip for secure key storage, AES128 execution and random number generation.
-Another serial chip like ESP8266 offers WiFi connectivity, typically used with Arduino, and can also perform AES encryption.
-For random number generation, a technique used with Contiki devices is to read from sensors aleatory data and use it as seed.
-
-All these alternatives depend on the target device, but are all valid, and would substitute OpenSSL with only a Serial communication library, so we can talk to the dedicated chips.
+These libraries use dynamic memory and offer more functionality than we need, although they are a good tool in this PoC, future versions should use more lightweight solutions, optimized fo the target device.
 
@@
 
-As we mentioned, every instruction handler must end, before the \texttt{return;} expression, with \texttt{mExit} or  \texttt{output\_large\_data()}, that will use \texttt{mExit}.
-This reimplemented MULTOS function will save the current status of the smart card, with the help of the serialization interface, and once the file is saved, it outputs the APDU Response back to the delegation server.
-
-After returning from \texttt{mExit}, the instruction handler and the APDU handler functions, the program listens again from the socket, until an $0xff$ byte is received.
-
-\subsection{Design and implementation conclusions}
-
-During this design and implementation process we had to study multiple technologies, including the ones mentioned in the state of the art, regarding Idemix, and those related to the IoT, we studied many systems (e.g.
-Contiki, Arduino) and hardware (e.g.
-MSP430, ESP8266), and after this, we had to take many decisions during development.
-
-The design is open in those aspects that depend on each specific project's requisites, and closed on the aspects involving the P2ABCE's core logic.
-
-However, the PoC had to make more concessions when implementing the cryptographic and modular arithmetic logic.
-But those are parts of the code that are more prone to be changed when developing for a new system.
-The core logic of the PoC is preserved as it is in every system and device.
-
-Even with the many points of future work left, we think we achieved the most extensible solution for Idemix+IoT at the time, our main objective of the project.
+Finally, every instruction handler must end, before the \texttt{return;} expression, with a call to \texttt{mExit}.
+This reimplemented MULTOS function will save the current status of the smart card and send the APDU Response back to the delegation server.
+
+After returning from the functions \texttt{mExit}, instruction handler and APDU handler, the program listens again from the socket.
 
@@
 
-In this chapter, we will describe the deployment of three testing scenarios: a laptop, a Raspberry Pi 3, and a Omega2 IoT device with the Raspberry Pi 3 as the delegation server.
+In this section, we will describe the deployment of three testing scenarios: a laptop, a Raspberry Pi 3, and a Omega2 IoT device with the Raspberry Pi 3 as the delegation server.
 We will measure and compare the results to determine if the proposed solution is feasible or must be submitted to revision.
@@
 First we will execute the test in our development machine (laptop).
-After asserting that the services work as expected, we then run the test in a Raspberry Pi 3, exactly like in the laptop.
-Finally, we will deploy the IoT smart card in a Omega2 and the delegation services in the Raspberry Pi 3.
+After asserting that the services work as expected, we then run the test in a Raspberry Pi 3\footnote{\url{https://www.raspberrypi.org/products/raspberry-pi-3-model-b/}}, exactly like in the laptop.
+Finally, we will deploy the IoT smart card in a Omega2\footnote{\url{https://docs.onion.io/omega2-docs/omega2.html}} and the delegation services in the Raspberry Pi 3.
 After every test, we checked that the issuing and proving were successful, in case a cryptographic error appeared in the implementation.
 
-Lets have a closer look at the hardware of each device:
-
-\paragraph{Development laptop}
-Our device is a DELL XPS 15, with a Core i7-6700HQ at 3.5GHz quad core processor and 32GB of DDR4 RAM, running Ubuntu 16.10.
-
-This is a powerful machine that can simulate the performance of many servers and clients that would implement P2ABCE in a real environment, giving a reference point for performance comparisons.
-
-\paragraph{Raspberry Pi 3} A familiar environment, powerful enough to debug and hold the delegated P2ABCE Java services of P2ABCE with its 1GB of RAM, and with two network interfaces, perfect to work as the gateway for the IoT devices to the Internet.
-
-Only a microSD with enough space to burn the OS is needed to plug\&play with the Raspberry Pi.
-We use Raspbian, a stable Debian based distro, recommended by the Raspberry Pi designers, and ready to use with the P2ABCE compiled \textit{self-contained .jar} services.
-
-\begin{table}[h]
-	\myfloatalign
-	\begin{tabular}{ll} \toprule
-		CPU & ARMv8 64bit quad-core @1.2GHz \\
-		RAM & 1GB \\
-		Storage & microSD \\
-		Firmware & Raspbian (Debian based distro) \\
-		Connectivity & Wifi n + Ethernet \\
-		Power & 5V 2A \\
-		\bottomrule
-	\end{tabular}
-	\caption[Raspberry Pi 3 Specifications]{Raspberry Pi 3 Specifications.}
-	\label{tab:RPi3Specs}
-\end{table}
-
-\paragraph{Onion Omega2} A device that falls inside the category of IoT, powerful enough to run a Linux embedded environment, such as LEDE, where we can develop and debug the first PoC without troubling ourselves with problems not related to the project itself.
-
-\begin{table}[h]
-	\myfloatalign
-	\begin{tabular}{ll} \toprule
-		MCU & Mediatek MT688\footnote{MediaTek MT7688 Datasheet - \url{https://goo.gl/kqD2gF}} \\
-		CPU & MIPS32 24KEc 580MHz \\
-		RAM & 64MB \\
-		Storage & 16MB \\
-		Firmware & LEDE (OpenWRT fork distro) \\
-		Connectivity & Wifi b/g/n \\
-		Power & 3.3V 300mA \\
-		\bottomrule
-	\end{tabular}
-	\caption[Onion Omega 2 Specifications]{Onion Omega2 Specifications.}
-	\label{tab:Omega2Specs}
-\end{table}
-
-Nonetheless, the Omega2 needs fine tunning to start operating, and basic knowledge of electronics to make it work. The two main things to begin with Omega2 are:
-\begin{itemize}
-	\item A reliable 3.3V with a maximum of 800mA power supply, e.g.
-a USB2.0 with a step-down circuit, with quality soldering and wires to avoid unwanted resistances.
-The Omega2 will usually use up to 350mA, when the WiFi module is booting up.
-The mean consumption is about 200mA.
-	\item A Serial to USB adapter wired to the TX and RX UART pins to use the Serial Terminal, to avoid the use of SSH over WiFi.
-\end{itemize}
-
-A downside of using LEDE and the Omega2 is the lack of hardware acceleration for cryptographic operations, unlike the MULTOS applications, because the smart cards hardware and MULTOS API include support for such common operations in smart cards.
+A downside of using the Omega2 is the lack of hardware acceleration for cryptographic operations, unlike the MULTOS applications, because the smart cards hardware and MULTOS API include support for such common operations in smart cards.
 
@@
 To ensure the delay wasn't significant, we measured 6000 APDU messages, and the results show that the mean transmission time is less than half a millisecond per APDU.
-As we will see in the results section, this network time is negligible.
-
-\paragraph{Future work for tests} The lack of a physical MULTOS smart card precludes us to load and test the ABC4Trust Card Lite's code and measure the time P2ABCE would need when using the \textit{HardwareSmartcard} class.
-This would be really interesting because for a single method from the \textit{Smartcard} interface, \textit{HardwareSmartcard} implementation needs to send multiple APDU Commands, but \textit{SoftwareSmartcard} can perform the operations immediately, with the full computer's resources.
-Also, physical smart cards benefit from hardware acceleration in most of the cryptographic operations, unlike our software implementations in the PoC.
-
-\subsection{The test code}
-
-In this section we present the scripts and binaries used during the tests.
-
-There are three pieces of software that conform the test: the P2ABCE services, the IoT smart card, and shell scripts automatizing the REST calls, from the terminal.
-
-\paragraph{P2ABCE services}\hfil
-
-This is a common part to our three sets.
-The services are compiled in a self-contained Jetty web server, or in WAR format, ready to be deployed in a server like Tomcat.
-We use the same JAR files with the embedded Jetty web server for the PC and Raspberry Pi.
-
-We modified the User Service code to measure the execution elapsed time for each REST method.
-We don't measure the time the web server spends processing the HTTP protocol and deciding which Java method to call.
-
-\paragraph{IoT smart card}\hfil
-
-Our C implementation of the P2ABCE smart card, compiled for the Omega2, with BIOSC listening on port 8888 for the APDU messages.
-
-We tested the execution in two modes, a full logging where every step was printed in terminal, and another one with no logging.
-With the first mode, we can check a proper execution, every byte exchanged, and with the second one, we measure the execution without unnecessary I/O.
-
-\paragraph{Shell scripts} \hfil
-
-To orchestrate all the services we use shell scripts that execute the REST calls using  \texttt{curl}.
-Here we perform the mentioned steps: setup of the P2ABCE system, issuance of the credential and proving for the presentation policy.
-
-In the setup, the system parameters are generated, indicating key sizes of 1024 bits, the ones currently supported by the ABC4Trust and IoT smart cards.
-Then the system parameters and public keys from the services (issuer, inspector, revocation authority) are also exchanged and stored in each Service.
-
-During the issuance and proving, the User and the Issuer and Verifier exchange multiple XML files, with the cryptographic information of each step of the Idemix protocol.
-
-We offer two versions for the scripts.
-The first one is a single shell script file, to be executed from a single terminal.
-The advantage of this version is that, because every XML file is stored in the machine running the script, we don't have to transfer these files by other means, e.g.
-\texttt{scp}.
-An excerpt of this script showcases how we receive an XML file from the User Service, e.g.
-\texttt{secondIssuanceMessage.xml}, and send it to the Issuer Service, referencing the file stored in the script's working directory:
-
-\begin{lstlisting}[language=bash]
-[...]
-# First issuance protocol step - UI (first step for the user).
-echo "Second issuance protocol step (first step for the user)"
-curl -X POST --header 'Content-Type: text/xml' -d @uiIssuanceReturn.xml 'http://localhost:9200/user/issuanceProtocolStepUi/' > secondIssuanceMessage.xml
-
-# Second issuance protocol step (second step for the issuer).
-echo "Second issuance protocol step (second step for the issuer)"
-curl -X POST --header 'Content-Type: text/xml' -d @secondIssuanceMessage.xml 'http://localhost:9100/issuer/issuanceProtocolStep/' > thirdIssuanceMessageAndBoolean.xml
-[...]
-\end{lstlisting}
-
-The second script version is intended for a more realistic execution.
-In a real scenario, the User and the Issuer, for example, would call the REST methods of their respective Services, but not the other actor's Service, exchanging the XML files through other means.
-Instead of one script running in a single machine, orchestrating the Omega2, Raspberry Pi 3 and laptop, we offer three scripts, one per machine, with \textit{pauses} where one device must send an XML file to another machine.
-An excerpt from the Omega2's script shows the same step as above, but instead of calling the Issuer's REST Service, waits until we send the corresponding files.
-
-\begin{lstlisting}[language=bash]	
-# First issuance protocol step - UI (first step for the user).
-echo "Second issuance protocol step (first step for the user)"
-curl -X POST --header 'Content-Type: text/xml' -d @uiIssuanceReturn.xml "http://$Rpi3IP:9200/user/issuanceProtocolStepUi/" > secondIssuanceMessage.xml
-
-read -p "Send => secondIssuanceMessage.xml <= back to the Issuer. Then press any key to continue... " -n1 -s
-
-read -p "Receive => thirdIssuanceMessageAndBoolean.xml <= from the Issuer. Then press any key to continue... " -n1 -s
-\end{lstlisting}
-
-\hfil
-
-The first version is recommended for tests, and the second version only for instructive purposes.
+From our analysis of the tests we conclude that the network time is negligible compared to the rest of the computations.
 
@@
 \begin{figure}[bth]
-	\myfloatalign
-	\subfloat[Times and relative speedup]
-	{\includegraphics[width=\linewidth]{gfx/graphics/setuptable}} \quad
-	\subfloat[Comparison graph]
-	{\includegraphics[width=.8\linewidth]{gfx/graphics/setup}} \\
-	\caption{Setup times (milliseconds)}
+	\includegraphics[width=\linewidth]{gfx/graphics/setuptable}
+	\caption{Setup times (milliseconds). Times and relative speedup}
 	\label{fig:setup:graph}
@@
 
+%\begin{figure}[bth]
+%	\caption{Setup times (milliseconds). Comparison graph}
+%	\includegraphics[width=0.8\linewidth]{gfx/graphics/setup}
+%	\label{fig:setup:graph2}
+%\end{figure}
+
 As we can see in \autoref{fig:setup:graph}, the laptop is about ten times faster than Raspberry Pi 3, but considering that the highest time is less than two and a half seconds, and that the setup is done only once, this isn't a worrisome problem.
@@
 \begin{figure}[bth]
-	\myfloatalign
-	\subfloat[Times and relative speedup]
-	{\includegraphics[width=0.5\linewidth]{gfx/graphics/createSCtable}} \quad
-	\subfloat[Comparison graph]
-	{\includegraphics[width=0.45\linewidth]{gfx/graphics/createSC}} \\
-	\caption{Create smart card times (seconds)}
+	\caption{Create smart card times (seconds). Times and relative speedup}
+	\includegraphics[width=0.7\linewidth]{gfx/graphics/createSCtable}
 	\label{fig:createSmartCard:graph}
 \end{figure}
+
+%\begin{figure}[bth]
+%	\includegraphics[width=0.8\linewidth]{gfx/graphics/createSC}
+%	\caption{Create smart card times (seconds). Comparison graph}
+%	\label{fig:createSmartCard:graph2}
+%\end{figure}
 
@@
 This is the first interaction between the RPi3 and the IoT smart card running in the Omega2.
-To setup the smart card \textbf{30 APDU Commands}, and their respective Responses, are exchanged, as shown in \autoref{ch:resultsdiagrams}, \autoref{fig:APDUsInitIoTSC}, with a total of $1109$ bytes.
+To setup the smart card 30 APDU Commands, and their respective Responses, are exchanged, with a total of $1109$ bytes.
 From our network benchmark, using TCP sockets, the delay in the transmission is only around 15 and 20 ms, negligible, as we said, compared to the almost 20 seconds the operation lasts.
@@
 Our credential will have 5 attributes and key sizes of 1024 bits, as specified during the setup process.
-
-The issuance is done in three steps for the User service, shown in \autoref{fig:IssuanceInteraction} with a red note showing the start of each step for the User delegation, in green for the interactions between Issuer and User, and the darker red is the Identity service, choosing the first available identity to use.
-The arrows in the figure show the REST calls performed during the test, where the IoT device acts as User, the RPi3 hosts the P2ABCE delegation services and the laptop is the Issuer.
 
@@
 
+%\begin{figure}[bth]
+%	\begin{center}
+%		\includegraphics[width=\linewidth]{gfx/UML/IssuanceInteraction}
+%	\end{center}
+%	\caption{Issuance interaction.}
+%	\label{fig:IssuanceInteraction}
+%\end{figure}
+
+As we can see, the three REST calls to the delegation User service involve communication with the smart card.
+There are 45 APDU Commands in total, $3197$ bytes exchanged, that would introduce a latency of 45ms in the network, negligible.
+
 \begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/UML/IssuanceInteraction}
-	\end{center}
-	\caption{Issuance interaction.}
-	\label{fig:IssuanceInteraction}
-\end{figure}
-
-As we can see, the three REST calls to the delegation User service involve communication with the smart card.
-We show in \autoref{ch:resultsdiagrams}, \autoref{fig:IssuanceAPDUs}, the APDU Commands used for each REST call in the issuance.
-There are 45 APDU Commands in total, $3197$ bytes exchanged, that would introduce a latency of 45ms in the network, negligible.
-
-\begin{figure}[bth]
-	\myfloatalign
-	\subfloat[Times (ms) and relative speedup]
-	{\includegraphics[width=\linewidth]{gfx/graphics/issuancetable}} \quad
-	\subfloat[Comparison graph]
-	{\includegraphics[width=.8\linewidth]{gfx/graphics/issuance}} \\
-	\caption{Issuance times (milliseconds)}
+	\includegraphics[width=\linewidth]{gfx/graphics/issuancetable}
+	\caption{Issuance times (milliseconds). Times (ms) and relative speedup}
 	\label{fig:issuance:graph}
 \end{figure}
+
+%\begin{figure}[bth]
+%	\includegraphics[width=\linewidth]{gfx/graphics/issuance}
+%	\caption{Issuance times (milliseconds). Comparison graph}
+%	\label{fig:issuance:graph2}
+%\end{figure}
 
@@
 The final step of the test involves a Prove, or Presentation in P2ABCE, where the Verifier sends the User or Prover the Presentation Policy, and the User answers with the Presentation Token, without more steps.
-In \autoref{fig:ProvingInteraction}, using the same colors as in the Issuance interaction, we can see the delegation messages done by the Omega2.
+%In \autoref{fig:ProvingInteraction}, using the same colors as in the Issuance interaction, we can see the delegation messages done by the Omega2.
 
@@
 
-In \autoref{ch:resultsdiagrams}, \autoref{fig:APDUsProving}, we provide the APDU Commands for each step, $28$ in total, with $1939$ bytes, giving us about $27$ms of delay in the network transmission.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/UML/ProvingInteraction}
-	\end{center}
-	\caption{Proving interaction.}
-	\label{fig:ProvingInteraction}
-\end{figure}
+The APDU Commands and Responses pairs sent are $28$ in total, with $1939$ bytes, giving us about $27$ms of delay in the network transmission.
+
+%\begin{figure}[bth]
+%	\begin{center}
+%		\includegraphics[width=\linewidth]{gfx/UML/ProvingInteraction}
+%	\end{center}
+%	\caption{Proving interaction.}
+%	\label{fig:ProvingInteraction}
+%\end{figure}
 
@@
 \begin{figure}[bth]
-	\myfloatalign
-	\subfloat[Times (ms) and relative speedup]
-	{\includegraphics[width=.8\linewidth]{gfx/graphics/provingtable}} \quad
-	\subfloat[Comparison graph]
-	{\includegraphics[width=.7\linewidth]{gfx/graphics/proving}} \\
-	\caption{Proving times (milliseconds)}
+	\includegraphics[width=.8\linewidth]{gfx/graphics/provingtable}
+	\caption{Proving times (milliseconds). Times (ms) and relative speedup}
 	\label{fig:proving:graph}
 \end{figure}
+
+%\begin{figure}[bth]
+%	\includegraphics[width=\linewidth]{gfx/graphics/proving}
+%	\caption{Proving times (milliseconds). Comparison graph}
+%	\label{fig:proving:graph2}
+%\end{figure}
 
@@
 In the implementation chapter we give guidelines to port MULTOS applications, considering the particularities we encountered, that's why we can't consider it an \textit{instruction manual to port MULTOS apps}, but an interesting reading on how to confront a similar project.
-For example, if we wanted the Idemix implementation from \citep{vullers2013efficient}, previous to P2ABCE, in a IoT device, we could apply almost the same steps, obtaining a core functionality of Idemix for constrained devices in C.
-
-Our PoC implementation demonstrates that this project is actually feasible, not by performing a simulation of an IoT device, like in \citep{vanet}.
+For example, if we wanted the Idemix implementation from \cite{vullers2013efficient}, previous to P2ABCE, in a IoT device, we could apply almost the same steps, obtaining a core functionality of Idemix for constrained devices in C.
+
+Our PoC implementation demonstrates that this project is actually feasible, not by performing a simulation of an IoT device, like in \cite{vanet}.
 However, the use of third party libraries and no hardware acceleration support, makes the PoC too slow for certain cases, e.g., a Real-Time system requires almost immediate operations, like a car warning that another one is approaching too fast.
@@
 
-Recalling the objectives listed in \autoref{objectives:section}, we think we accomplish them, except the implementation of a PoC in the most constrained device possible, where we used LEDE to ease our development, and the last objective, where we could have given a use to the PoC, and we leave as part of the future work.
+%Recalling the objectives listed in \autoref{objectives:section}, we think we accomplish them, except the implementation of a PoC in the most constrained device possible, where we used LEDE to ease our development, and the last objective, where we could have given a use to the PoC, and we leave as part of the future work.
 In spite of this negative auto-critic, we have to acknowledge that this is the first privacy-preserving ABC implementation for IoT, designed for extensibility, interoperability and maintenance.
@@
 Finally, we are very thankful to IBM's grant for the \textit{Privacy Preserving Identity Management applied to IoT} project, which allowed us to get this far.
-Also, in \autoref{ch:JCR} we show the version for publication in a JCR magazine.
 
@@
 \begin{itemize}
-	\item Second PoC for Arduino systems
+	\item Second PoC for more constrained systems
 
 	We have already developed a working PoC for Linux based systems, and from our objective of using the most constrained devices available, we can follow an iterative process, building for devices a little more constrained than the last one.
-The next natural step, after building with LEDE/OpenWrt, is to aim for Arduino.
-
-	The critical changes to the actual PoC are: 
-	\subitem The transmission channel, if the device has no TCP/IP stack, it could use a Serial connection to transmit BIOSC and delegation messages.
-	\subitem The use of a Json file to store the smart card status depends on whether the device has enough space or a file system.
-The current Json parser also uses too much dynamic memory.
-At this stage the serialization should be performed in a binary format, in a file or writing variables directly to the EEPROM memory.
-	\subitem GMPLib and OpenSSL libraries must be replaced by others with similar capabilities, or special implementations for the particular device in use.
-
-	All these changes affect only the \textit{external utilities} and \textit{interfaces} sections of the code, leaving the \textit{core} untouched, because the smart card logic doesn't change.
+The next natural step, after building with LEDE/OpenWrt, would be aiming for Arduino.
 
@@
 
-	\item Improve the \textit{core smart card} code
-
-	From the simple refactoring of removing the MULTOS reimplemented API for direct calls to the \textit{interfaces}, to identify all the \texttt{memcpy} copies of multiple variables with one call, removing the compiler directive for not using memory padding in the defined structures, letting the compiler manage the memory automatically.
-
 	\item Use cryptographic hardware
@@
 
-	\item Integrate the IoT+Idemix solution in FIWARE
-
-	Our sixth objective, that we couldn't get to design in depth.
-The original idea was to integrate in a FIWARE\footnote{FIWARE \url{https://www.fiware.org/}} environment the Fiware Keyrock and Idemix Issuer, so both, users and IoT devices' identities are managed in the IdM\footnote{Identity Management - KeyRock \url{https://catalogue.fiware.org/enablers/identity-management-keyrock}} system.
+	\item Integrate the IoT+Idemix solution in FIWARE\footnote{\url{https://www.fiware.org/}}
+
+	The idea is to integrate inside the Fiware Keyrock an Idemix Issuer, so both users and IoT's identities are managed in the IdM\footnote{Identity Management - KeyRock \url{https://catalogue.fiware.org/enablers/identity-management-keyrock}} system.
 The Issuer generates credentials based on the user’s attributes held in the Keyrock, and IoT devices can authenticate against Fi-ware services (using Idemix), or authenticate against other IoT devices (M2M).
@@
 		\begin{center}
-			\includegraphics[width=0.8\linewidth]{gfx/fiware}
+			\includegraphics[width=\linewidth]{gfx/fiware}
+			\caption{IoT+Idemix Fi-Ware integration.}
 		\end{center}
-		\caption{IoT+Idemix Fi-Ware integration.}
 		\label{fig:fiware}
@@
 \end{itemize}
-
-\appendices
-
-\section{Test: APDU Commands exchanged}\label{ch:resultsdiagrams}
-
-In this appendix we provide 3 sequence diagrams highlighting the APDU Commands exchanged during our testbed.
-
-The first one is the setup of the IoT smart card, storing the system parameters needed to work in the deployed P2ABCE system.
-The first Command is called \texttt{isAndroid} because the \texttt{HardwareSmartcard} class distinguishes Android phones acting as smart cards from MULTOS smart cards, in order to use Extended APDUs.
-Then configures the smart card, setting the PIN (``\texttt{1234}'' by default) and PUK.
-Finally, it copies the cryptographic system parameters to the smart card with multiple \texttt{SET} instructions.
-
-The second image shows the issuance of the credential, divided in the three REST calls needed during the delegation.
-If we recall Idemix's issuance protocol, the User performed, during his first step, an exponentiation with his secret key and a ZKP, which can be identified by the \texttt{COMMITMENT} and \texttt{ISSUANCE RESPONSE} commands.
-During the second step, the User only has to verify the Issuer's ZKP and store the credential if the signature is valid.
-Because the P2ABC Engine can perform Verification of ZKP without the need of secret keys stored in the smart card, the APDU Commands in this phase are only for storing the credential information inside the IoT smart card's BLOB\footnote{\textbf{B}inary \textbf{L}arge \textbf{Ob}jects} database, with the \texttt{STORE BLOB} instruction.
-
-The last diagram represents the proving for the Presentation Policy received from a Verifier, using two REST calls.
-During the first REST call, the Service stores data for the proving in the smart card, but before starting the proving itself, the device must choose an identity for the proving.
-This is done with the Identity Service, that in the PoC chooses the first identity available.
-After this, the second REST call starts the proving, creating the \textit{commitment} and \textit{responses} that integrate any ZKP, and depends on the device's secret keys.
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=0.9\linewidth]{gfx/UML/APDUsInitIoTSC}
-	\end{center}
-	\caption{Init IoT Smart Card APDU Commands exchanged.}
-	\label{fig:APDUsInitIoTSC}
-\end{figure}
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=0.78\linewidth]{gfx/UML/IssuanceAPDUs}
-	\end{center}
-	\caption{Issuance APDU Commands.}
-	\label{fig:IssuanceAPDUs}
-\end{figure}
-
-\begin{figure}[bth]
-	\begin{center}
-		\includegraphics[width=\linewidth]{gfx/UML/APDUsProving}
-	\end{center}
-	\caption{Proving APDU Commands.}
-	\label{fig:APDUsProving}
-\end{figure}
 
@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
-
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-\end{thebibliography}
+\bibliography{Bibliography}
+\bibliographystyle{plain}
 

### 2017-06-21 Doc/ieee/bare_jrnl.tex
reduce ieee future work

@@
 
-Due to the nature of the project, there exist many options to continue researching in this area. We list below a collection of the tasks we couldn't accomplish but would have been our next steps:
-
-\begin{itemize}
-	\item Second PoC for more constrained systems
-
-	We have already developed a working PoC for Linux based systems, and from our objective of using the most constrained devices available, we can follow an iterative process, building for devices a little more constrained than the last one.
-The next natural step, after building with LEDE/OpenWrt, would be aiming for Arduino.
-
-	\item Implement more P2ABCE functionality in the device
-
-	If instead of aiming for more constrained devices, we work with similar devices to the Omega2, the computational power of the device could assume all the functionality of P2ABCE, at least for some roles.
-In a M2M (machine to machine) environment, the devices could act as Provers and Verifiers in the same interaction, and the verifying logic could be implemented entirely in the device (generate a Presentation Policy and checking a prove in a Presentation Token is valid), reducing the need of a delegation server.
-
-	\item Use cryptographic hardware
-
-	Atmel offers many solutions for secure memory and cryptographic operations chips, with serial interfaces for the IoT devices.
-We could benefit from this chips by storing the secret keys of the credentials inside the secure memory, and speed up the SHA256 and AES128 calculations, currently software implemented and slow compared to any hardware implementation.
-
-	\item Support Extended APDUs
-
-	Currently, the APDU instruction handlers use the GET RESPONSE command to send large APDU Response Payloads, but we can reduce the number of APDUs exchanged supporting the Extended APDU format.
-Because the P2ABCE already sent some Extended APDUs, the APDU I/O interface in the PoC has support for Extended APDU Commands, but not for Extended Responses.
-
-	\item Integrate the IoT+Idemix solution in FIWARE\footnote{\url{https://www.fiware.org/}}
-
-	The idea is to integrate inside the Fiware Keyrock an Idemix Issuer, so both users and IoT's identities are managed in the IdM\footnote{Identity Management - KeyRock \url{https://catalogue.fiware.org/enablers/identity-management-keyrock}} system.
-The Issuer generates credentials based on the user’s attributes held in the Keyrock, and IoT devices can authenticate against Fi-ware services (using Idemix), or authenticate against other IoT devices (M2M).
+Due to the nature of the project, there exist many options to continue researching in this area.
+
+In the implementation side, we would continue with a second PoC which aimed for more constrained systems, maybe starting with Arduino, or another line of development would be to implement more P2ABCE functionality inside the IoT device.
+It would be interesting to compare the execution of the current pure software PoC to that with cryptographic hardware support, e.g.
+using Atmel's chips for SHA, AES and secure memory.
+
+On the design aspect, now that an IoT device can perform P2ABCE operations, we should integrate it in a bigger project, like inside the IdM\footnote{Identity Management - KeyRock \url{https://catalogue.fiware.org/enablers/identity-management-keyrock}} system of FIWARE\footnote{\url{https://www.fiware.org/}}, to issue Idemix credentials to both users and IoT devices.
+And once the identification is achieved in a privacy-preserving fashion, more applications should be researched, like how the IoT device could benefit from the ZKP cryptography to prove information not issued in its credential, by expanding the P2ABCE presentation policies.
 
@@
 \end{figure}
-
-	\item Research P2ABCE policies for IoT applications
-
-	The next step, after integrating Idemix with an Identity Management system, should be to design more complex applications to IoT scenarios.
-We talked about a smart building in the introduction, but the straightforward proofs we can present depend on the signed values in the credentials.
-The Zero-Knowledge Proofs cryptography is already implemented, we only have to extend the protocols to benefit from them.
-
-	For example, a combination of authentication and arbitrary proof could be achieved by a combined proof of knowledge of a signed credential (current authentication system) and said arbitrary proof, although the values used in it aren't actually signed by the Issuer, the first proof gives credibility.
-
-	Another solution for the same problem, similar to a PKI scheme, the IoT device receives a signed credential including an attribute which is his own public key.
-The IoT device can issue new credentials with the parameters it needs for the arbitrary proofs.
-The Verifier then certifies the new credential is signed by the IoT device's public key.
-
-\end{itemize}
 

### 2017-07-24 Doc/classicthesis/Chapters/Introduction.tex
tercer borrador
fix

@@
 \url{https://en.wikipedia.org/wiki/2011_PlayStation_Network_outage}} is an example of the trust people had in Sony to store their billing information, and because the security of that information depended on both them and Sony, people found themselves compromised unable to do anything about it.
+
+In traditional M2M (Machine to Machine) environments the issues about security and privacy have already been treated deeply, but in the IoT world, due to it's recent and fast growth, lacks of those tools to solve autonomously these problems.
 
@@
 
-	Victor then asks to repeat the experiment. With $20$ repetitions, the chances Peggy fools Victor in all of them is only  $2^{-20}$,
+	Victor then asks to repeat the experiment. With $20$ repetitions, the chances Peggy fools Victor in all of them is only  $2^{-20}$.
 	\marginpar{\includegraphics[width=1.\linewidth]{gfx/graficoJL_ZKP_3}\\The cave. Peggy returns by the requested path.}
